### 2020-11-13 lab_2.Rmd
adding variables to use

---
title: "Lab 2: Regression to Study the Spread of Covid-19"
author: 'w203: Statistics for Data Science'
date: "10/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r lib,include=FALSE}
library(lmtest)
library(sandwich)
library(tidyverse)
library(magrittr)

```

# Introduction

In this lab, you will apply what you are learning about linear regression to study the spread of COVID-19.  Your task is to select a research question, then conduct a regression study to analyze it.

Your research question must focus attention on a specific measurement goal.  It is not enough to say "we are looking for policies that help stop COVID-19" (That would be a fishing expedition).  Instead, use your introduction to motivate a specific effect for measurement.

Once you have a measurement goal, you will build a set of linear models that are tailored to that goal, and display them in a well formatted regression table.  You will finally use your conclusion to distill important conclusions from your estimates.

This is a group assignment.  Your live session instructor will coordinate the formation of groups. We would like to encourage teams to focus on using the lab as a way to learn how to work as a team of collaborating data scientists on shared code; how to clean and organize data; and, how to present work in a compelling way. As a result, we encourage teams to allow individuals to take risks and be supportive in the face of successes and failures. Create an opportunity for people who want to improve a particular skill to do so -- this might be project coordination, management of code through git, plotting, or any of the many aspects that you'll work on. *We hope that you can support and learn from one another through this team-based project.*

# Deliverables

| Deliverable Name       | Week Due | Grade Weight |  
|------------------------|----------|--------------| 
| [Draft Report]         | Week 12  | 10%          | 
| [Within-Team Review]   | Week 12  | 5%           | 
| [Cross-Team Review]    | Week 13  | 10%          | 
| [Final Report]         | Week 14  | 65%          | 
| [Final Presentation]   | Week 14  | 10%          |

# The Data

The data is provided in a file within this repository.  Majid Maki-Nayeri compiled the data, drawing many variables from the COVID-19 US state policy database (Raifman J, Nocka K, Jones D, Bor J, Lipson S, Jay J, and Chan P.). Supporting documents like the specific legal language used by states, additional data sources, and much more are available in unstructured format [here](https://tinyurl.com/statepolicysources).

The dataset includes:

1. Variables representing the spread of the disease; 
2. Variables representing state-level policy responses; and, 
3. General state-level characteristics.

If you want to, you are allowed to add extra variables from external sources.  However, this is not necessary, and we will not assign any bonus points to teams that derive unique data, as our focus is on statistics and statistical writing.

```{r variables}
df <- read.csv("covid-19.csv", header = TRUE)

df<-df%>%
  rename(case_rate_100k = 'Case.Rate.per.100000',
         population_density = 'Population.density.per.square.miles',
         white_pct = 'White...of.Cases',
         black_pct = 'Black...of.Cases',
         hispanic_pct = 'Hispanic...of.Cases',
         other_pct = 'Other...of.Cases',
         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
         mask_legal='No.legal.enforcement.of.face.mask.mandate') %>%
  select(case_rate_100k, population_density, white_pct, black_pct, hispanic_pct, 
         other_pct, mask_public, mask_legal)
head(df)
```

# Final Project Components {.tabset}

## Draft Report

In the first stage of the lab, you will create a draft report.  You should aim to make this report as complete as possible, so that you get the best possible feedback from the Cross-Team Review.

The exact format of your report is flexible, but it should include the following elements.

### 1. An Introduction

Your introduction should present a research question and explain the concept that you're attempting to measure and how it will be operationalized.  This section should pave the way for the body of the report, preparing the reader to understand why the models are constructed the way that they are.  It is not enough to simply say "We are looking for policies that help against COVID"  Your introduction must do work for you, focusing the reader on a specific measurement goal, making them care about it, and propelling the narrative forward.  This is also good time to put your work into context, discuss cross-cutting issues, and assess the overall appropriateness of the data.

### 2. A Model Building Process

You will next build a set of models to investigate your research question, documenting your decisions.  Here are some things to keep in mind during your model building process:

1. *What do you want to measure*?  Make sure you identify one, or a few, variables that will allow you to derive conclusions relevant to your research question, and include those variables in all model specifications.
2. Is your modeling goal one of description or explanation? 
3. What [covariates](https://en.wikipedia.org/wiki/Dependent_and_independent_variables#Statistics_synonyms) help you achieve your modeling goals?  What covariates are problematic, either due to *collinearity*, or because they are outcomes that will absorb some of a causal effect you want to measure?
4. What *transformations*, if any, should you apply to each variable?  These transformations might reveal linearities in scatterplots, make your results relevant, or help you meet model assumptions.
5. Are your choices supported by exploratory data analysis (*EDA*)?  You will likely start with some general EDA to *detect anomalies* (missing values, top-coded variables, etc.).  From then on, your EDA should be interspersed with your model building.  Use visual tools to *guide* your decisions.  You can also leverage statistical *tests* to help assess whether variables, or groups of variables, are improving model fit.

At the same time, it is important to remember that you are not trying to create one perfect model.  You will create several specifications, giving the reader a sense of how robust (or sensitive) your results are to modeling choices, and to show that you're not just cherry-picking the specification that leads to the largest effects.

At a minimum, you should include the following three specifications:

1. **Model 1**: One model with *only the key variables* you want to measure (possibly transformed, as determined by your EDA), and no other covariates (or perhaps one, or at most two, covariates if they are so crucial that it would be unreasonable to omit them)
1. **Model 2**: One model that includes *key explanatory variables and covariates that you believe advance your modeling* goals without introducing too much collinearity or causing other issues.  This model should strike a balance between accuracy and parsimony and reflect your best understanding of the relationships among key variables.
1. **Model 3**: One model that includes the *previous covariates, and many other covariates*, erring on the side of inclusion.  A key purpose of this model is to demonstrate the robustness of your results to model specification.  (However, you should still not include variables that are clearly unreasonable. For example, don't include outcome variables that will absorb some of the causal effect you are interested in measuring.)

Guided by your background knowledge and your EDA, other specifications may make sense.  You are trying to choose points that encircle the space of reasonable modeling choices, to give an overall understanding of how these choices impact results.

### 3. Limitations of your Model

As a team, evaluate all of the CLM assumptions that must hold for your models. However, do not report an exhaustive examination all 5 CLM assumption. Instead, bring forward only those assumptions that you think pose significant problems for your analysis. For each problem that you identify, describe the statistical consequences. If you are able to identify any strategies to mitigate the consequences, explain these strategies.

Note that you may need to change your model specifications in response to violations of the CLM.

### 4. A Regression Table

You should display all of your model specifications in a regression table, using a package like [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) to format your output.  It should be easy for the reader to find the coefficients that represent key effects near the top of the regression table, and scan horizontally to see how they change from specification to specification.  Make sure that you display the most appropriate standard errors in your table, along with significance stars.

In your text, comment on both *statistical significance and practical significance*.  You may want to include statistical tests besides the standard t-tests for regression coefficients.

### 5. Discussion of Omitted Variables

If your team has chosen an explanatory (i.e. causal) question to evaluate, then identify what you think are the 5 most important *omitted variables* that bias the coefficients you care about.  For each variable, you should *reason about the direction of bias* caused by omitting this variable.  If you can argue whether the bias is large or small, that is even better.  State whether you have any variables available that may proxy (even imperfectly) for the omitted variable. Pay particular attention to whether each omitted variable bias is *towards zero or away from zero*.  You will use this information to judge whether the effects you find are likely to be real, or whether they might be entirely an artifact of omitted variable bias.

### 6. Conclusion

Make sure that you end your report with a discussion that distills key-takeaways from your estimates, addresses your research question, and draws attention to larger contexts.

### Submission

- Submit your draft report via ISVC; please do not submit via email.
- Submit 2 files:
    1. A pdf, html or md file including the summary, the details of your analysis, and all the R codes used to produce the analysis. *Please show code in your compiled file.*
    2. The Rmd or source file used to produce the pdf file.
- Only one group member needs to submit the files.
- Be sure to include the names of all team members in your report.  Place the word 'draft' in the file names.
- Please limit your submission to 8000 words, excluding code cells and R output.

## Within-Team Review

Being an effective team member is a crucial part of data science work.  Your performance on this lab includes the role you play in supporting your teammates.  This includes being responsive, creating an environment in which all members feel included, and above all treating each other with respect.  In line with this perspective, we will ask each team member to write to two paragraphs to their instructor about the progress they have made individually, and the team has made as a whole toward completing their report. This self-assessment should:

- Reflect on the strengths and weaknesses of the team and the team's process. Where your collaboration has worked well, how will you work to ensure that these successful practices continue to be employed? If there are places where collaboration has been challenging, what can the team do jointly to improve? 
- If there are any individual performances that deserve special recognition, please let me know in this evaluation. As well, if there are any individual performances that require special attention, please also let me know. I will treat these reviews as confidential and will not take any action without first consulting you.

You will submit this through ISVC, and like all parts of your educational record, this will be treated confidentially by the instructional team.

## Cross-Team Review

In the cross-team review, you will provide feedback on another team's draft report.  We will ask you to comment separately on different sections.  The following list is very similar to the rubric we will use when grading your final report.

- **Introduction.**  Is the introduction clear? Is the research question specific and well defined?  Does the introduction motivate a specific concept to be measured and explain how it will be operationalized.  Does it do a good job of preparing the reader to understand the model specifications?

- **The Initial Data Loading and Cleaning.**  Did the team notice any anomalous values?  Is there a sufficient justification for any data points that are removed?  Did the report note any coding features that affect the meaning of variables (e.g. top-coding or bottom-coding)?  Overall, does the report demonstrate a thorough understanding of the data?

- **The Model Building Process.** Overall, is each step in the model building process supported by EDA?  Is the outcome variable (or variables) appropriate? Is there a thorough univariate analysis of the outcome variable? Did the team identify one, or very small number of explanatory variables and perform a thorough univariate analysis of each one? Did the team clearly state why they chose these explanatory variables, does this explanation make sense in term of their research question? Did the team consider available variable transformations and select them with an eye towards model plausibility and interpretability?  Are transformations used to expose linear relationships in scatterplots?  Is there enough explanation in the text to understand the meaning of each visualization?

- **Regression Models:**
    - **Model 1** Does this model only include key explanatory variables?  Do the variables make sense given the measurement goals?  Did the team apply reasonable transformations to these variables, to capture the nature of the relationships?
    - **Model 2** Does this model represent a balanced approach, including variables that advance modeling goals without causing major issues?  Does the model succeed in reducing standard errors of the key variables compared to the base model?  Does it capture major non-linearities in the joint distribution of the variables?
    - **Model 3**  Does this model represent a maximalist approach, erring on the side of including most variables?  Is it still a reasonable model?  Are there any variables that are outcomes, and should therefore still be excluded?  Is there too much multicolinearity, to the point that the key causal effects cannot be measured?

```{r}

```

- **Plots, Figures, and Table** Do the plots, figures and tables that the team has chosen to include successfully move forward the argument that they are making? Do they have a good ratio of "Information to Ink" (Tufte)? Has the team chosen the most effective method (a table or a chart) to display their evidence? Is that table or chart as communicative as it can be? Is every single plot, figure, or table that is included in the report referenced in the main text?

- **Assessment of the CLM.**  Has the team assessed each of the CLM assumptions (including random sampling)?  Did they use visual tools or statistical tests, as appropriate?  Did they respond appropriately to any violations?

- **A  Regression Table.**  Are the model specifications properly chosen to outline the boundary of reasonable choices?  Is it easy to find key coefficients in the regression table?  Does the text include a discussion of practical significance for key effects?

- **An Omitted Variables Discussion.**  Did the report miss any important sources of omitted variable bias?  Are the estimated directions of bias correct? Was their explanation clear?  Is the discussion connected to whether the key effects are real or whether they may be solely an artifact of omitted variable bias?

- **Conclusion.**  Does the conclusion address the research question?  Does it raise interesting points beyond numerical estimates?  Does it place relevant context around the results?

- Are there any other errors, faulty logic, unclear or unpersuasive writing, or other elements that leave you less convinced by the conclusions?

Please be thorough and read the report critically, actively trying to find areas that the report could be improved.  Your comments will directly help your peers get the most value out of the project.

## Final Report

In the final stage of the project, you will incorporate the feedback you receive, and use what you've learned about OLS inference to create a final report.

We will assess your final report using a rubric that includes the elements listed above.  We will also consider whether you have correctly applied the statistical techniques in your report.

Please limit your submission to 8000 words, excluding code cells and R output.

As above, you must submit both the source and pdf files.  Be sure to include the names of all team members in your report.  Place the word 'final' in the file names.

## Final Presentation

During the Unit 14 live session, each team will give a slide presentation of their work to their classmates -- i.e. collaborating data scientists. This audience is generally aware of the project that you're working on, but they will need to be reminded of the specific research question that you are addressing. 
**The time limit for your slide presentation is 10 minutes.**  We will budget a few extra minutes for questions after you are done. We'd like to emphasize that this is an *incredibly* limited amount of time to present. The materials that you present should reflect these serious constraints!

1. There should be no more than two slides -- probably one is best -- that set-up your research problem and these slides should take no more than two minutes to present. (1 minute)
2. There should be at least one, and not more than two, slides that describe the one or two most important variables that you're using in your models. These slides should cover the important features of the variables that you're using: how are they are measured, the unit of observation, and why these *particular* variables are appropriate to use to answer your research question. (3 minutes) 
    - Do not present R code, discuss data wrangling, or discuss normality - details like this are best left to the full analysis.

3. There should then be several slides that provide what you've learned from your models. If you show model results, you need to provide your audience with enough time to read and engage with these models; not flash past them. Any model you show will take at least one minute to talk about.

Finally, a few more general thoughts:

- Practice your talk with a timer. We saw the 2020 debates, and we're going to end your talk 10 minutes in -- we have a mute button! :joy:
- If you divide your talk with your teammates, practice your section with a timer so that you do not spill over into your teammates' time.
- We *strongly* recommend having no more than 5 slides total.

### 2020-11-13 DraftReport.Rmd
Added draft report Rmd file.

---
title: "Draft Report"
author: "203 Section 6, Team 1"
output: pdf_document
---

2. A Model Building Process

  The primary variable of interest will be the total covid case rate per 100,000 residents. The covid case rate was judged to give a better understanding than the death rate because of the potential for other, unrecorded variables that could be correlated with a person dying from covid versus just becoming infected. For example, the availability and quality of medical care in a state may impact its ability to keep covid infected patients alive, although this is not represented in the data set. The health status of the residents in one state compared to another state may also affect the death rate, but this is also not included. Instead, the covid infection rate was chosen so that these other variables which may correlate with the death rate would not have to be considered. Finally, the covid infection rate was also chosen on a per capita basis so that potential total population differences between states would already be accounted for in the variable.

  The modeling goal for the covid case rate will be one of description. A modelf of causation may be created out of this model if potential causal relationships are examined and incorporated into the model's structure. This may later be expanded to predict the covid case rate in the past seven days in order for a state to use it to potentially judge what effect a modeled policy may have on its current situation.

  There are two broad groups of covariates which will be examined in building the model. The first is demographic information on the state, such as population density, the unemployment rate, or rate of people living in poverty. The second general group is that of policy decisions taken by the state, such as mask mandates, legal enforcements of policies, or the political party of the state's governor (which is assumed to be related to the policies they may support). Problematic covariates would include any of the other direct measures of covid severity in the state, such as total infection rate not on a per capita basis or death rates. It can be assumed that these variables would be colinear with the primary variable of interest since more covid cases per capita would imply more total cases not adjusted for population or be strongly correlated with more deaths per capita, etc.

### 2020-11-13 DraftReport.Rmd
Updated the draft report with the introduction

@@
 
-2. A Model Building Process
+### 1. An Introduction
+
+For this report, we will be investigating the question, "How is the COVID case rate related to the population density and distribution of demographics of a state?
+How does the effect of population make-up on case rate compare with the effect of policy decisions made in each state?".
+This research question aims to measure the case rate per 100,000 residents within each state based on the make-up of the population, as well as the policy decisions that were or were not put into place in order to combat the rise of COVID cases.
+In this sense, we aim to measure how dependent the COVID case rate is on that of which cannot be controlled (i.e., population and demographics) vs that of which can be controlled to a certain degree (i.e., implementation of policies to attempt to combat cases).
+With this information, we may be able to suggest whether or not the proliferation of COVID can be controlled, and to the level at which these policies help with this control.
+
+This question can be broken up into two distinctive areas of investigation.
+The first and primary area of investigation we are interested in is how the COVID case rate is affected by population metrics.
+For this, our key variables are COVID case rate per 100,000 residents and the population density per square mile.
+This will give us an initial understanding of our interest in the relationship between the spread of the virus and how densely populated a state is.
+Following this, we will also consider variables that address the make-up of these populations, specifically focusing on the distribution of demographics such as race/ethnicity and age.
+In order to account for the varying population sizes across the states, we will operationalize these variables by specifically looking at variables that contain a rate, in order to standardize across the samples.
+In this sense, we will be quantifying the spread of COVID by leveraging the COVID case rate per 100,000 rate, and for population demographics we will be investigating the percentage distribution of the race/ethnicity and age categories (e.g., White % of Total Population).
+
+The secondary area of investigation will be to also include variables that measure policy implementation within states, and to compare the effect of these as opposed to the primary area of investigation that is population metrics.
+Specifically, we will be focusing on policies that address the mandates of wearing masks within the state.
+Through this, we aim to analyze whether the implementation of mask related policies will have an effect on case rate, as well as the degree of this effect relative to population metrics.
+To operationalize these policies and simplifying the measurements, we will only be considering whether or not these policies were implemented through indicator variables (1 = implemented, 0 = not implemented).
+
+We preface our analysis by identifying certain assumptions and assessing the appropriateness of the data.
+Although there may be other considerations against the appropriateness of the data, we highlight three arguments that must be taken into account when interpreting the results of our analysis.
+
+Firstly, it is important to note that given we are treating each state (with the addition of the District of Columbia) as a data point, our sample contains 51 data points.
+Although we do meet the general rule of sample size, as we begin to factor in the wide range of potential population distributions and demographics within the states, we see that we have large variance our sample.
+For example, the population density can have large variability depending on the population concentration among a few large areas as well as the level of uninhabited or sparsely inhabited land within a state, while the demographics of population can depend on a wide range of factors such as geographic location and employment opportunities.
+We highlight this to show that the sample cannot be assumed to be IID.
+
+Secondly, we note that there are many internal and external aspects of our selected variables that have not been included within our models.
+Just addressing the internal information that is lost, we can see that much of the information is not captured given our method of operationalization.
+Among others, the loss of date-related information within our policy implementation variables strips us of any contextual knowledge regarding the length of policy implementation, and precludes us from identifying how long a specific policy was implemented.
+If we assume that policy implementation has some effect on case rate, this difference could play a large part in the effect.
+There are also other external aspects that cannot be included, such as the level of population density broken out among different population demographics.
+
+Finally, with regard to our policy variables, for this report we have focused solely on mask mandates.
+Given this, we do not capture the effect of other policies that may or may not have a greater effect on the case rate (e.g., implementation of stay-at-home orders).
+This follows for our population-related variables, as we capture race/ethnicity and age, but not other characteristics such as proportions of residents with pre-existing conditions and economic factors.
+In justification, these choices were largely made due to there either being an appropriate amount of samples in each category (e.g., most states have implemented basic policies such as the closure of non-essential businesses and the implementation of stay-at-home orders, so we chose not to include such policies given thelack of samples for the states that have not), or that data on other policies or demographics was not readily available in our dataset.
+
+### 2. A Model Building Process
 

### 2020-11-14 model_1.Rmd
Adding incomplete model 1 analysis

--- 2020-11-13 lab_2.Rmd
@@
 ---
-title: "Lab 2: Regression to Study the Spread of Covid-19"
-author: 'w203: Statistics for Data Science'
-date: "10/28/2020"
-output: html_document
+title: "model_1"
+author: "Sam Temlock"
+date: "13/11/2020"
+output: pdf_document
 ---
@@
 knitr::opts_chunk$set(echo = TRUE)
-```
-```{r lib,include=FALSE}
+
 library(lmtest)
@@
 
+df <- read.csv("covid-19.csv", header = TRUE)
 ```
 
-# Introduction
+For our first model, we will analyze the relationship between the COVID case rate per 100,000 and the population density per square mile of states.
+First, we take a look at the distribution of population density per square mile.
 
-In this lab, you will apply what you are learning about linear regression to study the spread of COVID-19.  Your task is to select a research question, then conduct a regression study to analyze it.
+```{r}
+histogram_of_pdensity <- df %>%
+  ggplot(aes(x = Population.density.per.square.miles)) + 
+  geom_histogram(bins = 20) + 
+  labs(
+    title = 'Figure 1. Distribution of Population Density per Square Mile', 
+    x = 'Population density per square mile', y = 'Count')
 
-Your research question must focus attention on a specific measurement goal.
-It is not enough to say "we are looking for policies that help stop COVID-19" (That would be a fishing expedition).
-Instead, use your introduction to motivate a specific effect for measurement.
+histogram_of_pdensity
+```
 
-Once you have a measurement goal, you will build a set of linear models that are tailored to that goal, and display them in a well formatted regression table.
-You will finally use your conclusion to distill important conclusions from your estimates.
+As we can see from Figure 1, although most of the population density is concentrated in the 0 to 1500 range, we have some grouping of outliers that are very far from this concentration.
+When we perform an analysis, we see that there is only one data sample that is the outlier, which is the District of Columbia (D.C.).
+This is given due to the fact that D.C.
+is a district that solely consists of a large city, as mentioned as a possibility in the introduction.
+Given that this causes the data to be skewed, we choose to remove this data point from the sample.
+Once removed, we see that we now have a relatively normal distribution of population densities (see Figure 2).
 
-This is a group assignment.
-Your live session instructor will coordinate the formation of groups.
-We would like to encourage teams to focus on using the lab as a way to learn how to work as a team of collaborating data scientists on shared code; how to clean and organize data; and, how to present work in a compelling way.
-As a result, we encourage teams to allow individuals to take risks and be supportive in the face of successes and failures.
-Create an opportunity for people who want to improve a particular skill to do so -- this might be project coordination, management of code through git, plotting, or any of the many aspects that you'll work on.
-*We hope that you can support and learn from one another through this team-based project.*
+```{r}
+# Find the outlier data points
+outliers <- subset(df,Population.density.per.square.miles > 4000)
+outliers
 
-# Deliverables
+# Remove the outlier from the dataset
+df_new <- df[ !(df$State %in% outliers$State), ]
 
-| Deliverable Name       | Week Due | Grade Weight |  
-|------------------------|----------|--------------| 
-| [Draft Report]         | Week 12  | 10%          | 
-| [Within-Team Review]   | Week 12  | 5%           | 
-| [Cross-Team Review]    | Week 13  | 10%          | 
-| [Final Report]         | Week 14  | 65%          | 
-| [Final Presentation]   | Week 14  | 10%          |
+# Check to see that D.C. has been removed
+paste("Outliers in the new dataframe:", outliers$State %in% df_new$State)
 
-# The Data
+histogram_of_pdensity <- df_new %>%
+  ggplot(aes(x = Population.density.per.square.miles)) + 
+  geom_histogram(bins = 20) + 
+  labs(
+    title = 'Figure 2. Distribution of Population Density per Square Mile without D.C.', 
+    x = 'Population density per square mile', y = 'Count')
 
-The data is provided in a file within this repository.
-Majid Maki-Nayeri compiled the data, drawing many variables from the COVID-19 US state policy database (Raifman J, Nocka K, Jones D, Bor J, Lipson S, Jay J, and Chan P.).
-Supporting documents like the specific legal language used by states, additional data sources, and much more are available in unstructured format [here](https://tinyurl.com/statepolicysources).
+histogram_of_pdensity
+```
 
-The dataset includes:
-
-1. Variables representing the spread of the disease; 
-2. Variables representing state-level policy responses; and, 
-3. General state-level characteristics.
-
-If you want to, you are allowed to add extra variables from external sources.
-However, this is not necessary, and we will not assign any bonus points to teams that derive unique data, as our focus is on statistics and statistical writing.
+Next, we evaluate the values for case_rate
 
 ```{r variables}
-df <- read.csv("covid-19.csv", header = TRUE)
+
+df_new %>% 
+  ggplot(aes(x = Case.Rate.per.100000, y = Population.density.per.square.miles)) + 
+  geom_point()
+
+df_new %>% 
+  ggplot(aes(x = Death.Rate.per.100000, y = Population.density.per.square.miles)) + 
+  geom_point() 
+```
+```{r}
+model_current <- lm(Case.Rate.per.100000 ~ Population.density.per.square.miles , data = df_new, na.action = na.omit)
+coeftest(model_current, vcov = vcovHC)
+
+model_alternative <- lm(Death.Rate.per.100000 ~ Population.density.per.square.miles , data = df_new, na.action = na.omit)
+coeftest(model_alternative, vcov = vcovHC)
+```
+
+```{r}
 
@@
 ```
-
-# Final Project Components {.tabset}
-
-## Draft Report
-
-In the first stage of the lab, you will create a draft report.  You should aim to make this report as complete as possible, so that you get the best possible feedback from the Cross-Team Review.
-
-The exact format of your report is flexible, but it should include the following elements.
-
-### 1. An Introduction
-
-Your introduction should present a research question and explain the concept that you're attempting to measure and how it will be operationalized.
-This section should pave the way for the body of the report, preparing the reader to understand why the models are constructed the way that they are.
-It is not enough to simply say "We are looking for policies that help against COVID"  Your introduction must do work for you, focusing the reader on a specific measurement goal, making them care about it, and propelling the narrative forward.
-This is also good time to put your work into context, discuss cross-cutting issues, and assess the overall appropriateness of the data.
-
-### 2. A Model Building Process
 
@@
 **Model 1**: One model with *only the key variables* you want to measure (possibly transformed, as determined by your EDA), and no other covariates (or perhaps one, or at most two, covariates if they are so crucial that it would be unreasonable to omit them)
-1.
-**Model 2**: One model that includes *key explanatory variables and covariates that you believe advance your modeling* goals without introducing too much collinearity or causing other issues.
-This model should strike a balance between accuracy and parsimony and reflect your best understanding of the relationships among key variables.
-1.
-**Model 3**: One model that includes the *previous covariates, and many other covariates*, erring on the side of inclusion.
-A key purpose of this model is to demonstrate the robustness of your results to model specification.
-(However, you should still not include variables that are clearly unreasonable.
-For example, don't include outcome variables that will absorb some of the causal effect you are interested in measuring.)
-
-Guided by your background knowledge and your EDA, other specifications may make sense.
-You are trying to choose points that encircle the space of reasonable modeling choices, to give an overall understanding of how these choices impact results.
-
-### 3. Limitations of your Model
-
-As a team, evaluate all of the CLM assumptions that must hold for your models.
-However, do not report an exhaustive examination all 5 CLM assumption.
-Instead, bring forward only those assumptions that you think pose significant problems for your analysis.
-For each problem that you identify, describe the statistical consequences.
-If you are able to identify any strategies to mitigate the consequences, explain these strategies.
-
-Note that you may need to change your model specifications in response to violations of the CLM.
-
-### 4. A Regression Table
-
-You should display all of your model specifications in a regression table, using a package like [`stargazer`](https://cran.r-project.org/web/packages/stargazer/vignettes/stargazer.pdf) to format your output.
-It should be easy for the reader to find the coefficients that represent key effects near the top of the regression table, and scan horizontally to see how they change from specification to specification.
-Make sure that you display the most appropriate standard errors in your table, along with significance stars.
-
-In your text, comment on both *statistical significance and practical significance*.  You may want to include statistical tests besides the standard t-tests for regression coefficients.
-
-### 5. Discussion of Omitted Variables
-
-If your team has chosen an explanatory (i.e.
-causal) question to evaluate, then identify what you think are the 5 most important *omitted variables* that bias the coefficients you care about.
-For each variable, you should *reason about the direction of bias* caused by omitting this variable.
-If you can argue whether the bias is large or small, that is even better.
-State whether you have any variables available that may proxy (even imperfectly) for the omitted variable.
-Pay particular attention to whether each omitted variable bias is *towards zero or away from zero*.
-You will use this information to judge whether the effects you find are likely to be real, or whether they might be entirely an artifact of omitted variable bias.
-
-### 6. Conclusion
-
-Make sure that you end your report with a discussion that distills key-takeaways from your estimates, addresses your research question, and draws attention to larger contexts.
-
-### Submission
-
-- Submit your draft report via ISVC; please do not submit via email.
-- Submit 2 files:
-    1. A pdf, html or md file including the summary, the details of your analysis, and all the R codes used to produce the analysis. *Please show code in your compiled file.*
-    2. The Rmd or source file used to produce the pdf file.
-- Only one group member needs to submit the files.
-- Be sure to include the names of all team members in your report.  Place the word 'draft' in the file names.
-- Please limit your submission to 8000 words, excluding code cells and R output.
-
-## Within-Team Review
-
-Being an effective team member is a crucial part of data science work.
-Your performance on this lab includes the role you play in supporting your teammates.
-This includes being responsive, creating an environment in which all members feel included, and above all treating each other with respect.
-In line with this perspective, we will ask each team member to write to two paragraphs to their instructor about the progress they have made individually, and the team has made as a whole toward completing their report.
-This self-assessment should:
-
-- Reflect on the strengths and weaknesses of the team and the team's process.
-Where your collaboration has worked well, how will you work to ensure that these successful practices continue to be employed?
-If there are places where collaboration has been challenging, what can the team do jointly to improve?
-- If there are any individual performances that deserve special recognition, please let me know in this evaluation.
-As well, if there are any individual performances that require special attention, please also let me know.
-I will treat these reviews as confidential and will not take any action without first consulting you.
-
-You will submit this through ISVC, and like all parts of your educational record, this will be treated confidentially by the instructional team.
-
-## Cross-Team Review
-
-In the cross-team review, you will provide feedback on another team's draft report.
-We will ask you to comment separately on different sections.
-The following list is very similar to the rubric we will use when grading your final report.
-
-- **Introduction.**  Is the introduction clear?
-Is the research question specific and well defined?
-Does the introduction motivate a specific concept to be measured and explain how it will be operationalized.
-Does it do a good job of preparing the reader to understand the model specifications?
-
-- **The Initial Data Loading and Cleaning.**  Did the team notice any anomalous values?
-Is there a sufficient justification for any data points that are removed?
-Did the report note any coding features that affect the meaning of variables (e.g.
-top-coding or bottom-coding)?
-Overall, does the report demonstrate a thorough understanding of the data?
-
-- **The Model Building Process.** Overall, is each step in the model building process supported by EDA?
-Is the outcome variable (or variables) appropriate?
-Is there a thorough univariate analysis of the outcome variable?
-Did the team identify one, or very small number of explanatory variables and perform a thorough univariate analysis of each one?
-Did the team clearly state why they chose these explanatory variables, does this explanation make sense in term of their research question?
-Did the team consider available variable transformations and select them with an eye towards model plausibility and interpretability?
-Are transformations used to expose linear relationships in scatterplots?
-Is there enough explanation in the text to understand the meaning of each visualization?
-
-- **Regression Models:**
-    - **Model 1** Does this model only include key explanatory variables?
-Do the variables make sense given the measurement goals?
-Did the team apply reasonable transformations to these variables, to capture the nature of the relationships?
-    - **Model 2** Does this model represent a balanced approach, including variables that advance modeling goals without causing major issues?
-Does the model succeed in reducing standard errors of the key variables compared to the base model?
-Does it capture major non-linearities in the joint distribution of the variables?
-    - **Model 3**  Does this model represent a maximalist approach, erring on the side of including most variables?
-Is it still a reasonable model?
-Are there any variables that are outcomes, and should therefore still be excluded?
-Is there too much multicolinearity, to the point that the key causal effects cannot be measured?
 
 ```{r}
-
+df %>% 
+  ggplot(aes(x = case_rate_100k, y = log(population))) + 
+  geom_point() 
 ```
-
-- **Plots, Figures, and Table** Do the plots, figures and tables that the team has chosen to include successfully move forward the argument that they are making?
-Do they have a good ratio of "Information to Ink" (Tufte)?
-Has the team chosen the most effective method (a table or a chart) to display their evidence?
-Is that table or chart as communicative as it can be?
-Is every single plot, figure, or table that is included in the report referenced in the main text?
-
-- **Assessment of the CLM.**  Has the team assessed each of the CLM assumptions (including random sampling)?
-Did they use visual tools or statistical tests, as appropriate?
-Did they respond appropriately to any violations?
-
-- **A  Regression Table.**  Are the model specifications properly chosen to outline the boundary of reasonable choices?
-Is it easy to find key coefficients in the regression table?
-Does the text include a discussion of practical significance for key effects?
-
-- **An Omitted Variables Discussion.**  Did the report miss any important sources of omitted variable bias?
-Are the estimated directions of bias correct?
-Was their explanation clear?
-Is the discussion connected to whether the key effects are real or whether they may be solely an artifact of omitted variable bias?
-
-- **Conclusion.**  Does the conclusion address the research question?  Does it raise interesting points beyond numerical estimates?  Does it place relevant context around the results?
-
-- Are there any other errors, faulty logic, unclear or unpersuasive writing, or other elements that leave you less convinced by the conclusions?
-
-Please be thorough and read the report critically, actively trying to find areas that the report could be improved.  Your comments will directly help your peers get the most value out of the project.
-
-## Final Report
-
-In the final stage of the project, you will incorporate the feedback you receive, and use what you've learned about OLS inference to create a final report.
-
-We will assess your final report using a rubric that includes the elements listed above.  We will also consider whether you have correctly applied the statistical techniques in your report.
-
-Please limit your submission to 8000 words, excluding code cells and R output.
-
-As above, you must submit both the source and pdf files.  Be sure to include the names of all team members in your report.  Place the word 'final' in the file names.
-
-## Final Presentation
-
-During the Unit 14 live session, each team will give a slide presentation of their work to their classmates -- i.e.
-collaborating data scientists.
-This audience is generally aware of the project that you're working on, but they will need to be reminded of the specific research question that you are addressing.
-**The time limit for your slide presentation is 10 minutes.**  We will budget a few extra minutes for questions after you are done.
-We'd like to emphasize that this is an *incredibly* limited amount of time to present.
-The materials that you present should reflect these serious constraints!
-
-1. There should be no more than two slides -- probably one is best -- that set-up your research problem and these slides should take no more than two minutes to present. (1 minute)
-2.
-There should be at least one, and not more than two, slides that describe the one or two most important variables that you're using in your models.
-These slides should cover the important features of the variables that you're using: how are they are measured, the unit of observation, and why these *particular* variables are appropriate to use to answer your research question.
-(3 minutes)
-    - Do not present R code, discuss data wrangling, or discuss normality - details like this are best left to the full analysis.
-
-3.
-There should then be several slides that provide what you've learned from your models.
-If you show model results, you need to provide your audience with enough time to read and engage with these models; not flash past them.
-Any model you show will take at least one minute to talk about.
-
-Finally, a few more general thoughts:
-
-- Practice your talk with a timer. We saw the 2020 debates, and we're going to end your talk 10 minutes in -- we have a mute button! :joy:
-- If you divide your talk with your teammates, practice your section with a timer so that you do not spill over into your teammates' time.
-- We *strongly* recommend having no more than 5 slides total.

### 2020-11-14 Model_2.rmd
model 2

---
title: "Untitled"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(lmtest)
library(sandwich)
library(stargazer)
library(magrittr)
library(ggplot2)
library(patchwork)

knitr::opts_chunk$set(echo = TRUE)
```

```{r}
D = read.csv("covid-19.csv")
```
## List of Demographic Variables  
COVID-19 Cases by Race/Ethnicity:
 - white % of Total Population,
 - Black % of Total Population, 
 - Hispanic % of Total Population, 
 - Other % of Total Population

State Characteristics:
 - Governor,       
 - Population density per square miles, 
 - Population 2018, 
 - Nonelderly Adults Who Have A Pre-Existing Condition
 - Percent at risk for serious illness due to COVID	
 - All-cause deaths 2018, 
 - Number Homeless (2019), 
 - Medicaid Expenditures as a Percent of Total State Expenditures by Fund, 
 - Life Expectancy at Birth (years),
 - Percent Unemployed (2018)
 - Percent living under the federal poverty line (2018), 
 - Weekly UI maximum amount with extra stimulus (through July 31, 2020) (dollars), 
 - Median Annual Household Income

Demographics:
	➤Children 0-18, ➤Adults 19-25, ➤Adults 26-34,  ➤Adults 35-54, ➤Adults 55-64, ➤65+

## Data Explorary

a) Varialbe of Case Rate per 100000 is of nearly normal distribution.

```{r}
summary(D$Case.Rate.per.100000)

```

```{r}
ggplot(data = D,
  mapping = aes(x= Case.Rate.per.100000))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,6000,600)) +
  labs(title =  "Histogram of Case rate per 100000", x = "Case per 100000", y = 'Count')+
  scale_x_continuous(breaks=seq(0, 6000, 600))
```

b) Varialbe of Population density per squres 
The distribution of the Population density per squres clusters at the two ends, which is caused by the extremely high population density of District of Columbia (11496.81) compared to the 50 states.
While, the logrithm of population density is close to a normal distribution. In addtion, an analysis excluding the District of Columbia is another option. 
```{r}
summary(D$Population.density.per.square.miles)
```
District of Columbia is extremely high population density than all 50 states
```{r}
hist_population_dens <- ggplot(data = D,
  mapping = aes(x= Population.density.per.square.miles))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,12000,500)) +
  labs(title =  "Histogram of Population density", x = "Population density per square miles", y = 'Count')+
  scale_x_continuous(breaks=seq(0, 12000, 1200))

hist_log_population_dens <- ggplot(data = D, aes(x = log(Population.density.per.square.miles))) + 
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,10,1)) + 
  labs(title = 'Histogram of Population density', x = 'Log Population density per square miles', y = 'Count')

hist_population_dens/hist_log_population_dens

```
c) White % of Total Population/ Black % of Total Population

Three entries of Black % of total population is "<0.01", which is to be dropped or replaced by values determined by extra resources for further analysis.

In addition, white % of total population and black % of total population could be highly related. (Correlation of these two vairalbes need to be investigated after cleaning the black % variables)

The distribution of the white % is skewed, not a ideal normal distribution. While the distribution of the square of white % is more close to a normal distribution. 
```{r}
D$Black...of.Total.Population
#cor(D$White...of.Total.Population, D$Black...of.Total.Population)
```

```{r}
summary(D$White...of.Total.Population)
```

```{r}
ggplot(data = D,
  mapping = aes(x= White...of.Total.Population))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,1,0.05)) +
  labs(title =  "Histogram of White Percentage", x = "White % of Total Population", y = 'Count')+
  scale_x_continuous(breaks=seq(0, 1, 0.1))
```
```{r}
ggplot(data = D,
  mapping = aes(x= (White...of.Total.Population)^2))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,1,0.05)) +
  labs(title =  "Histogram of White Percentage", x = "White % of Total Population", y = 'Count')+
  scale_x_continuous(breaks=seq(0, 1, 0.1))
```

d) Percent living under the federal poverty line 2018

Correlation of pecent living under poverty line and white percentage of total population shows that they are related.

The distribution is not heavily skewed.

```{r}
summary(D$Percent.living.under.the.federal.poverty.line..2018.)
cor(D$Percent.living.under.the.federal.poverty.line..2018., D$White...of.Total.Population)
```

```{r}
ggplot(data = D,
  mapping = aes(x= Percent.living.under.the.federal.poverty.line..2018.))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(6,21,1)) +
  labs(title =  "Histogram of Percent of poverty 2018", x = "Percent.living.under.the.federal.poverty.line", y = 'Count')+
  scale_x_continuous(breaks=seq(6, 21, 1))
```

e)
```

```

## Model 2

No obvious trend could be observed or concluded from the plots.

```{r}
D %>% 
  ggplot(aes(log(Population.density.per.square.miles), Case.Rate.per.100000, color = White...of.Total.Population)) + 
  geom_point() +   
  labs(
    title = 'Relation of Case.Rate.per.100000 to Population Density', 
    x = 'Log Population Density', 
    y = 'Case.Rate.per.100000',
    color = 'White %'
  )
```
```{r}
D %>% 
  ggplot(aes(x = log(Population.density.per.square.miles), y = Case.Rate.per.100000, color = Percent.living.under.the.federal.poverty.line..2018.)) + 
  geom_point() +   
  labs(
    title = 'Relation of Case.Rate.per.100000 to Population Density', 
    x = 'log Population Density', 
    y = 'Case.Rate.per.100000',
    color = 'Poverty Percentage'
  )
```

```{r}
D %>% 
  ggplot(aes(x = White...of.Total.Population, y = Case.Rate.per.100000, color = Percent.living.under.the.federal.poverty.line..2018.)) + 
  geom_point() +   
  labs(
    title = 'Relation of Case.Rate.per.100000 to White %', 
    x = 'White % of Total Population', 
    y = 'Case.Rate.per.100000',
    color = 'Poverty Percentage'
  )
```

```{r}
D %>% 
  ggplot(aes(x = Percent.living.under.the.federal.poverty.line..2018., y = Case.Rate.per.100000, color = log(Population.density.per.square.miles))) + 
  geom_point() +   
  labs(
    title = 'Relation of Case.Rate.per.100000 to Poverty %', 
    x = 'Percent.living.under.the.federal.poverty.line..2018.', 
    y = 'Case.Rate.per.100000',
    color = 'log(Population Density)'
  )
```

## Model 1-a
$$
\begin{aligned}
 Case.Rate.Per.100000 &=\beta_0 +\beta_1 Population.Density.Per.Square.Miles
\end{aligned}
$$

```{r}
model1a <- lm(Case.Rate.per.100000 ~ Population.density.per.square.miles, data = D)
coeftest(model1a, vcov = vcovHC)
```
## Model 1-b
$$
\begin{aligned}
 Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)
\end{aligned}
$$
```{r}
model1b <- lm(Case.Rate.per.100000 ~ log(Population.density.per.square.miles) , data = D)
coeftest(model1b, vcov = vcovHC)

```
## Model 2-a
$$
\begin{aligned}
 Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.total. population)
\end{aligned}
$$

```{r}
model2a <- lm(Case.Rate.per.100000 ~ log(Population.density.per.square.miles)+ White...of.Total.Population, data = D)
coeftest(model2a, vcov = vcovHC)

#model2a <- lm(Case.Rate.per.100000 ~  White...of.Total.Population, data = D)
#coeftest(model2a, vcov = vcovHC)
```
$$
\begin{aligned}
 Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.total. population)^2
\end{aligned}
$$
```{r}
model2a <- lm(Case.Rate.per.100000 ~ log(Population.density.per.square.miles)+ I((White...of.Total.Population)^2), data = D)
coeftest(model2a, vcov = vcovHC)

#model2a <- lm(Case.Rate.per.100000 ~  White...of.Total.Population, data = D)
#coeftest(model2a, vcov = vcovHC)
```

## Model 2-b
$$
\begin{aligned}
 Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.total. population)\\ &+\beta_3(Percen.living.under.the.federal.poverty.line.2018)
\end{aligned}
$$

```{r}
model2b <- lm(Case.Rate.per.100000 ~log(Population.density.per.square.miles)+ White...of.Total.Population+Percent.living.under.the.federal.poverty.line..2018., data = D)
coeftest(model2b, vcov = vcovHC)

#model2b <- lm(Case.Rate.per.100000 ~Percent.living.under.the.federal.poverty.line..2018., data = D)
#coeftest(model2b, vcov = vcovHC)
```

$$
\begin{aligned}
 Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.total. population)^2\\ &+\beta_3(Percen.living.under.the.federal.poverty.line.2018)
\end{aligned}
$$
```{r}
model2b <- lm(Case.Rate.per.100000 ~log(Population.density.per.square.miles)+ I((White...of.Total.Population)^2) + Percent.living.under.the.federal.poverty.line..2018., data = D)
coeftest(model2b, vcov = vcovHC)

#model2b <- lm(Case.Rate.per.100000 ~Percent.living.under.the.federal.poverty.line..2018., data = D)
#coeftest(model2b, vcov = vcovHC)
```

## Limitations

I.I.D assumption for CLM is not satisfied perfectly for all the models discussed above, since the states are not isolated from each other. It is impossible states suffered from pandmice at randomly different levels. Actually, the pandemic status across states may be highly related to each other.

In addition, all states are involved in the same market economy system even though there are econoy perfermance divisions across the states. Moreover, in most states white race is still dominant in the population. Many states share similar racial composition.

### 2020-11-14 DraftReport.Rmd
Updated introduction per discussion during meeting

--- 2020-11-13 DraftReport.Rmd
@@
 
-2. A Model Building Process
+### 1. An Introduction
+
+**Research Question:** How is the COVID case rate related to the population density and distribution of demographics of a state?
+How does the effect of population make-up on case rate compare with the effect of policy decisions made in each state?"
+
+For this report, the investigation will be centered around the effect of population metrics such as density and demographics and policy on the COVID case rates across the United States (US), which is grouped by the 50 states plus the District of Columbia (D.C.).
+This research question aims to measure the case rate per 100,000 residents within each state based on the make-up of the population, as well as the policy decisions that were or were not put into place in order to combat the rise of COVID cases.
+In this sense, the aim is to measure how dependent the COVID case rate is on that of which cannot be controlled (i.e., population and demographics) vs that of which can be controlled to a certain degree (i.e., implementation of policies to attempt to combat cases).
+With this information, one could suggest whether or not the proliferation of COVID can be controlled, and to the level at which these policies help with this control.
+
+This question can be broken up into two distinctive areas of investigation.
+The first and primary area of investigation involves how the COVID case rate is affected by population metrics.
+For this, the key variables are COVID case rate per 100,000 residents and the population density per square mile.
+This will provide an initial understanding in the relationship between the spread of the virus and how densely populated a state is.
+Following this, the analysis considers variables that address the make-up of these populations, specifically focusing on the distribution of demographics such as race/ethnicity and age.
+In order to account for the varying population sizes across the states, the variables are operationalized by specifically looking at variables that contain a rate, in order to standardize across the samples.
+In this sense, the analysis quantifies the spread of COVID by leveraging the COVID case rate per 100,000 rate, and for population demographics it investigates the percentage distribution of the race/ethnicity and age categories (e.g., White % of Total Population).
+
+The secondary area of investigation will be to also include variables that measure policy implementation within states, and to compare the effect of these as opposed to the primary area of investigation that is population metrics.
+Specifically, there is a focus on policies that address the mandates of wearing masks within the state.
+Through this, the aim is to analyze whether the implementation of mask related policies will have an effect on case rate, as well as the degree of this effect relative to population metrics.
+To operationalize these policies and simplifying the measurements, the models will only consider whether or not these policies were implemented through indicator variables (1 = implemented, 0 = not implemented).
+
+Prior to the analysis, it is important to identify a set of assumptions that have been made throughout the report and to assess the appropriateness of the data.
+Although there may be other considerations against the appropriateness of the data, the following highlights three particular arguments that must be taken into account when interpreting the results of the analysis.
+
+Firstly, it is important to note that given each state (with the addition of the District of Columbia) is treated as a unique data point, the sample contains 51 data points.
+Although this size meets the general rule of an adequate sample size, as the analysis begins to factor in the wide range of potential population distributions and demographics within the states, it is clear that there is large variation within the sample.
+For example, the population density can have large variability depending on the population concentration among a few large areas as well as the level of uninhabited or sparsely inhabited land within a state, while the demographics of population can depend on a wide range of factors such as geographic location and employment opportunities.
+It is also important to note that these data points cannot be considered independent of each other, as there is a lot of interstate travel and regional similarities in terms of population and policy.
+This indicates that the sample cannot be assumed to be IID.
+
+Secondly, note that there are many internal and external aspects of the selected variables that have not been included within the models.
+Just addressing the internal information that is lost, it can be seen that much of the information is not captured given the methods of operationalization.
+Among others, the loss of date-related information within policy implementation variables strips out any contextual knowledge regarding the length of policy implementation, and precludes the identification of how long a specific policy was implemented.
+Assuming that policy implementation has some effect on case rate, this difference in length of time could play a large part in the effect.
+There are also other external aspects that cannot be included, such as the level of population density broken out among different population demographics.
+
+Finally, with regard to the policy variables, this report focuses solely on mask mandates.
+Given this, it does not capture the effect of other policies that may or may not have a greater effect on the case rate (e.g., implementation of stay-at-home orders).
+This follows for the population-related variables, as they capture race/ethnicity and poverty, but not other characteristics such as proportions of residents with pre-existing conditions.
+In justification, these choices were largely made due to there either being an appropriate amount of samples in each category (e.g., most states have implemented basic policies such as the closure of non-essential businesses and the implementation of stay-at-home orders, and thus these were not included given the lack of samples for the states that have not implemented them), or that data on other policies or demographic measures were simply not readily available in the dataset.
+
+### 2. A Model Building Process
 

### 2020-11-15 lab_2.Rmd
Update lab_2.Rmd

@@
 library(magrittr)
-
+library(ggplot2)
+library(patchwork)
 ```
@@
 df <- read.csv("covid-19.csv", header = TRUE)
-
+names(df)
 df<-df%>%
   rename(case_rate_100k = 'Case.Rate.per.100000',
+         death_rate_100k = 'Death.Rate.per.100000',
          population_density = 'Population.density.per.square.miles',
@@
          other_pct = 'Other...of.Cases',
+         state_emergency = 'State.of.emergency',
+         business_closed = 'Closed.other.non.essential.businesses',
+         business_reopen = 'Began.to.reopen.businesses.statewide',
          mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
          mask_legal='No.legal.enforcement.of.face.mask.mandate') %>%
-  select(case_rate_100k, population_density, white_pct, black_pct, hispanic_pct, 
-         other_pct, mask_public, mask_legal)
+  select(case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
+         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal)
+# Assign correct data types
+cols.num <- c("white_pct","black_pct","hispanic_pct","other_pct")
+df[cols.num] <- sapply(df[cols.num],as.numeric)
+df$state_emergency=as.Date(df$state_emergency, format = "%m/%d/%Y")
+df$business_closed=as.Date(df$business_closed, format = "%m/%d/%Y")
+df$business_reopen=as.Date(df$business_reopen, format = "%m/%d/%Y")
 head(df)
@@
 
+model2b <- lm(case_rate_100k ~log(population_density) + I((white_pct)^2) , data = df)
+coeftest(model2b, vcov = vcovHC)
+
+```
+
+For Model 3. Policy related variables will be introduced for the analytics.
+
+```{r}
+
+df <- df %>% 
+  mutate(
+    mask_public_bool = case_when(
+      mask_public == 0 ~ 0, 
+      !(mask_public == 0) ~ 1
+      )
+  )
+head(df)
+
+hist(df$case_rate_100k[df$mask_legal == 1], xlab='Case Rate', ylab='Count', main='Case Rate - Mask Legal')
+hist(df$case_rate_100k[df$mask_legal == 0], xlab='Case Rate', ylab='Count', main='Dem Voter - Mask not legal')
+
+hist(df$case_rate_100k[df$mask_public_bool == 1], xlab='Case Rate', ylab='Count', main='Case Rate - Mask Mandate in Public')
+hist(df$case_rate_100k[df$mask_public_bool == 0], xlab='Case Rate', ylab='Count', main='Dem Voter - No Mask Mandate in Public')
+
+model3 <- lm(case_rate_100k ~log(population_density) + I((white_pct)^2) + mask_legal, data = df)
+model3 <- lm(case_rate_100k ~ log(population_density) +white_pct+ mask_public_bool, data = df)
+#model3 <- lm(death_rate_100k ~ log(population_density) + mask_public_bool, data = df)
+coeftest(model3, vcov = vcovHC)
+
+df %>% 
+  ggplot(aes(x = mask_public_bool, y = case_rate_100k, color = case_rate_100k)) + 
+  geom_point(aes(size=case_rate_100k)) + 
+  labs(
+    title = 'Case Rate vs Mask Public Mandate', 
+    x = 'Mask Mandate in Public', 
+    y = 'Case Rate', 
+    color = 'Case Rate'
+  )
+
+df$close_period = as.integer( df$business_reopen-df$business_closed)
+head(df)
+hist(df$close_period, xlab='close period', ylab='Count', main='close length')
+df %>% 
+  ggplot(aes(x = close_period, y = case_rate_100k, color = close_period)) + 
+  geom_point() + 
+  labs(
+    title = 'Case Rate vs Mask Public Mandate', 
+    x = 'Mask Mandate in Public', 
+    y = 'Case Rate', 
+    color = 'Case Rate'
+  )
+
+model3 <- lm(case_rate_100k ~ close_period, data = df)
+coeftest(model3, vcov = vcovHC)
+
+df[is.na(df$business_closed),]
+df[42,'business_closed'] = max(df$business_closed[!is.na(df$business_reopen)])
+mean(df$business_closed)
+df$business_close_diff = df$business_closed-mean(df$business_closed)
+head(df)
+
+df %>% 
+  ggplot(aes(x = business_close_diff, y = case_rate_100k, color = close_period)) + 
+  geom_point() + 
+  labs(
+    title = 'Case Rate vs Mask Public Mandate', 
+    x = 'Mask Mandate in Public', 
+    y = 'Case Rate', 
+    color = 'Case Rate'
+  )
+
+model3 <- lm(case_rate_100k ~ business_close_diff + mask_public_bool, data = df)
+coeftest(model3, vcov = vcovHC)
+
 ```

### 2020-11-16 DraftReport.Rmd
Updated Section 3 of the report.

@@
 It can be assumed that these variables would be colinear with the primary variable of interest since more covid cases per capita would imply more total cases not adjusted for population or be strongly correlated with more deaths per capita, etc.
+
+### 3. Limitations of the Model
+
+The most general requirement for the linear model is that data points be independent and identically distributed (IID).
+While for the most part the events of one state will primarily affect that single state, travel by individuals spreading the virus across state lines is an inevitable occurence.
+This may cause the spread of a virus within one state to be influenced by those around it, or generally by its position in the country.
+While this is acknowledged to occur, the effect is likely minor compared to the demographic and policy data collected state by state which would affect the spread of the virus.
+Therefore it is relatively safe for this assumption to be met.
+
+Secondly, the process being modeled must be described by a linear conditional expectation function of the variables which are included, in order for the linear regression to be valid.
+This can be assessed after the fact by visualizing the model over the range of each variable alongside the data, or also by visualizing the residuals over the range of each variable.
+
+Related to the relationship of the residuals to the data, homoskedasticity can be assessed by visualizing the size of residuals across the range of each variable.
+Alternatively, it can also be done via the Breusch-Pagan test to more objectively determine whether homoskedasticity is not present in the data.
+
+Perfect colinearity is automatically detected by the R regression algorithms, which drop potential variables if they have complete colinearity with another variable.
+Near perfect colinearity is more problematic to detect, but a common hint to its existence is large standard errors on colinear features.
+Conceptually, the variables which may have an important colinearity in the final model are
+
+Finally, normally distributed errors can be easily assessed by a histogram of the residuals around zero.

### 2020-11-16 Model_2.rmd
model 2 update, Please enter the commit message for your changes. Lines starting

@@
 ---
-title: "Untitled"
-output: pdf_document
+title: "Model2"
+output:
+  pdf_document: default
+  html_document:
+    df_print: paged
 ---
@@
 ```{r setup, include=FALSE, warning=FALSE, message=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+library(tinytex)
 library(tidyverse)
@@
 library(patchwork)
-
-knitr::opts_chunk$set(echo = TRUE)
 ```
@@
 ```{r}
-D = read.csv("covid-19.csv")
+df = read.csv("covid-19.csv")
 ```
-## List of Demographic Variables  
+
+## Model 2 Building Process
+Part of the model aims to investigate the relationship between the covid-19 pandemic status and the state demographic information.
+The list of demographic variables included in the current dataset is as below.
+
 COVID-19 Cases by Race/Ethnicity:
@@
 Demographics:
-	➤Children 0-18, ➤Adults 19-25, ➤Adults 26-34,  ➤Adults 35-54, ➤Adults 55-64, ➤65+
+ -Children 0-18
+ - Adults 19-25
+ - Adults 26-34
+ - Adults 35-54
+ - Adults 55-64
+ - 65+
 
-## Data Explorary
+Currently, Case Rate per 100000 is chosen as a dependent variable to represent the pandemic status in each state.
+Variables as Population density per square miles, Percent Unemployed (2018)/Percent living under the federal poverty line (2018), White % of of Total Population/Black % of Total Population, Percentage of 65+ are chosen as candidate covariates to explore potential connection between pandemic status demographic characteristics, economy status, racial composition.
 
-a) Varialbe of Case Rate per 100000 is of nearly normal distribution.
+Through data explorary, most data investigated is clean as numberica type except the black percentage of total population.
+This information of some states is reported vaguely as less than 0.01, which is replaced as 0 for further work.
+This could be updated with more precise data with the help of external resources.
+
+Several transformations have been made for variables a Population density per square miles and White % of Total Population.
+
+### Data Explorary
+Through data explorary, most data investigated is clean as numberica type except the black percentage of total population.
+This information of some states is reported vaguely as less than 0.01, which is replaced as 0 for further work.
+This could be updated with more precise data with the help of external resources.
+
+Several transformations have been made for variables a Population density per square miles and White % of Total Population.
+Population density is extremely skewed due to the high population density of the District of Columbia.
+A logarithm transformation of this variable is of more uniform distribution.
+In addition, the square of the white percentage of the total population is of a distribution closer to the normal distribution.
+It also represents the chance that the interaction happens between two white people in practice.
+
+In addition, correlation values of White% and Black %, Percent Unemployed and Percent living under the federal poverty line shows these two groups of variables are correlated respectively. 
+#### 1-a) Varialbe of Case Rate per 100000 is of nearly normal distribution.
 
 ```{r}
-summary(D$Case.Rate.per.100000)
+summary(df$Case.Rate.per.100000)
 
@@
 ```{r}
-ggplot(data = D,
+ggplot(data = df,
   mapping = aes(x= Case.Rate.per.100000))+
@@
 
-b) Varialbe of Population density per squres 
-The distribution of the Population density per squres clusters at the two ends, which is caused by the extremely high population density of District of Columbia (11496.81) compared to the 50 states.
-While, the logrithm of population density is close to a normal distribution. In addtion, an analysis excluding the District of Columbia is another option. 
+#### 1-a) Varialbe of Population density per squres 
+The distribution of the Population density per squares clusters at the two ends, which is caused by the extremely high population density of District of Columbia (11496.81) compared to the 50 states.
+While, the logarithm of population density is close to a normal distribution. In addition, an analysis excluding the District of Columbia is another option.
+
 ```{r}
-summary(D$Population.density.per.square.miles)
+summary(df$Population.density.per.square.miles)
 ```
@@
 ```{r}
-hist_population_dens <- ggplot(data = D,
+hist_population_dens <- ggplot(data = df,
   mapping = aes(x= Population.density.per.square.miles))+
@@
 
-hist_log_population_dens <- ggplot(data = D, aes(x = log(Population.density.per.square.miles))) + 
+hist_log_population_dens <- ggplot(data = df, aes(x = log(Population.density.per.square.miles))) + 
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,10,1)) + 
@@
 ```
-c) White % of Total Population/ Black % of Total Population
+#### 2-a) White % of Total Population/ Black % of Total Population
 
-Three entries of Black % of total population is "<0.01", which is to be dropped or replaced by values determined by extra resources for further analysis.
+Variable of Black % is stored as string variables.
+Moreover, three entries of Black % of total population is "<0.01", which is to be dropped or replaced by values determined by extra resources for further analysis.
 
-In addition, white % of total population and black % of total population could be highly related. (Correlation of these two vairalbes need to be investigated after cleaning the black % variables)
+Here, the value of Black % is replaced by 0 when it is "<0.01".
+Afterwards, the correlation of white % and black % is -0.42, which shows these two variables are correlated.
+Considering that the value of other race groups is much smaller, it is proper to only include White% in the regression model to explore the relation of care rate and racial composition.
 
-The distribution of the white % is skewed, not a ideal normal distribution. While the distribution of the square of white % is more close to a normal distribution. 
+The distribution of the white % is skewed, not an ideal normal distribution.
+While the distribution of the square of white % is more close to a normal distribution.
+In practice, this square could reflect the chance that the interaction happens between two white people.
+
 ```{r}
-D$Black...of.Total.Population
-#cor(D$White...of.Total.Population, D$Black...of.Total.Population)
+black_num <- as.numeric(df$Black...of.Total.Population)
+black_num[is.na(black_num)]<- 0
+cor(df$White...of.Total.Population, black_num)
+
 ```
@@
 ```{r}
-summary(D$White...of.Total.Population)
+summary(df$White...of.Total.Population)
 ```
@@
 ```{r}
-ggplot(data = D,
+ggplot(data = df,
   mapping = aes(x= White...of.Total.Population))+
@@
 ```{r}
-ggplot(data = D,
+ggplot(data = df,
   mapping = aes(x= (White...of.Total.Population)^2))+
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,1,0.05)) +
-  labs(title =  "Histogram of White Percentage", x = "White % of Total Population", y = 'Count')+
+  labs(title =  "Histogram of White Percentage", x = "White % of Total Population^2", y = 'Count')+
   scale_x_continuous(breaks=seq(0, 1, 0.1))
@@
 
-d) Percent living under the federal poverty line 2018
+#### 2-b) Percent living under the federal poverty line 2018
 
-Correlation of pecent living under poverty line and white percentage of total population shows that they are related.
-
-The distribution is not heavily skewed.
+Correlation of pecent living under poverty line and white percentage of total population shows that they are related. The distribution is not heavily skewed.
 
 ```{r}
-summary(D$Percent.living.under.the.federal.poverty.line..2018.)
-cor(D$Percent.living.under.the.federal.poverty.line..2018., D$White...of.Total.Population)
+summary(df$Percent.living.under.the.federal.poverty.line..2018.)
+cor(df$Percent.living.under.the.federal.poverty.line..2018., df$White...of.Total.Population)
 ```
@@
 ```{r}
-ggplot(data = D,
+ggplot(data = df,
   mapping = aes(x= Percent.living.under.the.federal.poverty.line..2018.))+
@@
 ```
+#### 2-c) Percent Unemployed 2018
+Although, the distribution is of high peak in the middle of the range.
+Overall, it is not highly skewed or heavily tailed.
+In addition, the unemployed rate is correlated to the variable to the poverty rate, which is in an agreement with intuitive expectation.
 
-e)
+```{r}
+summary(df$Percent.Unemployed..2018.)
+cor(df$Percent.Unemployed..2018., df$Percent.living.under.the.federal.poverty.line..2018.)
+```
+```{r}
+ggplot(data = df,
+  mapping = aes(x= Percent.Unemployed..2018.))+
+  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,10,1)) +
+  labs(title =  "Histogram of Unemployed percentage", x = "Percent.Unemployed..2018.", y = 'Count')+
+  scale_x_continuous(breaks=seq(0, 10, 1))
 ```
 
+#### 2-d) Age 65+ Percentage
+The 65+ percentage is of a nearly normal distribution.
+
+```{r}
+summary(df$X65.)
+```
+```{r}
+ggplot(data = df,
+  mapping = aes(x= X65.))+
+  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0.10,0.22,0.01)) +
+  labs(title =  "Histogram of Percent of 65+", x = "65+ Percentage", y = 'Count')+
+  scale_x_continuous(breaks=seq(0.1, 0.25, 0.05))
 ```
 
-## Model 2
-
-No obvious trend could be observed or concluded from the plots.
+### Plot 
+No obvious trend could be easily observed or concluded from the plots regarding the dependent variables and candidate covariates.
 
 ```{r}
-D %>% 
+df %>% 
   ggplot(aes(log(Population.density.per.square.miles), Case.Rate.per.100000, color = White...of.Total.Population)) + 
@@
 ```{r}
-D %>% 
+df %>% 
   ggplot(aes(x = log(Population.density.per.square.miles), y = Case.Rate.per.100000, color = Percent.living.under.the.federal.poverty.line..2018.)) + 
@@
 ```{r}
-D %>% 
+df %>% 
   ggplot(aes(x = White...of.Total.Population, y = Case.Rate.per.100000, color = Percent.living.under.the.federal.poverty.line..2018.)) + 
@@
 ```{r}
-D %>% 
+df %>% 
   ggplot(aes(x = Percent.living.under.the.federal.poverty.line..2018., y = Case.Rate.per.100000, color = log(Population.density.per.square.miles))) + 
@@
 ```
+```{r}
+df %>% 
+  ggplot(aes(X65., Case.Rate.per.100000, color = log(Population.density.per.square.miles))) + 
+  geom_point() +   
+  labs(
+    title = 'Relation of Case.Rate.per.100000 to 65+ Percentage', 
+    x = 'X65+.', 
+    y = 'Case.Rate.per.100000',
+    color = 'log(Population density)'
+  )
+```
 
-## Model 1-a
+```{r}
+df %>% 
+  ggplot(aes(Percent.Unemployed..2018., Case.Rate.per.100000, color = log(Population.density.per.square.miles))) + 
+  geom_point() +   
+  labs(
+    title = 'Relation of Case.Rate.per.100000 to Unemployed Percentage', 
+    x = 'Percent.Unemployed..2018.', 
+    y = 'Case.Rate.per.100000',
+    color = 'log(Population density)'
+  )
+```
+
+### Regression
+
+According to the regression results, no coefficients regarding the demographic variables investigated in model 2 are statistically significant.
+There is no statistically meaningful relationship that could be concluded by now.
+
+#### Model 1-a
 $$
@@
 ```{r}
-model1a <- lm(Case.Rate.per.100000 ~ Population.density.per.square.miles, data = D)
+model1a <- lm(Case.Rate.per.100000 ~ Population.density.per.square.miles, data = df)
 coeftest(model1a, vcov = vcovHC)
 ```
-## Model 1-b
+#### Model 1-b
 $$
-\begin{aligned}
- Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)
-\end{aligned}
+Case.Rate.Per.100000 =\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)
 $$
 ```{r}
-model1b <- lm(Case.Rate.per.100000 ~ log(Population.density.per.square.miles) , data = D)
+model1b <- lm(Case.Rate.per.100000 ~ log(Population.density.per.square.miles) , data = df)
 coeftest(model1b, vcov = vcovHC)
@@
 ```
-## Model 2-a
+#### Model 2-a
 $$
 \begin{aligned}
- Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.total. population)
+ Case.Rate.Per.100000 =\beta_0 &+\beta_1 log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.total. population)
 \end{aligned}
@@
 ```{r}
-model2a <- lm(Case.Rate.per.100000 ~ log(Population.density.per.square.miles)+ White...of.Total.Population, data = D)
+model2a <- lm(Case.Rate.per.100000 ~ log(Population.density.per.square.miles)+ White...of.Total.Population, data = df)
 coeftest(model2a, vcov = vcovHC)
 
-#model2a <- lm(Case.Rate.per.100000 ~  White...of.Total.Population, data = D)
-#coeftest(model2a, vcov = vcovHC)
 ```
+
 $$
@@
 ```{r}
-model2a <- lm(Case.Rate.per.100000 ~ log(Population.density.per.square.miles)+ I((White...of.Total.Population)^2), data = D)
+model2a <- lm(Case.Rate.per.100000 ~ log(Population.density.per.square.miles)+ I((White...of.Total.Population)^2), data = df)
+coeftest(model2a, vcov = vcovHC)
+```
+$$
+\begin{aligned}
+ Case.Rate.Per.100000 &=\beta_0 + \beta_2(White.Percentage.of.total. population)^2
+\end{aligned}
+$$
+```{r}
+model2a <- lm(Case.Rate.per.100000 ~ I((White...of.Total.Population)^2), data = df)
 coeftest(model2a, vcov = vcovHC)
@@
 
-## Model 2-b
+#### Model 2-b
 $$
@@
 ```{r}
-model2b <- lm(Case.Rate.per.100000 ~log(Population.density.per.square.miles)+ White...of.Total.Population+Percent.living.under.the.federal.poverty.line..2018., data = D)
+model2b <- lm(Case.Rate.per.100000 ~log(Population.density.per.square.miles)+ White...of.Total.Population+Percent.living.under.the.federal.poverty.line..2018., data = df)
 coeftest(model2b, vcov = vcovHC)
@@
 ```{r}
-model2b <- lm(Case.Rate.per.100000 ~log(Population.density.per.square.miles)+ I((White...of.Total.Population)^2) + Percent.living.under.the.federal.poverty.line..2018., data = D)
+model2b <- lm(Case.Rate.per.100000 ~log(Population.density.per.square.miles)+ I((White...of.Total.Population)^2) + Percent.living.under.the.federal.poverty.line..2018., data = df)
 coeftest(model2b, vcov = vcovHC)
 
-#model2b <- lm(Case.Rate.per.100000 ~Percent.living.under.the.federal.poverty.line..2018., data = D)
-#coeftest(model2b, vcov = vcovHC)
+```
+$$
+\begin{aligned}
+ Case.Rate.Per.100000 &=\beta_0 +\beta_3(Percen.living.under.the.federal.poverty.line.2018)
+\end{aligned}
+$$
+```{r}
+model2b <- lm(Case.Rate.per.100000 ~ Percent.living.under.the.federal.poverty.line..2018., data = df)
+coeftest(model2b, vcov = vcovHC)
 ```
 
-## Limitations
+#### Model 2-c
 
-I.I.D assumption for CLM is not satisfied perfectly for all the models discussed above, since the states are not isolated from each other.
-It is impossible states suffered from pandmice at randomly different levels.
-Actually, the pandemic status across states may be highly related to each other.
+$$
+\begin{aligned}
+ Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.total. population)^2\\ &+\beta_4(Percent.Unemployed..2018.)
+\end{aligned}
+$$
+```{r}
+model2c <- lm(Case.Rate.per.100000 ~log(Population.density.per.square.miles)+ I((White...of.Total.Population)^2) + Percent.Unemployed..2018., data = df)
+coeftest(model2c, vcov = vcovHC)
 
-In addition, all states are involved in the same market economy system even though there are econoy perfermance divisions across the states.
+```
+$$
+\begin{aligned}
+ Case.Rate.Per.100000 &=\beta_0+\beta_4(Percent.Unemployed..2018.)
+\end{aligned}
+$$
+```{r}
+model2c <- lm(Case.Rate.per.100000 ~ Percent.Unemployed..2018., data = df)
+coeftest(model2c, vcov = vcovHC)
+
+```
+
+#### Model 2-d
+
+$$
+\begin{aligned}
+Case.Rate.Per.100000=\beta_0 &+\beta_1log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.total.population)\\&+\beta_3(Percen.living.under.the.federal.poverty.line.2018)\\&+\beta_5(65+)
+\end{aligned}
+$$
+```{r}
+model2d <- lm(Case.Rate.per.100000 ~log(Population.density.per.square.miles)+ White...of.Total.Population+Percent.living.under.the.federal.poverty.line..2018.+X65., data = df)
+coeftest(model2d, vcov = vcovHC)
+```
+## Model 2-d
+$$
+\begin{aligned}
+Case.Rate.Per.100000=\beta_0 &+\beta_1log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.total.population)^2\\&+\beta_3(Percen.living.under.the.federal.poverty.line.2018)\\&+\beta_4(65+)
+\end{aligned}
+$$
+```{r}
+model2d <- lm(Case.Rate.per.100000 ~log(Population.density.per.square.miles)+ I((White...of.Total.Population)^2) + Percent.living.under.the.federal.poverty.line..2018.+X65., data = df)
+coeftest(model2d, vcov = vcovHC)
+
+```
+
+$$
+\begin{aligned}
+Case.Rate.Per.100000=\beta_0+\beta_4(65+)
+\end{aligned}
+$$
+```{r}
+model2d <- lm(Case.Rate.per.100000 ~X65., data = df)
+coeftest(model2d, vcov = vcovHC)
+
+```
+
+## Limitations of Model 2
+
+The I.I.D assumption for CLM is not satisfied perfectly for all the models discussed above, since the states are not isolated from each other.
+It is impossible that states suffered from pandemic at different levels independently.
+Actually, the pandemic status across states tends to be highly related to each other.
+
+In addition, all states are involved in the same market economy system even though there are economic performance divisions across the states.
 Moreover, in most states white race is still dominant in the population.
-Many states share similar racial composition.
+Many states share the similar racial composition.

### 2020-11-16 DraftReport.Rmd

@@
 
+- **Common Data across models:**
+```{r lib,include=FALSE}
+library(lmtest)
+library(sandwich)
+library(tidyverse)
+library(magrittr)
+library(ggplot2)
+library(patchwork)
+```
+
+```{r variables, warning=FALSE}
+df <- read.csv("covid-19.csv", header = TRUE)
+df<-df%>%
+  rename(case_rate_100k = 'Case.Rate.per.100000',
+         death_rate_100k = 'Death.Rate.per.100000',
+         population_density = 'Population.density.per.square.miles',
+         white_pct = 'White...of.Cases',
+         black_pct = 'Black...of.Cases',
+         hispanic_pct = 'Hispanic...of.Cases',
+         other_pct = 'Other...of.Cases',
+         state_emergency = 'State.of.emergency',
+         business_closed = 'Closed.other.non.essential.businesses',
+         business_reopen = 'Began.to.reopen.businesses.statewide',
+         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
+         mask_legal='No.legal.enforcement.of.face.mask.mandate') %>%
+  select(case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
+         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal)
+# Assign correct data types
+cols.num <- c("white_pct","black_pct","hispanic_pct","other_pct")
+df[cols.num] <- sapply(df[cols.num],as.numeric)
+df$state_emergency=as.Date(df$state_emergency, format = "%m/%d/%Y")
+df$business_closed=as.Date(df$business_closed, format = "%m/%d/%Y")
+df$business_reopen=as.Date(df$business_reopen, format = "%m/%d/%Y")
+head(df)
+```
+
+- **Regression Models:**
+    - **Model 1** Does this model only include key explanatory variables?
+Do the variables make sense given the measurement goals?
+Did the team apply reasonable transformations to these variables, to capture the nature of the relationships?
+    - **Model 2** Does this model represent a balanced approach, including variables that advance modeling goals without causing major issues?
+Does the model succeed in reducing standard errors of the key variables compared to the base model?
+Does it capture major non-linearities in the joint distribution of the variables?
+    - **Model 3**  Does this model represent a maximalist approach, erring on the side of including most variables?
+Is it still a reasonable model?
+Are there any variables that are outcomes, and should therefore still be excluded?
+Is there too much multicolinearity, to the point that the key causal effects cannot be measured?
+
+    - For Model 3.
+Policy related variables will be introduced for the analytics.
+Policy variable "Mandate face mask use by all individuals in public spaces" (renamed as mask_public) is selected as control variable to predict the case rate.
+Given it has been proved that mask is effective in preventing the transmission of disease, it is expected that the mask policy mandate in public would have an effect in reducing the case rate.
+There are other variables related to mask such as "No legal enforcement of face mask mandate" and "Mandate face mask use by employees in public-facing businesses", we believe these variables have a lower impact than the policy that is enforced to public.
+Also, it is very likely that if a state enforce public mask policy, they will by default enforce mask policy for employees in public-facing business.
+As a result, we believe mask policy for all individuals in public space can be used to effectively represents the policy impact.
+Also, it is worth to note that, the mask_public is in date format, which has value of zero or a actual date when the policy is enforced.
+According to the documentation, zero represents "the absence of an order or directive", which can be interpreted as the policy is not enforced by the state explicitly.
+For a leaner regression, the date values in mask_public column are transformed as value 1, so that we can distinguish whether states has public mask policy or not.
+
+Firstly, "mask_public_bool" column is created from "mask_public", with 1 represents that the state has a pulbic mask policy, and 0 means there's no explicit public mask policy from the state.
+```{r}
+
+df <- df %>% 
+  mutate(
+    mask_public_bool = case_when(
+      mask_public == 0 ~ 0, 
+      !(mask_public == 0) ~ 1
+      )
+  )
+head(df)
+```
+
+Histograms below that shows the distribution of states that has public mask policy or not. Both distributions exhibits certain degree of normality. 
+```{r}
+hist(df$case_rate_100k[df$mask_public_bool == 1], xlab='Case Rate', ylab='Count', main='Case Rate - Mask Mandate in Public')
+hist(df$case_rate_100k[df$mask_public_bool == 0], xlab='Case Rate', ylab='Count', main='Case Rate  - No Mask Mandate in Public')
+```
+
+A plot that shows the relationship between different policies vs case rate.
+```{r}
+df %>% 
+  ggplot(aes(x = mask_public_bool, y = case_rate_100k, color = case_rate_100k)) + 
+  geom_point(aes(size=case_rate_100k)) + 
+  labs(
+    title = 'Case Rate vs Mask Public Mandate', 
+    x = 'Mask Mandate in Public', 
+    y = 'Case Rate', 
+    color = 'Case Rate'
+  )
+```
+
+Next, the linear model is created from model 2 with the mask policy variables added for the regression.
+
+```{r}
+model3 <- lm(case_rate_100k ~ log(population_density) +white_pct+ mask_public_bool, data = df)
+coeftest(model3, vcov = vcovHC)
+```
+
+Evaluation of the business closed variable.
+```{r}
+
+df$close_period = as.integer( df$business_reopen-df$business_closed)
+head(df)
+hist(df$close_period, xlab='close period', ylab='Count', main='close length')
+df %>% 
+  ggplot(aes(x = close_period, y = case_rate_100k, color = close_period)) + 
+  geom_point() + 
+  labs(
+    title = 'Case Rate vs Mask Public Mandate', 
+    x = 'Mask Mandate in Public', 
+    y = 'Case Rate', 
+    color = 'Case Rate'
+  )
+
+model3 <- lm(case_rate_100k ~ close_period, data = df)
+coeftest(model3, vcov = vcovHC)
+
+df[is.na(df$business_closed),]
+df[42,'business_closed'] = max(df$business_closed[!is.na(df$business_reopen)])
+mean(df$business_closed)
+df$business_close_diff = df$business_closed-mean(df$business_closed)
+head(df)
+
+df %>% 
+  ggplot(aes(x = business_close_diff, y = case_rate_100k, color = close_period)) + 
+  geom_point() + 
+  labs(
+    title = 'Case Rate vs Mask Public Mandate', 
+    x = 'Mask Mandate in Public', 
+    y = 'Case Rate', 
+    color = 'Case Rate'
+  )
+
+model3 <- lm(case_rate_100k ~ business_close_diff + mask_public_bool, data = df)
+coeftest(model3, vcov = vcovHC)
+
+```
+
 ### 3. Limitations of the Model
@@
 Finally, normally distributed errors can be easily assessed by a histogram of the residuals around zero.
+Finally, normally distributed errors can be easily assessed by a histogram of the residuals around zero.

### 2020-11-16 DraftReport.Rmd
update model 3 intro

@@
 As a result, we believe mask policy for all individuals in public space can be used to effectively represents the policy impact.
-Also, it is worth to note that, the mask_public is in date format, which has value of zero or a actual date when the policy is enforced.
+    It is worth to note that, the mask_public is in date format, which has value of zero or a actual date when the policy is enforced.
 According to the documentation, zero represents "the absence of an order or directive", which can be interpreted as the policy is not enforced by the state explicitly.
 For a leaner regression, the date values in mask_public column are transformed as value 1, so that we can distinguish whether states has public mask policy or not.
+It is also noted that by transforming the variable, some important information will be lost because the actual date (early or late) to enforce the policy can also have an impact to the case rate.
+However, it is hard to measure the time effect as different states that enforced the public mask policy may have different threshold.
 

### 2020-11-16 DraftReport.Rmd
model 2 merge to draft

@@
 author: "203 Section 6, Team 1"
-output: pdf_document
+output:
+  pdf_document: default
+  html_document:
+    df_print: paged
 ---
+
+- **Common Data across models:**
+```{r setup, include=FALSE, warning=FALSE, message=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+library(lmtest)
+library(sandwich)
+library(tidyverse)
+library(magrittr)
+library(ggplot2)
+library(patchwork)
+library(stargazer)
+library(tinytex)
+```
+
+```{r variables, warning=FALSE}
+df <- read.csv("covid-19.csv", header = TRUE)
+df<-df%>%
+  rename(case_rate_100k = 'Case.Rate.per.100000',
+         death_rate_100k = 'Death.Rate.per.100000',
+         population_density = 'Population.density.per.square.miles',
+         white_pct = 'White...of.Cases',
+         black_pct = 'Black...of.Cases',
+         hispanic_pct = 'Hispanic...of.Cases',
+         other_pct = 'Other...of.Cases',
+         state_emergency = 'State.of.emergency',
+         business_closed = 'Closed.other.non.essential.businesses',
+         business_reopen = 'Began.to.reopen.businesses.statewide',
+         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
+         mask_legal='No.legal.enforcement.of.face.mask.mandate') %>%
+  select(case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
+         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, Black...of.Total.Population, White...of.Total.Population, Percent.living.under.the.federal.poverty.line..2018., Percent.Unemployed..2018., X65.)
+# Assign correct data types
+cols.num <- c("white_pct","black_pct","hispanic_pct","other_pct")
+df[cols.num] <- sapply(df[cols.num],as.numeric)
+df$state_emergency=as.Date(df$state_emergency, format = "%m/%d/%Y")
+df$business_closed=as.Date(df$business_closed, format = "%m/%d/%Y")
+df$business_reopen=as.Date(df$business_reopen, format = "%m/%d/%Y")
+head(df)
+```
 
@@
 
-- **Common Data across models:**
-```{r lib,include=FALSE}
-library(lmtest)
-library(sandwich)
-library(tidyverse)
-library(magrittr)
-library(ggplot2)
-library(patchwork)
+### Model 1 & Model 2 Building Process
+
+Part of the model aims to investigate the relationship between the covid-19 pandemic status and the state demographic information.
+The list of demographic variables included in the current dataset is as below.
+
+COVID-19 Cases by Race/Ethnicity:
+ - white % of Total Population,
+ - Black % of Total Population, 
+ - Hispanic % of Total Population, 
+ - Other % of Total Population
+
+State Characteristics:
+ - Governor,       
+ - Population density per square miles, 
+ - Population 2018, 
+ - Nonelderly Adults Who Have A Pre-Existing Condition
+ - Percent at risk for serious illness due to COVID	
+ - All-cause deaths 2018, 
+ - Number Homeless (2019), 
+ - Medicaid Expenditures as a Percent of Total State Expenditures by Fund, 
+ - Life Expectancy at Birth (years),
+ - Percent Unemployed (2018)
+ - Percent living under the federal poverty line (2018), 
+ - Weekly UI maximum amount with extra stimulus (through July 31, 2020) (dollars), 
+ - Median Annual Household Income
+
+Demographics:
+ -Children 0-18
+ - Adults 19-25
+ - Adults 26-34
+ - Adults 35-54
+ - Adults 55-64
+ - 65+
+
+Currently, Case Rate per 100000 is chosen as a dependent variable to represent the pandemic status in each state.
+Variables as Population density per square miles, Percent Unemployed (2018)/Percent living under the federal poverty line (2018), White % of of Total Population/Black % of Total Population, Percentage of 65+ are chosen as candidate covariates to explore potential connection between pandemic status demographic characteristics, economy status, racial composition.
+
+Through data explorary, most data investigated is clean as numberica type except the black percentage of total population.
+This information of some states is reported vaguely as less than 0.01, which is replaced as 0 for further work.
+This could be updated with more precise data with the help of external resources.
+
+Several transformations have been made for variables a Population density per square miles and White % of Total Population.
+
+### Data Explorary
+Through data explorary, most data investigated is clean as numberica type except the black percentage of total population.
+This information of some states is reported vaguely as less than 0.01, which is replaced as 0 for further work.
+This could be updated with more precise data with the help of external resources.
+
+Several transformations have been made for variables a Population density per square miles and White % of Total Population.
+Population density is extremely skewed due to the high population density of the District of Columbia.
+A logarithm transformation of this variable is of more uniform distribution.
+In addition, the square of the white percentage of the total population is of a distribution closer to the normal distribution.
+It also represents the chance that the interaction happens between two white people in practice.
+
+In addition, correlation values of White% and Black %, Percent Unemployed and Percent living under the federal poverty line shows these two groups of variables are correlated respectively. 
+#### 1-a) Varialbe of Case Rate per 100000 is of nearly normal distribution.
+
+```{r}
+summary(df$case_rate_100k)
+
 ```
 
-```{r variables, warning=FALSE}
-df <- read.csv("covid-19.csv", header = TRUE)
-df<-df%>%
-  rename(case_rate_100k = 'Case.Rate.per.100000',
-         death_rate_100k = 'Death.Rate.per.100000',
-         population_density = 'Population.density.per.square.miles',
-         white_pct = 'White...of.Cases',
-         black_pct = 'Black...of.Cases',
-         hispanic_pct = 'Hispanic...of.Cases',
-         other_pct = 'Other...of.Cases',
-         state_emergency = 'State.of.emergency',
-         business_closed = 'Closed.other.non.essential.businesses',
-         business_reopen = 'Began.to.reopen.businesses.statewide',
-         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
-         mask_legal='No.legal.enforcement.of.face.mask.mandate') %>%
-  select(case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
-         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal)
-# Assign correct data types
-cols.num <- c("white_pct","black_pct","hispanic_pct","other_pct")
-df[cols.num] <- sapply(df[cols.num],as.numeric)
-df$state_emergency=as.Date(df$state_emergency, format = "%m/%d/%Y")
-df$business_closed=as.Date(df$business_closed, format = "%m/%d/%Y")
-df$business_reopen=as.Date(df$business_reopen, format = "%m/%d/%Y")
-head(df)
+```{r}
+ggplot(data = df,
+  mapping = aes(x= case_rate_100k))+
+  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,6000,600)) +
+  labs(title =  "Histogram of Case rate per 100000", x = "Case per 100000", y = 'Count')+
+  scale_x_continuous(breaks=seq(0, 6000, 600))
 ```
 
-- **Regression Models:**
-    - **Model 1** Does this model only include key explanatory variables?
-Do the variables make sense given the measurement goals?
-Did the team apply reasonable transformations to these variables, to capture the nature of the relationships?
-    - **Model 2** Does this model represent a balanced approach, including variables that advance modeling goals without causing major issues?
-Does the model succeed in reducing standard errors of the key variables compared to the base model?
-Does it capture major non-linearities in the joint distribution of the variables?
-    - **Model 3**  Does this model represent a maximalist approach, erring on the side of including most variables?
-Is it still a reasonable model?
-Are there any variables that are outcomes, and should therefore still be excluded?
-Is there too much multicolinearity, to the point that the key causal effects cannot be measured?
+#### 1-a) Varialbe of Population density per squres 
+The distribution of the Population density per squares clusters at the two ends, which is caused by the extremely high population density of District of Columbia (11496.81) compared to the 50 states.
+While, the logarithm of population density is close to a normal distribution. In addition, an analysis excluding the District of Columbia is another option.
+
+```{r}
+summary(df$population_density)
+```
+```{r}
+hist_population_dens <- ggplot(data = df,
+  mapping = aes(x= population_density))+
+  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,12000,500)) +
+  labs(title =  "Histogram of Population density", x = "Population density per square miles", y = 'Count')+
+  scale_x_continuous(breaks=seq(0, 12000, 1200))
+
+hist_log_population_dens <- ggplot(data = df, aes(x = log(population_density))) + 
+  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,10,1)) + 
+  labs(title = 'Histogram of Population density', x = 'Log Population density per square miles', y = 'Count')
+
+hist_population_dens/hist_log_population_dens
+
+```
+#### 2-a) White Percentage of Total Population/ Black Percentage of Total Population
+
+Variable of black percentage is stored as string variables.
+Moreover, three entries of black percentage of total population is "<0.01", which is to be dropped or replaced by values determined by extra resources for further analysis.
+
+Here, the value of black percentage is replaced by 0 when it is "<0.01".
+Afterwards, the correlation of white percentage and black percentage is -0.42, which shows these two variables are correlated.
+Considering that the value of other race groups is much smaller, it is proper to only include white percentage in the regression model to explore the relation of care rate and racial composition.
+
+The distribution of the white % is skewed, not an ideal normal distribution.
+While the distribution of the square of white % is more close to a normal distribution.
+In practice, this square could reflect the chance that the interaction happens between two white people.
+
+```{r}
+black_num <- as.numeric(df$Black...of.Total.Population)
+black_num[is.na(black_num)]<- 0
+cor(df$White...of.Total.Population, black_num)
+```
+
+```{r}
+summary(df$White...of.Total.Population)
+```
+
+```{r}
+ggplot(data = df,
+  mapping = aes(x= White...of.Total.Population))+
+  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,1,0.05)) +
+  labs(title =  "Histogram of White Percentage", x = "White % of Total Population", y = 'Count')+
+  scale_x_continuous(breaks=seq(0, 1, 0.1))
+```
+```{r}
+ggplot(data = df,
+  mapping = aes(x= (White...of.Total.Population)^2))+
+  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,1,0.05)) +
+  labs(title =  "Histogram of White Percentage", x = "White % of Total Population^2", y = 'Count')+
+  scale_x_continuous(breaks=seq(0, 1, 0.1))
+```
+
+#### 2-b) Percent living under the federal poverty line 2018
+
+Correlation of pecent living under poverty line and white percentage of total population shows that they are not significantly related. The distribution is not heavily skewed.
+
+```{r}
+summary(df$Percent.living.under.the.federal.poverty.line..2018.)
+cor(df$Percent.living.under.the.federal.poverty.line..2018., df$White...of.Total.Population)
+```
+
+```{r}
+ggplot(data = df,
+  mapping = aes(x= Percent.living.under.the.federal.poverty.line..2018.))+
+  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(6,21,1)) +
+  labs(title =  "Histogram of Percent of poverty 2018", x = "Percent.living.under.the.federal.poverty.line", y = 'Count')+
+  scale_x_continuous(breaks=seq(6, 21, 1))
+```
+#### 2-c) Percent Unemployed 2018
+Although, the distribution is of high peak in the middle of the range.
+Overall, it is not highly skewed or heavily tailed.
+In addition, the unemployed rate is correlated to the variable to the poverty rate, which is in an agreement with intuitive expectation.
+
+```{r}
+summary(df$Percent.Unemployed..2018.)
+cor(df$Percent.Unemployed..2018., df$Percent.living.under.the.federal.poverty.line..2018.)
+```
+```{r}
+ggplot(data = df,
+  mapping = aes(x= Percent.Unemployed..2018.))+
+  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,10,1)) +
+  labs(title =  "Histogram of Unemployed percentage", x = "Percent.Unemployed..2018.", y = 'Count')+
+  scale_x_continuous(breaks=seq(0, 10, 1))
+```
+
+#### 2-d) Age 65+ Percentage
+The 65+ percentage is of a nearly normal distribution.
+
+```{r}
+summary(df$X65.)
+```
+```{r}
+ggplot(data = df,
+  mapping = aes(x= X65.))+
+  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0.10,0.22,0.01)) +
+  labs(title =  "Histogram of Percent of 65+", x = "65+ Percentage", y = 'Count')+
+  scale_x_continuous(breaks=seq(0.1, 0.25, 0.05))
+```
+
+### Plots
+No obvious trend could be easily observed or concluded from the plots regarding the dependent variables and candidate covariates.
+
+```{r}
+df %>% 
+  ggplot(aes(log(population_density), case_rate_100k, color = White...of.Total.Population)) + 
+  geom_point() +   
+  labs(
+    title = 'Relation of Case.Rate.per.100000 to Population Density', 
+    x = 'Log Population Density', 
+    y = 'Case.Rate.per.100000',
+    color = 'White %'
+  )
+```
+```{r}
+df %>% 
+  ggplot(aes(x = log(population_density), y = case_rate_100k, color = Percent.living.under.the.federal.poverty.line..2018.)) + 
+  geom_point() +   
+  labs(
+    title = 'Relation of Case.Rate.per.100000 to Population Density', 
+    x = 'log Population Density', 
+    y = 'Case.Rate.per.100000',
+    color = 'Poverty Percentage'
+  )
+```
+
+```{r}
+df %>% 
+  ggplot(aes(x = White...of.Total.Population, y = case_rate_100k, color = Percent.living.under.the.federal.poverty.line..2018.)) + 
+  geom_point() +   
+  labs(
+    title = 'Relation of Case.Rate.per.100000 to White %', 
+    x = 'White % of Total Population', 
+    y = 'Case.Rate.per.100000',
+    color = 'Poverty Percentage'
+  )
+```
+
+```{r}
+df %>% 
+  ggplot(aes(x = Percent.living.under.the.federal.poverty.line..2018., y = case_rate_100k, color = log(population_density))) + 
+  geom_point() +   
+  labs(
+    title = 'Relation of Case.Rate.per.100000 to Poverty %', 
+    x = 'Percent.living.under.the.federal.poverty.line..2018.', 
+    y = 'Case.Rate.per.100000',
+    color = 'log(Population Density)'
+  )
+```
+```{r}
+df %>% 
+  ggplot(aes(X65., case_rate_100k, color = log(population_density))) + 
+  geom_point() +   
+  labs(
+    title = 'Relation of Case.Rate.per.100000 to 65+ Percentage', 
+    x = 'X65+.', 
+    y = 'Case.Rate.per.100000',
+    color = 'log(Population density)'
+  )
+```
+
+```{r}
+df %>% 
+  ggplot(aes(Percent.Unemployed..2018., case_rate_100k, color = log(population_density))) + 
+  geom_point() +   
+  labs(
+    title = 'Relation of Case.Rate.per.100000 to Unemployed Percentage', 
+    x = 'Percent.Unemployed..2018.', 
+    y = 'Case.Rate.per.100000',
+    color = 'log(Population density)'
+  )
+```
+
+### Model Regression
+
+According to the regression results, no coefficients regarding the demographic variables investigated in model 2 are statistically significant.
+There is no statistically meaningful relationship that could be concluded by now.
+
+#### Model 1-a
+$$
+\begin{aligned}
+ Case.Rate.Per.100000 &=\beta_0 +\beta_1 Population.Density.Per.Square.Miles
+\end{aligned}
+$$
+
+```{r}
+model1a <- lm(case_rate_100k ~ population_density, data = df)
+coeftest(model1a, vcov = vcovHC)
+```
+#### Model 1-b
+$$
+Case.Rate.Per.100000 =\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)
+$$
+```{r}
+model1b <- lm(case_rate_100k ~ log(population_density) , data = df)
+coeftest(model1b, vcov = vcovHC)
+```
+
+### Model-2 Regression
+#### Model 2-a
+$$
+\begin{aligned}
+ Case.Rate.Per.100000 =\beta_0 &+\beta_1 log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.total. population)
+\end{aligned}
+$$
+
+```{r}
+model2a <- lm(case_rate_100k ~ log(population_density)+ White...of.Total.Population, data = df)
+coeftest(model2a, vcov = vcovHC)
+
+```
+
+$$
+\begin{aligned}
+ Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.total. population)^2
+\end{aligned}
+$$
+```{r}
+model2a <- lm(case_rate_100k ~ log(population_density)+ I((White...of.Total.Population)^2), data = df)
+coeftest(model2a, vcov = vcovHC)
+```
+$$
+\begin{aligned}
+ Case.Rate.Per.100000 &=\beta_0 + \beta_2(White.Percentage.of.total. population)^2
+\end{aligned}
+$$
+```{r}
+model2a <- lm(case_rate_100k ~ I((White...of.Total.Population)^2), data = df)
+coeftest(model2a, vcov = vcovHC)
+```
+
+#### Model 2-b
+$$
+\begin{aligned}
+ Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.total. population)\\ &+\beta_3(Percen.living.under.the.federal.poverty.line.2018)
+\end{aligned}
+$$
+
+```{r}
+model2b <- lm(case_rate_100k ~log(population_density)+ White...of.Total.Population+Percent.living.under.the.federal.poverty.line..2018., data = df)
+coeftest(model2b, vcov = vcovHC)
+```
+
+$$
+\begin{aligned}
+ Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.total. population)^2\\ &+\beta_3(Percen.living.under.the.federal.poverty.line.2018)
+\end{aligned}
+$$
+```{r}
+model2b <- lm(case_rate_100k ~log(population_density)+ I((White...of.Total.Population)^2) + Percent.living.under.the.federal.poverty.line..2018., data = df)
+coeftest(model2b, vcov = vcovHC)
+```
+$$
+\begin{aligned}
+ Case.Rate.Per.100000 &=\beta_0 +\beta_3(Percen.living.under.the.federal.poverty.line.2018)
+\end{aligned}
+$$
+```{r}
+model2b <- lm(case_rate_100k ~ Percent.living.under.the.federal.poverty.line..2018., data = df)
+coeftest(model2b, vcov = vcovHC)
+```
+
+#### Model 2-c
+
+$$
+\begin{aligned}
+ Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.total. population)^2\\ &+\beta_4(Percent.Unemployed..2018.)
+\end{aligned}
+$$
+```{r}
+model2c <- lm(case_rate_100k ~log(population_density)+ I((White...of.Total.Population)^2) + Percent.Unemployed..2018., data = df)
+coeftest(model2c, vcov = vcovHC)
+
+```
+$$
+\begin{aligned}
+ Case.Rate.Per.100000 &=\beta_0+\beta_4(Percent.Unemployed..2018.)
+\end{aligned}
+$$
+```{r}
+model2c <- lm(case_rate_100k ~ Percent.Unemployed..2018., data = df)
+coeftest(model2c, vcov = vcovHC)
+```
+
+#### Model 2-d
+
+$$
+\begin{aligned}
+Case.Rate.Per.100000=\beta_0 &+\beta_1log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.total.population)\\&+\beta_3(Percen.living.under.the.federal.poverty.line.2018)\\&+\beta_5(65+)
+\end{aligned}
+$$
+```{r}
+model2d <- lm(case_rate_100k ~log(population_density)+ White...of.Total.Population+Percent.living.under.the.federal.poverty.line..2018.+X65., data = df)
+coeftest(model2d, vcov = vcovHC)
+```
+## Model 2-d
+$$
+\begin{aligned}
+Case.Rate.Per.100000=\beta_0 &+\beta_1log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.total.population)^2\\&+\beta_3(Percen.living.under.the.federal.poverty.line.2018)\\&+\beta_4(65+)
+\end{aligned}
+$$
+```{r}
+model2d <- lm(case_rate_100k ~log(population_density)+ I((White...of.Total.Population)^2) + Percent.living.under.the.federal.poverty.line..2018.+X65., data = df)
+coeftest(model2d, vcov = vcovHC)
+```
+
+$$
+\begin{aligned}
+Case.Rate.Per.100000=\beta_0+\beta_4(65+)
+\end{aligned}
+$$
+```{r}
+model2d <- lm(case_rate_100k ~X65., data = df)
+coeftest(model2d, vcov = vcovHC)
+```
+
+### Model 3 Building Process
 
@@
 Finally, normally distributed errors can be easily assessed by a histogram of the residuals around zero.
+
+#### Limitations of Model 2
+
+The I.I.D assumption for CLM is not satisfied perfectly for all the models discussed above, since the states are not isolated from each other.
+It is impossible that states suffered from pandemic at different levels independently.
+Actually, the pandemic status across states tends to be highly related to each other.
+
+In addition, all states are involved in the same market economy system even though there are economic performance divisions across the states.
+Moreover, in most states white race is still dominant in the population.
+Many states share the similar racial composition.

### 2020-11-17 DraftReport.Rmd
Added model 1, draft regression table, and conclusion to report

@@
 output:
-  pdf_document: default
   html_document:
     df_print: paged
+  pdf_document: default
 ---
 
-- **Common Data across models:**
+ **Initial reassignment of common data used across models:**
 ```{r setup, include=FALSE, warning=FALSE, message=FALSE}
@@
 
-```{r variables, warning=FALSE}
+```{r variables, warning=FALSE, results='hide'}
 df <- read.csv("covid-19.csv", header = TRUE)
@@
          mask_legal='No.legal.enforcement.of.face.mask.mandate') %>%
-  select(case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
-         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, Black...of.Total.Population, White...of.Total.Population, Percent.living.under.the.federal.poverty.line..2018., Percent.Unemployed..2018., X65.)
+  select(State, case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
+         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, 
+         Black...of.Total.Population, White...of.Total.Population,
+         Percent.living.under.the.federal.poverty.line..2018., Percent.Unemployed..2018., X65.)
 # Assign correct data types
@@
 
-### 1. An Introduction
+## 1. An Introduction
 
@@
 
-### 2. A Model Building Process
+## 2. A Model Building Process
 
@@
 
-### Model 1 & Model 2 Building Process
+### Model Variables
 
-Part of the model aims to investigate the relationship between the covid-19 pandemic status and the state demographic information.
-The list of demographic variables included in the current dataset is as below.
+The three models aim to investigate the relationship between the COVID rate, state demographic information, and mask mandates.
+The list of variables included in the current dataset that pertain to these concepts are listed below.
 
-COVID-19 Cases by Race/Ethnicity:
+State Characteristics:
+
+  - Governor
+  - Population density per square miles
+  - Population 2018
+  - Nonelderly Adults Who Have A Pre-Existing Condition
+  - Percent at risk for serious illness due to COVID
+  - All-cause deaths 2018 
+  - Number Homeless (2019)
+  - Medicaid Expenditures as a Percent of Total State Expenditures by Fund
+  - Life Expectancy at Birth (years)
+  - Percent Unemployed (2018)
+  - Percent living under the federal poverty line (2018)
+  - Weekly UI maximum amount with extra stimulus (through July 31, 2020) (dollars)
+  - Median Annual Household Income
+  - Closed other non-essential businesses
+  - Stay at home/ shelter in place
+  - Mandate face mask use by all individuals in public spaces
+  - No legal enforcement of face mask mandate
+
+Demographics:
+
+ - Children 0-18
+ - Adults 19-25
+ - Adults 26-34
+ - Adults 35-54
+ - Adults 55-64
+ - 65+
  - white % of Total Population,
@@
 
-State Characteristics:
- - Governor,       
- - Population density per square miles, 
- - Population 2018, 
- - Nonelderly Adults Who Have A Pre-Existing Condition
- - Percent at risk for serious illness due to COVID	
- - All-cause deaths 2018, 
- - Number Homeless (2019), 
- - Medicaid Expenditures as a Percent of Total State Expenditures by Fund, 
- - Life Expectancy at Birth (years),
- - Percent Unemployed (2018)
- - Percent living under the federal poverty line (2018), 
- - Weekly UI maximum amount with extra stimulus (through July 31, 2020) (dollars), 
- - Median Annual Household Income
+Currently, Case Rate per 100,000 is chosen as a dependent variable to represent the pandemic status in each state.
+Variables such as Population density per square miles, Percent Unemployed (2018)/Percent living under the federal poverty line (2018), White % of of Total Population/Black % of Total Population, Percentage of 65+, and mask mandate policies are chosen as candidate covariates to explore potential relationships between pandemic status demographic characteristics, economic status, racial composition, and state policy.
 
-Demographics:
- -Children 0-18
- - Adults 19-25
- - Adults 26-34
- - Adults 35-54
- - Adults 55-64
- - 65+
+### Data Exploration
 
-Currently, Case Rate per 100000 is chosen as a dependent variable to represent the pandemic status in each state.
-Variables as Population density per square miles, Percent Unemployed (2018)/Percent living under the federal poverty line (2018), White % of of Total Population/Black % of Total Population, Percentage of 65+ are chosen as candidate covariates to explore potential connection between pandemic status demographic characteristics, economy status, racial composition.
-
-Through data explorary, most data investigated is clean as numberica type except the black percentage of total population.
-This information of some states is reported vaguely as less than 0.01, which is replaced as 0 for further work.
-This could be updated with more precise data with the help of external resources.
-
-Several transformations have been made for variables a Population density per square miles and White % of Total Population.
-
-### Data Explorary
-Through data explorary, most data investigated is clean as numberica type except the black percentage of total population.
+Through data exploration, most data investigated is clean as numerical type except the black percentage of total population.
 This information of some states is reported vaguely as less than 0.01, which is replaced as 0 for further work.
@@
 In addition, correlation values of White% and Black %, Percent Unemployed and Percent living under the federal poverty line shows these two groups of variables are correlated respectively.
-#### 1-a) Varialbe of Case Rate per 100000 is of nearly normal distribution.
+
+### Model 1
+
+For the first model, the relationship between the COVID case rate per 100,000 and the population density per square mile of states was analyzed.
+First, the distribution of COVID case rate per 100,000 dependent variable is examined.
 
@@
 summary(df$case_rate_100k)
-
 ```
@@
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,6000,600)) +
-  labs(title =  "Histogram of Case rate per 100000", x = "Case per 100000", y = 'Count')+
+  labs(title =  "Histogram of COVID Case Rate per 100,000", x = "Cases per 100,000", y = 'Count')+
   scale_x_continuous(breaks=seq(0, 6000, 600))
@@
 
-#### 1-a) Varialbe of Population density per squres 
-The distribution of the Population density per squares clusters at the two ends, which is caused by the extremely high population density of District of Columbia (11496.81) compared to the 50 states.
-While, the logarithm of population density is close to a normal distribution. In addition, an analysis excluding the District of Columbia is another option.
+As can be seen above, the distribution is fairly normal, and given that it has already been standardized as a rate across all states, there is no need to perform any transformations on this variable.
+Thus, the case rate per 100,000 variable can be leveraged as is as the dependent variable for all three models.
+
+Next, the distribution of population density per square mile variable is examined.
 
 ```{r}
-summary(df$population_density)
+histogram_of_pdensity <- df %>%
+  ggplot(aes(x = population_density)) + 
+  geom_histogram(fill = 'skyblue', color = 'grey30', bins = 20) + 
+  labs(
+    title = 'Distribution of Population Density per Square Mile', 
+    x = 'Population density per square mile', y = 'Count')
+
+histogram_of_pdensity
 ```
+
+As can be seen from the histogram, although most of the population density is concentrated in the 0 to 1500 range, there are some grouping of outliers that are very far from this concentration.
+When an analysis is performed, it can be seen that there is only one data sample that is the outlier, which is the District of Columbia (D.C.) with a value of 11,496.
+This is given due to the fact that D.C.
+is a district that solely consists of a large city, as mentioned as a possibility in the introduction.
+Given that this causes the data to be skewed, the logarithm is taken to scale the variable.
+Alternatively, the data point could have been droppped from the sample, but given the already small sample size, it was determined that a better approach would be to keep it within the sample.
+Once this transformation was performed, it is shown that there is a relatively normal distribution of population densities (see figure below).
+
 ```{r}
-hist_population_dens <- ggplot(data = df,
-  mapping = aes(x= population_density))+
-  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,12000,500)) +
-  labs(title =  "Histogram of Population density", x = "Population density per square miles", y = 'Count')+
-  scale_x_continuous(breaks=seq(0, 12000, 1200))
+# Find the outlier data points
+outliers <- subset(df, population_density > 4000)
+paste(outliers$State, '=', outliers$population_density)
 
-hist_log_population_dens <- ggplot(data = df, aes(x = log(population_density))) + 
-  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,10,1)) + 
-  labs(title = 'Histogram of Population density', x = 'Log Population density per square miles', y = 'Count')
+# Transform the variable by taking the logarithm and assign it to a new variable
+df <- df %>% 
+  mutate(l_population_density = log(population_density))
 
-hist_population_dens/hist_log_population_dens
+# Plot the new distribution in a histogram
+histogram_of_pdensity <- df %>%
+  ggplot(aes(x = l_population_density)) + 
+  geom_histogram(fill = 'skyblue', color = 'grey30', bins = 20) + 
+  labs(
+    title = 'Figure 2. Distribution of Population Density per Square Mile without D.C.', 
+    x = 'Log of population density per square mile', y = 'Count')
 
+histogram_of_pdensity
 ```
+
+With the appropriate variables transformed, a plot is created to show the relationship between them.
+
+```{r}
+df %>% 
+  ggplot(aes(l_population_density, case_rate_100k)) + 
+  geom_smooth(se = FALSE) +
+  geom_point() +   
+  labs(
+    title = 'COVID Case Rate due to Population Density', 
+    x = 'Log of population density per square mile', 
+    y = 'COVID case rate per 100,000'
+  )
+```
+
+From the above plot, it can be posited that there is no discernible relationship between the variables given the non-linear relationship.
+In order to test this, the following equation is used to create a regression model to determine the true relationship between the case rate and population density variables.
+
+$$
+Case.Rate.Per.100000 =\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)
+$$
+
+```{r}
+model1 <- lm(case_rate_100k ~ log(population_density) , data = df)
+coeftest(model1, vcov = vcovHC)
+```
+
+From the results of the regression, two things are identified.
+Firstly, the result of the coefficient is nowhere near significant, and therefore we fail to reject the null hypothesis that there is no correlation between case rate and population density, and there is no detectable effect of the independent variable population density.
+However, the coefficient of population density appears to be negative, which would concur with the assumption that the higher the population density, the lower the COVID case rate in a state.
+
+### Model 2
+
 #### 2-a) White Percentage of Total Population/ Black Percentage of Total Population
@@
 
-### Plots
+### Model 2 Plots
+
 No obvious trend could be easily observed or concluded from the plots regarding the dependent variables and candidate covariates.
@@
 
-### Model Regression
-
 According to the regression results, no coefficients regarding the demographic variables investigated in model 2 are statistically significant.
@@
 
-#### Model 1-a
-$$
-\begin{aligned}
- Case.Rate.Per.100000 &=\beta_0 +\beta_1 Population.Density.Per.Square.Miles
-\end{aligned}
-$$
+### Model-2 Regression
 
-```{r}
-model1a <- lm(case_rate_100k ~ population_density, data = df)
-coeftest(model1a, vcov = vcovHC)
-```
-#### Model 1-b
-$$
-Case.Rate.Per.100000 =\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)
-$$
-```{r}
-model1b <- lm(case_rate_100k ~ log(population_density) , data = df)
-coeftest(model1b, vcov = vcovHC)
-```
-
-### Model-2 Regression
 #### Model 2-a
@@
 ```{r}
-model2b <- lm(case_rate_100k ~log(population_density)+ White...of.Total.Population+Percent.living.under.the.federal.poverty.line..2018., data = df)
-coeftest(model2b, vcov = vcovHC)
+model2bi <- lm(case_rate_100k ~log(population_density)+ white_pct+Percent.living.under.the.federal.poverty.line..2018., data = df)
+coeftest(model2bi, vcov = vcovHC)
 ```
@@
 ```{r}
-model2b <- lm(case_rate_100k ~log(population_density)+ I((White...of.Total.Population)^2) + Percent.living.under.the.federal.poverty.line..2018., data = df)
-coeftest(model2b, vcov = vcovHC)
+model2bii <- lm(case_rate_100k ~log(population_density)+ I((White...of.Total.Population)^2) + Percent.living.under.the.federal.poverty.line..2018., data = df)
+coeftest(model2bii, vcov = vcovHC)
 ```
@@
 ```{r}
-model2b <- lm(case_rate_100k ~ Percent.living.under.the.federal.poverty.line..2018., data = df)
-coeftest(model2b, vcov = vcovHC)
+model2biii <- lm(case_rate_100k ~ Percent.living.under.the.federal.poverty.line..2018., data = df)
+coeftest(model2biii, vcov = vcovHC)
 ```
@@
 ```
-## Model 2-d
+
 $$
@@
 
-### Model 3 Building Process
+### Model 3
 
-    - For Model 3.
-Policy related variables will be introduced for the analytics.
+For Model 3, policy related variables will be introduced for the analytics.
 Policy variable "Mandate face mask use by all individuals in public spaces" (renamed as mask_public) is selected as control variable to predict the case rate.
@@
 ```{r}
-model3 <- lm(case_rate_100k ~ log(population_density) +white_pct+ mask_public_bool, data = df)
-coeftest(model3, vcov = vcovHC)
+model3a <- lm(case_rate_100k ~ log(population_density) + white_pct 
+              + Percent.living.under.the.federal.poverty.line..2018.
+              + mask_public_bool, data = df)
+coeftest(model3a, vcov = vcovHC)
 ```
@@
 
-model3 <- lm(case_rate_100k ~ close_period, data = df)
-coeftest(model3, vcov = vcovHC)
+model3b <- lm(case_rate_100k ~ close_period, data = df)
+coeftest(model3b, vcov = vcovHC)
 
@@
 
-model3 <- lm(case_rate_100k ~ business_close_diff + mask_public_bool, data = df)
-coeftest(model3, vcov = vcovHC)
+model3c <- lm(case_rate_100k ~ business_close_diff + mask_public_bool, data = df)
+coeftest(model3c, vcov = vcovHC)
 
@@
 
-### 3. Limitations of the Model
+## 3. Limitations of the Models
 
-The most general requirement for the linear model is that data points be independent and identically distributed (IID).
+Firstly, the most general requirement for the linear model is that data points be independent and identically distributed (IID).
 While for the most part the events of one state will primarily affect that single state, travel by individuals spreading the virus across state lines is an inevitable occurence.
 This may cause the spread of a virus within one state to be influenced by those around it, or generally by its position in the country.
+In addition in terms of demographics, all states are involved in the same market economy system even though there are economic performance divisions across the states.
+Moreover, in most states white race is still dominant in the population.
+Many states share the similar racial composition.
+
 While this is acknowledged to occur, the effect is likely minor compared to the demographic and policy data collected state by state which would affect the spread of the virus.
 Therefore it is relatively safe for this assumption to be met.
-
 Secondly, the process being modeled must be described by a linear conditional expectation function of the variables which are included, in order for the linear regression to be valid.
 This can be assessed after the fact by visualizing the model over the range of each variable alongside the data, or also by visualizing the residuals over the range of each variable.
+This can be seen across the three models.
 
-Related to the relationship of the residuals to the data, homoskedasticity can be assessed by visualizing the size of residuals across the range of each variable.
-Alternatively, it can also be done via the Breusch-Pagan test to more objectively determine whether homoskedasticity is not present in the data.
+Thirdly, related to the relationship of the residuals to the data, homoscedasticity can be assessed by visualizing the size of residuals across the range of each variable.
+Alternatively, it can also be done via the Breusch-Pagan test to more objectively determine whether homoscedasticity is not present in the data.
 
-Perfect colinearity is automatically detected by the R regression algorithms, which drop potential variables if they have complete colinearity with another variable.
+Next, perfect colinearity is automatically detected by the R regression algorithms, which drop potential variables if they have complete colinearity with another variable.
 Near perfect colinearity is more problematic to detect, but a common hint to its existence is large standard errors on colinear features.
-Conceptually, the variables which may have an important colinearity in the final model are
+Conceptually, the variables which may have an important colinearity in the final model are variables such as different percentage make ups of race and ethnicity, or related public policy.
+These variables are appropriately evaluatea and excluded throughout the model building process.
 
 Finally, normally distributed errors can be easily assessed by a histogram of the residuals around zero.
-Finally, normally distributed errors can be easily assessed by a histogram of the residuals around zero.
 
-#### Limitations of Model 2
+## 4. Regression Table
 
-The I.I.D assumption for CLM is not satisfied perfectly for all the models discussed above, since the states are not isolated from each other.
-It is impossible that states suffered from pandemic at different levels independently.
-Actually, the pandemic status across states tends to be highly related to each other.
+```{r warning=FALSE}
+se.model1 = coeftest(model1, vcov = vcovHC)[ , "Std. Error"]
+se.model2 = coeftest(model2bi, vcov = vcovHC)[ , "Std. Error"]
+#se.model3 = coeftest(model2b, vcov = vcovHC)[ , "Std. Error"]
+#se.model4 = coeftest(model2c, vcov = vcovHC)[ , "Std. Error"]
+#se.model5 = coeftest(model2d, vcov = vcovHC)[ , "Std. Error"]
+se.model6 = coeftest(model3a, vcov = vcovHC)[ , "Std. Error"]
 
-In addition, all states are involved in the same market economy system even though there are economic performance divisions across the states.
-Moreover, in most states white race is still dominant in the population.
-Many states share the similar racial composition.
+stargazer(model1, model2bi, model3a, type = "text", omit.stat = "f",
+          se = list(se.model6),
+          star.cutoffs = c(0.05, 0.01, 0.001), title = "Table 1: The relationship between COVID case rate and population density")
+
+```
+
+## 5. Conclusion
+
+In conclusion, although there is no evidence to show that although there is a direct correlation between the COVID case rate and population density as originally hypothesized in our research question, it can be seen from further development of the models that there is some correlation in other variables examined, namely the mask mandate variable.
+As can be seen from model 3, the regression model gives a mask mandate variable coefficient with p-value of 0.003058, which is highly significant.
+The negative coefficient also suggests that the original conjecture that the implementation of mask policies within a state would decrease the COVID case rate was indeed accurate.
+
+Conversely, the hypotheses that population and demographics would have an effect on the COVID case rate was unable to be accepted.
+Although upon investigating the poverty line percentage variable in model 2 there was a detection of notable adjacency to a significant result with a coefficient value of 0.06286, it was not enough to be truly significant and reject the null hypothesis.
+Similarly, the relationship of race/ethnicity as well as age on case rate was also insignificant with no apparent correlation.
+
+Given these results in conjunction with the original aim of the research question, it can be suggested that variables that can be controlled (in this case, mask policy) has a greater effect on stymieing the spread of COVID-19 than the variables that may be inherent to a state, such as population density and age/ethnicity demographics.
+This is a promising insight, as it provides opportunity for states to implement and apply policies to intervene and control the case rate without being worried about its varying effects depending on population characteristics.

### 2020-11-17 model_1.Rmd
Added model 1, draft regression table, and conclusion to report

@@
 date: "13/11/2020"
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
 ---
@@
 
-For our first model, we will analyze the relationship between the COVID case rate per 100,000 and the population density per square mile of states.
-First, we take a look at the distribution of population density per square mile.
+For the first model, the relationship between the COVID case rate per 100,000 and the population density per square mile of states was analyzed.
+First, the distribution of population density per square mile variable is examined.
 
@@
 
-As we can see from Figure 1, although most of the population density is concentrated in the 0 to 1500 range, we have some grouping of outliers that are very far from this concentration.
-When we perform an analysis, we see that there is only one data sample that is the outlier, which is the District of Columbia (D.C.).
+As can be seen from Figure 1, although most of the population density is concentrated in the 0 to 1500 range, there are some grouping of outliers that are very far from this concentration.
+When an analysis is performed, it can be seen that there is only one data sample that is the outlier, which is the District of Columbia (D.C.).
 This is given due to the fact that D.C.
 is a district that solely consists of a large city, as mentioned as a possibility in the introduction.
-Given that this causes the data to be skewed, we choose to remove this data point from the sample.
-Once removed, we see that we now have a relatively normal distribution of population densities (see Figure 2).
+Given that this causes the data to be skewed, the logarithm is taken to scale the variable.
+Once this transformation is performed, it is shown that there is a relatively normal distribution of population densities (see Figure 2).
 
@@
 
-# Remove the outlier from the dataset
-df_new <- df[ !(df$State %in% outliers$State), ]
+# Transform the variable by taking the logarithm and assign it to a new variable
+df <- df %>% 
+  mutate(l_population_density = log(Population.density.per.square.miles))
 
-# Check to see that D.C. has been removed
-paste("Outliers in the new dataframe:", outliers$State %in% df_new$State)
-
-histogram_of_pdensity <- df_new %>%
-  ggplot(aes(x = Population.density.per.square.miles)) + 
+# Plot the new distribution in a histogram
+histogram_of_pdensity <- df %>%
+  ggplot(aes(x = l_population_density)) + 
   geom_histogram(bins = 20) + 
@@
     title = 'Figure 2. Distribution of Population Density per Square Mile without D.C.', 
-    x = 'Population density per square mile', y = 'Count')
+    x = 'Log of population density per square mile', y = 'Count')
 
@@
 
-Next, we evaluate the values for case_rate
+Next, the distribution for the case rate of the states is examined for normality.
+
+```{r}
+histogram_of_case_rate <- df %>%
+  ggplot(aes(x = Case.Rate.per.100000)) + 
+  geom_histogram(bins = 20) + 
+  labs(
+    title = 'Figure 3. Distribution of Case Rate per 100,000', 
+    x = 'Case rate per 100,000', y = 'Count')
+
+histogram_of_case_rate
+```
+
+As can be seen above in Figure 3, the distribution is fairly normal, and given that it has already been standardized as a rate across all states, there is no need to perform any transformations on this variable.
 

### 2020-11-17 DraftReport.Rmd
update model 3 details

@@
 Given it has been proved that mask is effective in preventing the transmission of disease, it is expected that the mask policy mandate in public would have an effect in reducing the case rate.
+
 There are other variables related to mask such as "No legal enforcement of face mask mandate" and "Mandate face mask use by employees in public-facing businesses", we believe these variables have a lower impact than the policy that is enforced to public.
@@
 As a result, we believe mask policy for all individuals in public space can be used to effectively represents the policy impact.
+
 It is worth to note that, the mask_public is in date format, which has value of zero or a actual date when the policy is enforced.
@@
 ```
-
+#### Model 3-a
 Next, the linear model is created from model 2 with the mask policy variables added for the regression.
+The regression coefficient shows that Mask_public_bool variable is highly significant.
+The result suggests that public mask mandate policy has postive effect in bringing the case number down.
 
@@
 ```
-
-Evaluation of the business closed variable.
+#### Model 3-b&c
+Considering that there could be multiple policy variables that are relevant to predicting the case number.
+We have also examined the "Closed.other.non.essential.businesses" variable (renamed as "business_closed") which represents the date when the state issued closure of non-essential businesses.
+As "business_closed" is a date variable, a transformation was performed to convert the date into a variable "business_close_diff" which represents the days difference between the business close date of the state compared to the average date of all the states.
 ```{r}
 
-df$close_period = as.integer( df$business_reopen-df$business_closed)
+df$business_close_diff =as.numeric(df$business_closed-mean(df$business_closed, na.rm=TRUE))
 head(df)
-hist(df$close_period, xlab='close period', ylab='Count', main='close length')
+```
+
+From the plot of business close date difference against case rate, the graph shows that there is some degree of linearity between the two variables.
+```{r warning=FALSE}
 df %>% 
-  ggplot(aes(x = close_period, y = case_rate_100k, color = close_period)) + 
+  ggplot(aes(x = business_close_diff, y = case_rate_100k, color = business_close_diff)) + 
   geom_point() + 
+  geom_smooth(se = FALSE) +
   labs(
     title = 'Case Rate vs Mask Public Mandate', 
-    x = 'Mask Mandate in Public', 
+    x = 'Non-Essential Business Close Date difference compared to Average', 
     y = 'Case Rate', 
@@
   )
+```
 
-model3b <- lm(case_rate_100k ~ close_period, data = df)
+By running the model with "business_close_diff" against case rate variable, the coefficient shows that the variable is significant.
+```{r}
+model3b <- lm(case_rate_100k ~ business_close_diff, data = df)
 coeftest(model3b, vcov = vcovHC)
-
-df[is.na(df$business_closed),]
-df[42,'business_closed'] = max(df$business_closed[!is.na(df$business_reopen)])
-mean(df$business_closed)
-df$business_close_diff = df$business_closed-mean(df$business_closed)
-head(df)
-
-df %>% 
-  ggplot(aes(x = business_close_diff, y = case_rate_100k, color = close_period)) + 
-  geom_point() + 
-  labs(
-    title = 'Case Rate vs Mask Public Mandate', 
-    x = 'Mask Mandate in Public', 
-    y = 'Case Rate', 
-    color = 'Case Rate'
-  )
-
+```
+However, the regression with both policy variables "mask_pubic_bool" and "business_close_diff" shows that neither the the variable is significant. 
+```
 model3c <- lm(case_rate_100k ~ business_close_diff + mask_public_bool, data = df)
 coeftest(model3c, vcov = vcovHC)
+```
 
-```
+The findings above suggests that there is colinearity between the two variables "mask_pubic_bool" and "business_close_diff".
+Therefore, we decided to use "mask_public_bool" variable which has higher significance to represent our policy variable.
+As a result, model3a is selected as the final model for model 3.
 

### 2020-11-17 model_1.Rmd

@@
 
-```{r setup, include=FALSE}
+```{r setup, include=FALSE, warning=FALSE, message=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
-
 library(lmtest)
@@
 library(magrittr)
-
-df <- read.csv("covid-19.csv", header = TRUE)
+library(ggplot2)
+library(patchwork)
+library(stargazer)
+library(tinytex)
 ```
 
+```{r variables, warning=FALSE, results='hide'}
+df <- read.csv("covid-19.csv", header = TRUE)
+df<-df%>%
+  rename(case_rate_100k = 'Case.Rate.per.100000',
+         death_rate_100k = 'Death.Rate.per.100000',
+         population_density = 'Population.density.per.square.miles',
+         white_pct = 'White...of.Cases',
+         black_pct = 'Black...of.Cases',
+         hispanic_pct = 'Hispanic...of.Cases',
+         other_pct = 'Other...of.Cases',
+         state_emergency = 'State.of.emergency',
+         business_closed = 'Closed.other.non.essential.businesses',
+         business_reopen = 'Began.to.reopen.businesses.statewide',
+         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
+         mask_legal='No.legal.enforcement.of.face.mask.mandate') %>%
+  select(State, case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
+         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, 
+         Black...of.Total.Population, White...of.Total.Population,
+         Percent.living.under.the.federal.poverty.line..2018., Percent.Unemployed..2018., X65.)
+# Assign correct data types
+cols.num <- c("white_pct","black_pct","hispanic_pct","other_pct")
+df[cols.num] <- sapply(df[cols.num],as.numeric)
+df$state_emergency=as.Date(df$state_emergency, format = "%m/%d/%Y")
+df$business_closed=as.Date(df$business_closed, format = "%m/%d/%Y")
+df$business_reopen=as.Date(df$business_reopen, format = "%m/%d/%Y")
+head(df)
+```
+
+### Model 1
+
 For the first model, the relationship between the COVID case rate per 100,000 and the population density per square mile of states was analyzed.
-First, the distribution of population density per square mile variable is examined.
+First, the distribution of COVID case rate per 100,000 dependent variable is examined.
+
+```{r}
+summary(df$case_rate_100k)
+```
+
+```{r}
+ggplot(data = df,
+  mapping = aes(x= case_rate_100k))+
+  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,6000,600)) +
+  labs(title =  "Histogram of COVID Case Rate per 100,000", x = "Cases per 100,000", y = 'Count')+
+  scale_x_continuous(breaks=seq(0, 6000, 600))
+```
+
+As can be seen above, the distribution is fairly normal, and given that it has already been standardized as a rate across all states, there is no need to perform any transformations on this variable.
+Thus, the case rate per 100,000 variable can be leveraged as is as the dependent variable for all three models.
+
+Next, the distribution of population density per square mile variable is examined.
 
@@
 histogram_of_pdensity <- df %>%
-  ggplot(aes(x = Population.density.per.square.miles)) + 
-  geom_histogram(bins = 20) + 
+  ggplot(aes(x = population_density)) + 
+  geom_histogram(fill = 'skyblue', color = 'grey30', bins = 20) + 
   labs(
-    title = 'Figure 1. Distribution of Population Density per Square Mile', 
+    title = 'Distribution of Population Density per Square Mile', 
     x = 'Population density per square mile', y = 'Count')
@@
 
-As can be seen from Figure 1, although most of the population density is concentrated in the 0 to 1500 range, there are some grouping of outliers that are very far from this concentration.
-When an analysis is performed, it can be seen that there is only one data sample that is the outlier, which is the District of Columbia (D.C.).
+As can be seen from the histogram, although most of the population density is concentrated in the 0 to 1500 range, there are some grouping of outliers that are very far from this concentration.
+When an analysis is performed, it can be seen that there is only one data sample that is the outlier, which is the District of Columbia (D.C.) with a value of 11,496.
 This is given due to the fact that D.C.
@@
 Given that this causes the data to be skewed, the logarithm is taken to scale the variable.
-Once this transformation is performed, it is shown that there is a relatively normal distribution of population densities (see Figure 2).
+Alternatively, the data point could have been droppped from the sample, but given the already small sample size, it was determined that a better approach would be to keep it within the sample.
+Once this transformation was performed, it is shown that there is a relatively normal distribution of population densities (see figure below).
 
@@
 # Find the outlier data points
-outliers <- subset(df,Population.density.per.square.miles > 4000)
-outliers
+outliers <- subset(df, population_density > 4000)
+paste(outliers$State, '=', outliers$population_density)
 
@@
 df <- df %>% 
-  mutate(l_population_density = log(Population.density.per.square.miles))
+  mutate(l_population_density = log(population_density))
 
@@
   ggplot(aes(x = l_population_density)) + 
-  geom_histogram(bins = 20) + 
+  geom_histogram(fill = 'skyblue', color = 'grey30', bins = 20) + 
   labs(
@@
 
-Next, the distribution for the case rate of the states is examined for normality.
-
-```{r}
-histogram_of_case_rate <- df %>%
-  ggplot(aes(x = Case.Rate.per.100000)) + 
-  geom_histogram(bins = 20) + 
-  labs(
-    title = 'Figure 3. Distribution of Case Rate per 100,000', 
-    x = 'Case rate per 100,000', y = 'Count')
-
-histogram_of_case_rate
-```
-
-As can be seen above in Figure 3, the distribution is fairly normal, and given that it has already been standardized as a rate across all states, there is no need to perform any transformations on this variable.
-
-```{r variables}
-
-df_new %>% 
-  ggplot(aes(x = Case.Rate.per.100000, y = Population.density.per.square.miles)) + 
-  geom_point()
-
-df_new %>% 
-  ggplot(aes(x = Death.Rate.per.100000, y = Population.density.per.square.miles)) + 
-  geom_point() 
-```
-```{r}
-model_current <- lm(Case.Rate.per.100000 ~ Population.density.per.square.miles , data = df_new, na.action = na.omit)
-coeftest(model_current, vcov = vcovHC)
-
-model_alternative <- lm(Death.Rate.per.100000 ~ Population.density.per.square.miles , data = df_new, na.action = na.omit)
-coeftest(model_alternative, vcov = vcovHC)
-```
-
-```{r}
-
-df<-df%>%
-  rename(case_rate_100k = 'Case.Rate.per.100000',
-         population_density = 'Population.density.per.square.miles',
-         white_pct = 'White...of.Cases',
-         black_pct = 'Black...of.Cases',
-         hispanic_pct = 'Hispanic...of.Cases',
-         other_pct = 'Other...of.Cases',
-         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
-         mask_legal='No.legal.enforcement.of.face.mask.mandate') %>%
-  select(case_rate_100k, population_density, white_pct, black_pct, hispanic_pct, 
-         other_pct, mask_public, mask_legal)
-head(df)
-```
-
-You will next build a set of models to investigate your research question, documenting your decisions.  Here are some things to keep in mind during your model building process:
-
-1.
-*What do you want to measure*?
-Make sure you identify one, or a few, variables that will allow you to derive conclusions relevant to your research question, and include those variables in all model specifications.
-2. Is your modeling goal one of description or explanation? 
-3.
-What [covariates](https://en.wikipedia.org/wiki/Dependent_and_independent_variables#Statistics_synonyms) help you achieve your modeling goals?
-What covariates are problematic, either due to *collinearity*, or because they are outcomes that will absorb some of a causal effect you want to measure?
-4. What *transformations*, if any, should you apply to each variable?  These transformations might reveal linearities in scatterplots, make your results relevant, or help you meet model assumptions.
-5.
-Are your choices supported by exploratory data analysis (*EDA*)?
-You will likely start with some general EDA to *detect anomalies* (missing values, top-coded variables, etc.).
-From then on, your EDA should be interspersed with your model building.
-Use visual tools to *guide* your decisions.
-You can also leverage statistical *tests* to help assess whether variables, or groups of variables, are improving model fit.
-
-At the same time, it is important to remember that you are not trying to create one perfect model.
-You will create several specifications, giving the reader a sense of how robust (or sensitive) your results are to modeling choices, and to show that you're not just cherry-picking the specification that leads to the largest effects.
-
-At a minimum, you should include the following three specifications:
-
-1.
-**Model 1**: One model with *only the key variables* you want to measure (possibly transformed, as determined by your EDA), and no other covariates (or perhaps one, or at most two, covariates if they are so crucial that it would be unreasonable to omit them)
+With the appropriate variables transformed, a plot is created to show the relationship between them.
 
@@
 df %>% 
-  ggplot(aes(x = case_rate_100k, y = log(population))) + 
-  geom_point() 
+  ggplot(aes(l_population_density, case_rate_100k)) + 
+  geom_smooth(se = FALSE) +
+  geom_point() +   
+  labs(
+    title = 'COVID Case Rate due to Population Density', 
+    x = 'Log of population density per square mile', 
+    y = 'COVID case rate per 100,000'
+  )
 ```
+
+From the above plot, it can be posited that there is no discernible relationship between the variables given the non-linear relationship.
+In order to test this, the following equation is used to create a regression model to determine the true relationship between the case rate and population density variables.
+
+$$
+Case.Rate.Per.100000 =\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)
+$$
+
+```{r}
+model1 <- lm(case_rate_100k ~ log(population_density) , data = df)
+coeftest(model1, vcov = vcovHC)
+```
+
+From the results of the regression, two things are identified.
+Firstly, the result of the coefficient is nowhere near significant, and therefore we fail to reject the null hypothesis that there is no correlation between case rate and population density, and there is no detectable effect of the independent variable population density.
+However, the coefficient of population density appears to be negative, which would concur with the assumption that the higher the population density, the lower the COVID case rate in a state.

### 2020-11-17 DraftReport.Rmd
Updated the draft report with meeting discussions

@@
 df <- df %>% 
-  mutate(l_population_density = log(population_density))
+  mutate(l_population_density = log10(population_density))
 
@@
 ```{r}
-model1 <- lm(case_rate_100k ~ log(population_density) , data = df)
+model1 <- lm(case_rate_100k ~ log10(population_density) , data = df)
 coeftest(model1, vcov = vcovHC)
@@
 df %>% 
-  ggplot(aes(log(population_density), case_rate_100k, color = White...of.Total.Population)) + 
+  ggplot(aes(log10(population_density), case_rate_100k, color = White...of.Total.Population)) + 
   geom_point() +   
@@
 df %>% 
-  ggplot(aes(x = log(population_density), y = case_rate_100k, color = Percent.living.under.the.federal.poverty.line..2018.)) + 
+  ggplot(aes(x = log10(population_density), y = case_rate_100k, color = Percent.living.under.the.federal.poverty.line..2018.)) + 
   geom_point() +   
@@
 df %>% 
-  ggplot(aes(x = Percent.living.under.the.federal.poverty.line..2018., y = case_rate_100k, color = log(population_density))) + 
+  ggplot(aes(x = Percent.living.under.the.federal.poverty.line..2018., y = case_rate_100k, color = log10(population_density))) + 
   geom_point() +   
@@
 df %>% 
-  ggplot(aes(X65., case_rate_100k, color = log(population_density))) + 
+  ggplot(aes(X65., case_rate_100k, color = log10(population_density))) + 
   geom_point() +   
@@
 df %>% 
-  ggplot(aes(Percent.Unemployed..2018., case_rate_100k, color = log(population_density))) + 
+  ggplot(aes(Percent.Unemployed..2018., case_rate_100k, color = log10(population_density))) + 
   geom_point() +   
@@
 ```{r}
-model2a <- lm(case_rate_100k ~ log(population_density)+ White...of.Total.Population, data = df)
+model2a <- lm(case_rate_100k ~ log10(population_density)+ White...of.Total.Population, data = df)
 coeftest(model2a, vcov = vcovHC)
@@
 ```{r}
-model2a <- lm(case_rate_100k ~ log(population_density)+ I((White...of.Total.Population)^2), data = df)
-coeftest(model2a, vcov = vcovHC)
-```
-$$
-\begin{aligned}
- Case.Rate.Per.100000 &=\beta_0 + \beta_2(White.Percentage.of.total. population)^2
-\end{aligned}
-$$
-```{r}
-model2a <- lm(case_rate_100k ~ I((White...of.Total.Population)^2), data = df)
+model2a <- lm(case_rate_100k ~ log10(population_density)+ I((White...of.Total.Population)^2), data = df)
 coeftest(model2a, vcov = vcovHC)
@@
 #### Model 2-b
-$$
-\begin{aligned}
- Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.total. population)\\ &+\beta_3(Percen.living.under.the.federal.poverty.line.2018)
-\end{aligned}
-$$
-
-```{r}
-model2bi <- lm(case_rate_100k ~log(population_density)+ white_pct+Percent.living.under.the.federal.poverty.line..2018., data = df)
-coeftest(model2bi, vcov = vcovHC)
-```
 
@@
 \begin{aligned}
- Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.total. population)^2\\ &+\beta_3(Percen.living.under.the.federal.poverty.line.2018)
+ Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles) + \beta_3(Percen.living.under.the.federal.poverty.line.2018)
 \end{aligned}
@@
 ```{r}
-model2bii <- lm(case_rate_100k ~log(population_density)+ I((White...of.Total.Population)^2) + Percent.living.under.the.federal.poverty.line..2018., data = df)
-coeftest(model2bii, vcov = vcovHC)
-```
-$$
-\begin{aligned}
- Case.Rate.Per.100000 &=\beta_0 +\beta_3(Percen.living.under.the.federal.poverty.line.2018)
-\end{aligned}
-$$
-```{r}
-model2biii <- lm(case_rate_100k ~ Percent.living.under.the.federal.poverty.line..2018., data = df)
-coeftest(model2biii, vcov = vcovHC)
+model2b <- lm(case_rate_100k ~ log10(population_density) + Percent.living.under.the.federal.poverty.line..2018., data = df)
+coeftest(model2b, vcov = vcovHC)
 ```
@@
 \begin{aligned}
- Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.total. population)^2\\ &+\beta_4(Percent.Unemployed..2018.)
+ Case.Rate.Per.100000 &=\beta_0+\beta_1 log(Population.Density.Per.Square.Miles)+\beta_4(Percent.Unemployed..2018.)
 \end{aligned}
@@
 ```{r}
-model2c <- lm(case_rate_100k ~log(population_density)+ I((White...of.Total.Population)^2) + Percent.Unemployed..2018., data = df)
-coeftest(model2c, vcov = vcovHC)
-
-```
-$$
-\begin{aligned}
- Case.Rate.Per.100000 &=\beta_0+\beta_4(Percent.Unemployed..2018.)
-\end{aligned}
-$$
-```{r}
-model2c <- lm(case_rate_100k ~ Percent.Unemployed..2018., data = df)
+model2c <- lm(case_rate_100k ~ log10(population_density) + Percent.Unemployed..2018., data = df)
 coeftest(model2c, vcov = vcovHC)
@@
 \begin{aligned}
-Case.Rate.Per.100000=\beta_0 &+\beta_1log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.total.population)\\&+\beta_3(Percen.living.under.the.federal.poverty.line.2018)\\&+\beta_5(65+)
+Case.Rate.Per.100000=\beta_0+\beta_1log(Population.Density.Per.Square.Miles)+\beta_4(65+)
 \end{aligned}
@@
 ```{r}
-model2d <- lm(case_rate_100k ~log(population_density)+ White...of.Total.Population+Percent.living.under.the.federal.poverty.line..2018.+X65., data = df)
-coeftest(model2d, vcov = vcovHC)
-```
-
-$$
-\begin{aligned}
-Case.Rate.Per.100000=\beta_0 &+\beta_1log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.total.population)^2\\&+\beta_3(Percen.living.under.the.federal.poverty.line.2018)\\&+\beta_4(65+)
-\end{aligned}
-$$
-```{r}
-model2d <- lm(case_rate_100k ~log(population_density)+ I((White...of.Total.Population)^2) + Percent.living.under.the.federal.poverty.line..2018.+X65., data = df)
-coeftest(model2d, vcov = vcovHC)
-```
-
-$$
-\begin{aligned}
-Case.Rate.Per.100000=\beta_0+\beta_4(65+)
-\end{aligned}
-$$
-```{r}
-model2d <- lm(case_rate_100k ~X65., data = df)
+model2d <- lm(case_rate_100k ~log10(population_density)+ X65., data = df)
 coeftest(model2d, vcov = vcovHC)
@@
 ```{r}
-model3a <- lm(case_rate_100k ~ log(population_density) + white_pct 
+model3a <- lm(case_rate_100k ~ log10(population_density)
               + Percent.living.under.the.federal.poverty.line..2018.
@@
 se.model1 = coeftest(model1, vcov = vcovHC)[ , "Std. Error"]
-se.model2 = coeftest(model2bi, vcov = vcovHC)[ , "Std. Error"]
-#se.model3 = coeftest(model2b, vcov = vcovHC)[ , "Std. Error"]
-#se.model4 = coeftest(model2c, vcov = vcovHC)[ , "Std. Error"]
-#se.model5 = coeftest(model2d, vcov = vcovHC)[ , "Std. Error"]
-se.model6 = coeftest(model3a, vcov = vcovHC)[ , "Std. Error"]
+se.model2 = coeftest(model2b, vcov = vcovHC)[ , "Std. Error"]
+se.model3 = coeftest(model3a, vcov = vcovHC)[ , "Std. Error"]
 
-stargazer(model1, model2bi, model3a, type = "text", omit.stat = "f",
-          se = list(se.model6),
+stargazer(model1, model2b, model3a, type = "text", omit.stat = "f",
+          se = list(se.model3),
           star.cutoffs = c(0.05, 0.01, 0.001), title = "Table 1: The relationship between COVID case rate and population density")
@@
 ```
+
+- Discuss variables included and not included in the model
+- Talk about Adjusted R2
+- Talk about significant variables in the table
+- Provide a statement that the table may change in the final report with more variables and maybe different key variables (case rate and population density)
 

### 2020-11-17 DraftReport.Rmd
variable rename update model2, model3

@@
          mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
-         mask_legal='No.legal.enforcement.of.face.mask.mandate') %>%
+         mask_legal='No.legal.enforcement.of.face.mask.mandate',
+         black_population_pct ="Black...of.Total.Population",
+         white_population_pct = "White...of.Total.Population",
+         poverty_pct = "Percent.living.under.the.federal.poverty.line..2018.",
+         unemployed_pct = "Percent.Unemployed..2018.",
+         senior_pct = "X65.") %>%
   select(State, case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
          other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, 
-         Black...of.Total.Population, White...of.Total.Population,
-         Percent.living.under.the.federal.poverty.line..2018., Percent.Unemployed..2018., X65.)
+         black_population_pct, white_population_pct, poverty_pct, unemployed_pct, senior_pct)
 # Assign correct data types
@@
 
- - Children 0-18
- - Adults 19-25
- - Adults 26-34
- - Adults 35-54
- - Adults 55-64
- - 65+
- - white % of Total Population,
- - Black % of Total Population, 
- - Hispanic % of Total Population, 
- - Other % of Total Population
+ - Children 0-18 Percentage
+ - Adults 19-25 Percentage
+ - Adults 26-34 Percentage
+ - Adults 35-54 Percentage
+ - Adults 55-64 Percentage
+ - 65+ Percentage
+ - white Percentage of Total Population,
+ - Black Percentage of Total Population, 
+ - Hispanic Percentage of Total Population, 
+ - Other Percentage of Total Population
 
-Currently, Case Rate per 100,000 is chosen as a dependent variable to represent the pandemic status in each state.
-Variables such as Population density per square miles, Percent Unemployed (2018)/Percent living under the federal poverty line (2018), White % of of Total Population/Black % of Total Population, Percentage of 65+, and mask mandate policies are chosen as candidate covariates to explore potential relationships between pandemic status demographic characteristics, economic status, racial composition, and state policy.
+Currently, Case rate per 100,000 is chosen as a dependent variable to represent the pandemic status in each state.
+Variables such as population density per square miles, unemployed (2018)/percentage living under the federal poverty line (2018), white/black percentage of of total Population, percentage of seniors(65+), and mask mandate policies are chosen as candidate covariates to explore potential relationships between pandemic status demographic characteristics, economic status, racial composition, and state policy.
 
@@
 
-In addition, correlation values of White% and Black %, Percent Unemployed and Percent living under the federal poverty line shows these two groups of variables are correlated respectively.
+In addition, correlation values of white percentage and black percentage, unemployed percentage and percentage living under the federal poverty line shows these two groups of variables are correlated respectively.
 
@@
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,6000,600)) +
-  labs(title =  "Histogram of COVID Case Rate per 100,000", x = "Cases per 100,000", y = 'Count')+
+  labs(title =  "Histogram of COVID Case Rate per 100,000", x = "COVID case rate per 100,000", y = 'Count')+
   scale_x_continuous(breaks=seq(0, 6000, 600))
@@
   labs(
-    title = 'Distribution of Population Density per Square Mile', 
+    title = 'Distribution of Population Density per Square Miles', 
     x = 'Population density per square mile', y = 'Count')
@@
   labs(
-    title = 'Figure 2. Distribution of Population Density per Square Mile without D.C.', 
-    x = 'Log of population density per square mile', y = 'Count')
+    title = 'Figure 2. Distribution of Population Density per Square Miles', 
+    x = 'Log 10 of population density per square mile', y = 'Count')
 
@@
     title = 'COVID Case Rate due to Population Density', 
-    x = 'Log of population density per square mile', 
+    x = 'Log 10 of population density per square miles', 
     y = 'COVID case rate per 100,000'
@@
 ```{r}
-black_num <- as.numeric(df$Black...of.Total.Population)
+black_num <- as.numeric(df$black_population_pct)
 black_num[is.na(black_num)]<- 0
-cor(df$White...of.Total.Population, black_num)
+cor(df$white_population_pct, black_num)
 ```
@@
 ```{r}
-summary(df$White...of.Total.Population)
+summary(df$white_population_pct)
 ```
@@
 ggplot(data = df,
-  mapping = aes(x= White...of.Total.Population))+
+  mapping = aes(x= white_population_pct))+
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,1,0.05)) +
-  labs(title =  "Histogram of White Percentage", x = "White % of Total Population", y = 'Count')+
+  labs(title =  "Histogram of White Population Percentage", x = "White percentage of total population", y = 'Count')+
   scale_x_continuous(breaks=seq(0, 1, 0.1))
@@
 ggplot(data = df,
-  mapping = aes(x= (White...of.Total.Population)^2))+
+  mapping = aes(x= (white_population_pct)^2))+
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,1,0.05)) +
-  labs(title =  "Histogram of White Percentage", x = "White % of Total Population^2", y = 'Count')+
+  labs(title =  "Histogram of Square of White Percentage", x = "White percentage of total population^2", y = 'Count')+
   scale_x_continuous(breaks=seq(0, 1, 0.1))
@@
 
-#### 2-b) Percent living under the federal poverty line 2018
+#### 2-b) Percentage living under the federal poverty line(2018)
 
-Correlation of pecent living under poverty line and white percentage of total population shows that they are not significantly related. The distribution is not heavily skewed.
+Correlation of pecentage living under poverty line and white percentage of total population shows that they are not significantly related.
+The distribution of the poverty percentage is not heavily skewed.
 
 ```{r}
-summary(df$Percent.living.under.the.federal.poverty.line..2018.)
-cor(df$Percent.living.under.the.federal.poverty.line..2018., df$White...of.Total.Population)
+summary(df$poverty_pct)
+cor(df$poverty_pct, df$white_population_pct)
 ```
@@
 ggplot(data = df,
-  mapping = aes(x= Percent.living.under.the.federal.poverty.line..2018.))+
+  mapping = aes(x= poverty_pct))+
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(6,21,1)) +
-  labs(title =  "Histogram of Percent of poverty 2018", x = "Percent.living.under.the.federal.poverty.line", y = 'Count')+
+  labs(title =  "Histogram of Poverty Percentage", x = "Percentage living under poverty line(2018)", y = 'Count')+
   scale_x_continuous(breaks=seq(6, 21, 1))
 ```
-#### 2-c) Percent Unemployed 2018
+#### 2-c) Unemployed Percentage (2018)
 Although, the distribution is of high peak in the middle of the range.
@@
 ```{r}
-summary(df$Percent.Unemployed..2018.)
-cor(df$Percent.Unemployed..2018., df$Percent.living.under.the.federal.poverty.line..2018.)
+summary(df$unemployed_pct)
+cor(df$unemployed_pct, df$poverty_pct)
 ```
@@
 ggplot(data = df,
-  mapping = aes(x= Percent.Unemployed..2018.))+
+  mapping = aes(x= unemployed_pct))+
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,10,1)) +
-  labs(title =  "Histogram of Unemployed percentage", x = "Percent.Unemployed..2018.", y = 'Count')+
+  labs(title =  "Histogram of Unemployed Percentage", x = "Unemployed percentage (2018)", y = 'Count')+
   scale_x_continuous(breaks=seq(0, 10, 1))
@@
 
-#### 2-d) Age 65+ Percentage
-The 65+ percentage is of a nearly normal distribution.
+#### 2-d) Seniors Percentage(Age 65+)
+The seniors percentage (age 65+) is of a nearly normal distribution.
 
 ```{r}
-summary(df$X65.)
+summary(df$senior_pct)
 ```
@@
 ggplot(data = df,
-  mapping = aes(x= X65.))+
+  mapping = aes(x= senior_pct))+
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0.10,0.22,0.01)) +
-  labs(title =  "Histogram of Percent of 65+", x = "65+ Percentage", y = 'Count')+
+  labs(title =  "Histogram of Seniors Percentage (65+)", x = "Seniors percentage", y = 'Count')+
   scale_x_continuous(breaks=seq(0.1, 0.25, 0.05))
@@
 df %>% 
-  ggplot(aes(log10(population_density), case_rate_100k, color = White...of.Total.Population)) + 
+  ggplot(aes(white_population_pct, case_rate_100k, color = log10(population_density))) + 
   geom_point() + 
+  geom_smooth(se=FALSE)+
   labs(
-    title = 'Relation of Case.Rate.per.100000 to Population Density', 
-    x = 'Log Population Density', 
-    y = 'Case.Rate.per.100000',
-    color = 'White %'
+    title = 'Relation of Case Rate Per 100000 to White Population Percentage', 
+    x = 'White percentage of total population', 
+    y = 'Case rate per 100000',
+    color = 'Log 10 of population density'
   )
@@
 df %>% 
-  ggplot(aes(x = log10(population_density), y = case_rate_100k, color = Percent.living.under.the.federal.poverty.line..2018.)) + 
+  ggplot(aes(x = white_population_pct, y = case_rate_100k, color = poverty_pct)) + 
   geom_point() +
+  geom_smooth(se=FALSE)+
   labs(
-    title = 'Relation of Case.Rate.per.100000 to Population Density', 
-    x = 'log Population Density', 
-    y = 'Case.Rate.per.100000',
-    color = 'Poverty Percentage'
+    title = 'Relation of Case Rate per 100000 to White Population Percentage', 
+    x = 'White percentage of total population', 
+    y = 'Case rate per 100000',
+    color = 'Poverty percentage'
+  )
+```
+```{r}
+df %>% 
+  ggplot(aes(x = poverty_pct , y = case_rate_100k, color = log10(population_density))) + 
+  geom_point() + 
+  geom_smooth(se=FALSE)+
+  labs(
+    title = 'Relation of Case Rate per 100000 to Poverty Percentage', 
+    x = 'Poverty Percentage', 
+    y = 'Case rate per 100000',
+    color = 'Log 10 of population density'
   )
@@
 df %>% 
-  ggplot(aes(x = White...of.Total.Population, y = case_rate_100k, color = Percent.living.under.the.federal.poverty.line..2018.)) + 
+  ggplot(aes(unemployed_pct, case_rate_100k, color = log10(population_density))) + 
   geom_point() + 
+  geom_smooth(se=FALSE)+
   labs(
-    title = 'Relation of Case.Rate.per.100000 to White %', 
-    x = 'White % of Total Population', 
-    y = 'Case.Rate.per.100000',
-    color = 'Poverty Percentage'
+    title = 'Relation of Case Rate per 100000 to Unemployed Percentage', 
+    x = 'Unemployed percentage(2018)', 
+    y = 'Case rate per 100000',
+    color = 'log 10 of population density'
   )
@@
 df %>% 
-  ggplot(aes(x = Percent.living.under.the.federal.poverty.line..2018., y = case_rate_100k, color = log10(population_density))) + 
+  ggplot(aes(senior_pct, case_rate_100k, color = log10(population_density))) + 
   geom_point() +   
+  geom_smooth(se=FALSE)+
   labs(
-    title = 'Relation of Case.Rate.per.100000 to Poverty %', 
-    x = 'Percent.living.under.the.federal.poverty.line..2018.', 
-    y = 'Case.Rate.per.100000',
-    color = 'log(Population Density)'
-  )
-```
-```{r}
-df %>% 
-  ggplot(aes(X65., case_rate_100k, color = log10(population_density))) + 
-  geom_point() +   
-  labs(
-    title = 'Relation of Case.Rate.per.100000 to 65+ Percentage', 
-    x = 'X65+.', 
-    y = 'Case.Rate.per.100000',
-    color = 'log(Population density)'
-  )
-```
-
-```{r}
-df %>% 
-  ggplot(aes(Percent.Unemployed..2018., case_rate_100k, color = log10(population_density))) + 
-  geom_point() +   
-  labs(
-    title = 'Relation of Case.Rate.per.100000 to Unemployed Percentage', 
-    x = 'Percent.Unemployed..2018.', 
-    y = 'Case.Rate.per.100000',
-    color = 'log(Population density)'
+    title = 'Relation of Case Rate per 100000 to Seniors Percentage(65+)', 
+    x = 'Seniors percentage', 
+    y = 'Case rate per 100000',
+    color = 'log 10 of population density'
   )
@@
 \begin{aligned}
- Case.Rate.Per.100000 =\beta_0 &+\beta_1 log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.total. population)
+ Case.Rate.Per.100000 =\beta_0 &+\beta_1 log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.Total.Population)
 \end{aligned}
@@
 ```{r}
-model2a <- lm(case_rate_100k ~ log10(population_density)+ White...of.Total.Population, data = df)
+model2a <- lm(case_rate_100k ~ log10(population_density)+ white_population_pct, data = df)
 coeftest(model2a, vcov = vcovHC)
@@
 \begin{aligned}
- Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.total. population)^2
+ Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.Total.Population)^2
 \end{aligned}
@@
 ```{r}
-model2a <- lm(case_rate_100k ~ log10(population_density)+ I((White...of.Total.Population)^2), data = df)
+model2a <- lm(case_rate_100k ~ log10(population_density)+ I((white_population_pct)^2), data = df)
 coeftest(model2a, vcov = vcovHC)
@@
 \begin{aligned}
- Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles) + \beta_3(Percen.living.under.the.federal.poverty.line.2018)
+ Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles) + \beta_3(Percentage.Under.Poverty.Line.2018)
 \end{aligned}
@@
 ```{r}
-model2b <- lm(case_rate_100k ~ log10(population_density) + Percent.living.under.the.federal.poverty.line..2018., data = df)
+model2b <- lm(case_rate_100k ~ log10(population_density) + poverty_pct, data = df)
 coeftest(model2b, vcov = vcovHC)
@@
 \begin{aligned}
- Case.Rate.Per.100000 &=\beta_0+\beta_1 log(Population.Density.Per.Square.Miles)+\beta_4(Percent.Unemployed..2018.)
+ Case.Rate.Per.100000 &=\beta_0+\beta_1 log(Population.Density.Per.Square.Miles)+\beta_4(Percentage.Unemployed.2018.)
 \end{aligned}
@@
 ```{r}
-model2c <- lm(case_rate_100k ~ log10(population_density) + Percent.Unemployed..2018., data = df)
+model2c <- lm(case_rate_100k ~ log10(population_density) + unemployed_pct, data = df)
 coeftest(model2c, vcov = vcovHC)
@@
 \begin{aligned}
-Case.Rate.Per.100000=\beta_0+\beta_1log(Population.Density.Per.Square.Miles)+\beta_4(65+)
+Case.Rate.Per.100000=\beta_0+\beta_1log(Population.Density.Per.Square.Miles)+\beta_4(Seniors.Percentage)
 \end{aligned}
@@
 ```{r}
-model2d <- lm(case_rate_100k ~log10(population_density)+ X65., data = df)
+model2d <- lm(case_rate_100k ~log10(population_density)+ senior_pct, data = df)
 coeftest(model2d, vcov = vcovHC)
@@
 model3a <- lm(case_rate_100k ~ log10(population_density)
-              + Percent.living.under.the.federal.poverty.line..2018.
+              + poverty_pct
               + mask_public_bool, data = df)

### 2020-11-18 DraftReport.Rmd
Added model assumptions based on Model 3A.

@@
 For example, the population density can have large variability depending on the population concentration among a few large areas as well as the level of uninhabited or sparsely inhabited land within a state, while the demographics of population can depend on a wide range of factors such as geographic location and employment opportunities.
-It is also important to note that these data points cannot be considered independent of each other, as there is a lot of interstate travel and regional similarities in terms of population and policy.
-This indicates that the sample cannot be assumed to be IID.
 
@@
 This may cause the spread of a virus within one state to be influenced by those around it, or generally by its position in the country.
-In addition in terms of demographics, all states are involved in the same market economy system even though there are economic performance divisions across the states.
-Moreover, in most states white race is still dominant in the population.
-Many states share the similar racial composition.
+In addition, all states are involved in the same market economy system of the country to varying degrees, even though there are economic performance divisions across the states.
+Likewise, certain regions of the country may also have similar racial compositions due to a shared history across general geographic regions.
+While this is acknowledged to occur, the effect is likely minor compared to the policy data collected state by state which would affect the spread of the virus.
+Therefore it is relatively safe for this assumption to be met.
 
-While this is acknowledged to occur, the effect is likely minor compared to the demographic and policy data collected state by state which would affect the spread of the virus.
-Therefore it is relatively safe for this assumption to be met.
 Secondly, the process being modeled must be described by a linear conditional expectation function of the variables which are included, in order for the linear regression to be valid.
 This can be assessed after the fact by visualizing the model over the range of each variable alongside the data, or also by visualizing the residuals over the range of each variable.
-This can be seen across the three models.
+For the purpose of this assumption and other model or data related assumptions, only the Model 3A will be used as it is the most inclusive.
+```{r}
+avg_pop <- log10(mean(df$population_density))
+avg_pov <- mean(df$poverty_pct)
+avg_mask <- mean(df$mask_public_bool)
+y_pop <- model3a$coefficients[1] + model3a$coefficients[2]*log10(df$population_density) + model3a$coefficients[3]*avg_pov + model3a$coefficients[4]*avg_mask
+y_pov <- model3a$coefficients[1] + model3a$coefficients[2]*avg_pop + model3a$coefficients[3]*df$poverty_pct + model3a$coefficients[4]*avg_mask
+y_pop_nomask <- model3a$coefficients[1] + model3a$coefficients[2]*avg_pop + model3a$coefficients[3]*avg_pov
+y_pov_nomask <- model3a$coefficients[1] + model3a$coefficients[2]*avg_pop + model3a$coefficients[3]*df$poverty_pct
+
+df %>% 
+  ggplot() + 
+  geom_point(aes(x = log10(population_density), y = case_rate_100k, group = factor(mask_public_bool), color = factor(mask_public_bool))) + 
+  geom_line(aes(x = log10(population_density), y = y_pop, color = "red")) +
+  geom_line(aes(x = log10(population_density), y = y_pop_nomask, color = "blue")) +
+  scale_color_manual(labels = c("No", "Yes", "No", "Yes"), values = c("blue", "red", "blue", "red")) +
+  labs(
+    title = 'Linearity of Variables, Population Density', 
+    x = 'Base 10 Log of Population Density', 
+    y = 'Case Rate', 
+    color = 'Mask Mandate'
+  )
+df %>% 
+  ggplot() + 
+  geom_point(aes(x = poverty_pct, y = case_rate_100k, group = factor(mask_public_bool), color = factor(mask_public_bool))) + 
+  geom_line(aes(x = poverty_pct, y = y_pov, color = "red")) +
+  geom_line(aes(x = poverty_pct, y = y_pov_nomask, color = "blue")) +
+  scale_color_manual(labels = c("No", "Yes", "No", "Yes"), values = c("blue", "red", "blue", "red")) +
+  labs(
+    title = 'Linearity of Variables, Poverty Percentage', 
+    x = 'Poverty Percentage', 
+    y = 'Case Rate', 
+    color = 'Mask Mandate'
+  )
+```
+Linearity for the three regressed variables is demonstrated above.
+As can be seen when mask policy and population density are varied, the data generally follows a linear trend.
+This is similar when the poverty percentage is varied, although at higher levels of poverty the presence of a mask mandate made the case rate unexpectedly low.
+It appears that possible a negative quadratice function may fit the data better, although it appears close enough to linear that this assumption may hold.
 
@@
 Alternatively, it can also be done via the Breusch-Pagan test to more objectively determine whether homoscedasticity is not present in the data.
+Using the first method with the plots above, the residuals to each of the four regressions appear to be evenly spread throughout the range of the x axis.
+Therefore, this assumption can also be reasonably assumed to be met.
+The normal distribution of these errors also appears to be well met by the linear models, with most data points close to their respective line causing most residuals to be centered around zero.
 
-Next, perfect colinearity is automatically detected by the R regression algorithms, which drop potential variables if they have complete colinearity with another variable.
+Finally, perfect colinearity is automatically detected by the R regression algorithms, which drop potential variables if they have complete colinearity with another variable.
 Near perfect colinearity is more problematic to detect, but a common hint to its existence is large standard errors on colinear features.
+In Model 3A, all of the standard errors for model coefficients are at least less than half of the value of the coefficient itself.
+Therefore, near-perfect colinearity can be assumed to not be taking place between these variables.
 Conceptually, the variables which may have an important colinearity in the final model are variables such as different percentage make ups of race and ethnicity, or related public policy.
-These variables are appropriately evaluatea and excluded throughout the model building process.
-
-Finally, normally distributed errors can be easily assessed by a histogram of the residuals around zero.
+These variables are appropriately evaluated and excluded throughout the model building process.
 

### 2020-11-18 DraftReport.Rmd
update on model 3 spacing etc

@@
 ```
+
 #### Model 3-a
@@
 ```
+
 #### Model 3-b&c
@@
   labs(
-    title = 'Case Rate vs Mask Public Mandate', 
+    title = 'Case Rate vs Non-Essential Business Close Date difference compared to Average', 
     x = 'Non-Essential Business Close Date difference compared to Average', 
@@
 By running the model with "business_close_diff" against case rate variable, the coefficient shows that the variable is significant.
+
 ```{r}
@@
 ```
+
 However, the regression with both policy variables "mask_pubic_bool" and "business_close_diff" shows that neither the the variable is significant. 
-```
+```{r}
 model3c <- lm(case_rate_100k ~ business_close_diff + mask_public_bool, data = df)

### 2020-11-18 DraftReport.Rmd
Updated draft report with regression table analysis and updated conclusion

@@
 output:
+  pdf_document: default
   html_document:
     df_print: paged
-  pdf_document: default
 ---
@@
 This research question aims to measure the case rate per 100,000 residents within each state based on the make-up of the population, as well as the policy decisions that were or were not put into place in order to combat the rise of COVID cases.
-In this sense, the aim is to measure how dependent the COVID case rate is on that of which cannot be controlled (i.e., population and demographics) vs that of which can be controlled to a certain degree (i.e., implementation of policies to attempt to combat cases).
+In this sense, the aim is to measure how dependent the COVID case rate is on that of which cannot be controlled (i.e., population and demographics) vs that of which can be controlled to a certain degree (i.e., implementation of policies to attempt to combat case rate).
 With this information, one could suggest whether or not the proliferation of COVID can be controlled, and to the level at which these policies help with this control.
@@
 
-Finally, with regard to the policy variables, this report focuses solely on mask mandates.
+Finally, with regard to the policy variables, this report focuses mainly on mask mandates.
 Given this, it does not capture the effect of other policies that may or may not have a greater effect on the case rate (e.g., implementation of stay-at-home orders).
@@
 
-#### 2-b) Percentage living under the federal poverty line(2018)
+#### 2-b) Percentage living under the federal poverty line (2018)
 
@@
 ```
+
 #### 2-c) Unemployed Percentage (2018)
+
 Although, the distribution is of high peak in the middle of the range.
@@
 
-#### 2-d) Seniors Percentage(Age 65+)
+#### 2-d) Seniors Percentage (Age 65+)
+
 The seniors percentage (age 65+) is of a nearly normal distribution.
@@
 ```
+
 ```{r}
@@
 #### Model 2-a
+
 $$
@@
 Firstly, "mask_public_bool" column is created from "mask_public", with 1 represents that the state has a pulbic mask policy, and 0 means there's no explicit public mask policy from the state.
+
 ```{r}
@@
 Histograms below that shows the distribution of states that has public mask policy or not. Both distributions exhibits certain degree of normality.
+
 ```{r}
@@
 A plot that shows the relationship between different policies vs case rate.
+
 ```{r}
@@
 #### Model 3-a
+
 Next, the linear model is created from model 2 with the mask policy variables added for the regression.
@@
 #### Model 3-b&c
+
 Considering that there could be multiple policy variables that are relevant to predicting the case number.
@@
 As "business_closed" is a date variable, a transformation was performed to convert the date into a variable "business_close_diff" which represents the days difference between the business close date of the state compared to the average date of all the states.
+
 ```{r}
-
 df$business_close_diff =as.numeric(df$business_closed-mean(df$business_closed, na.rm=TRUE))
@@
 From the plot of business close date difference against case rate, the graph shows that there is some degree of linearity between the two variables.
+
 ```{r warning=FALSE}
@@
 However, the regression with both policy variables "mask_pubic_bool" and "business_close_diff" shows that neither the the variable is significant.
+
 ```{r}
@@
 For the purpose of this assumption and other model or data related assumptions, only the Model 3A will be used as it is the most inclusive.
+
 ```{r}
@@
 ```
+
 Linearity for the three regressed variables is demonstrated above.
@@
 
+Although there were several versions of models created for model 2 and 3, through the analysis of the appropriate variables respective to each model a single model version was selected to be included in the regression table.
+This was determined based on factors of significance when investigating the effect relationship of these variables on the outcome variable, which is detailed in section 2.
+
+For the regression table below, model 1, model 2b, and model 3a were selected as the most significant models to include.
+
 ```{r warning=FALSE}
@@
 
-- Discuss variables included and not included in the model
-- Talk about Adjusted R2
-- Talk about significant variables in the table
-- Provide a statement that the table may change in the final report with more variables and maybe different key variables (case rate and population density)
+When observing the table above, three elements stand out.
+
+Firstly, there is a difference in the calculated standard error between the coefficient test run on the linear models and the regression table.
+Taking a look at the standard error for the poverty percentage coefficient, the value in the table returns a significant result with a standard error of 55.125, as opposed to the insignificant result of a standard error of 57.684.
+This follows for model 3, with the table showing a significant result with standard error of 50.872, as opposed to an insignificant result of a standard error of 59.583.
+
+Secondly, following the first observation, when designating the poverty percentage variable as significant, the regression table suggests that both covariate variables (including the mask mandate variable) are indeed significant with the COVID case rate.
+For the poverty percentage effect, the coefficient highlights that there is a positive correlation between the poverty percentage and case rate (as poverty percentage increases, so does case rate).
+The practical significance of this shows that for every percentage increase of population below the federal poverty line, there is an increase of roughly 112 cases per 100k for model 2, and roughly 104 cases per 100k for model 3.
+As for the mask mandate variable, the coeeficient highlights a negative correlation between it and the case rate (if the mask mandate is implemented, the case rate decreases).
+The practical significance is such that if the mask mandate has been implemented, the case rate per 100k will decrease by roughly 1000 cases.
+
+Thirdly, we notice that the R2 value increases by a large value for every additional variable that is included within the model.
+Initially, we start with a very low R2 value, suggesting that almost no variance is explained in model 1.
+However, although still low, there is close to a 9 times increase in the amount of variance that is explained in model 2.
+Finally, in model 3, the R2 value increases by roughly 2.75 times the previous value, and a decent amount of the variance can be seen to be explained by the model.
+The adjusted R2 that accounts for the increase in number of variable terms in the model also correspondingly increases as the covariates are added.
+
+As a note, it should be stressed that the models and variables included in this regression table may change when the final report is created to provide a more descriptive analysis of the original research question.
 
@@
 Conversely, the hypotheses that population and demographics would have an effect on the COVID case rate was unable to be accepted.
-Although upon investigating the poverty line percentage variable in model 2 there was a detection of notable adjacency to a significant result with a coefficient value of 0.06286, it was not enough to be truly significant and reject the null hypothesis.
+Although upon investigating the poverty line percentage variable in model 2 there was a detection of notable adjacency to a significant result (and an actual significant result when using the `r stargazer` to compute the regression table) with a coefficient value of 0.06286, it was not enough to be truly significant and reject the null hypothesis.
 Similarly, the relationship of race/ethnicity as well as age on case rate was also insignificant with no apparent correlation.
@@
 This is a promising insight, as it provides opportunity for states to implement and apply policies to intervene and control the case rate without being worried about its varying effects depending on population characteristics.
+However, it should once again be stressed that the final report may include and analyze alternative variables and apply them to different models based on the findings from this draft report, which as a result may also lead to more concrete or slightly varied insights.

### 2020-11-18 DraftReport.Rmd
Updated draft report for consistency and adjusted for knitted PDF output

@@
 library(tinytex)
+library(dplyr)
 ```
@@
 **Research Question:** How is the COVID case rate related to the population density and distribution of demographics of a state?
-How does the effect of population make-up on case rate compare with the effect of policy decisions made in each state?"
+How does the effect of population make-up on case rate compare with the effect of policy decisions made in each state?
 
@@
 ggplot(data = df,
-  mapping = aes(x= case_rate_100k))+
+  mapping = aes(x= case_rate_100k)) +
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,6000,600)) +
-  labs(title =  "Histogram of COVID Case Rate per 100,000", x = "COVID case rate per 100,000", y = 'Count')+
+  labs(title =  "Histogram of COVID Case Rate per 100,000", 
+       x = "COVID case rate per 100,000", y = 'Count') +
   scale_x_continuous(breaks=seq(0, 6000, 600))
@@
   labs(
-    title = 'Figure 2. Distribution of Population Density per Square Miles', 
-    x = 'Log 10 of population density per square mile', y = 'Count')
+    title = 'Distribution of Population Density per Square Miles', 
+    x = 'Log of population density per square mile', y = 'Count')
 
@@
 
-```{r}
+```{r message=FALSE}
 df %>% 
@@
     title = 'COVID Case Rate due to Population Density', 
-    x = 'Log 10 of population density per square miles', 
+    x = 'Log of population density per square miles', 
     y = 'COVID case rate per 100,000'
@@
 
-```{r}
+```{r warning=FALSE}
 black_num <- as.numeric(df$black_population_pct)
@@
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,1,0.05)) +
-  labs(title =  "Histogram of White Population Percentage", x = "White percentage of total population", y = 'Count')+
+  labs(title =  "Histogram of White Population Percentage", 
+       x = "White percentage of total population", y = 'Count')+
   scale_x_continuous(breaks=seq(0, 1, 0.1))
@@
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,1,0.05)) +
-  labs(title =  "Histogram of Square of White Percentage", x = "White percentage of total population^2", y = 'Count')+
+  labs(title =  "Histogram of Square of White Percentage", 
+       x = "White percentage of total population^2", y = 'Count')+
   scale_x_continuous(breaks=seq(0, 1, 0.1))
@@
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(6,21,1)) +
-  labs(title =  "Histogram of Poverty Percentage", x = "Percentage living under poverty line(2018)", y = 'Count')+
+  labs(title =  "Histogram of Poverty Percentage", 
+       x = "Percentage living under poverty line (2018)", y = 'Count')+
   scale_x_continuous(breaks=seq(6, 21, 1))
@@
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,10,1)) +
-  labs(title =  "Histogram of Unemployed Percentage", x = "Unemployed percentage (2018)", y = 'Count')+
+  labs(title =  "Histogram of Unemployed Percentage", 
+       x = "Unemployed percentage (2018)", y = 'Count')+
   scale_x_continuous(breaks=seq(0, 10, 1))
@@
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0.10,0.22,0.01)) +
-  labs(title =  "Histogram of Seniors Percentage (65+)", x = "Seniors percentage", y = 'Count')+
+  labs(title =  "Histogram of Seniors Percentage (65+)", x = "Seniors percentage", 
+       y = 'Count')+
   scale_x_continuous(breaks=seq(0.1, 0.25, 0.05))
@@
     title = 'Relation of Case Rate per 100000 to Unemployed Percentage', 
-    x = 'Unemployed percentage(2018)', 
+    x = 'Unemployed percentage (2018)', 
     y = 'Case rate per 100000',
@@
   labs(
-    title = 'Relation of Case Rate per 100000 to Seniors Percentage(65+)', 
+    title = 'Relation of Case Rate per 100000 to Seniors Percentage (65+)', 
     x = 'Seniors percentage', 
@@
 
-### Model-2 Regression
+### Model 2 Regression
 
@@
   )
-head(df)
 ```
@@
 ```{r}
-hist(df$case_rate_100k[df$mask_public_bool == 1], xlab='Case Rate', ylab='Count', main='Case Rate - Mask Mandate in Public')
-hist(df$case_rate_100k[df$mask_public_bool == 0], xlab='Case Rate', ylab='Count', main='Case Rate  - No Mask Mandate in Public')
+hist(df$case_rate_100k[df$mask_public_bool == 1], xlab='Case Rate', ylab='Count', 
+     main='Case Rate - Mask Mandate in Public')
+hist(df$case_rate_100k[df$mask_public_bool == 0], xlab='Case Rate', ylab='Count', 
+     main='Case Rate  - No Mask Mandate in Public')
 ```
@@
 model3a <- lm(case_rate_100k ~ log10(population_density)
-              + poverty_pct
-              + mask_public_bool, data = df)
+              + poverty_pct + mask_public_bool, data = df)
 coeftest(model3a, vcov = vcovHC)
@@
 
-#### Model 3-b&c
+#### Model 3-b & 3-c
 
@@
 df$business_close_diff =as.numeric(df$business_closed-mean(df$business_closed, na.rm=TRUE))
-head(df)
 ```
@@
 
-```{r warning=FALSE}
+```{r warning=FALSE, message=FALSE}
 df %>% 
@@
   labs(
-    title = 'Case Rate vs Non-Essential Business Close Date difference compared to Average', 
+    title = 'Case Rate vs Non-Essential Business Close Date difference \ncompared to Average', 
     x = 'Non-Essential Business Close Date difference compared to Average', 
@@
 Therefore, we decided to use "mask_public_bool" variable which has higher significance to represent our policy variable.
-As a result, model3a is selected as the final model for model 3.
+As a result, model 3-a is selected as the final model for model 3.
 
@@
 This can be assessed after the fact by visualizing the model over the range of each variable alongside the data, or also by visualizing the residuals over the range of each variable.
-For the purpose of this assumption and other model or data related assumptions, only the Model 3A will be used as it is the most inclusive.
+For the purpose of this assumption and other model or data related assumptions, only the Model 3-a will be used as it is the most inclusive.
 
@@
   ggplot() + 
-  geom_point(aes(x = log10(population_density), y = case_rate_100k, group = factor(mask_public_bool), color = factor(mask_public_bool))) + 
+  geom_point(aes(x = log10(population_density), y = case_rate_100k, 
+                 group = factor(mask_public_bool), color = factor(mask_public_bool))) + 
   geom_line(aes(x = log10(population_density), y = y_pop, color = "red")) +
   geom_line(aes(x = log10(population_density), y = y_pop_nomask, color = "blue")) +
-  scale_color_manual(labels = c("No", "Yes", "No", "Yes"), values = c("blue", "red", "blue", "red")) +
+  scale_color_manual(labels = c("No", "Yes", "No", "Yes"), 
+                     values = c("blue", "red", "blue", "red")) +
   labs(
@@
   ggplot() + 
-  geom_point(aes(x = poverty_pct, y = case_rate_100k, group = factor(mask_public_bool), color = factor(mask_public_bool))) + 
+  geom_point(aes(x = poverty_pct, y = case_rate_100k, group = factor(mask_public_bool), 
+                 color = factor(mask_public_bool))) + 
   geom_line(aes(x = poverty_pct, y = y_pov, color = "red")) +
   geom_line(aes(x = poverty_pct, y = y_pov_nomask, color = "blue")) +
-  scale_color_manual(labels = c("No", "Yes", "No", "Yes"), values = c("blue", "red", "blue", "red")) +
+  scale_color_manual(labels = c("No", "Yes", "No", "Yes"), 
+                     values = c("blue", "red", "blue", "red")) +
   labs(
@@
 Near perfect colinearity is more problematic to detect, but a common hint to its existence is large standard errors on colinear features.
-In Model 3A, all of the standard errors for model coefficients are at least less than half of the value of the coefficient itself.
+In Model 3-a, all of the standard errors for model coefficients are at least less than half of the value of the coefficient itself.
 Therefore, near-perfect colinearity can be assumed to not be taking place between these variables.
@@
 
-For the regression table below, model 1, model 2b, and model 3a were selected as the most significant models to include.
+For the regression table below, model 1, model 2-b, and model 3-a were selected as the most significant models to include.
 
@@
           se = list(se.model3),
-          star.cutoffs = c(0.05, 0.01, 0.001), title = "Table 1: The relationship between COVID case rate and population density")
+          star.cutoffs = c(0.05, 0.01, 0.001), 
+          title = "Table 1: The relationship between COVID case rate and population density")
 
@@
 Thirdly, we notice that the R2 value increases by a large value for every additional variable that is included within the model.
-Initially, we start with a very low R2 value, suggesting that almost no variance is explained in model 1.
+Initially, we start with a very low R2 value, suggesting that almost no variance is explained in model 1 and that the model is a poor fit and predictor for the data.
 However, although still low, there is close to a 9 times increase in the amount of variance that is explained in model 2.
@@
 The adjusted R2 that accounts for the increase in number of variable terms in the model also correspondingly increases as the covariates are added.
+It also holds a negative value for model 1, signifying that there is no variance that can be explained by the model, further highlighting the poor fit.
 
@@
 Conversely, the hypotheses that population and demographics would have an effect on the COVID case rate was unable to be accepted.
-Although upon investigating the poverty line percentage variable in model 2 there was a detection of notable adjacency to a significant result (and an actual significant result when using the `r stargazer` to compute the regression table) with a coefficient value of 0.06286, it was not enough to be truly significant and reject the null hypothesis.
+Although upon investigating the poverty line percentage variable in model 2 there was a detection of notable adjacency to a significant result (and an actual significant result when using the `stargazer` package to compute the regression table) with a coefficient value of 0.06286, it was not enough to be truly significant and reject the null hypothesis.
 Similarly, the relationship of race/ethnicity as well as age on case rate was also insignificant with no apparent correlation.

### 2020-11-18 DraftReport.Rmd
comment add in model2, some polishing. p-value of poverty percentage update in conclusion part, 0.06** to 0.5755

@@
 In order to account for the varying population sizes across the states, the variables are operationalized by specifically looking at variables that contain a rate, in order to standardize across the samples.
-In this sense, the analysis quantifies the spread of COVID by leveraging the COVID case rate per 100,000 rate, and for population demographics it investigates the percentage distribution of the race/ethnicity and age categories (e.g., White % of Total Population).
+In this sense, the analysis quantifies the spread of COVID by leveraging the COVID case rate per 100,000 rate, and for population demographics it investigates the percentage distribution of the race/ethnicity and age categories (e.g., white percentage of total population).
 
@@
 Currently, Case rate per 100,000 is chosen as a dependent variable to represent the pandemic status in each state.
-Variables such as population density per square miles, unemployed (2018)/percentage living under the federal poverty line (2018), white/black percentage of of total Population, percentage of seniors(65+), and mask mandate policies are chosen as candidate covariates to explore potential relationships between pandemic status demographic characteristics, economic status, racial composition, and state policy.
+Variables such as population density per square miles, unemployed percentage(2018)/percentage living under the federal poverty line(2018), white/black percentage of total population, percentage of seniors(65+), and mask mandate policies are chosen as candidate covariates to explore potential relationships between pandemic status demographic characteristics, economic status, racial composition, and state policy.
 
@@
 
-Several transformations have been made for variables a Population density per square miles and White % of Total Population.
+Several transformations have been made for variables as population density per square miles and white percentage of total population.
 Population density is extremely skewed due to the high population density of the District of Columbia.
@@
 ### Model 2
+
+For model 2, state demographic information regarding racial composition(white percentage of total population /Black Percentage of Total Population), economic status(unemployed percentage/poverty percentage), and population seniors percentage was investigated to explore its potential connection to the COVID pandemic status.
 
@@
 
-According to the regression results, no coefficients regarding the demographic variables investigated in model 2 are statistically significant.
-There is no statistically meaningful relationship that could be concluded by now.
-
 ### Model 2 Regression
+The regression models regarding the demographic variables discussed previously are listed as below.However, no coefficients regarding these variables in model 2 are statistically significant.
 
@@
 ```
+
+From the results of the regression, none of the coefficients of investigated variables in model 2 are significant (p-value < 0.05).
+Among these regression coefficient tests,  the p-value regarding the poverty percentage variable is the lowest (0.05755).
+In addition, its positive estimated coefficient is consistent with the statement that low-income communities are at high pandemic risk claimed by mainstream media these days.
 
@@
 Conversely, the hypotheses that population and demographics would have an effect on the COVID case rate was unable to be accepted.
-Although upon investigating the poverty line percentage variable in model 2 there was a detection of notable adjacency to a significant result (and an actual significant result when using the `stargazer` package to compute the regression table) with a coefficient value of 0.06286, it was not enough to be truly significant and reject the null hypothesis.
+Although upon investigating the poverty percentage variable in model 2 there was a detection of notable adjacency to a significant result (and an actual significant result when using the `stargazer` package to compute the regression table) with a p-value of 0.05755, it was not enough to be truly significant and reject the null hypothesis.
 Similarly, the relationship of race/ethnicity as well as age on case rate was also insignificant with no apparent correlation.

### 2020-11-18 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Updated title and authors and changed name for submission

--- 2020-11-18 DraftReport.Rmd
@@
 ---
-title: "Draft Report"
-author: "203 Section 6, Team 1"
+title: "Lab 2 Draft Report: COVID-19 Case Rate vs Population and Policy"
+author: "Aidan Jackson, Frank Liu, Sam Temlock, Haoyu Zhang"
 output:

### 2020-11-29 model_1.Rmd
Added test models with 65+ and Pre-existing Condition variables

@@
 df <- read.csv("covid-19.csv", header = TRUE)
+head(df)
 df<-df%>%
@@
          death_rate_100k = 'Death.Rate.per.100000',
+         population_18 = 'Population.2018',
          population_density = 'Population.density.per.square.miles',
@@
          mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
-         mask_legal='No.legal.enforcement.of.face.mask.mandate') %>%
-  select(State, case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
+         mask_legal='No.legal.enforcement.of.face.mask.mandate',
+         nonelderly_preexisting_condition = 'Nonelderly.Adults.Who.Have.A.Pre.Existing.Condition'
+         ) %>%
+  select(State, case_rate_100k, death_rate_100k, population_18, population_density, white_pct, black_pct, hispanic_pct, 
          other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, 
          Black...of.Total.Population, White...of.Total.Population,
-         Percent.living.under.the.federal.poverty.line..2018., Percent.Unemployed..2018., X65.)
+         Percent.living.under.the.federal.poverty.line..2018., Percent.Unemployed..2018., X65., 
+         nonelderly_preexisting_condition, Children.0.18
+         )
 # Assign correct data types
@@
 head(df)
+```
+
+```{r}
+df <- df %>% 
+  mutate(nonelderly_preexisting_condition_100k = nonelderly_preexisting_condition/(population_18*(1-X65.-Children.0.18)) * 100000,
+         X65_100k = X65.*100000)
+head(df)
+```
+
+```{r}
+summary(df$nonelderly_preexisting_condition_100k)
+```
+
+```{r}
+ggplot(data = df,
+  mapping = aes(x= nonelderly_preexisting_condition_100k ))+
+  geom_histogram(fill = 'skyblue', color = 'grey30') +
+  labs(title =  "Histogram of Nonelderly Adults with Pre-Existing Conditions per 100,000", x = "Adults per 100,000", y = 'Count')
+```
+
+```{r}
+ggplot(data = df,
+  mapping = aes(x= X65_100k ))+
+  geom_histogram(fill = 'skyblue', color = 'grey30', bins = 30) +
+  labs(title =  "Histogram of Adults Over 65 per 100,000", x = "Adults over 65 per 100,000", y = 'Count')
+```
+
+```{r}
+df %>% 
+  ggplot(aes(nonelderly_preexisting_condition_100k, case_rate_100k)) + 
+  geom_smooth(se = FALSE) +
+  geom_point() +   
+  labs(
+    title = 'COVID Case Rate due to Nonelderly Adults with Pre-existing Conditions', 
+    x = 'Nonelderly adults with pre-existing conditions per 100,000', 
+    y = 'COVID case rate per 100,000'
+  )
+```
+
+```{r}
+df %>% 
+  ggplot(aes(X65_100k, case_rate_100k)) + 
+  geom_smooth(se = FALSE) +
+  geom_point() +   
+  labs(
+    title = 'COVID Case Rate due to Adults Over 65', 
+    x = 'Adults over 65 per 100,000', 
+    y = 'COVID case rate per 100,000'
+  )
+```
+
+```{r}
+model_test <- lm(case_rate_100k ~ nonelderly_preexisting_condition_100k + X65_100k, data = df)
+car::vif(model_test)
+coeftest(model_test, vcov = vcovHC)
+```
+
+```{r}
+model_test_65 <- lm(case_rate_100k ~ X65_100k, data = df)
+coeftest(model_test_65, vcov = vcovHC)
 ```

### 2020-11-29 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Added more plots and explanation to CLM assumptions.

@@
 It appears that possible a negative quadratice function may fit the data better, although it appears close enough to linear that this assumption may hold.
+```{r}
+df %>% {
+   ggplot(mapping = aes(x = log(df$population_density), y = model3a$residuals)) +
+      geom_point() + stat_smooth(se = TRUE) +
+      ggtitle("Model Residuals versus Population Density") +
+      xlab("Logarithm of Population Density") +
+      ylab("Residual")
+}
+df %>% {
+   ggplot(mapping = aes(x = df$poverty_pct, y = model3a$residuals)) +
+      geom_point() + stat_smooth(se = TRUE) +
+      ggtitle("Model Residuals versus Poverty Percentage") +
+      xlab("Poverty Percentage (%)") +
+      ylab("Residual")
+}
+df %>% {
+   ggplot(mapping = aes(x = log(df$mask_public_bool), y = model3a$residuals)) +
+      geom_point() + stat_smooth(se = TRUE) +
+      ggtitle("Model Residuals versus Mask Mandate") +
+      xlab("Public Mask Mandate") +
+      ylab("Residual")
+}
+```
 
@@
 
-Finally, perfect colinearity is automatically detected by the R regression algorithms, which drop potential variables if they have complete colinearity with another variable.
+Perfect colinearity is automatically detected by the R regression algorithms, which drop potential variables if they have complete colinearity with another variable.
 Near perfect colinearity is more problematic to detect, but a common hint to its existence is large standard errors on colinear features.
@@
 Therefore, near-perfect colinearity can be assumed to not be taking place between these variables.
-Conceptually, the variables which may have an important colinearity in the final model are variables such as different percentage make ups of race and ethnicity, or related public policy.
-These variables are appropriately evaluated and excluded throughout the model building process.
+```{r}
+correlation_one <- cor(log(df$population_density), df$poverty_pct)
+vif_one <- 1/(1-(correlation_one^2))
+correlation_two <- cor(log(df$population_density), df$mask_public_bool)
+vif_two <- 1/(1-(correlation_two^2))
+correlation_three <- cor(df$poverty_pct, df$mask_public_bool)
+vif_three <- 1/(1-(correlation_three^2))
+```
+A more methodical approach for evaluating if there is too high of colinearity between features is through the calculation of the Variance Inflation Factor (VIF).
+For the three variables in Model 3-A, the VIF values between the three combinations of the three variables are `r vif_one`, `r vif_two`, and `r vif_three`.
+Generally, as long as the VIF is below four there can be assumed not enough colinearity to violate the assumption as is the case in this model.
+
+Finally, the linear model is also assumed to generate residual error that is normally distributed away from the predicted values.
+A simple and straightforward way to assess this is through a histogram of the residuals of the model.
+```{r}
+hist(model3a$residuals, main = "Model Residuals", xlab = "Residual")
+```
+As seen above, this assumption is generally well met.
 

### 2020-11-29 test.Rmd
test.rmd upload, for models update 11/29

---
title: "TEST"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lmtest)
library(sandwich)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(stargazer)
library(tinytex)
library(dplyr)
library(car)
```

```{r variables, warning=FALSE, results='hide'}
df <- read.csv("covid-19.csv", header = TRUE)
df<-df%>%
  rename(case_rate_100k = 'Case.Rate.per.100000',
         death_rate_100k = 'Death.Rate.per.100000',
         population_density = 'Population.density.per.square.miles',
         white_pct = 'White...of.Cases',
         black_pct = 'Black...of.Cases',
         hispanic_pct = 'Hispanic...of.Cases',
         other_pct = 'Other...of.Cases',
         state_emergency = 'State.of.emergency',
         business_closed = 'Closed.other.non.essential.businesses',
         business_reopen = 'Began.to.reopen.businesses.statewide',
         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
         mask_legal='No.legal.enforcement.of.face.mask.mandate',
         black_population_pct ="Black...of.Total.Population",
         white_population_pct = "White...of.Total.Population",
         poverty_pct = "Percent.living.under.the.federal.poverty.line..2018.",
         unemployed_pct = "Percent.Unemployed..2018.",
         senior_pct = "X65.",
         population_2018 = "Population.2018", ## extra variable in test
         non_elderly_pre_existing = "Nonelderly.Adults.Who.Have.A.Pre.Existing.Condition",
         homeless = "Number.Homeless..2019."
         ) %>%
  select(State, case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, 
         black_population_pct, white_population_pct, poverty_pct, unemployed_pct, senior_pct,
         population_2018, non_elderly_pre_existing, homeless)## extra variable in test
# Assign correct data types
cols.num <- c("white_pct","black_pct","hispanic_pct","other_pct")
df[cols.num] <- sapply(df[cols.num],as.numeric)
df$state_emergency=as.Date(df$state_emergency, format = "%m/%d/%Y")
df$business_closed=as.Date(df$business_closed, format = "%m/%d/%Y")
df$business_reopen=as.Date(df$business_reopen, format = "%m/%d/%Y")
head(df)
```
"Average social distance": average area of the square shape each people occupies, the length of such square is assumed to the average social distance(meter)
1 square miles = 2589988.11 square meters
```{r}
df<-df %>%
  mutate(
    homeless_pct = homeless/population_2018,
    noneld_pre_existing_pct = non_elderly_pre_existing/population_2018,
    avg_distance = sqrt(2589988.11/population_density),
    area_per_capita = 2589911.11/population_density
  )
```

```{r}
ggplot(data = df,
  mapping = aes(x= homeless_pct))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks=seq(0,0.05,0.001)) +
  labs(title =  "Homeless percentage", x = "Homeless percentage", 
       y = 'Count')#+
  #scale_x_continuous(breaks=seq(0.1, 0.25, 0.05)
```
```{r}
ggplot(data = df,
  mapping = aes(x= noneld_pre_existing_pct))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks=seq(0.1,0.25,0.01)) +
  labs(title =  "Non-elderly pre-existing condition percentage", x = "percentage", y = 'Count')
  #+scale_x_continuous(breaks=seq(0.1, 0.25, 0.05))
```
```{r}
hist_log_distance <- ggplot(data = df,
  mapping = aes(x= log(avg_distance)))+
  geom_histogram(fill = 'skyblue', color = 'grey30') +
  labs(title =  "log of Average distabce", x = "log of Meter", 
       y = 'Count')

hist_distance <- ggplot(data = df,
  mapping = aes(x= avg_distance))+
  geom_histogram(fill = 'skyblue', color = 'grey30') +
  labs(title =  "Average distabce", x = "Meter", 
       y = 'Count')

hist_log_distance/hist_distance
```
```{r}
hist_log_area <- ggplot(data = df,
  mapping = aes(x= log(area_per_capita)))+
  geom_histogram(fill = 'skyblue', color = 'grey30') +
  labs(title =  "log of Average area", x = "log of Meter^2", 
       y = 'Count')

hist_area <- ggplot(data = df,
  mapping = aes(x= area_per_capita))+
  geom_histogram(fill = 'skyblue', color = 'grey30') +
  labs(title =  "Average area", x = "Meter^2", 
       y = 'Count')

hist_log_area/hist_area
```
```{r}
model_test <- lm(case_rate_100k ~ log(avg_distance) + poverty_pct + noneld_pre_existing_pct + senior_pct + unemployed_pct+ white_population_pct, data = df)
coeftest(model_test)
vif(model_test)
plot(model_test, which = 3)
```

```{r}
model_test_b <- lm(case_rate_100k ~ log(population_density) + poverty_pct + noneld_pre_existing_pct + senior_pct + unemployed_pct+ white_population_pct, data = df)
coeftest(model_test_b)
vif(model_test_b)
plot(model_test_b, which = 3)
```
```{r}
model_test_b <- lm(case_rate_100k ~   poverty_pct + senior_pct + unemployed_pct, data = df)
coeftest(model_test_b, vcov = vcovHC)
vif(model_test_b)
plot(model_test_b, which = 3)
```
```{r}
model_test_b <- lm(case_rate_100k ~   noneld_pre_existing_pct+poverty_pct + senior_pct + unemployed_pct, data = df)
coeftest(model_test_b, vcov = vcovHC)
vif(model_test_b)
plot(model_test_b, which = 3)
```

### 2020-12-01 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Added omitted variable bias 1-3.

@@
 
-## 5. Conclusion
+## 5. Omitted Variables
+
+Omitted variables from the model may affect both the internal specifications of included variables as well as the general applicability of the model to explain the output variable.
+For this analysis, some of these variables are within the dataset but not used while others are hypothetical.
+
+# 5-1: Urban vs Rural Population Ratio
+
+One of the variables that was hypothesized as being potentially influential is a ratio of a state's urban to rural population.
+Currently, the dataset in use contains population information as averaged across each state as a whole.
+While population density is potentially useful, it was found to not be significant in the models.
+This may be due to states which large urban centers, where population density woukd be high, losing their impact due to also containing large areas of land that is sparsely populated.
+One example of this would be New York, which contains several of the most densely populated counties in the country, but also large rural areas.
+This variable would also be explanatory of the population density variable as a whole, where a higher urban to rural ratio would result in a higher statewide population density generally.
+With a greater fraction of urban residents, people would also be less likely to be able to social distance as effectively, increasing the per capita case rate.
+Therefore, the coefficient for population density in the original model would be larger than a model where this ratio is included, resulting in a bias that is positive and away from zero.
+
+# 5-2: Median Income
+
+Knowledge of the median income for a state would give an indication of the relative job status for a working person in that state.
+This would be important for the case rate to understand about how many people work in low paying jobs that are more frequently in close contact with others (such as the service industry, gig industry, etc.).
+Higher median income may imply that more workers are able to function remotely with greater social distancing.
+A higher median income would then be negatively related to the covid case rate.
+It would also be explanatory in a way that makes the poverty percentage variable less impactful, where either of these variables may be considered as proxies for one another.
+Because inclusion of the median income variable would led to a lower magnitude poverty percentage variable, this would cause a positive omitted variable bias away from zero.
+
+# 5-3: Gender Ratio
+
+The ratio of gender between women and men may impact the covid case rate in that men are more suscepitble to viral infections and generally have lower life expectancies from lifestyle differences.
+Greater susceptibility may relate to the case rate in that men may more often become infected when exposed to COVID under similar circumstances as a women.
+Their shorter life expentancy may also impact the case rate in that more COVID cases may result in greater or more serious symptoms, with less occurances of an infection that is asymptomatic or goes undetected.
+In either of these cases, a higher fraction of men among the population would lead to an overall higher covid case rate.
+
+This variable could also be partially explained by the senior percentage variable, because of the gap in gender ratios across generations and generally as people get older.
+Because men have lower life expectancies, older groups of people living in the US would have higher ratios of women to men.
+Therefore, introducing the gender ratio variable into the model would lower the value of the coefficient to the senior percentage variable.
+This would be a positive omitted variable bias away from zero.
+
+However, the magnitude of this change is likely to not be large due to other causal mechanisms between seniors and covid case rate.
+While men may more frequently present COVID symptoms, older people in general are also more susceptible to the virus due to their physical condition and would also present more symptoms than non seniors.
+When comparing the two, generally seniors are considered as more at-risk of COVID complications than men across all ages, and so the senior percentage would still have a greater impact on the overall covid case rate than the gender ratio.
+
+# 5-4:
+
+## 6. Conclusion
 

### 2020-12-01 Model_2_final_ver.Rmd
model_2 _final, regression & clm assumptions

---
title: "Model_2_final_version"
output: pdf_document
---
```{r}
install.packages("GGally")
```

 **Initial reassignment of common data used across models:**
```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lmtest)
library(sandwich)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(patchwork)
library(stargazer)
library(tinytex)
library(dplyr)
library(car)
```

```{r variables, warning=FALSE, results='hide'}
df <- read.csv("covid-19.csv", header = TRUE)
df<-df%>%
  rename(case_rate_100k = 'Case.Rate.per.100000',
         death_rate_100k = 'Death.Rate.per.100000',
         population_density = 'Population.density.per.square.miles',
         white_pct = 'White...of.Cases',
         black_pct = 'Black...of.Cases',
         hispanic_pct = 'Hispanic...of.Cases',
         other_pct = 'Other...of.Cases',
         state_emergency = 'State.of.emergency',
         business_closed = 'Closed.other.non.essential.businesses',
         business_reopen = 'Began.to.reopen.businesses.statewide',
         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
         mask_legal='No.legal.enforcement.of.face.mask.mandate',
         black_population_pct ="Black...of.Total.Population",
         white_population_pct = "White...of.Total.Population",
         poverty_pct = "Percent.living.under.the.federal.poverty.line..2018.",
         unemployed_pct = "Percent.Unemployed..2018.",
         senior_pct = "X65.",
         population_2018 = "Population.2018", ## extra variable in test
         non_elderly_pre_existing = "Nonelderly.Adults.Who.Have.A.Pre.Existing.Condition",
         homeless = "Number.Homeless..2019."
         ) %>%
  select(State, case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, 
         black_population_pct, white_population_pct, poverty_pct, unemployed_pct, senior_pct,
         population_2018, non_elderly_pre_existing, homeless)## extra variable in test
# Assign correct data types
cols.num <- c("white_pct","black_pct","hispanic_pct","other_pct")
df[cols.num] <- sapply(df[cols.num],as.numeric)
df$state_emergency=as.Date(df$state_emergency, format = "%m/%d/%Y")
df$business_closed=as.Date(df$business_closed, format = "%m/%d/%Y")
df$business_reopen=as.Date(df$business_reopen, format = "%m/%d/%Y")
head(df)
```

```{r}
df<-df %>%
  mutate(
    noneld_pre_existing_per_100k = 100000*non_elderly_pre_existing/population_2018,
    white_per_100k = 100000*white_population_pct,
    senior_per_100k = 100000*senior_pct,
    poverty_per_100k = 100000*poverty_pct, 
    unemployed_per_100k = 100000*unemployed_pct,
    log_population_density = log(population_density),
    avg_distance = sqrt(2589988.11/population_density),
    area_per_capita = 2589911.11/population_density
  )
```
## Second Model

### Model Regression

Case Rate per 100000 = senior_pct + nonelderly + poverty_pct + population_density + white_pct

$$
\begin{aligned}
 Case.Rate.Per.100K =\beta_0 &+\beta_1Senior.Rate.Per.100K \\&+\beta_2Nonelderly.Pre.Existing.Rate.Per.100K\\&+\beta_3Poverty.Rate.Per.100K+\beta_4ln(Population.Density)\\&+\beta_5White.Rate.Per.100K
\end{aligned}
$$
```{r}
model_2 <- lm(case_rate_100k ~ senior_per_100k + log(population_density) + noneld_pre_existing_per_100k + poverty_per_100k + white_per_100k, data = df)
coeftest(model_2)
```

## Limits
###i.IID Sampling

###ii.Linear Conditional Expectation
Assessing non-linear in higher-dimensional space: look at the predicted vs. residuals of the model.

```{r, echo=FALSE}
df<- df%>% 
  mutate(
    model_2_preds = predict(model_2), 
    model_2_resids = resid(model_2)
  ) 
df %>% 
  ggplot(aes(model_2_preds, model_2_resids)) + 
  geom_point() + 
  stat_smooth()
```

###iii.No Perfect Collinearity

First, check if any variables were dropped by R. No coefficient is missing, which mean there is no perfect collinearity.

```{r}
model_2$coefficients
```
In addition, all the variance inflation factors are less than 4, which doesn't indicate the exisitence of collinearity. 
```{r}
vif(model_2)
```
###iv.Homoskedastic Errors
32 is New Mexico, 33 is New York, 35 is North Dakota

```{r, echo=FALSE}
plot(model_2, which=3)
```
###v.Normally Distributed Errors
Check the normality of error distribution based on the histogram of residuals and the qqplot.

```{r, echo=FALSE}
plot_one <- df %>% 
  ggplot(aes(x = model_2_resids)) + 
  stat_bin(bins = 25) +
  geom_histogram()

plot_two <- df %>% 
  ggplot(aes(sample = model_2_resids)) + 
  stat_qq() + stat_qq_line()

plot_one / plot_two
```
### Cook Distance 
9-DC, 12-Hawaii, 32-New Mexico.

```{r, echo =FALSE}
plot(model_2, which = 5)
```

$$
\begin{aligned}
 Case.Rate.Per.100K =\beta_0 &+\beta_1Senior.Rate.Per.100K \\&+\beta_2Nonelderly.Pre.Existing.Rate.Per.100K\\&+\beta_3Poverty.Rate.Per.100K+\beta_4ln(Population.Density)\\&+\beta_5White.Rate.Per.100K+?\beta_6Unemployed.per.100k
\end{aligned}
$$
```{r}
model_2b <- lm(case_rate_100k ~ senior_per_100k + log(population_density) + noneld_pre_existing_per_100k + poverty_per_100k + white_per_100k + unemployed_per_100k, data = df)
coeftest(model_2b)
```
## i.IID Sampling

###ii.Linear Conditional Expectation
Assessing non-linear in higher-dimensional space: look at the predicted vs. residuals of the model.

```{r, echo=FALSE}
df<- df%>% 
  mutate(
    model_2b_preds = predict(model_2b), 
    model_2b_resids = resid(model_2b)
  ) 
df %>% 
  ggplot(aes(model_2b_preds, model_2b_resids)) + 
  geom_point() + 
  stat_smooth()
```
###iii.No Perfect Collinearity
First, check if any variables were dropped by R. No coefficient is missing, which mean there is no perfect collinearity.

```{r}
model_2b$coefficients
```
In addition, all the variance inflation factors are less than 4, which doesn't indicate the exisitence of collinearity. 
```{r}
vif(model_2b)
```
###iv.Homoskedastic Errors
32 is New Mexico, 33 is New York, 35 is North Dakota

```{r, echo=FALSE}
plot(model_2b, which=3)
```
###v.Normally Distributed Errors
Check the normality of error distribution based on the histogram of residuals and the qqplot.

```{r, echo=FALSE}
plot_one <- df %>% 
  ggplot(aes(x = model_2b_resids)) + 
  stat_bin(bins = 25)
  geom_histogram()

plot_two <- df %>% 
  ggplot(aes(sample = model_2b_resids)) + 
  stat_qq() + stat_qq_line()

plot_one / plot_two
```

```{r, echo=FALSE}
df %>% 
  select(case_rate_100k, senior_per_100k, log_population_density, noneld_pre_existing_per_100k, poverty_per_100k, white_per_100k) %>% 
  GGally::ggpairs()
```
### Cook Distance

```{r, echo =FALSE}
plot(model_2b, which = 5)
```

### 2020-12-01 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
model2_final merge to report draft

@@
 knitr::opts_chunk$set(echo = TRUE)
+install.packages("GGally")
 library(lmtest)
@@
 library(dplyr)
+library(GGally)
+library(car)
 ```
@@
          unemployed_pct = "Percent.Unemployed..2018.",
-         senior_pct = "X65.") %>%
+         senior_pct = "X65.",
+         population_2018 = "Population.2018", ## extra variable in test
+         non_elderly_pre_existing = "Nonelderly.Adults.Who.Have.A.Pre.Existing.Condition",
+         homeless = "Number.Homeless..2019.") %>%
   select(State, case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
          other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, 
-         black_population_pct, white_population_pct, poverty_pct, unemployed_pct, senior_pct)
+         black_population_pct, white_population_pct, poverty_pct, unemployed_pct, senior_pct,
+         population_2018, non_elderly_pre_existing, homeless)
 # Assign correct data types
@@
 
+```{r}
+df<-df %>%
+  mutate(
+    noneld_pre_existing_per_100k = 100000*non_elderly_pre_existing/population_2018,
+    white_per_100k = 100000*white_population_pct,
+    senior_per_100k = 100000*senior_pct,
+    poverty_per_100k = 100000*poverty_pct, 
+    unemployed_per_100k = 100000*unemployed_pct,
+    log_population_density = log(population_density),
+    avg_distance = sqrt(2589988.11/population_density),
+    area_per_capita = 2589911.11/population_density
+  )
+```
+
 #### Model 2-a
@@
 \begin{aligned}
- Case.Rate.Per.100000 =\beta_0 &+\beta_1 log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.Total.Population)
+ Case.Rate.Per.100K =\beta_0 &+\beta_1Senior.Rate.Per.100K \\&+\beta_2Nonelderly.Pre.Existing.Rate.Per.100K\\&+\beta_3Poverty.Rate.Per.100K+\beta_4ln(Population.Density)\\&+\beta_5White.Rate.Per.100K
 \end{aligned}
@@
 ```{r}
-model2a <- lm(case_rate_100k ~ log10(population_density)+ white_population_pct, data = df)
-coeftest(model2a, vcov = vcovHC)
-
-```
-
-$$
-\begin{aligned}
- Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.Total.Population)^2
-\end{aligned}
-$$
-```{r}
-model2a <- lm(case_rate_100k ~ log10(population_density)+ I((white_population_pct)^2), data = df)
-coeftest(model2a, vcov = vcovHC)
+model2a <- lm(case_rate_100k ~ senior_per_100k + log(population_density) + noneld_pre_existing_per_100k + poverty_per_100k + white_per_100k, data = df)
+coeftest(model2a)
 ```
@@
 \begin{aligned}
- Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles) + \beta_3(Percentage.Under.Poverty.Line.2018)
+ Case.Rate.Per.100K =\beta_0 &+\beta_1Senior.Rate.Per.100K \\&+\beta_2Nonelderly.Pre.Existing.Rate.Per.100K\\&+\beta_3Poverty.Rate.Per.100K+\beta_4ln(Population.Density)\\&+\beta_5White.Rate.Per.100K+\beta_6Unemployed.Per.100k
 \end{aligned}
@@
 ```{r}
-model2b <- lm(case_rate_100k ~ log10(population_density) + poverty_pct, data = df)
-coeftest(model2b, vcov = vcovHC)
+model2b <- lm(case_rate_100k ~ senior_per_100k + log(population_density) + noneld_pre_existing_per_100k + poverty_per_100k + white_per_100k + unemployed_per_100k, data = df)
+coeftest(model2b)
 ```
 
-#### Model 2-c
-
-$$
-\begin{aligned}
- Case.Rate.Per.100000 &=\beta_0+\beta_1 log(Population.Density.Per.Square.Miles)+\beta_4(Percentage.Unemployed.2018.)
-\end{aligned}
-$$
-```{r}
-model2c <- lm(case_rate_100k ~ log10(population_density) + unemployed_pct, data = df)
-coeftest(model2c, vcov = vcovHC)
-```
-
-#### Model 2-d
-
-$$
-\begin{aligned}
-Case.Rate.Per.100000=\beta_0+\beta_1log(Population.Density.Per.Square.Miles)+\beta_4(Seniors.Percentage)
-\end{aligned}
-$$
-```{r}
-model2d <- lm(case_rate_100k ~log10(population_density)+ senior_pct, data = df)
-coeftest(model2d, vcov = vcovHC)
-```
-
+#### need to update
 From the results of the regression, none of the coefficients of investigated variables in model 2 are significant (p-value < 0.05).
@@
 
+### Model 2 limits
+
+#### 2-a
+#### i. IID Sampling 
+( Discussed in the common section)
+
+#### ii.Linear Conditional Expectation
+Assessing non-linear in higher-dimensional space: look at the predicted vs. residuals of the model.
+
+```{r, echo=FALSE}
+df<- df%>% 
+  mutate(
+    model2a_preds = predict(model2a), 
+    model2a_resids = resid(model2a)
+  ) 
+df %>% 
+  ggplot(aes(model2a_preds, model2a_resids)) + 
+  geom_point() + 
+  stat_smooth()
+```
+
+####iii.No Perfect Collinearity
+First, check if any variables were dropped by R. No coefficient is missing, which mean there is no perfect collinearity.
+
+```{r}
+model2a$coefficients
+```
+
+In addition, all the variance inflation factors are less than 4, which doesn't indicate the exisitence of collinearity.
+
+```{r}
+vif(model2a)
+```
+
+####iv.Homoskedastic Errors
+32 is New Mexico, 33 is New York, 35 is North Dakota
+
+```{r, echo=FALSE}
+plot(model2a, which=3)
+```
+
+####v.Normally Distributed Errors
+Check the normality of error distribution based on the histogram of residuals and the qqplot.
+
+```{r, echo=FALSE}
+plot_one <- df %>% 
+  ggplot(aes(x = model2a_resids)) + 
+  stat_bin(bins = 25)
+  geom_histogram()
+
+plot_two <- df %>% 
+  ggplot(aes(sample = model2a_resids)) + 
+  stat_qq() + stat_qq_line()
+
+plot_one / plot_two
+```
+
+#### GGPairs
+
+```{r, echo=FALSE}
+df %>% 
+  select(case_rate_100k, senior_per_100k, log_population_density, noneld_pre_existing_per_100k, poverty_per_100k, white_per_100k) %>% 
+  GGally::ggpairs()
+```
+
+#### Cook Distance 
+9-DC, 12-Hawaii, 32-New Mexico.
+
+```{r, echo =FALSE}
+plot(model2a, which = 5)
+```
+
+#### 2-b
+####i.IID Sampling
+
+#####ii.Linear Conditional Expectation
+Assessing non-linear in higher-dimensional space: look at the predicted vs. residuals of the model.
+
+```{r, echo=FALSE}
+df<- df%>% 
+  mutate(
+    model2b_preds = predict(model2b), 
+    model2b_resids = resid(model2b)
+  ) 
+df %>% 
+  ggplot(aes(model2b_preds, model2b_resids)) + 
+  geom_point() + 
+  stat_smooth()
+```
+
+###iii.No Perfect Collinearity
+First, check if any variables were dropped by R. No coefficient is missing, which mean there is no perfect collinearity.
+
+```{r}
+model2b$coefficients
+```
+
+In addition, all the variance inflation factors are less than 4, which doesn't indicate the exisitence of collinearity.
+
+```{r}
+vif(model2b)
+```
+
+###iv.Homoskedastic Errors
+32 is New Mexico, 33 is New York, 35 is North Dakota
+
+```{r, echo=FALSE}
+plot(model2b, which=3)
+```
+
+###v.Normally Distributed Errors
+Check the normality of error distribution based on the histogram of residuals and the qqplot.
+
+```{r, echo=FALSE}
+plot_one <- df %>% 
+  ggplot(aes(x = model2b_resids)) + 
+  stat_bin(bins = 25)
+  geom_histogram()
+
+plot_two <- df %>% 
+  ggplot(aes(sample = model2b_resids)) + 
+  stat_qq() + stat_qq_line()
+
+plot_one / plot_two
+```
+
+#### GGPairs
+```{r, echo=FALSE}
+df %>% 
+  select(case_rate_100k, senior_per_100k, log_population_density, noneld_pre_existing_per_100k, poverty_per_100k, white_per_100k) %>% 
+  GGally::ggpairs()
+```
+
+### Cook Distance 
+9-DC, 12-Hawaii, 32-New Mexico.
+
+```{r, echo =FALSE}
+plot(model2b, which = 5)
+```
+
 ## 4. Regression Table

### 2020-12-01 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Added omitted variables 4 and 5.

--- 2020-12-01 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
@@
 
-# 5-4:
+# 5-4: Mask Use Among Population
+
+Currently, the model incorporates state policy around masks with a variable indicating whether or not a state mandated a mask policy for the general public.
+While this is useful for describing an action a state government can make and its affect on the case rate, it doesn't full capture whether how much the population actively wears a mask to drive down transmission.
+Instead, a variable which would indicate the fraction of the population actively wearing a mask would better capture the mechanism of how the virus is transmitted among people.
+
+A variable capturing the mask use among the population would likely reduce the explanatory power of the state mask mandate variable.
+This would cause the bias of the omitted variable to be positive and away from zero.
+
+# 5-5: GDP Percentage of Tourism
+
+The economic standing of a state not only affects the individual households of its population, but also includes how it interacts with other states and how its peoples' lives might have been economically changed by the pandemic.
+One way that a state may increase its number of covid cases is both from tourists or others visiting during the pandemic, but also if its own population must work in close contact with others.
+In addition, if certain key industries to a state, such as tourism, layed off many workers, they would likely have to take lower paying close contact jobs such as in the service or gig economy in order to continue having a livelihood.
+
+This effect can be included with a variable that demonstrates how much of a state's GDP is from tourism.
+It would be related to other economic indicator variables, such as the percent of the population living in poverty.
+While poverty is a broader phenomena, the number of people in poverty in the pandemic when the data was taken would be potentially higher due to negative effects in the tourism industry.
+Therefore, it is expected that states with a greater focus on tourism may also have more people in poverty in the pandemic.
+This would take some of the explanatory power of the poverty variable away, resulting in a positive omitted variable bias away from zero.
 

### 2020-12-01 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Update Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
model 3 update from model 2

@@
 
-Firstly, "mask_public_bool" column is created from "mask_public", with 1 represents that the state has a pulbic mask policy, and 0 means there's no explicit public mask policy from the state.
+Firstly, "mask_public_bool" column is created from "mask_public", with 1 represents that the state has a pubic mask policy, and 0 means there's no explicit public mask policy from the state.
 
@@
 ```{r}
-model3a <- lm(case_rate_100k ~ log10(population_density)
-              + poverty_pct + mask_public_bool, data = df)
+model3a <- lm(case_rate_100k ~ senior_per_100k + log(population_density) 
+              + noneld_pre_existing_per_100k + poverty_per_100k + white_per_100k 
+              + unemployed_per_100k + mask_public_bool, data = df)
 coeftest(model3a, vcov = vcovHC)
+
+model3a <- lm(case_rate_100k ~ senior_per_100k + log(population_density) 
+              + poverty_per_100k 
+              + unemployed_per_100k + mask_public_bool, data = df)
+coeftest(model3a, vcov = vcovHC)
+```
+
+CLM assumptions check for Model 3:
+```{r}
+#Zero Conditional Mean
+plot(model3a, which=1)
+
+#Normality of errors
+plot(model3a, which=2)
+
+#Homoskedasticity
+plot(model3a, which=3)
+
+#Collinearity
+vif(model3a)>4
 ```

### 2020-12-03 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Update Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
update cook distance for model3
model_2 update, 2-a, 2-b deleted

@@
 In addition, correlation values of white percentage and black percentage, unemployed percentage and percentage living under the federal poverty line shows these two groups of variables are correlated respectively.
+
+```{r}
+
+# percentage varialbe re-scale to number_per_100k (need to double check where to put this part in EDA or model regression)
+
+df<-df %>%
+  mutate(
+    noneld_pre_existing_per_100k = 100000*non_elderly_pre_existing/population_2018,
+    white_per_100k = 100000*white_population_pct,
+    senior_per_100k = 100000*senior_pct,
+    poverty_per_100k = 100000*poverty_pct, 
+    unemployed_per_100k = 100000*unemployed_pct,
+    log_population_density = log(population_density),
+    avg_distance = sqrt(2589988.11/population_density),
+    area_per_capita = 2589911.11/population_density
+  )
+```
 
@@
 ### Model 2 Regression
-The regression models regarding the demographic variables discussed previously are listed as below.However, no coefficients regarding these variables in model 2 are statistically significant.
-
-```{r}
-df<-df %>%
-  mutate(
-    noneld_pre_existing_per_100k = 100000*non_elderly_pre_existing/population_2018,
-    white_per_100k = 100000*white_population_pct,
-    senior_per_100k = 100000*senior_pct,
-    poverty_per_100k = 100000*poverty_pct, 
-    unemployed_per_100k = 100000*unemployed_pct,
-    log_population_density = log(population_density),
-    avg_distance = sqrt(2589988.11/population_density),
-    area_per_capita = 2589911.11/population_density
-  )
-```
-
-#### Model 2-a
+The regression models regarding the demographic variables discussed previously are listed as below.
 
@@
 \begin{aligned}
- Case.Rate.Per.100K =\beta_0 &+\beta_1Senior.Rate.Per.100K \\&+\beta_2Nonelderly.Pre.Existing.Rate.Per.100K\\&+\beta_3Poverty.Rate.Per.100K+\beta_4ln(Population.Density)\\&+\beta_5White.Rate.Per.100K
+ Case.Rate.Per.100K =\beta_0 &+\beta_1Senior.Rate.Per.100K +\beta_2Poverty.Rate.Per.100K +\beta_3Unemployed.Per.100K
 \end{aligned}
@@
 ```{r}
-model2a <- lm(case_rate_100k ~ senior_per_100k + log(population_density) + noneld_pre_existing_per_100k + poverty_per_100k + white_per_100k, data = df)
-coeftest(model2a)
+model2 <- lm(case_rate_100k ~ senior_per_100k + poverty_per_100k + unemployed_per_100k, data = df)
+coeftest(model2)
 ```
 
-#### Model 2-b
-
-$$
-\begin{aligned}
- Case.Rate.Per.100K =\beta_0 &+\beta_1Senior.Rate.Per.100K \\&+\beta_2Nonelderly.Pre.Existing.Rate.Per.100K\\&+\beta_3Poverty.Rate.Per.100K+\beta_4ln(Population.Density)\\&+\beta_5White.Rate.Per.100K+\beta_6Unemployed.Per.100k
-\end{aligned}
-$$
-```{r}
-model2b <- lm(case_rate_100k ~ senior_per_100k + log(population_density) + noneld_pre_existing_per_100k + poverty_per_100k + white_per_100k + unemployed_per_100k, data = df)
-coeftest(model2b)
-```
-
-#### need to update
+#### (need to update)
 From the results of the regression, none of the coefficients of investigated variables in model 2 are significant (p-value < 0.05).
@@
 plot(model3a, which=3)
+
+#Cook distance
+plot(model3a, which=5)
 
@@
 
-#### 2-a
 #### i. IID Sampling 
@@
   mutate(
-    model2a_preds = predict(model2a), 
-    model2a_resids = resid(model2a)
+    model2_preds = predict(model2), 
+    model2_resids = resid(model2)
   ) 
 df %>% 
-  ggplot(aes(model2a_preds, model2a_resids)) + 
+  ggplot(aes(model2_preds, model2_resids)) + 
   geom_point() + 
@@
 ```{r}
-model2a$coefficients
+model2$coefficients
 ```
@@
 ```{r}
-vif(model2a)
+vif(model2)
 ```
@@
 ####iv.Homoskedastic Errors
-32 is New Mexico, 33 is New York, 35 is North Dakota
+10 is Florida, 33 is New York, 35 is North Dakota
 
 ```{r, echo=FALSE}
-plot(model2a, which=3)
+plot(model2, which=3)
 ```
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, warninng=False}
 plot_one <- df %>% 
-  ggplot(aes(x = model2a_resids)) + 
+  ggplot(aes(x = model2_resids)) + 
   stat_bin(bins = 25)
@@
 plot_two <- df %>% 
-  ggplot(aes(sample = model2a_resids)) + 
+  ggplot(aes(sample = model2_resids)) + 
   stat_qq() + stat_qq_line()
@@
 df %>% 
-  select(case_rate_100k, senior_per_100k, log_population_density, noneld_pre_existing_per_100k, poverty_per_100k, white_per_100k) %>% 
+  select(case_rate_100k, senior_per_100k, poverty_per_100k, unemployed_per_100k) %>% 
   GGally::ggpairs()
@@
 #### Cook Distance 
-9-DC, 12-Hawaii, 32-New Mexico.
+10 Florida, 20 Maine, 35 North Dakota
 
 ```{r, echo =FALSE}
-plot(model2a, which = 5)
-```
-
-#### 2-b
-####i.IID Sampling
-
-#####ii.Linear Conditional Expectation
-Assessing non-linear in higher-dimensional space: look at the predicted vs. residuals of the model.
-
-```{r, echo=FALSE}
-df<- df%>% 
-  mutate(
-    model2b_preds = predict(model2b), 
-    model2b_resids = resid(model2b)
-  ) 
-df %>% 
-  ggplot(aes(model2b_preds, model2b_resids)) + 
-  geom_point() + 
-  stat_smooth()
-```
-
-###iii.No Perfect Collinearity
-First, check if any variables were dropped by R. No coefficient is missing, which mean there is no perfect collinearity.
-
-```{r}
-model2b$coefficients
-```
-
-In addition, all the variance inflation factors are less than 4, which doesn't indicate the exisitence of collinearity.
-
-```{r}
-vif(model2b)
-```
-
-###iv.Homoskedastic Errors
-32 is New Mexico, 33 is New York, 35 is North Dakota
-
-```{r, echo=FALSE}
-plot(model2b, which=3)
-```
-
-###v.Normally Distributed Errors
-Check the normality of error distribution based on the histogram of residuals and the qqplot.
-
-```{r, echo=FALSE}
-plot_one <- df %>% 
-  ggplot(aes(x = model2b_resids)) + 
-  stat_bin(bins = 25)
-  geom_histogram()
-
-plot_two <- df %>% 
-  ggplot(aes(sample = model2b_resids)) + 
-  stat_qq() + stat_qq_line()
-
-plot_one / plot_two
-```
-
-#### GGPairs
-```{r, echo=FALSE}
-df %>% 
-  select(case_rate_100k, senior_per_100k, log_population_density, noneld_pre_existing_per_100k, poverty_per_100k, white_per_100k) %>% 
-  GGally::ggpairs()
-```
-
-### Cook Distance 
-9-DC, 12-Hawaii, 32-New Mexico.
-
-```{r, echo =FALSE}
-plot(model2b, which = 5)
+plot(model2, which = 5)
 ```
@@
 se.model1 = coeftest(model1, vcov = vcovHC)[ , "Std. Error"]
-se.model2 = coeftest(model2b, vcov = vcovHC)[ , "Std. Error"]
+se.model2 = coeftest(model2, vcov = vcovHC)[ , "Std. Error"]
 se.model3 = coeftest(model3a, vcov = vcovHC)[ , "Std. Error"]
 
-stargazer(model1, model2b, model3a, type = "text", omit.stat = "f",
+stargazer(model1, model2, model3a, type = "text", omit.stat = "f",
           se = list(se.model3),

### 2020-12-03 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Updated introduction, assumptions, and intro to model buildling process

--- 2020-12-01 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
@@
 ---
-title: "Lab 2 Draft Report: COVID-19 Case Rate vs Population and Policy"
+title: "Lab 2 Final Report: COVID-19 Case Rate vs Population Demographics and Mask Policy"
 author: "Aidan Jackson, Frank Liu, Sam Temlock, Haoyu Zhang"
@@
 
-**Research Question:** How is the COVID case rate related to the population density and distribution of demographics of a state?
-How does the effect of population make-up on case rate compare with the effect of policy decisions made in each state?
+**Research Question:** How is the COVID-19 case rate related to the distribution of population demographics within a state, and how does the effect of these demographics compare with the effect of mask-related policy decisions made by that state?
 
-For this report, the investigation will be centered around the effect of population metrics such as density and demographics and policy on the COVID case rates across the United States (US), which is grouped by the 50 states plus the District of Columbia (D.C.).
-This research question aims to measure the case rate per 100,000 residents within each state based on the make-up of the population, as well as the policy decisions that were or were not put into place in order to combat the rise of COVID cases.
-In this sense, the aim is to measure how dependent the COVID case rate is on that of which cannot be controlled (i.e., population and demographics) vs that of which can be controlled to a certain degree (i.e., implementation of policies to attempt to combat case rate).
-With this information, one could suggest whether or not the proliferation of COVID can be controlled, and to the level at which these policies help with this control.
+For this report, the investigation will be centered around the effect of population demographics and policy on the COVID-19 case rates across the United States (US), which is grouped by the 50 states plus the District of Columbia (D.C.).
+Within the report, this collection will be referred to as the "states", and each member a "state", inclusive of D.C..
 
-This question can be broken up into two distinctive areas of investigation.
-The first and primary area of investigation involves how the COVID case rate is affected by population metrics.
-For this, the key variables are COVID case rate per 100,000 residents and the population density per square mile.
-This will provide an initial understanding in the relationship between the spread of the virus and how densely populated a state is.
-Following this, the analysis considers variables that address the make-up of these populations, specifically focusing on the distribution of demographics such as race/ethnicity and age.
-In order to account for the varying population sizes across the states, the variables are operationalized by specifically looking at variables that contain a rate, in order to standardize across the samples.
-In this sense, the analysis quantifies the spread of COVID by leveraging the COVID case rate per 100,000 rate, and for population demographics it investigates the percentage distribution of the race/ethnicity and age categories (e.g., white percentage of total population).
+The research question aims to measure the COVID-19 case rate per 100,000 residents within each state based on the make-up of the population, as well as the policy decisions that were or were not put into place in order to combat the rise of COVID-19 cases.
+In this sense, the aim is to examine how dependent the COVID-19 case rate is on that of which cannot be controlled (i.e., population demographics), as well as that of which can be controlled to a certain degree (i.e., implementation of policies to attempt to combat case rate).
+With this information, one could suggest whether or not the proliferation of COVID-19 seems to be related to either or both controllable and uncontrollable variables, and to the level at which the effect of these variables differ help.
+This modeling goal of this research question will be one of description, and will be broken up into three phases of investigation
 
-The secondary area of investigation will be to also include variables that measure policy implementation within states, and to compare the effect of these as opposed to the primary area of investigation that is population metrics.
-Specifically, there is a focus on policies that address the mandates of wearing masks within the state.
-Through this, the aim is to analyze whether the implementation of mask related policies will have an effect on case rate, as well as the degree of this effect relative to population metrics.
-To operationalize these policies and simplifying the measurements, the models will only consider whether or not these policies were implemented through indicator variables (1 = implemented, 0 = not implemented).
+The first and primary phase of investigation involves how the COVID-19 case rate is affected by key population demographics.
+For this, the key variables are COVID-19 case rate per 100,000 residents and the percentage population distribution of seniors (defined as 65 years old or older) within a state.
+The case rate per 100,000 was selected as the key output/dependent variable as it provides a standardized measure of the spread of COVID-19 across the states, taking into account the absolute population of the states.
+The senior percentage was selected as the key input/independent variable given that it is posited in external research that seniors make up a significant number of the reported cases.
+This is likely due to the both the fact that seniors in this age group were more likely to be or had greater access to be tested given that they are categorized in the high risk category for COVID-19, and the fact that they would have a greater likelihood of exhibiting detectable COVID-19 related symptoms that would prompt them to get tested.
+An additional rationale for measuring the senior percentage is that they may also be more likely to contract COVID-19 at lower viral loads, making it easier for them more susceptible to the virus.
+These key variables will provide an initial understanding of the relationship between the spread of the virus and a key metric of susceptibility and detectability of COVID-19.
+
+Following this, in the second phase of investigation the analysis considers other variables of state-level population that are considered factors of susceptibility, namely the rate of poverty (defined as the percent of individuals living under the federal poverty line in 2018) and the rate of unemployment (defined as the percent unemployment in 2018.).
+As mentioned, these variables operationalize population metrics that may lead to greater case rates, and are already standardized as rates to account for varying absolute populations across states.
+
+The tertiary and final phase of investigation will include a variable that measures the implementation of policy as a response to COVID-19, and to understand the added effect of this variable in conjunction with the first phase of investigation that is state population metrics.
+Specifically, there is a focus on a policy that mandates the wearing of masks within the state.
+Through this, the aim is to analyze whether the implementation of a mask-related policy will have an effect on case rate, as well as the degree of this effect relative to population metrics.
+To operationalize this policy and simplify the measurements, the models will only consider whether or not this policy was implemented through transformed indicator variables (1 = implemented, 0 = not implemented).
+As a result, factors of when and for how long the policy was implemented will be lost.
+Although this loss of information may fail to capture the impact of length of time of a policy on case rate (the policy may take time to be adopted/show meaningful efficacy), as there is no time-series data for case rate included in the data set, it was adjudged to be incongruent with the analysis.
+
+Additionally, the final phase of investigation will also continue to expand on the examination of state-level population demographics via the population density variable (defined as the population density in square miles, population/square miles).
+A key factor to the spread of COVID-19 is the idea of social distancing, where the virus is  considered to be more likely to spread when people are close in proximity.
+Thus, population density is measured as a conduit to indicate the level of proximity within each state and will operationalize the concept of social distancing.
+
+### Assumptions
 
@@
 
-Firstly, it is important to note that given each state (with the addition of the District of Columbia) is treated as a unique data point, the sample contains 51 data points.
-Although this size meets the general rule of an adequate sample size, as the analysis begins to factor in the wide range of potential population distributions and demographics within the states, it is clear that there is large variation within the sample.
-For example, the population density can have large variability depending on the population concentration among a few large areas as well as the level of uninhabited or sparsely inhabited land within a state, while the demographics of population can depend on a wide range of factors such as geographic location and employment opportunities.
+Firstly, it is important to note that given each state is treated as a unique data point, the sample contains 51 data points.
+Although this size meets the general rule that an adequate sample size is 30 data points or more, as the analysis begins to factor in the wide range of potential population distributions and demographics within the states, it is clear that there is large variation within the sample.
+For example, the population density can have large variability depending on the population concentration among a few large areas as well as the level of uninhabited or sparsely inhabited land within a state, while the demographics of population can depend on a wide range of things such as regional factors and employment opportunities.
 
 Secondly, note that there are many internal and external aspects of the selected variables that have not been included within the models.
-Just addressing the internal information that is lost, it can be seen that much of the information is not captured given the methods of operationalization.
-Among others, the loss of date-related information within policy implementation variables strips out any contextual knowledge regarding the length of policy implementation, and precludes the identification of how long a specific policy was implemented.
-Assuming that policy implementation has some effect on case rate, this difference in length of time could play a large part in the effect.
+Just addressing the internal information that is lost, it can be seen that some of the information is not captured given the methods of operationalization discussed previously.
+Among others, the loss of information on the date of implementation within the mask policy variable strips out any contextual knowledge regarding the length of policy implementation, and precludes the identification of how long a specific policy was implemented.
+Assuming that this has some effect on case rate, this difference in length of time may play a part in the effect.
 There are also other external aspects that cannot be included, such as the level of population density broken out among different population demographics.
 
-Finally, with regard to the policy variables, this report focuses mainly on mask mandates.
-Given this, it does not capture the effect of other policies that may or may not have a greater effect on the case rate (e.g., implementation of stay-at-home orders).
-This follows for the population-related variables, as they capture race/ethnicity and poverty, but not other characteristics such as proportions of residents with pre-existing conditions.
-In justification, these choices were largely made due to there either being an appropriate amount of samples in each category (e.g., most states have implemented basic policies such as the closure of non-essential businesses and the implementation of stay-at-home orders, and thus these were not included given the lack of samples for the states that have not implemented them), or that data on other policies or demographic measures were simply not readily available in the dataset.
+Finally, with regard to the policy variables, this report focuses on a form of mask mandate.
+Given this, it does not capture the effect of other policies that may or may not have a greater effect on the case rate (e.g., implementation of stay-at-home orders, closure of non-essential businesses, etc.).
+In justification, these choices were largely made due to there either being an appropriate amount of samples in each category (e.g., most states have implemented basic policies such as the closure of non-essential businesses and the implementation of stay-at-home orders, and thus these were not included given the lack of samples for the states that have not implemented them), or that data on other policies were simply not readily available in the dataset.
+The same argument can be applied in the selection of population demographics.
 
@@
 
-  The primary variable of interest will be the total covid case rate per 100,000 residents.
-The covid case rate was judged to give a better understanding than the death rate because of the potential for other, unrecorded variables that could be correlated with a person dying from covid versus just becoming infected.
-For example, the availability and quality of medical care in a state may impact its ability to keep covid infected patients alive, although this is not represented in the data set.
+The primary variable of interest will be the total COVID-19 case rate per 100,000 residents.
+The COVID-19 case rate was judged to be able to provide a better understanding than the related death rate because of the potential for other, unrecorded variables that could be correlated with a person dying from COVID-19 versus just becoming infected.
+For example, the availability and quality of medical care in a state may impact its ability to keep COVID-19 infected patients alive, and these variables are not included in the data set.
 The health status of the residents in one state compared to another state may also affect the death rate, but this is also not included.
-Instead, the covid infection rate was chosen so that these other variables which may correlate with the death rate would not have to be considered.
-Finally, the covid infection rate was also chosen on a per capita basis so that potential total population differences between states would already be accounted for in the variable.
+Instead, the COVID-19 case rate was chosen so that these other variables which may correlate with the death rate would not have to be considered.
 
-  The modeling goal for the covid case rate will be one of description.
-A modelf of causation may be created out of this model if potential causal relationships are examined and incorporated into the model's structure.
-This may later be expanded to predict the covid case rate in the past seven days in order for a state to use it to potentially judge what effect a modeled policy may have on its current situation.
+The covariates that will be examined in building the models fit into two broad categories.
+The first category is demographic information on the state, namely percentage population distribution of seniors, the unemployment rate, the rate of people living in poverty, and the population density.
+The second general group is that of policy decisions taken by the state, specifically a mask mandate.
+Problematic covariates would include any of the other direct measures of COVID-19 severity in the state, namely total infection rate not on a per capita basis and COVID-19 death rates.
+It can be assumed that these variables would be collinear with the primary variable of interest since COVID-19 case rate is simply a linear transformation of total cases not adjusted for population, and death rate is derived from case rate.
 
-  There are two broad groups of covariates which will be examined in building the model.
-The first is demographic information on the state, such as population density, the unemployment rate, or rate of people living in poverty.
-The second general group is that of policy decisions taken by the state, such as mask mandates, legal enforcements of policies, or the political party of the state's governor (which is assumed to be related to the policies they may support).
-Problematic covariates would include any of the other direct measures of covid severity in the state, such as total infection rate not on a per capita basis or death rates.
-It can be assumed that these variables would be colinear with the primary variable of interest since more covid cases per capita would imply more total cases not adjusted for population or be strongly correlated with more deaths per capita, etc.
+As mentioned in the introduction, the modeling goal for the COVID-19 case rate will be one of description.
+A model of causation may follow this research if potential causal relationships are examined and incorporated into the model's structure.
+This may later be expanded to predict the COVID-19 case rate in the past seven days in order for a state to use it to potentially judge what effect a mask policy could have.
 
@@
 
-The three models aim to investigate the relationship between the COVID rate, state demographic information, and mask mandates.
-The list of variables included in the current dataset that pertain to these concepts are listed below.
-
-State Characteristics:
-
-  - Governor
-  - Population density per square miles
-  - Population 2018
-  - Nonelderly Adults Who Have A Pre-Existing Condition
-  - Percent at risk for serious illness due to COVID
-  - All-cause deaths 2018 
-  - Number Homeless (2019)
-  - Medicaid Expenditures as a Percent of Total State Expenditures by Fund
-  - Life Expectancy at Birth (years)
-  - Percent Unemployed (2018)
-  - Percent living under the federal poverty line (2018)
-  - Weekly UI maximum amount with extra stimulus (through July 31, 2020) (dollars)
-  - Median Annual Household Income
-  - Closed other non-essential businesses
-  - Stay at home/ shelter in place
-  - Mandate face mask use by all individuals in public spaces
-  - No legal enforcement of face mask mandate
-
-Demographics:
-
- - Children 0-18 Percentage
- - Adults 19-25 Percentage
- - Adults 26-34 Percentage
- - Adults 35-54 Percentage
- - Adults 55-64 Percentage
- - 65+ Percentage
- - white Percentage of Total Population,
- - Black Percentage of Total Population, 
- - Hispanic Percentage of Total Population, 
- - Other Percentage of Total Population
-
-Currently, Case rate per 100,000 is chosen as a dependent variable to represent the pandemic status in each state.
-Variables such as population density per square miles, unemployed percentage(2018)/percentage living under the federal poverty line(2018), white/black percentage of total population, percentage of seniors(65+), and mask mandate policies are chosen as candidate covariates to explore potential relationships between pandemic status demographic characteristics, economic status, racial composition, and state policy.
-
-### Data Exploration
-
-Through data exploration, most data investigated is clean as numerical type except the black percentage of total population.
-This information of some states is reported vaguely as less than 0.01, which is replaced as 0 for further work.
-This could be updated with more precise data with the help of external resources.
-
-Several transformations have been made for variables as population density per square miles and white percentage of total population.
-Population density is extremely skewed due to the high population density of the District of Columbia.
-A logarithm transformation of this variable is of more uniform distribution.
-In addition, the square of the white percentage of the total population is of a distribution closer to the normal distribution.
-It also represents the chance that the interaction happens between two white people in practice.
-
-In addition, correlation values of white percentage and black percentage, unemployed percentage and percentage living under the federal poverty line shows these two groups of variables are correlated respectively.
+The three models aim to investigate the relationship between the COVID-19 case rate, state population demographics, and a mask mandates.
+COVID-19 case rate per 100,000 is chosen as the key dependent variable to represent the proliferation of the COVID-19 pandemic in each state.
+The population percentage of seniors is chosen as the key independent variable as a proxy for susceptibility by innate state population make-up.
+The variables of poverty percentage, unemployment percentage, mask mandate policy, and population density per square miles are chosen as candidate covariates to further explore the aforementioned potential relationship.
 

### 2020-12-03 model_1.Rmd
Updated introduction, assumptions, and intro to model buildling process

@@
 ```{r}
+df1 <- read.csv("../pop-urban-pct-historical.csv", header = TRUE)
+df1 <- df1 %>%
+  rename(State = 'Area.Name',
+         urban_percent_2010 = 'X2010'
+         ) %>%
+  select(State, urban_percent_2010
+         )
+head(df1)
+```
+
+```{r}
+df <- merge(df, df1, by="State")
+head(df)
+```
+
+```{r}
 df <- df %>% 
@@
 ggplot(data = df,
+  mapping = aes(x= urban_percent_2010 ))+
+  geom_histogram(fill = 'skyblue', color = 'grey30') +
+  labs(title =  "Histogram of Urban Population Percentage", x = "Urban percentage", y = 'Count')
+```
+
+```{r}
+ggplot(data = df,
   mapping = aes(x= X65_100k ))+
@@
   labs(title =  "Histogram of Adults Over 65 per 100,000", x = "Adults over 65 per 100,000", y = 'Count')
+```
+
+```{r}
+df %>% 
+  ggplot(aes(urban_percent_2010, case_rate_100k)) + 
+  geom_smooth(se = FALSE) +
+  geom_point() +   
+  labs(
+    title = 'COVID Case Rate due to Urban Population', 
+    x = 'Urban population percentage', 
+    y = 'COVID case rate per 100,000'
+  )
 ```
@@
 coeftest(model_test_65, vcov = vcovHC)
+```
+
+```{r}
+model_test_urban <- lm(case_rate_100k ~ urban_percent_2010, data = df)
+coeftest(model_test_urban, vcov = vcovHC)
+```
+
+```{r}
+plot(model1, which=5)
+```
+
+```{r}
+plot(model_test_urban, which=5)
 ```

### 2020-12-03 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd

@@
 The variables of poverty percentage, unemployment percentage, mask mandate policy, and population density per square miles are chosen as candidate covariates to further explore the aforementioned potential relationship.
+
+```{r}
+
+# percentage varialbe re-scale to number_per_100k (need to double check where to put this part in EDA or model regression)
+
+df<-df %>%
+  mutate(
+    noneld_pre_existing_per_100k = 100000*non_elderly_pre_existing/population_2018,
+    white_per_100k = 100000*white_population_pct,
+    senior_per_100k = 100000*senior_pct,
+    poverty_per_100k = 100000*poverty_pct, 
+    unemployed_per_100k = 100000*unemployed_pct,
+    log_population_density = log(population_density),
+    avg_distance = sqrt(2589988.11/population_density),
+    area_per_capita = 2589911.11/population_density
+  )
+```
 

### 2020-12-03 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Updated omitted variable section.

@@
 One of the variables that was hypothesized as being potentially influential is a ratio of a state's urban to rural population.
-Currently, the dataset in use contains population information as averaged across each state as a whole.
+Currently, the dataset contains population information as averaged across each state as a whole.
 While population density is potentially useful, it was found to not be significant in the models.
-This may be due to states which large urban centers, where population density woukd be high, losing their impact due to also containing large areas of land that is sparsely populated.
+This may be due to states which have large urban centers, where population density woukd be high, losing their impact due to also containing large areas of land that is sparsely populated.
 One example of this would be New York, which contains several of the most densely populated counties in the country, but also large rural areas.
-This variable would also be explanatory of the population density variable as a whole, where a higher urban to rural ratio would result in a higher statewide population density generally.
+
+An urban to rural ratio would be explanatory of the population density variable as a whole, where a higher urban to rural ratio would result in a higher statewide population density generally.
 With a greater fraction of urban residents, people would also be less likely to be able to social distance as effectively, increasing the per capita case rate.
-Therefore, the coefficient for population density in the original model would be larger than a model where this ratio is included, resulting in a bias that is positive and away from zero.
+Therefore, the coefficient for population density in the original model would be larger and more positive than a model where this ratio is included, resulting in a bias that is positive and away from zero.
 
@@
 It would also be explanatory in a way that makes the poverty percentage variable less impactful, where either of these variables may be considered as proxies for one another.
-Because inclusion of the median income variable would led to a lower magnitude poverty percentage variable, this would cause a positive omitted variable bias away from zero.
+Because inclusion of the median income variable would led to a lower magnitude poverty percentage variable, and that the coefficient of the poverty percentage variable is positive, this would cause a positive omitted variable bias away from zero.
 
@@
 
-This variable could also be partially explained by the senior percentage variable, because of the gap in gender ratios across generations and generally as people get older.
-Because men have lower life expectancies, older groups of people living in the US would have higher ratios of women to men.
-Therefore, introducing the gender ratio variable into the model would lower the value of the coefficient to the senior percentage variable.
-This would be a positive omitted variable bias away from zero.
+This variable could also be partially explained by the senior percentage variable, because of the increase of women in gender ratios generally as people get older.
+Therefore, introducing the gender ratio variable into the model would lower the value of the coefficient to the senior percentage variable to be less negative.
+This would be a negative omitted variable bias away towards zero.
 
@@
 
-A variable capturing the mask use among the population would likely reduce the explanatory power of the state mask mandate variable.
-This would cause the bias of the omitted variable to be positive and away from zero.
+A variable capturing the mask use among the population would likely reduce the explanatory power of the state mask mandate variable to be less negative.
+This would cause the bias of the omitted variable to be negative and towards from zero.
 
@@
 The economic standing of a state not only affects the individual households of its population, but also includes how it interacts with other states and how its peoples' lives might have been economically changed by the pandemic.
-One way that a state may increase its number of covid cases is both from tourists or others visiting during the pandemic, but also if its own population must work in close contact with others.
+One way that a state may increase its number of covid cases is both from tourists or others visiting during the pandemic and bringing disease, but also if its own population must work in close contact with others serving them in the tourism industry.
 In addition, if certain key industries to a state, such as tourism, layed off many workers, they would likely have to take lower paying close contact jobs such as in the service or gig economy in order to continue having a livelihood.
+This would further lead to a higher covid case rate
 
-This effect can be included with a variable that demonstrates how much of a state's GDP is from tourism.
+These effects may be included with a variable that demonstrates how much of a state's GDP is from tourism.
 It would be related to other economic indicator variables, such as the percent of the population living in poverty.
-While poverty is a broader phenomena, the number of people in poverty in the pandemic when the data was taken would be potentially higher due to negative effects in the tourism industry.
+While poverty is a broader phenomena, the number of people in poverty in the pandemic could be potentially higher due to negative growth of the tourism industry.
 Therefore, it is expected that states with a greater focus on tourism may also have more people in poverty in the pandemic.
-This would take some of the explanatory power of the poverty variable away, resulting in a positive omitted variable bias away from zero.
+This would take some of the explanatory power of the poverty variable away, resulting in its model coefficient becoming less positive.
+This would result in a positive omitted variable bias away from zero.
 

### 2020-12-03 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Update Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
updates in model 3

@@
 According to the documentation, zero represents "the absence of an order or directive", which can be interpreted as the policy is not enforced by the state explicitly.
-For a leaner regression, the date values in mask_public column are transformed as value 1, so that we can distinguish whether states has public mask policy or not.
+For linear regression, the date values in mask_public column are transformed as value 1, so that we can distinguish whether states has public mask policy or not.
 It is also noted that by transforming the variable, some important information will be lost because the actual date (early or late) to enforce the policy can also have an impact to the case rate.
-However, it is hard to measure the time effect as different states that enforced the public mask policy may have different threshold.
+However, it is difficult to measure the time effect as different states that enforced the public mask policy may depend on related situations or different thresholds.
 
 Firstly, "mask_public_bool" column is created from "mask_public", with 1 represents that the state has a pubic mask policy, and 0 means there's no explicit public mask policy from the state.
+The summary shows that there are 35 states that enforced public mask mandate, and 16 states that had no public mask policy.
+The ratio looks reasonable for the analysis considering sample size of 51 as there is decent number of samples for each state policy type.
 
@@
   )
+
+cat('Count for states that enforced mask mandate in public: ', 
+    length(df$mask_public_bool[df$mask_public_bool==1]))
+
+cat('Count for states did not enforced mask mandate in public: ', 
+    length(df$mask_public_bool[df$mask_public_bool==0]))
 ```
 
-Histograms below that shows the distribution of states that has public mask policy or not. Both distributions exhibits certain degree of normality.
+Histograms below that shows the distribution of states that has public mask policy or not. Both distributions are not very normal with relatively small skews toward both tails.
 
 ```{r}
+
 hist(df$case_rate_100k[df$mask_public_bool == 1], xlab='Case Rate', ylab='Count', 
-     main='Case Rate - Mask Mandate in Public')
+     col='skyblue', main='Case Rate - States Enforced Mask Mandate in Public')
 hist(df$case_rate_100k[df$mask_public_bool == 0], xlab='Case Rate', ylab='Count', 
-     main='Case Rate  - No Mask Mandate in Public')
+     col='skyblue', main='Case Rate  - States Enformced No Mask Mandate in Public')
 ```
@@
 
-#### Model 3-a
+#### Model 3 Regression
 
@@
 ```{r}
-model3a <- lm(case_rate_100k ~ senior_per_100k + log(population_density) 
-              + noneld_pre_existing_per_100k + poverty_per_100k + white_per_100k 
-              + unemployed_per_100k + mask_public_bool, data = df)
-coeftest(model3a, vcov = vcovHC)
 
-model3a <- lm(case_rate_100k ~ senior_per_100k + log(population_density) 
+model3 <- lm(case_rate_100k ~ senior_per_100k 
               + poverty_per_100k 
-              + unemployed_per_100k + mask_public_bool, data = df)
-coeftest(model3a, vcov = vcovHC)
+              + unemployed_per_100k 
+             + mask_public_bool
+             + log(population_density), data = df)
+coeftest(model3, vcov = vcovHC)
 ```
@@
 #Zero Conditional Mean
-plot(model3a, which=1)
+plot(model3, which=1)
 
 #Normality of errors
-plot(model3a, which=2)
+plot(model3, which=2)
 
 #Homoskedasticity
-plot(model3a, which=3)
+plot(model3, which=3)
 
 #Cook distance
-plot(model3a, which=5)
+plot(model3, which=5)
 
 #Collinearity
-vif(model3a)>4
+vif(model3)
 ```
-
-#### Model 3-b & 3-c
-
-Considering that there could be multiple policy variables that are relevant to predicting the case number.
-We have also examined the "Closed.other.non.essential.businesses" variable (renamed as "business_closed") which represents the date when the state issued closure of non-essential businesses.
-As "business_closed" is a date variable, a transformation was performed to convert the date into a variable "business_close_diff" which represents the days difference between the business close date of the state compared to the average date of all the states.
-
-```{r}
-df$business_close_diff =as.numeric(df$business_closed-mean(df$business_closed, na.rm=TRUE))
-```
-
-From the plot of business close date difference against case rate, the graph shows that there is some degree of linearity between the two variables.
-
-```{r warning=FALSE, message=FALSE}
-df %>% 
-  ggplot(aes(x = business_close_diff, y = case_rate_100k, color = business_close_diff)) + 
-  geom_point() + 
-  geom_smooth(se = FALSE) +
-  labs(
-    title = 'Case Rate vs Non-Essential Business Close Date difference \ncompared to Average', 
-    x = 'Non-Essential Business Close Date difference compared to Average', 
-    y = 'Case Rate', 
-    color = 'Case Rate'
-  )
-```
-
-By running the model with "business_close_diff" against case rate variable, the coefficient shows that the variable is significant.
-
-```{r}
-model3b <- lm(case_rate_100k ~ business_close_diff, data = df)
-coeftest(model3b, vcov = vcovHC)
-```
-
-However, the regression with both policy variables "mask_pubic_bool" and "business_close_diff" shows that neither the the variable is significant.
-
-```{r}
-model3c <- lm(case_rate_100k ~ business_close_diff + mask_public_bool, data = df)
-coeftest(model3c, vcov = vcovHC)
-```
-
-The findings above suggests that there is colinearity between the two variables "mask_pubic_bool" and "business_close_diff".
-Therefore, we decided to use "mask_public_bool" variable which has higher significance to represent our policy variable.
-As a result, model 3-a is selected as the final model for model 3.
 
@@
 avg_mask <- mean(df$mask_public_bool)
-y_pop <- model3a$coefficients[1] + model3a$coefficients[2]*log10(df$population_density) + model3a$coefficients[3]*avg_pov + model3a$coefficients[4]*avg_mask
-y_pov <- model3a$coefficients[1] + model3a$coefficients[2]*avg_pop + model3a$coefficients[3]*df$poverty_pct + model3a$coefficients[4]*avg_mask
-y_pop_nomask <- model3a$coefficients[1] + model3a$coefficients[2]*avg_pop + model3a$coefficients[3]*avg_pov
-y_pov_nomask <- model3a$coefficients[1] + model3a$coefficients[2]*avg_pop + model3a$coefficients[3]*df$poverty_pct
+y_pop <- model3$coefficients[1] + model3$coefficients[2]*log10(df$population_density) + model3$coefficients[3]*avg_pov + model3$coefficients[4]*avg_mask
+y_pov <- model3$coefficients[1] + model3$coefficients[2]*avg_pop + model3$coefficients[3]*df$poverty_pct + model3$coefficients[4]*avg_mask
+y_pop_nomask <- model3$coefficients[1] + model3$coefficients[2]*avg_pop + model3$coefficients[3]*avg_pov
+y_pov_nomask <- model3$coefficients[1] + model3$coefficients[2]*avg_pop + model3$coefficients[3]*df$poverty_pct
 
@@
 df %>% {
-   ggplot(mapping = aes(x = log(df$population_density), y = model3a$residuals)) +
+   ggplot(mapping = aes(x = log(df$population_density), y = model3$residuals)) +
       geom_point() + stat_smooth(se = TRUE) +
@@
 df %>% {
-   ggplot(mapping = aes(x = df$poverty_pct, y = model3a$residuals)) +
+   ggplot(mapping = aes(x = df$poverty_pct, y = model3$residuals)) +
       geom_point() + stat_smooth(se = TRUE) +
@@
 df %>% {
-   ggplot(mapping = aes(x = log(df$mask_public_bool), y = model3a$residuals)) +
+   ggplot(mapping = aes(x = log(df$mask_public_bool), y = model3$residuals)) +
       geom_point() + stat_smooth(se = TRUE) +
@@
 ```{r}
-hist(model3a$residuals, main = "Model Residuals", xlab = "Residual")
+hist(model3$residuals, main = "Model Residuals", xlab = "Residual")
 ```
@@
 
-```{r, echo=FALSE, warninng=False}
+```{r, echo=FALSE, warning=FALSE}
 plot_one <- df %>% 
@@
 se.model2 = coeftest(model2, vcov = vcovHC)[ , "Std. Error"]
-se.model3 = coeftest(model3a, vcov = vcovHC)[ , "Std. Error"]
+se.model3 = coeftest(model3, vcov = vcovHC)[ , "Std. Error"]
 
-stargazer(model1, model2, model3a, type = "text", omit.stat = "f",
+stargazer(model1, model2, model3, type = "text", omit.stat = "f",
           se = list(se.model3),

### 2020-12-03 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Updated model 1

@@
 With this information, one could suggest whether or not the proliferation of COVID-19 seems to be related to either or both controllable and uncontrollable variables, and to the level at which the effect of these variables differ help.
-This modeling goal of this research question will be one of description, and will be broken up into three phases of investigation
+The modeling goal of this research question will be one of description, and will be broken up into three phases of investigation
 
@@
 The case rate per 100,000 was selected as the key output/dependent variable as it provides a standardized measure of the spread of COVID-19 across the states, taking into account the absolute population of the states.
-The senior percentage was selected as the key input/independent variable given that it is posited in external research that seniors make up a significant number of the reported cases.
-This is likely due to the both the fact that seniors in this age group were more likely to be or had greater access to be tested given that they are categorized in the high risk category for COVID-19, and the fact that they would have a greater likelihood of exhibiting detectable COVID-19 related symptoms that would prompt them to get tested.
-An additional rationale for measuring the senior percentage is that they may also be more likely to contract COVID-19 at lower viral loads, making it easier for them more susceptible to the virus.
-These key variables will provide an initial understanding of the relationship between the spread of the virus and a key metric of susceptibility and detectability of COVID-19.
+The senior percentage was selected as the key input/independent variable given that it is posited in external research that seniors belong to the category for those at higher risk of COVID-19, and thus are a well studied and documented group.
+This is likely due to the both the fact that seniors in this age group are more likely to be or had greater access to be tested given their categorization and the fact that they would have a greater likelihood of exhibiting detectable COVID-19 related symptoms that would prompt them to get tested.
+An additional rationale for measuring the senior percentage is that they may also be more likely to contract COVID-19 at lower viral loads, making them more susceptible to the virus.
+Conversely, this may result in seniors being more wary of the threat of COVID-19, and taking additional precautions to prevent infection relative to other age categories, such as limiting social interactions and taking more preventative measures in terms of hygiene.
+These key variables will provide an initial understanding of the relationship between the spread of the virus and population demographics.
 
-Following this, in the second phase of investigation the analysis considers other variables of state-level population that are considered factors of susceptibility, namely the rate of poverty (defined as the percent of individuals living under the federal poverty line in 2018) and the rate of unemployment (defined as the percent unemployment in 2018.).
+Following this, in the second phase of investigation the analysis considers other variables of state-level population that are considered factors of contractability, namely the rate of poverty (defined as the percent of individuals living under the federal poverty line in 2018) and the rate of unemployment (defined as the percent unemployment in 2018).
+Those living below the poverty line may have less access/are unable to afford to preventative controls such as sanitation products and masks that help prevent the spread of the virus, while those who are designated as unemployed may not have the flexibility of sheltering at home as well as access to the aforementioned preventative controls.
 As mentioned, these variables operationalize population metrics that may lead to greater case rates, and are already standardized as rates to account for varying absolute populations across states.
 
-The tertiary and final phase of investigation will include a variable that measures the implementation of policy as a response to COVID-19, and to understand the added effect of this variable in conjunction with the first phase of investigation that is state population metrics.
+The tertiary and final phase of investigation includes a variable that measures the implementation of policy as a response to COVID-19, and to understand the added effect of this variable in conjunction with the first phase of investigation that is state population metrics.
 Specifically, there is a focus on a policy that mandates the wearing of masks within the state.
 Through this, the aim is to analyze whether the implementation of a mask-related policy will have an effect on case rate, as well as the degree of this effect relative to population metrics.
-To operationalize this policy and simplify the measurements, the models will only consider whether or not this policy was implemented through transformed indicator variables (1 = implemented, 0 = not implemented).
+To operationalize this policy and simplify the measurements, the models only consider whether or not this policy was implemented through transformed indicator variables (1 = implemented, 0 = not implemented).
 As a result, factors of when and for how long the policy was implemented will be lost.
@@
 
-Additionally, the final phase of investigation will also continue to expand on the examination of state-level population demographics via the population density variable (defined as the population density in square miles, population/square miles).
+Additionally, the final phase of investigation continues to expand on the examination of state-level population demographics via the population density variable (defined as the population density in square miles, population/square miles).
 A key factor to the spread of COVID-19 is the idea of social distancing, where the virus is considered to be more likely to spread when people are close in proximity.
@@
 ```{r}  
-
-# percentage varialbe re-scale to number_per_100k (need to double check where to put this part in EDA or model regression)
+# percentage variable re-scale to number_per_100k (need to double check where to put this part in EDA or model regression)
 
@@
     unemployed_per_100k = 100000*unemployed_pct,
-    log_population_density = log(population_density),
-    avg_distance = sqrt(2589988.11/population_density),
-    area_per_capita = 2589911.11/population_density
+    log_population_density = log(population_density)
   )
@@
 
-For the first model, the relationship between the COVID case rate per 100,000 and the population density per square mile of states was analyzed.
-First, the distribution of COVID case rate per 100,000 dependent variable is examined.
+For the first model, the relationship between the COVID-19 case rate per 100,000 and the population percentage distribution of seniors was analyzed.
+First, the distribution of the case rate dependent variable is examined.
 
@@
 
-```{r}
+```{r, warning=FALSE, message=FALSE}
 ggplot(data = df,
@@
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,6000,600)) +
-  labs(title =  "Histogram of COVID Case Rate per 100,000", 
-       x = "COVID case rate per 100,000", y = 'Count') +
+  labs(title =  "Histogram of COVID-19 Case Rate per 100,000", 
+       x = "COVID-19 case rate per 100,000", y = 'Count') +
   scale_x_continuous(breaks=seq(0, 6000, 600))
@@
 
-Next, the distribution of population density per square mile variable is examined.
+Next, the distribution of the senior percentage variable is examined.
+
+```{r, warning=FALSE, message=FALSE}
+summary(df$senior_pct*100)
+
+ggplot(data = df,
+  mapping = aes(x= senior_pct*100)) +
+  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(10,21,1)) +
+  labs(title = 'Histogram of Percentage Population of Seniors', 
+  x = 'Percentage population of seniors (%)', y = 'Count') +
+  scale_x_continuous(breaks=seq(10,22,1))
+```
+
+As can be seen from the distribution above, again there is a fairly normal distribution in the percentage population of seniors across the state.
+Additionally, there do not seem to be any outliers within the distribution, given that the range of percentages fall between 11-21%.
+Therefore, there are no transformations that need to be made in order to restructure the distribution.
+However, it was determined that the percentages would be transformed into a rate per 100,000 value in order to allow for easier interpretation when analyzing the effect on the output variable which itself is based on scale of rate per 100,000.
+Given that this transformation is constant across all values, it has no effect on the normality of the distribution itself, as seen in the histogram below.
 
 ```{r}
-histogram_of_pdensity <- df %>%
-  ggplot(aes(x = population_density)) + 
-  geom_histogram(fill = 'skyblue', color = 'grey30', bins = 20) + 
-  labs(
-    title = 'Distribution of Population Density per Square Miles', 
-    x = 'Population density per square mile', y = 'Count')
+df<-df %>%
+  mutate(
+    senior_per_100k = 100000*senior_pct)
 
-histogram_of_pdensity
-```
-
-As can be seen from the histogram, although most of the population density is concentrated in the 0 to 1500 range, there are some grouping of outliers that are very far from this concentration.
-When an analysis is performed, it can be seen that there is only one data sample that is the outlier, which is the District of Columbia (D.C.) with a value of 11,496.
-This is given due to the fact that D.C.
-is a district that solely consists of a large city, as mentioned as a possibility in the introduction.
-Given that this causes the data to be skewed, the logarithm is taken to scale the variable.
-Alternatively, the data point could have been droppped from the sample, but given the already small sample size, it was determined that a better approach would be to keep it within the sample.
-Once this transformation was performed, it is shown that there is a relatively normal distribution of population densities (see figure below).
-
-```{r}
-# Find the outlier data points
-outliers <- subset(df, population_density > 4000)
-paste(outliers$State, '=', outliers$population_density)
-
-# Transform the variable by taking the logarithm and assign it to a new variable
-df <- df %>% 
-  mutate(l_population_density = log10(population_density))
-
-# Plot the new distribution in a histogram
-histogram_of_pdensity <- df %>%
-  ggplot(aes(x = l_population_density)) + 
-  geom_histogram(fill = 'skyblue', color = 'grey30', bins = 20) + 
-  labs(
-    title = 'Distribution of Population Density per Square Miles', 
-    x = 'Log of population density per square mile', y = 'Count')
-
-histogram_of_pdensity
+ggplot(data = df,
+  mapping = aes(x= senior_per_100k)) +
+  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(10000,21000,1000)) +
+  labs(title = 'Histogram of Seniors per 100,000', 
+  x = 'Population of seniors per 100,000', y = 'Count') + 
+  scale_x_continuous(breaks=seq(10000,22000,1000))
 ```
@@
 df %>% 
-  ggplot(aes(l_population_density, case_rate_100k)) + 
+  ggplot(aes(senior_per_100k, case_rate_100k)) + 
   geom_smooth(se = FALSE) +
@@
   labs(
-    title = 'COVID Case Rate due to Population Density', 
-    x = 'Log of population density per square miles', 
+    title = 'COVID-19 Case Rate due to Senior Population', 
+    x = 'Population of senior per 100,000', 
     y = 'COVID case rate per 100,000'
@@
 
-From the above plot, it can be posited that there is no discernible relationship between the variables given the non-linear relationship.
+From the above plot, there is noticeable variation in the smoothed blue line.
+However, it can be posited that there is a inverse relationship between the variables given the roughly linear relationship in the overall downward trend in the line.
 In order to test this, the following equation is used to create a regression model to determine the true relationship between the case rate and population density variables.
@@
 $$
-Case.Rate.Per.100000 =\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)
+case{\_}rate{\_}100k =\beta_0 + \beta_1 senior{\_}per{\_}100k
 $$
@@
 ```{r}
-model1 <- lm(case_rate_100k ~ log10(population_density) , data = df)
+model1 <- lm(case_rate_100k ~ senior_per_100k , data = df)
 coeftest(model1, vcov = vcovHC)
@@
 
-From the results of the regression, two things are identified.
-Firstly, the result of the coefficient is nowhere near significant, and therefore we fail to reject the null hypothesis that there is no correlation between case rate and population density, and there is no detectable effect of the independent variable population density.
-However, the coefficient of population density appears to be negative, which would concur with the assumption that the higher the population density, the lower the COVID case rate in a state.
+From the results of the regression using robust standard errors, the p-value of the senior rate variable is 0.056, which is close but not quite within the significant level of >0.05 to reject the null hypothesis that there is no correlation between case rate and senior rate.
+To improve the precision of the t-test, classical standard errors can be used.
+To determine the applicability of using classical standard errors, in addition to the three Classical Linear Model (CLM) assumptions necessary for the robust standard errors, the CLM assumption for homoskedastic conditional errors should be evaluated.
+If this additional assumption is met, the robust standard errors can be replaced by the classical standard errors.
+The assumption of homoskedasticity is met if the data does not have large variance among the residuals.
+
+To evaluate this assumption, the square root of the residuals (to remove any negative values) is plotted against the fitted values of the model.
+Additionally, a Breusch-Pagan test is run to check the level of heteroskedasticity.
+
+```{r}
+# Check the homoskedasticity of errors in model 1
+plot(model1, which=3)
+
+# Run the Breusch-Pagan test for model 1
+bptest(model1)
+```
+
+From the above plot, the variance in residuals can be seen to be roughly constant given the relative flatness of the plotted red line, thus meeting the assumption of homoskedastic conditional errors.
+This is supported by the p-value of 0.497 from the Breusch-Pagan test, meaning there is a failure to reject that null hypothesis that the conditional errors are not heteroskedastic.
+With this additional assumption met, the regression model is rerun using a t-test with classical standard errors.
+
+```{r}
+coeftest(model1)
+```
+
+As can be seen from the above output, with the improved precision of the classical standard errors the variable for senior percentage is shown to be significant within the model specification with a p-value of 0.023.
+Furthermore, The coefficient of -0.174 for the variable is negative, supporting the earlier evaluation that there is an inverse relationship between the senior percentage and the case rate.
+In practical significance, this can be interpreted as for every additional senior per 100,000 in the population, there is a decrease of 0.174 COVID-19 cases per 100,000.
+More intuitively, for roughly every 6 additional seniors per 100,000, there is a decrease of 1 case per 100,000.
+
+Next, the other CLM assumptions are revisited to determine the whether the coefficients are unbiased and consistent.
+However, the assumption for IID is not detailed here given that it is evaluated more generally for all three models within section 3.
+
+The second CLM assumption to be evaluated here is that of a linear conditional expectation relationship within the data, so that the model is accurately estimating the relationship between the variables.
+To test for this, the residuals of model 1 are plotted against its fitted values.
+
+```{r}
+# Check for linearty in model 1
+plot(model1, which=1)
+```
+
+When examining the plot above for linearity in the conditional expectation, there is a noticeable slightly quadratic shape to the relationship of the conditional expectation.
+This may suggest that the relationship between the variables may not be as linear as the model assumes.
+However, as discussed previously in the scatter plot of the two variables, the relationship can be approximated as roughly linear.
+It is also difficult to explain the complexity behind the COVID-19 case rate with just a single variable that does not fully capture the broader population demographics.
+Additionally, there is hope that given the addition of other input variables within models 2 and 3, the relationship within the data will be better explained and result in more linearity.
+
+The third CLM assumption examined for model 1 is the assumption of normality of errors.
+This is a necessary assumption to ascertain that the errors used in the regression are drawn from a normal distribution, so that they can be accepted when used to calculate the significance levels.
+Both a QQ plot and histogram of the distribution of the residuals is plotted to observe the normality.
+
+```{r, message = FALSE}
+# Create a QQ plot for the model 1 residuals
+plot(model1, which=2)
+
+# Create a histogram of the distribution of the model 1 residuals
+df %>%
+  ggplot(aes(x = resid(model1))) +
+  stat_bin() +
+  geom_histogram(fill = 'skyblue', color = 'grey30', bins=30) +
+  labs(title =  "Histogram of the Model 1 Residuals", 
+       x = "Model 1 residuals", y = 'Count')
+```
+
+From the QQ plot, normality in the residuals can be detected given the proximity of the points to the normal line.
+Additionally, although the shape is a little harder to observe, the histogram of the residuals also gives a fairly normal distribution.
+
+Finally the Cook's distance is examined by a plot of the residuals vs the model 1 leverage to look for any outliers in the values. From the plot below, no obvious outliers are detected.
+
+```{r, message = FALSE}
+# Investigate outliers using Cook's distance
+plot(model1, which=5)
+```
+
+It is also noted that there is no test conducted for collinearity given there is only one input variable in model 1.
 
@@
 
+### Sam - @Frank, this was my previous analysis of the POPULATION DENSITY variable, feel free to use this
+
+Next, the distribution of the population density variable is examined.
+
+```{r}
+# Plot the distribution in a histogram
+histogram_of_pdensity <- df %>%
+  ggplot(aes(x = population_density)) + 
+  geom_histogram(fill = 'skyblue', color = 'grey30', bins = 20) + 
+  labs(
+    title = 'Distribution of Population Density per Square Miles', 
+    x = 'Population density per square mile', y = 'Count')
+
+histogram_of_pdensity
+```
+
+As can be seen from the histogram, although most of the population density is concentrated in the 0 to 1500 range, there are some grouping of outliers that are very far from this concentration.
+When an analysis is performed, it can be seen that there is only one data sample that is the outlier, which is D.C.
+with a value of 11,496.
+This is given due to the fact that D.C.
+is a district that solely consists of a large city, as mentioned as a possibility in the introduction.
+Given that this causes the data to be skewed, the logarithm is taken to scale the variable.
+Alternatively, the data point could have been droppped from the sample, but given the already small sample size, it was determined that a better approach would be to keep it within the sample.
+Once this transformation was performed, it is shown that there is a relatively normal distribution of population densities (see figure below).
+
+```{r}
+# Find the outlier data points
+outliers <- subset(df, population_density > 4000)
+paste(outliers$State, '=', outliers$population_density)
+
+# Transform the variable by taking the logarithm and assign it to a new variable
+df <- df %>% 
+  mutate(l_population_density = log10(population_density))
+
+# Plot the new distribution in a histogram
+histogram_of_pdensity <- df %>%
+  ggplot(aes(x = l_population_density)) + 
+  geom_histogram(fill = 'skyblue', color = 'grey30', bins = 20) + 
+  labs(
+    title = 'Distribution of Population Density per Square Miles', 
+    x = 'Log of population density per square mile', y = 'Count')
+
+histogram_of_pdensity
+```
+
+With the appropriate variables transformed, a plot is created to show the relationship between them.
+
+```{r message=FALSE}
+df %>% 
+  ggplot(aes(l_population_density, case_rate_100k)) + 
+  geom_smooth(se = FALSE) +
+  geom_point() +   
+  labs(
+    title = 'COVID Case Rate due to Population Density', 
+    x = 'Log of population density per square miles', 
+    y = 'COVID case rate per 100,000'
+  )
+```
+
 #### Model 3 Regression
@@
 ```{r warning=FALSE}
-se.model1 = coeftest(model1, vcov = vcovHC)[ , "Std. Error"]
-se.model2 = coeftest(model2, vcov = vcovHC)[ , "Std. Error"]
-se.model3 = coeftest(model3, vcov = vcovHC)[ , "Std. Error"]
+se.model1 = coeftest(model1)[ , "Std. Error"]
+se.model2 = coeftest(model2)[ , "Std. Error"]
+se.model3 = coeftest(model3)[ , "Std. Error"]
 
-stargazer(model1, model2, model3, type = "text", omit.stat = "f",
+stargazer(model1, model2, model3, type = "text",
           se = list(se.model3),

### 2020-12-04 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Added bptest to model 2
Updated intro, initial model buildling process section, and model 1

@@
 The case rate per 100,000 was selected as the key output/dependent variable as it provides a standardized measure of the spread of COVID-19 across the states, taking into account the absolute population of the states.
-The senior percentage was selected as the key input/independent variable given that it is posited in external research that seniors belong to the category for those at higher risk of COVID-19, and thus are a well studied and documented group.
-This is likely due to the both the fact that seniors in this age group are more likely to be or had greater access to be tested given their categorization and the fact that they would have a greater likelihood of exhibiting detectable COVID-19 related symptoms that would prompt them to get tested.
+The senior percentage was selected as the key input/independent variable given that guidance had been released by the Centers for Disease Control and Prevention (CDC) that seniors belong to the age category for those at higher risk of COVID-19, and thus are a well studied and documented group.
+This is likely due to the both the fact that seniors in this age group are more likely or had greater access to be tested given their categorization and the fact that they would have a greater likelihood of exhibiting detectable COVID-19 related symptoms that would prompt them to get tested.
 An additional rationale for measuring the senior percentage is that they may also be more likely to contract COVID-19 at lower viral loads, making them more susceptible to the virus.
@@
 
+In order to allow for more simplistic and direct interpretations when analyzing the effect of the input variables on the output, it was determined that all percentage variables (senior, poverty, and unemployment) would be transformed into rates per 100,000 to align with the case rate variable.
+It will also serve to ease the burden on the readers of this report as like-for-like scaled relationships can be drawn between the variables.
+Given that this transformation is constant for all values (scalar multiplication of 100,000), it will have no effect on the distributions of the variables.
+
 ```{r}  
-# percentage variable re-scale to number_per_100k (need to double check where to put this part in EDA or model regression)
+# Percentage variables are re-scaled to raters per 100k
 
-df<-df %>%
+df <-df %>%
   mutate(    
-    noneld_pre_existing_per_100k = 100000*non_elderly_pre_existing/population_2018,
-    white_per_100k = 100000*white_population_pct,
     senior_per_100k = 100000*senior_pct,
@@
     unemployed_per_100k = 100000*unemployed_pct,
-    log_population_density = log(population_density)
   )
@@
 
-Next, the distribution of the senior percentage variable is examined.
+Next, the distribution of the senior percentage transformed into the rate per 100,000 (as discussed in the Model Variables section) is examined.
 
 ```{r, warning=FALSE, message=FALSE}
-summary(df$senior_pct*100)
-
-ggplot(data = df,
-  mapping = aes(x= senior_pct*100)) +
-  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(10,21,1)) +
-  labs(title = 'Histogram of Percentage Population of Seniors', 
-  x = 'Percentage population of seniors (%)', y = 'Count') +
-  scale_x_continuous(breaks=seq(10,22,1))
-```
-
-As can be seen from the distribution above, again there is a fairly normal distribution in the percentage population of seniors across the state.
-Additionally, there do not seem to be any outliers within the distribution, given that the range of percentages fall between 11-21%.
-Therefore, there are no transformations that need to be made in order to restructure the distribution.
-However, it was determined that the percentages would be transformed into a rate per 100,000 value in order to allow for easier interpretation when analyzing the effect on the output variable which itself is based on scale of rate per 100,000.
-Given that this transformation is constant across all values, it has no effect on the normality of the distribution itself, as seen in the histogram below.
-
-```{r}
-df<-df %>%
-  mutate(
-    senior_per_100k = 100000*senior_pct)
+summary(df$senior_per_100k)
 
@@
 
-With the appropriate variables transformed, a plot is created to show the relationship between them.
+As can be seen from the distribution above, again there is a fairly normal distribution in the population of seniors per 100,000 across the state.
+Additionally, there do not seem to be any outliers within the distribution, given that the range of rates fall between 11,000 and 21,000.
+Therefore, there are no transformations that need to be made in order to restructure the distribution, and this variable can be used across all three models.
+
+With the appropriate variable distributions investigated, a plot is created to analyze the relationship between them.
 
@@
 ```{r}
+# Build the regression model for Model 1
 model1 <- lm(case_rate_100k ~ senior_per_100k , data = df)
-coeftest(model1, vcov = vcovHC)
 ```
 
-From the results of the regression using robust standard errors, the p-value of the senior rate variable is 0.056, which is close but not quite within the significant level of >0.05 to reject the null hypothesis that there is no correlation between case rate and senior rate.
-To improve the precision of the t-test, classical standard errors can be used.
+To improve the precision of the t-test of coefficients, classical standard errors can be used over robust standard errors to test the regression.
 To determine the applicability of using classical standard errors, in addition to the three Classical Linear Model (CLM) assumptions necessary for the robust standard errors, the CLM assumption for homoskedastic conditional errors should be evaluated.
@@
 From the above plot, the variance in residuals can be seen to be roughly constant given the relative flatness of the plotted red line, thus meeting the assumption of homoskedastic conditional errors.
-This is supported by the p-value of 0.497 from the Breusch-Pagan test, meaning there is a failure to reject that null hypothesis that the conditional errors are not heteroskedastic.
-With this additional assumption met, the regression model is rerun using a t-test with classical standard errors.
+This is supported by the p-value of 0.497 from the Breusch-Pagan test, meaning it fails to reject the null hypothesis that the conditional errors are not heteroskedastic.
+With this additional assumption met, the regression model is run using a t-test with classical standard errors.
 
@@
 
-As can be seen from the above output, with the improved precision of the classical standard errors the variable for senior percentage is shown to be significant within the model specification with a p-value of 0.023.
+As can be seen from the above output, the variable for senior percentage is shown to be significant within the model specification with a p-value of 0.023.
 Furthermore, The coefficient of -0.174 for the variable is negative, supporting the earlier evaluation that there is an inverse relationship between the senior percentage and the case rate.
@@
 
-It is also noted that there is no test conducted for collinearity given there is only one input variable in model 1.
+It is also noted that there is no test conducted for collinearity in model 1 given there is only one input variable.
 
@@
 
+```{r, echo=FALSE}
+bptest(model2)
+```
+
 ####v.Normally Distributed Errors

### 2020-12-04 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Updated regression table and discussion

@@
 The research question aims to measure the COVID-19 case rate per 100,000 residents within each state based on the make-up of the population, as well as the policy decisions that were or were not put into place in order to combat the rise of COVID-19 cases.
-In this sense, the aim is to examine how dependent the COVID-19 case rate is on that of which cannot be controlled (i.e., population demographics), as well as that of which can be controlled to a certain degree (i.e., implementation of policies to attempt to combat case rate).
-With this information, one could suggest whether or not the proliferation of COVID-19 seems to be related to either or both controllable and uncontrollable variables, and to the level at which the effect of these variables differ help.
+In this sense, the aim is to examine how dependent the COVID-19 case rate is on that of which cannot be controlled (i.e., population demographics), as well as that of which can be controlled to a certain degree (i.e., implementation of policies to attempt to combat case rate) by each state.
+With this information, one could suggest whether or not the proliferation of COVID-19 within a state seems to be related to either or both controllable and uncontrollable variables, and to the level at which the effect of these variables differ help.
 The modeling goal of this research question will be one of description, and will be broken up into three phases of investigation
@@
 For example, the population density can have large variability depending on the population concentration among a few large areas as well as the level of uninhabited or sparsely inhabited land within a state, while the demographics of population can depend on a wide range of things such as regional factors and employment opportunities.
+This point is further discussed in the IID assumption addressed in section 3 of this report.
 
@@
 ```{r}  
-# Percentage variables are re-scaled to raters per 100k
+# Percentage variables are re-scaled to rates per 100k
 
@@
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,6000,600)) +
-  labs(title =  "Histogram of COVID-19 Case Rate per 100,000", 
+  labs(title =  "Histogram of COVID-19 Case Rate", 
        x = "COVID-19 case rate per 100,000", y = 'Count') +
@@
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(10000,21000,1000)) +
-  labs(title = 'Histogram of Seniors per 100,000', 
-  x = 'Population of seniors per 100,000', y = 'Count') + 
+  labs(title = 'Histogram of Seniors Distribution', 
+  x = 'Number of seniors per 100,000 people', y = 'Count') + 
   scale_x_continuous(breaks=seq(10000,22000,1000))
@@
     title = 'COVID-19 Case Rate due to Senior Population', 
-    x = 'Population of senior per 100,000', 
+    x = 'Number of seniors per 100,000', 
     y = 'COVID case rate per 100,000'
@@
 
-### Sam - @Frank, this was my previous analysis of the POPULATION DENSITY variable, feel free to use this
+### @Frank, this was my previous analysis of the POPULATION DENSITY variable, feel free to use any of this
 
@@
 
-Although there were several versions of models created for model 2 and 3, through the analysis of the appropriate variables respective to each model a single model version was selected to be included in the regression table.
-This was determined based on factors of significance when investigating the effect relationship of these variables on the outcome variable, which is detailed in section 2.
-
-For the regression table below, model 1, model 2-b, and model 3-a were selected as the most significant models to include.
+With the analysis for the selected variables, the three corresponding models, and the respective diagnostics of the CLM assumptions complete, the models can then be included for comparison in a regression table.
+Model 1, model 2, and model 3 are included in the regression table below,.
 
@@
           star.cutoffs = c(0.05, 0.01, 0.001), 
-          title = "Table 1: The relationship between COVID case rate and population density")
-
+          title = "Table 1: The effect of population demographics and mask policy on COVID-19 case rate")
 ```
@@
 
-# 5-1: Urban vs Rural Population Ratio
+### 5-1: Urban vs Rural Population Ratio
 
@@
 
-# 5-2: Median Income
+### 5-2: Median Income
 
@@
 
-# 5-3: Gender Ratio
+### 5-3: Gender Ratio
 
@@
 
-# 5-4: Mask Use Among Population
+### 5-4: Mask Use Among Population
 
@@
 
-# 5-5: GDP Percentage of Tourism
+### 5-5: GDP Percentage of Tourism
 

### 2020-12-04 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Finished regression table section and analysis

@@
 ```{r}
-
 model3 <- lm(case_rate_100k ~ senior_per_100k 
@@
              + log(population_density), data = df)
-coeftest(model3, vcov = vcovHC)
+coeftest(model3)
 ```
@@
 
-When observing the table above, three elements stand out.
+Initially addressing the progression of the models at a macro level, the regression table highlights the fact that as the models build upon one another, they seem to get better at explaining the relationship between the input and output variables.
+Between model 1, 2, and 3, both the R2 and stricter adjusted R2 value increases as variables are added.
+The R2 value reflects the amount of variation that can be explained in the model on a 0-1 scale, while the adjusted R2 value does the same but also takes into account the number of variables measured (as can be seen in the decreasing degrees of freedom, df).
+The adjusted R2 value jumps from 0.082 initially in model 1 to a value of 0.267 in model 2, which is more than a 200% increase.
+The value again increases, albeit by a smaller amount, between model 2 and a value of 0.376 in model 3, which represents a 40% jump.
+When the R2 value is not adjusted, this value increases to 0.439 for model 3.
+Although not extraordinarily high given the scale, the value shows that that model 3 still explains much of the variance and is a relatively good predictor of the data, especially when compared to model 1.
+An additional value that reflects this improvement is the fact that the F-statistic is significant for all three models.
+The F-statistic is used to test whether or not to use the full model (inclusive of all input variables) or a reduced one with fewer variables.
+Given that the null hypothesis for this test is that the reduced model should be over the full model, this can be safely rejected in favor of the full model.
 
-Firstly, there is a difference in the calculated standard error between the coefficient test run on the linear models and the regression table.
-Taking a look at the standard error for the poverty percentage coefficient, the value in the table returns a significant result with a standard error of 55.125, as opposed to the insignificant result of a standard error of 57.684.
-This follows for model 3, with the table showing a significant result with standard error of 50.872, as opposed to an insignificant result of a standard error of 59.583.
+When diving into each model more closely and investigating the individual variable coefficients themselves, there is further evidence for robustness in the results across the models.
 
-Secondly, following the first observation, when designating the poverty percentage variable as significant, the regression table suggests that both covariate variables (including the mask mandate variable) are indeed significant with the COVID case rate.
-For the poverty percentage effect, the coefficient highlights that there is a positive correlation between the poverty percentage and case rate (as poverty percentage increases, so does case rate).
-The practical significance of this shows that for every percentage increase of population below the federal poverty line, there is an increase of roughly 112 cases per 100k for model 2, and roughly 104 cases per 100k for model 3.
-As for the mask mandate variable, the coeeficient highlights a negative correlation between it and the case rate (if the mask mandate is implemented, the case rate decreases).
-The practical significance is such that if the mask mandate has been implemented, the case rate per 100k will decrease by roughly 1000 cases.
+In model 1, the coefficient of -0.174 for the senior rate per 100,000 is significant with a p-value of 0.0234 (per the regression results from the Model 1 section).
+When the covariate variables of unemployment rate and poverty rate are added in to model 2, the coefficient for senior rate adjusts to a greater magnitude of -0.231 while the p-value appreciably decreases to a value of 0.0015 (per the regression results from the Model 2 section).
+This explains that not only has the magnitude of effect of the senior rate on case rate increased, but that this increased effect is also more statistically significant.
+In addition, the coefficients of the covariate variables themselves are also highly significant with p-values of 0.0005 and 0.0059 for poverty rate and unemployment rate respectively.
+However, it should be noted that the magnitudes of these effects are very small, with poverty rate having a positive relationship of 0.002 per 100,000 with case rate while unemployment rate has a negative relationship of 0.005 per 100,000.
 
-Thirdly, we notice that the R2 value increases by a large value for every additional variable that is included within the model.
-Initially, we start with a very low R2 value, suggesting that almost no variance is explained in model 1 and that the model is a poor fit and predictor for the data.
-However, although still low, there is close to a 9 times increase in the amount of variance that is explained in model 2.
-Finally, in model 3, the R2 value increases by roughly 2.75 times the previous value, and a decent amount of the variance can be seen to be explained by the model.
-The adjusted R2 that accounts for the increase in number of variable terms in the model also correspondingly increases as the covariates are added.
-It also holds a negative value for model 1, signifying that there is no variance that can be explained by the model, further highlighting the poor fit.
+Moving onto model 3, the magnitude of the coefficient for senior rate increases once again to a value of -0.242 while the statistical significance increases with a decreased p-value of 0.0004 (per the regression results from the Model 3 section).
+As highlighted before, the more complex model does a better job of predicting the effect of the senior rate variable.
+Conversely, the p-values for the poverty rate and unemployment rate coefficients become less significant, as they increase to 0.002 and 0.049 respectively.
+However, there is almost no detectable change in the coefficient values, as poverty rate is still at 0.002 and unemployment rate increases to -0.004.
+The change in the p-values are likely a direct result of the added variables of mask policy and the logarithmic transform of population density, where these variables introduce some additional variance in the model 2 covariates.
+Notably, the coefficient for the mask policy covariate is significant with a p-value of 0.0027 and has by far the greatest effect on the outcome variable with a coefficient of -945.6.
+The populuation density covariate is not significant with a p-value of 0.3369.
 
-As a note, it should be stressed that the models and variables included in this regression table may change when the final report is created to provide a more descriptive analysis of the original research question.
+In terms of practical significance, there is great discrepancy between the effect of the mask policy variable compared to the other three significant features.
+With a coefficient of -945.6, this means that given the binary mask policy variable, when a mask policy is implemented (value of 1), there is an decrease in case rate of roughly 946 per 100,000 compared to when the policy is not in place.
+The next largest effect is from the senior rate variable, where in the final model, the coefficient of -0.242 translates as a decrease of 0.242 cases per 100,000 with every additional senior per 100,000.
+More intuitively, it can be said that there is an decrease of roughly 1 case per 100,000 with every 5 additional seniors per 100,000.
+This effect is roughly 190 times smaller than the effect of wearing a mask.
+Similarly, the effects of both poverty rate and unemployment rate are nominal.
+For every 500 additional people per 100,000 who are below the poverty line, there is 1 additional case per 100,000, and for every 250 additional people per 100,000 who are unemployed, there is 1 fewer case per 100,000.
+In comparison with likewise minimal effect of the senior rate, these effects are 100 and 50 times smaller respectively.
 

### 2020-12-04 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Cleaned up CLM section.

@@
 ```{r variables, warning=FALSE, results='hide'}
-df <- read.csv("covid-19.csv", header = TRUE)
+df <- read.csv("COVID-19.csv", header = TRUE)
 df<-df%>%
@@
 
+### Model 2 limits
+
+#### i. IID Sampling 
+( Discussed in the common section)
+
+#### ii.Linear Conditional Expectation
+Assessing non-linear in higher-dimensional space: look at the predicted vs. residuals of the model.
+
+```{r, echo=FALSE}
+df<- df%>% 
+  mutate(
+    model2_preds = predict(model2), 
+    model2_resids = resid(model2)
+  ) 
+df %>% 
+  ggplot(aes(model2_preds, model2_resids)) + 
+  geom_point() + 
+  stat_smooth()
+```
+
+####iii.No Perfect Collinearity
+First, check if any variables were dropped by R. No coefficient is missing, which mean there is no perfect collinearity.
+
+```{r}
+model2$coefficients
+```
+
+In addition, all the variance inflation factors are less than 4, which doesn't indicate the exisitence of collinearity.
+
+```{r}
+vif(model2)
+```
+
+####iv.Homoskedastic Errors
+10 is Florida, 33 is New York, 35 is North Dakota
+
+```{r, echo=FALSE}
+plot(model2, which=3)
+```
+
+```{r, echo=FALSE}
+bptest(model2)
+```
+
+####v.Normally Distributed Errors
+Check the normality of error distribution based on the histogram of residuals and the qqplot.
+
+```{r, echo=FALSE, warning=FALSE}
+plot_one <- df %>% 
+  ggplot(aes(x = model2_resids)) + 
+  stat_bin(bins = 25)
+  geom_histogram()
+
+plot_two <- df %>% 
+  ggplot(aes(sample = model2_resids)) + 
+  stat_qq() + stat_qq_line()
+
+plot_one / plot_two
+```
+
+#### GGPairs
+
+```{r, echo=FALSE}
+df %>% 
+  select(case_rate_100k, senior_per_100k, poverty_per_100k, unemployed_per_100k) %>% 
+  GGally::ggpairs()
+```
+
+#### Cook Distance 
+10 Florida, 20 Maine, 35 North Dakota
+
+```{r, echo =FALSE}
+plot(model2, which = 5)
+```
+
 ### Model 3
@@
 
-Firstly, the most general requirement for the linear model is that data points be independent and identically distributed (IID).
+The most general requirement for all the linear models is that data points be independent and identically distributed (IID) which they are built off of.
 While for the most part the events of one state will primarily affect that single state, travel by individuals spreading the virus across state lines is an inevitable occurence.
@@
 In addition, all states are involved in the same market economy system of the country to varying degrees, even though there are economic performance divisions across the states.
-Likewise, certain regions of the country may also have similar racial compositions due to a shared history across general geographic regions.
+Likewise, certain regions of the country may also have similar demographics due to a shared history across general geographic regions.
 While this is acknowledged to occur, the effect is likely minor compared to the policy data collected state by state which would affect the spread of the virus.
@@
 
-Secondly, the process being modeled must be described by a linear conditional expectation function of the variables which are included, in order for the linear regression to be valid.
-This can be assessed after the fact by visualizing the model over the range of each variable alongside the data, or also by visualizing the residuals over the range of each variable.
-For the purpose of this assumption and other model or data related assumptions, only the Model 3-a will be used as it is the most inclusive.
+In examining the results of various diagnostics across different models, generally the CLM assumptions are best met with the second model.
+In the first model, deviations away from certains assumptions are the strongest.
+This is seen mostly with regards to the linear conditonal expectation assumption, where the residuals against the fitted values have a noticeable inverted "V" shape as opposed to being constant around zero.
+In the second model, the same diagnostic shows that the assumption is better met.
 
-```{r}
-avg_pop <- log10(mean(df$population_density))
-avg_pov <- mean(df$poverty_pct)
-avg_mask <- mean(df$mask_public_bool)
-y_pop <- model3$coefficients[1] + model3$coefficients[2]*log10(df$population_density) + model3$coefficients[3]*avg_pov + model3$coefficients[4]*avg_mask
-y_pov <- model3$coefficients[1] + model3$coefficients[2]*avg_pop + model3$coefficients[3]*df$poverty_pct + model3$coefficients[4]*avg_mask
-y_pop_nomask <- model3$coefficients[1] + model3$coefficients[2]*avg_pop + model3$coefficients[3]*avg_pov
-y_pov_nomask <- model3$coefficients[1] + model3$coefficients[2]*avg_pop + model3$coefficients[3]*df$poverty_pct
-
-df %>% 
-  ggplot() + 
-  geom_point(aes(x = log10(population_density), y = case_rate_100k, 
-                 group = factor(mask_public_bool), color = factor(mask_public_bool))) + 
-  geom_line(aes(x = log10(population_density), y = y_pop, color = "red")) +
-  geom_line(aes(x = log10(population_density), y = y_pop_nomask, color = "blue")) +
-  scale_color_manual(labels = c("No", "Yes", "No", "Yes"), 
-                     values = c("blue", "red", "blue", "red")) +
-  labs(
-    title = 'Linearity of Variables, Population Density', 
-    x = 'Base 10 Log of Population Density', 
-    y = 'Case Rate', 
-    color = 'Mask Mandate'
-  )
-df %>% 
-  ggplot() + 
-  geom_point(aes(x = poverty_pct, y = case_rate_100k, group = factor(mask_public_bool), 
-                 color = factor(mask_public_bool))) + 
-  geom_line(aes(x = poverty_pct, y = y_pov, color = "red")) +
-  geom_line(aes(x = poverty_pct, y = y_pov_nomask, color = "blue")) +
-  scale_color_manual(labels = c("No", "Yes", "No", "Yes"), 
-                     values = c("blue", "red", "blue", "red")) +
-  labs(
-    title = 'Linearity of Variables, Poverty Percentage', 
-    x = 'Poverty Percentage', 
-    y = 'Case Rate', 
-    color = 'Mask Mandate'
-  )
-```
-
-Linearity for the three regressed variables is demonstrated above.
-As can be seen when mask policy and population density are varied, the data generally follows a linear trend.
-This is similar when the poverty percentage is varied, although at higher levels of poverty the presence of a mask mandate made the case rate unexpectedly low.
-It appears that possible a negative quadratice function may fit the data better, although it appears close enough to linear that this assumption may hold.
-```{r}
-df %>% {
-   ggplot(mapping = aes(x = log(df$population_density), y = model3$residuals)) +
-      geom_point() + stat_smooth(se = TRUE) +
-      ggtitle("Model Residuals versus Population Density") +
-      xlab("Logarithm of Population Density") +
-      ylab("Residual")
-}
-df %>% {
-   ggplot(mapping = aes(x = df$poverty_pct, y = model3$residuals)) +
-      geom_point() + stat_smooth(se = TRUE) +
-      ggtitle("Model Residuals versus Poverty Percentage") +
-      xlab("Poverty Percentage (%)") +
-      ylab("Residual")
-}
-df %>% {
-   ggplot(mapping = aes(x = log(df$mask_public_bool), y = model3$residuals)) +
-      geom_point() + stat_smooth(se = TRUE) +
-      ggtitle("Model Residuals versus Mask Mandate") +
-      xlab("Public Mask Mandate") +
-      ylab("Residual")
-}
-```
-
-Thirdly, related to the relationship of the residuals to the data, homoscedasticity can be assessed by visualizing the size of residuals across the range of each variable.
-Alternatively, it can also be done via the Breusch-Pagan test to more objectively determine whether homoscedasticity is not present in the data.
-Using the first method with the plots above, the residuals to each of the four regressions appear to be evenly spread throughout the range of the x axis.
-Therefore, this assumption can also be reasonably assumed to be met.
-The normal distribution of these errors also appears to be well met by the linear models, with most data points close to their respective line causing most residuals to be centered around zero.
-
-Perfect colinearity is automatically detected by the R regression algorithms, which drop potential variables if they have complete colinearity with another variable.
-Near perfect colinearity is more problematic to detect, but a common hint to its existence is large standard errors on colinear features.
-In Model 3-a, all of the standard errors for model coefficients are at least less than half of the value of the coefficient itself.
-Therefore, near-perfect colinearity can be assumed to not be taking place between these variables.
-```{r}
-correlation_one <- cor(log(df$population_density), df$poverty_pct)
-vif_one <- 1/(1-(correlation_one^2))
-correlation_two <- cor(log(df$population_density), df$mask_public_bool)
-vif_two <- 1/(1-(correlation_two^2))
-correlation_three <- cor(df$poverty_pct, df$mask_public_bool)
-vif_three <- 1/(1-(correlation_three^2))
-```
-A more methodical approach for evaluating if there is too high of colinearity between features is through the calculation of the Variance Inflation Factor (VIF).
-For the three variables in Model 3-A, the VIF values between the three combinations of the three variables are `r vif_one`, `r vif_two`, and `r vif_three`.
-Generally, as long as the VIF is below four there can be assumed not enough colinearity to violate the assumption as is the case in this model.
-
-Finally, the linear model is also assumed to generate residual error that is normally distributed away from the predicted values.
-A simple and straightforward way to assess this is through a histogram of the residuals of the model.
-```{r}
-hist(model3$residuals, main = "Model Residuals", xlab = "Residual")
-```
-As seen above, this assumption is generally well met.
-
-### Model 2 limits
-
-#### i. IID Sampling 
-( Discussed in the common section)
-
-#### ii.Linear Conditional Expectation
-Assessing non-linear in higher-dimensional space: look at the predicted vs. residuals of the model.
-
-```{r, echo=FALSE}
-df<- df%>% 
-  mutate(
-    model2_preds = predict(model2), 
-    model2_resids = resid(model2)
-  ) 
-df %>% 
-  ggplot(aes(model2_preds, model2_resids)) + 
-  geom_point() + 
-  stat_smooth()
-```
-
-####iii.No Perfect Collinearity
-First, check if any variables were dropped by R. No coefficient is missing, which mean there is no perfect collinearity.
-
-```{r}
-model2$coefficients
-```
-
-In addition, all the variance inflation factors are less than 4, which doesn't indicate the exisitence of collinearity.
-
-```{r}
-vif(model2)
-```
-
-####iv.Homoskedastic Errors
-10 is Florida, 33 is New York, 35 is North Dakota
-
-```{r, echo=FALSE}
-plot(model2, which=3)
-```
-
-```{r, echo=FALSE}
-bptest(model2)
-```
-
-####v.Normally Distributed Errors
-Check the normality of error distribution based on the histogram of residuals and the qqplot.
-
-```{r, echo=FALSE, warning=FALSE}
-plot_one <- df %>% 
-  ggplot(aes(x = model2_resids)) + 
-  stat_bin(bins = 25)
-  geom_histogram()
-
-plot_two <- df %>% 
-  ggplot(aes(sample = model2_resids)) + 
-  stat_qq() + stat_qq_line()
-
-plot_one / plot_two
-```
-
-#### GGPairs
-
-```{r, echo=FALSE}
-df %>% 
-  select(case_rate_100k, senior_per_100k, poverty_per_100k, unemployed_per_100k) %>% 
-  GGally::ggpairs()
-```
-
-#### Cook Distance 
-10 Florida, 20 Maine, 35 North Dakota
-
-```{r, echo =FALSE}
-plot(model2, which = 5)
-```
+However, when adding more variables in the third model, other assumptions begin to be less well met.
+For example, the normality of the error term in the model is less well fit than it was in the earlier models.
+It can be seen in the QQ plot for the model how at fitted values towards the edges of the model's range, the is the greatest deviation away from normality.
+This assumption is generally less important than that of the linear conditional expectation, however, where the latter is a fundamental part of CLM.
 
@@
 Higher median income may imply that more workers are able to function remotely with greater social distancing.
-A higher median income would then be negatively related to the covid case rate.
+A higher median income would then be negatively related to the COVID case rate.
 It would also be explanatory in a way that makes the poverty percentage variable less impactful, where either of these variables may be considered as proxies for one another.
@@
 
-The ratio of gender between women and men may impact the covid case rate in that men are more suscepitble to viral infections and generally have lower life expectancies from lifestyle differences.
+The ratio of gender between women and men may impact the COVID case rate in that men are more suscepitble to viral infections and generally have lower life expectancies from lifestyle differences.
 Greater susceptibility may relate to the case rate in that men may more often become infected when exposed to COVID under similar circumstances as a women.
 Their shorter life expentancy may also impact the case rate in that more COVID cases may result in greater or more serious symptoms, with less occurances of an infection that is asymptomatic or goes undetected.
-In either of these cases, a higher fraction of men among the population would lead to an overall higher covid case rate.
+In either of these cases, a higher fraction of men among the population would lead to an overall higher COVID case rate.
 
@@
 
-However, the magnitude of this change is likely to not be large due to other causal mechanisms between seniors and covid case rate.
+However, the magnitude of this change is likely to not be large due to other causal mechanisms between seniors and COVID case rate.
 While men may more frequently present COVID symptoms, older people in general are also more susceptible to the virus due to their physical condition and would also present more symptoms than non seniors.
-When comparing the two, generally seniors are considered as more at-risk of COVID complications than men across all ages, and so the senior percentage would still have a greater impact on the overall covid case rate than the gender ratio.
+When comparing the two, generally seniors are considered as more at-risk of COVID complications than men across all ages, and so the senior percentage would still have a greater impact on the overall COVID case rate than the gender ratio.
 
@@
 The economic standing of a state not only affects the individual households of its population, but also includes how it interacts with other states and how its peoples' lives might have been economically changed by the pandemic.
-One way that a state may increase its number of covid cases is both from tourists or others visiting during the pandemic and bringing disease, but also if its own population must work in close contact with others serving them in the tourism industry.
+One way that a state may increase its number of COVID cases is both from tourists or others visiting during the pandemic and bringing disease, but also if its own population must work in close contact with others serving them in the tourism industry.
 In addition, if certain key industries to a state, such as tourism, layed off many workers, they would likely have to take lower paying close contact jobs such as in the service or gig economy in order to continue having a livelihood.
-This would further lead to a higher covid case rate
+This would further lead to a higher COVID case rate
 

### 2020-12-04 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Update Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
update model 3 on population density discussion

@@
 
-Next, the linear model is created from model 2 with the mask policy variables added for the regression.
+Next, the regression model for model 3 is based on the existing work from model 2.
+Two additional variables will be introduced as maximalist approach.
+The first variable is the "mask_public_bool" (whether the state enforced public mask mandate or not) that represents the policy effects.
+The second variable introduced is "population_density" (the value is logged as discussed earlier) which is considered an important factor considering nature of spread in disease.
+
 The regression coefficient shows that Mask_public_bool variable is highly significant.
-The result suggests that public mask mandate policy has postive effect in bringing the case number down.
+And the result suggests that public mask mandate policy has positive effect in bringing the case number down.
+In detail, the COVID-19 case rate can be reduced by 945 out of every 100k people if the state enforces public mask mandate keeping all other variables the same.
+
+The population density variable however is not significant in the regression even it is considered as one of the major factors to the spread of disease.
+So according to the regression, we fail to reject the null hypothesis that population density have an direct impact to the COVID-19 case rate.
+There could be multiple factors causing the population density to be insignificant.
+For example, the population density is calculated by population divided by total square miles.
+As a result the numbers can be diluted for states that have major cities with high population density, but very little urbanization in the majority part of the state suburb area.
+In addition, as stated earlier in the research, Washington, D.C.
+is considered as a state however the demography structure and geographical related characteristics are different from other states.
+And due to a sample sample we have, individual variables that exhibit unique characteristics could lead to certain degree of bias.
+
+After careful discussion, the team still decided to keep "population_density" variable in model 3 even it is insignificant for the following reasons.
+First, adding population density variable has little effect on the change of coefficients of control variables in model 2.
+Second, we don't see evidence that the variable worsens the validation of OLS assumptions compared to model 2.
+Finally, the variable do offer meaningful information that is relevant to the main research question.
 
@@
 coeftest(model3)
+
 ```

### 2020-12-04 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Update Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
udpate model 3

@@
 
-### @Frank, this was my previous analysis of the POPULATION DENSITY variable, feel free to use any of this
-
 Next, the distribution of the population density variable is examined.
@@
 Next, the regression model for model 3 is based on the existing work from model 2.
-Two additional variables will be introduced as maximalist approach.
+Two additional variables discussed in Model 3 will be introduced as maximalist approach.
 The first variable is the "mask_public_bool" (whether the state enforced public mask mandate or not) that represents the policy effects.
@@
 is considered as a state however the demography structure and geographical related characteristics are different from other states.
-And due to a sample sample we have, individual variables that exhibit unique characteristics could lead to certain degree of bias.
+And due to a small sample we have, individual variables that exhibit unique characteristics could lead to certain degree of bias.
 
@@
 CLM assumptions check for Model 3:
+
+####Zero Conditional Mean
+
+The Residual vs Fitted chat shows a almost flat line that is close to zero, with the tails a bit skewed.
+There is no strong evidence that there are unincoorperated linear relationship that we are not capturing from the control variables.
 ```{r}
-#Zero Conditional Mean
+
 plot(model3, which=1)
+```
 
-#Normality of errors
+####Normality of errors
+From the Normal Q-Q plot below, the points generally stay close to the normal line, with the tails exhibit certain degree of skew. 
+```{r}
 plot(model3, which=2)
+```
+####Homoskedasticity
+```{r}
 
-#Homoskedasticity
 plot(model3, which=3)
-
+```
+```{r}
 #Cook distance
 plot(model3, which=5)
-
+```
+```{r}
 #Collinearity

### 2020-12-04 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Minor updates made during meeting

@@
 ```{r variables, warning=FALSE, results='hide'}
-df <- read.csv("COVID-19.csv", header = TRUE)
+df <- read.csv("covid.csv", header = TRUE)
 df<-df%>%
@@
 For the first model, the relationship between the COVID-19 case rate per 100,000 and the population percentage distribution of seniors was analyzed.
+
 First, the distribution of the case rate dependent variable is examined.
-
-```{r}
-summary(df$case_rate_100k)
-```
 
@@
     x = 'Number of seniors per 100,000', 
-    y = 'COVID case rate per 100,000'
+    y = 'COVID-19 case rate per 100,000'
   )
@@
 
-cat('Count for states did not enforced mask mandate in public: ', 
+cat('\nCount for states did not enforced mask mandate in public: ', 
     length(df$mask_public_bool[df$mask_public_bool==0]))
@@
 
-```{r}
+```{r, message = FALSE}
 
@@
 Given that this causes the data to be skewed, the logarithm is taken to scale the variable.
-Alternatively, the data point could have been droppped from the sample, but given the already small sample size, it was determined that a better approach would be to keep it within the sample.
+Alternatively, the data point could have been dropped from the sample, but given the already small sample size, it was determined that a better approach would be to keep it within the sample.
 Once this transformation was performed, it is shown that there is a relatively normal distribution of population densities (see figure below).
@@
 histogram_of_pdensity
-```
-
-With the appropriate variables transformed, a plot is created to show the relationship between them.
-
-```{r message=FALSE}
-df %>% 
-  ggplot(aes(l_population_density, case_rate_100k)) + 
-  geom_smooth(se = FALSE) +
-  geom_point() +   
-  labs(
-    title = 'COVID Case Rate due to Population Density', 
-    x = 'Log of population density per square miles', 
-    y = 'COVID case rate per 100,000'
-  )
 ```
@@
 First, adding population density variable has little effect on the change of coefficients of control variables in model 2.
-Second, we don't see evidence that the variable worsens the validation of OLS assumptions compared to model 2.
+Second, we don't see evidence that the variable worsens the validation of CLM assumptions compared to model 2.
 Finally, the variable do offer meaningful information that is relevant to the main research question.
@@
 
-####Zero Conditional Mean
+#### Zero Conditional Mean
 
@@
 
-####Normality of errors
+#### Normality of errors
 From the Normal Q-Q plot below, the points generally stay close to the normal line, with the tails exhibit certain degree of skew. 
@@
 ```
-####Homoskedasticity
+
+#### Homoskedasticity
 ```{r}
@@
 In terms of practical significance, there is great discrepancy between the effect of the mask policy variable compared to the other three significant features.
-With a coefficient of -945.6, this means that given the binary mask policy variable, when a mask policy is implemented (value of 1), there is an decrease in case rate of roughly 946 per 100,000 compared to when the policy is not in place.
+With a coefficient of -945.6, this means that given the binary mask policy variable, when a mask policy is implemented (value of 1), there is an decrease in case rate of roughly 945 per 100,000 compared to when the policy is not in place.
 The next largest effect is from the senior rate variable, where in the final model, the coefficient of -0.242 translates as a decrease of 0.242 cases per 100,000 with every additional senior per 100,000.
-More intuitively, it can be said that there is an decrease of roughly 1 case per 100,000 with every 5 additional seniors per 100,000.
+More intuitively, it can be said that there is an decrease of roughly 1 case per 100,000 with every 4 additional seniors per 100,000.
 This effect is roughly 190 times smaller than the effect of wearing a mask.

### 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Updated omitted variable bias section.

@@
 ```{r variables, warning=FALSE, results='hide'}
-df <- read.csv("covid.csv", header = TRUE)
+df <- read.csv("covid-19.csv", header = TRUE)
 df<-df%>%
@@
 Omitted variables from the model may affect both the internal specifications of included variables as well as the general applicability of the model to explain the output variable.
-For this analysis, some of these variables are within the dataset but not used while others are hypothetical.
+For this analysis, the omitted variables mentioned are hypothetical.
+Therefore, without data to include in an follow up analysis with the omitted variable, only the direction of the bias may be estimated rather than the size effect as well.
 
@@
 Therefore, the coefficient for population density in the original model would be larger and more positive than a model where this ratio is included, resulting in a bias that is positive and away from zero.
+Assuming population density as an effective proxy for an urban to rural ratio variable, however, the original effect described in the models can be assumed to be real.
 
@@
 Because inclusion of the median income variable would led to a lower magnitude poverty percentage variable, and that the coefficient of the poverty percentage variable is positive, this would cause a positive omitted variable bias away from zero.
+Similar to the first omitted variable, however, poverty percentage may also be assumed to proxy median income in that they both describe the economic state of a location and what that implies for its population.
+Because of this, the original effect in the model can be assumed to be real.
 
@@
 Therefore, introducing the gender ratio variable into the model would lower the value of the coefficient to the senior percentage variable to be less negative.
-This would be a negative omitted variable bias away towards zero.
+This would be a negative omitted variable bias towards zero.
+With a negative bias, and no existing variable in the dataset that would serve as a good proxy for the gender ratio, the effects in the original model can be assumed to be real.
 
@@
 This would cause the bias of the omitted variable to be negative and towards from zero.
+It is also reasonable to assume that state mask policy may effectively enough proxy for the mask use among the population, in which case the original effect in the model may be assumed to be real.
 
@@
 This would result in a positive omitted variable bias away from zero.
+Without a clear proxy in the dataset, and with the nature of this variable in touching on the IID assumption between interaction among states, it should be further investigated to understand how much of the original model may be an effect of omitted variable bias.
 

### 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Update Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
update model 3

@@
          non_elderly_pre_existing = "Nonelderly.Adults.Who.Have.A.Pre.Existing.Condition",
-         homeless = "Number.Homeless..2019.") %>%
+         homeless = "Number.Homeless..2019.",
+         medical_expense = "Medicaid.Expenditures.as.a.Percent.of.Total.State.Expenditures.by.Fund",
+         retail_recreation = 'Retail...recreation',
+         transit = 'Transit.stations') %>%
   select(State, case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
@@
          black_population_pct, white_population_pct, poverty_pct, unemployed_pct, senior_pct,
-         population_2018, non_elderly_pre_existing, homeless)
+         population_2018, non_elderly_pre_existing, homeless, medical_expense, retail_recreation, transit)
 # Assign correct data types
@@
 
-#### Model 3 Regression
+### Model 3 Regression
 
-Next, the regression model for model 3 is based on the existing work from model 2.
+Next, the regression model for model 3 is created based on the existing work from model 2.
 Two additional variables discussed in Model 3 will be introduced as maximalist approach.
 The first variable is the "mask_public_bool" (whether the state enforced public mask mandate or not) that represents the policy effects.
-The second variable introduced is "population_density" (the value is logged as discussed earlier) which is considered an important factor considering nature of spread in disease.
+The second variable introduced is "population_density" (the value is logged as discussed earlier) which is considered an important factor related to the spread of disease.
 
@@
 
-```{r}
+```{r, warning=FALSE}
+
 model3 <- lm(case_rate_100k ~ senior_per_100k 
@@
 
-CLM assumptions check for Model 3:
+### CLM Assumptions Validation for Model 3:
+
+#### I.I.D Assumption
+
+( Discussed in the common section)
 
@@
 
-#### Normality of errors
+#### Normality of Errors
+
 From the Normal Q-Q plot below, the points generally stay close to the normal line, with the tails exhibit certain degree of skew.
+There is no significant evidence of concerning normality in the errros.
 ```{r}
@@
 #### Homoskedasticity
+
+According to the plot of Homoskedasticity, the standardized residual falls into a reasonable range and the line looks smooth without concerning skews.
+There is no significant evidence that the model violates homoskedasticity assumptions.
 ```{r}
@@
 plot(model3, which=3)
+
 ```
+
+#### Collinearity
+
+Based on the Variance Inflation Factors check, all the variables have variance-inflation factors less than 4. As a result, there's no strong evidence of collinearity concern within the model. 
 ```{r}
-#Cook distance
+
+vif(model3)>4
+```
+
+#### Cook's Distance
+
+From the Cook's plot, there is no strong evidence of concerning outliers.
+
+```{r}
+
 plot(model3, which=5)
-```
-```{r}
-#Collinearity
-vif(model3)
 ```

### 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Updated transformation of percentage variables

@@
 It will also serve to ease the burden on the readers of this report as like-for-like scaled relationships can be drawn between the variables.
-Given that this transformation is constant for all values (scalar multiplication of 100,000), it will have no effect on the distributions of the variables.
+Given that this transformation is constant for all values (scalar multiplication of 100,000 for the seniors variable, and 1,000 for poverty and unemployment), it will have no effect on the distributions of the variables.
 
@@
     senior_per_100k = 100000*senior_pct,
-    poverty_per_100k = 100000*poverty_pct, 
-    unemployed_per_100k = 100000*unemployed_pct,
+    poverty_per_100k = 1000*poverty_pct, 
+    unemployed_per_100k = 1000*unemployed_pct,
   )

### 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Updated introduction section and EDA to align with new transformation of percentage variables
model 2 update

--- 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
@@
     senior_per_100k = 100000*senior_pct,
-    poverty_per_100k = 100000*poverty_pct, 
-    unemployed_per_100k = 100000*unemployed_pct,
+    poverty_per_100k = 100000*poverty_pct/100, 
+    unemployed_per_100k = 100000*unemployed_pct/100,
   )
@@
 ```{r, warning=FALSE, message=FALSE}
-summary(df$senior_per_100k)
 
@@
   labs(title = 'Histogram of Seniors Distribution', 
-  x = 'Number of seniors per 100,000 people', y = 'Count') + 
+  x = 'Number of seniors per 100,000', y = 'Count') + 
   scale_x_continuous(breaks=seq(10000,22000,1000))
@@
 
-For model 2, state demographic information regarding racial composition(white percentage of total population /Black Percentage of Total Population), economic status(unemployed percentage/poverty percentage), and population seniors percentage was investigated to explore its potential connection to the COVID pandemic status.
+Based on the work of model 1, two extra state demographic information regarding poverty rate (percentage living under the federal poverty line in 2018) and unemployment rate (percentage of unemployed in 2018) would be included in the analysis of model 2.
+In addition, both newly-introduced percentage variables would be transformed into rates per 100,000 to align with the case rate variable.
 
-#### 2-a) White Percentage of Total Population/ Black Percentage of Total Population
-
-Variable of black percentage is stored as string variables.
-Moreover, three entries of black percentage of total population is "<0.01", which is to be dropped or replaced by values determined by extra resources for further analysis.
-
-Here, the value of black percentage is replaced by 0 when it is "<0.01".
-Afterwards, the correlation of white percentage and black percentage is -0.42, which shows these two variables are correlated.
-Considering that the value of other race groups is much smaller, it is proper to only include white percentage in the regression model to explore the relation of care rate and racial composition.
-
-The distribution of the white % is skewed, not an ideal normal distribution.
-While the distribution of the square of white % is more close to a normal distribution.
-In practice, this square could reflect the chance that the interaction happens between two white people.
-
-```{r warning=FALSE}
-black_num <- as.numeric(df$black_population_pct)
-black_num[is.na(black_num)]<- 0
-cor(df$white_population_pct, black_num)
-```
-
-```{r}
-summary(df$white_population_pct)
-```
+First, the distribution of the poverty rate is examined as below.
 
@@
 ggplot(data = df,
-  mapping = aes(x= white_population_pct))+
-  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,1,0.05)) +
-  labs(title =  "Histogram of White Population Percentage", 
-       x = "White percentage of total population", y = 'Count')+
-  scale_x_continuous(breaks=seq(0, 1, 0.1))
+  mapping = aes(x= poverty_per_100k))+
+  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(6000,21200,800)) +
+  labs(title =  "Histogram of Poverty Rate", 
+       x = "Number of living under poverty line per 100,000", y = 'Count')+
+  scale_x_continuous(breaks=seq(6000, 21200, 1600))
 ```
-```{r}
-ggplot(data = df,
-  mapping = aes(x= (white_population_pct)^2))+
-  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,1,0.05)) +
-  labs(title =  "Histogram of Square of White Percentage", 
-       x = "White percentage of total population^2", y = 'Count')+
-  scale_x_continuous(breaks=seq(0, 1, 0.1))
-```
+As shown in the above distribution, there is a fairly normal distribution in the poverty number 
+per 100,000 across the states. Additionally, there is no outliers observed within the distribution,
+given that the range of rates fall between 7,600 and 19,700. Therefore, there is no further transformation in need for this variable in the following analysis process for both model 2 and model 3.
 
-#### 2-b) Percentage living under the federal poverty line (2018)
-
-Correlation of pecentage living under poverty line and white percentage of total population shows that they are not significantly related.
-The distribution of the poverty percentage is not heavily skewed.
-
-```{r}
-summary(df$poverty_pct)
-cor(df$poverty_pct, df$white_population_pct)
-```
+Next, the distribution of the unemployment rate is examined as below.
 
@@
 ggplot(data = df,
-  mapping = aes(x= poverty_pct))+
-  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(6,21,1)) +
-  labs(title =  "Histogram of Poverty Percentage", 
-       x = "Percentage living under poverty line (2018)", y = 'Count')+
-  scale_x_continuous(breaks=seq(6, 21, 1))
+  mapping = aes(x= unemployed_per_100k))+
+  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(2000,8000,500)) +
+  labs(title =  "Histogram of Unemployment Rate", 
+       x = "Number of unemployed per 100,000", y = 'Count')+
+  scale_x_continuous(breaks=seq(2000, 8000, 1000))
 ```
+According to the histogram above, the distribution of unemployed number per 100,000 is of fairly normal distribution.
+Overall, it is not highly skewed or heavily tailed.
+Therefore, there is no further transformation in need for this variable in the following analysis process for both model 2 and model 3.
 
-#### 2-c) Unemployed Percentage (2018)
+#### Model 2 Regression
 
-Although, the distribution is of high peak in the middle of the range.
-Overall, it is not highly skewed or heavily tailed.
-In addition, the unemployed rate is correlated to the variable to the poverty rate, which is in an agreement with intuitive expectation.
-
-```{r}
-summary(df$unemployed_pct)
-cor(df$unemployed_pct, df$poverty_pct)
-```
-```{r}
-ggplot(data = df,
-  mapping = aes(x= unemployed_pct))+
-  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,10,1)) +
-  labs(title =  "Histogram of Unemployed Percentage", 
-       x = "Unemployed percentage (2018)", y = 'Count')+
-  scale_x_continuous(breaks=seq(0, 10, 1))
-```
-
-#### 2-d) Seniors Percentage (Age 65+)
-
-The seniors percentage (age 65+) is of a nearly normal distribution.
-
-```{r}
-summary(df$senior_pct)
-```
-
-```{r}
-ggplot(data = df,
-  mapping = aes(x= senior_pct))+
-  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0.10,0.22,0.01)) +
-  labs(title =  "Histogram of Seniors Percentage (65+)", x = "Seniors percentage", 
-       y = 'Count')+
-  scale_x_continuous(breaks=seq(0.1, 0.25, 0.05))
-```
-
-### Model 2 Plots
-
-No obvious trend could be easily observed or concluded from the plots regarding the dependent variables and candidate covariates.
+Compared to model 1, the relationship investigated here is about variables space with higher dimensions.
+It would be challenging to observe notable variation from variables distribution plots as below.
 
 ```{r message=FALSE}
-df %>% 
-  ggplot(aes(white_population_pct, case_rate_100k, color = log10(population_density))) + 
+case_poverty <- df %>% 
+  ggplot(aes(poverty_per_100k, case_rate_100k, color = senior_per_100k)) + 
   geom_point() + 
@@
   labs(
-    title = 'Relation of Case Rate Per 100000 to White Population Percentage', 
-    x = 'White percentage of total population', 
-    y = 'Case rate per 100000',
-    color = 'Log 10 of population density'
+    title = ' Association between Covid-19 ase rate and the poverty rate', 
+    x = 'Number of living under poverty line per 100,000',
+    y = 'Covid-19 care rate per 100,000', 
+    color = 'Senior num per 100k'
   )
-```
 
-```{r message=FALSE}
-df %>% 
-  ggplot(aes(x = white_population_pct, y = case_rate_100k, color = poverty_pct)) + 
+case_unemployed <- df %>% 
+  ggplot(aes(unemployed_per_100k, case_rate_100k, color = senior_per_100k)) + 
   geom_point() + 
@@
   labs(
-    title = 'Relation of Case Rate per 100000 to White Population Percentage', 
-    x = 'White percentage of total population', 
-    y = 'Case rate per 100000',
-    color = 'Poverty percentage'
+    title = 'Association between Covid-19 ase rate and the unemployment rate', 
+    x = 'Number of unemployed per 100,000', 
+    y = 'Covid-19 care rate per 100,000',
+    color = 'Senior num per 100k'
   )
+
+case_poverty / case_unemployed
 ```
+However, there is an overall increasing trend of COVID-19 care rate with increasing poverty rate, while there is a inverse relationship between the variables of the COVID-19 case rate and the unemployment rate.
 
-```{r message=FALSE}
-df %>% 
-  ggplot(aes(x = poverty_pct , y = case_rate_100k, color = log10(population_density))) + 
-  geom_point() + 
-  geom_smooth(se=FALSE)+
-  labs(
-    title = 'Relation of Case Rate per 100000 to Poverty Percentage', 
-    x = 'Poverty Percentage', 
-    y = 'Case rate per 100000',
-    color = 'Log 10 of population density'
-  )
-```
-
-```{r message=FALSE}
-df %>% 
-  ggplot(aes(unemployed_pct, case_rate_100k, color = log10(population_density))) + 
-  geom_point() + 
-  geom_smooth(se=FALSE)+
-  labs(
-    title = 'Relation of Case Rate per 100000 to Unemployed Percentage', 
-    x = 'Unemployed percentage (2018)', 
-    y = 'Case rate per 100000',
-    color = 'log 10 of population density'
-  )
-```
-
-```{r message=FALSE}
-df %>% 
-  ggplot(aes(senior_pct, case_rate_100k, color = log10(population_density))) + 
-  geom_point() +   
-  geom_smooth(se=FALSE)+
-  labs(
-    title = 'Relation of Case Rate per 100000 to Seniors Percentage (65+)', 
-    x = 'Seniors percentage', 
-    y = 'Case rate per 100000',
-    color = 'log 10 of population density'
-  )
-```
-
-### Model 2 Regression
-The regression models regarding the demographic variables discussed previously are listed as below.
+Based on above data exploratory result, the following equation is used to create a regression model
+to determine the relationship between the case rate and demographic features including the seniors rate, the rate of poverty and the rate of unemployment.
 
@@
 model2 <- lm(case_rate_100k ~ senior_per_100k + poverty_per_100k + unemployed_per_100k, data = df)
+```
+
+Accordingt to the assessment of homoskedasticity of model 2, which would be discussed in the following paragraph (Model 2 limits - iv.Homoskedastic Errors), it is plausible to apply the classical standard errors in the t-test of coefficients here.
+
+```{r}
 coeftest(model2)
@@
 
-#### (need to update)
-From the results of the regression, none of the coefficients of investigated variables in model 2 are significant (p-value < 0.05).
-Among these regression coefficient tests,  the p-value regarding the poverty percentage variable is the lowest (0.05755).
-In addition, its positive estimated coefficient is consistent with the statement that low-income communities are at high pandemic risk claimed by mainstream media these days.
+Based on the test results of the estimated coefficients in model 2, all three demographic features discussed here are significantly realated to the COVID-19 care rate, especially for the poverty rate (p-value < 0.01).
 
-### Model 2 limits
+The coefficient of the senior_per_100k is  -0.230970 and statistically significant, which implies that for every one thousand additional seniors per 100,000, there is a decrease of 231 COVID-19 cases per 100,000 holding other demographic features as constant.
+As discussed in mode1 1 section, although seniors are susceptible group, less mobility and social interaction with cautious self procetion may play a important role in this pandemic situation.
 
-#### i. IID Sampling 
-( Discussed in the common section)
+In addition, the coefficient of the poverty_per_100k (0.231750) indicats that for every one thousand additional people living under poverty line per 100,000, there is a increase of 232 COVID-19 case per 100,000 holding other demographic features as constant.
+People living under the poverty line have limited accesss to medical resourses.
+Moreover, most of them could only afford to live in the communites with poor hygiene.
+condition.
 
-#### ii.Linear Conditional Expectation
-Assessing non-linear in higher-dimensional space: look at the predicted vs. residuals of the model.
+Finally, the coefficient of the unemployed_per_100k is statiscally significant and negative(-0.486191), which implies that for every one thousand additional unemployed, there is a decrease of 486 COVID-19 case per 100,000 holding other demographic features as constant.
+Although the unemployment leads to less medical insurace coverage and impact on the incoming, the unempolyed people may have more flexibility in obeying stay-at-home.
+In addition, there is less expourse probability when no commuting or in-place working is required for the unemployed.
 
-```{r, echo=FALSE}
+#### Model 2 limits
+
+##### i. IID Sampling 
+(Would be discussed in the following section - 3. Limitations of the Models)
+
+##### ii.Linear Conditional Expectation
+To assess the linear conditional expectation assumption in higher-dimensional space, the plots of predicted versus residuals of the model 2 is investigated as follow.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 df<- df%>% 
@@
 
-####iii.No Perfect Collinearity
-First, check if any variables were dropped by R. No coefficient is missing, which mean there is no perfect collinearity.
+Overall, the residuals maintain around zero across the predicted range. There is no assumption violation phenomenon could be observed here.
+
+#####iii.No Perfect Collinearity
+
+Since there are multiple covariates included in this model compared to the model 1, there is more concern about the no perfect collinearity assumption here
+First, the list of variables coefficients shows that no variables were dropped, which mean there is no perfect collinearity detected by the R function.
 
@@
 
-In addition, all the variance inflation factors are less than 4, which doesn't indicate the exisitence of collinearity.
+In addition, all the variance inflation factors are less than 4 as follow, which doesn't indicate the exisitence of collinearity.
 
@@
 
-####iv.Homoskedastic Errors
-10 is Florida, 33 is New York, 35 is North Dakota
+#####iv.Homoskedastic Errors
+To evaluate the homoskedastic errors assumption, the square root of the residuals is plotted against the fitted values of the model 2 (scale-location plot).
+Although, the curve is not perfect flat smooth, there is no obvious variance of the errors either.
+For reference, the data points marked in the plot represent Florida(10), New York (33) and North Dakota (35).
 
@@
 
+Additionally, a Breusch-Pagan test is run to check the level of heteroskedasticity.
+The test result shows it fails to reject the null hypotheis, which means there is no evidence for heteroskedasticity.
+Overall, the assumption of homoskedastic errors is satisfied for model 2.
+
 ```{r, echo=FALSE}
@@
 
-####v.Normally Distributed Errors
-Check the normality of error distribution based on the histogram of residuals and the qqplot.
+#####v.Normally Distributed Errors
+To access the assumption of normality of error distribution, both the histogram and the qqplot of the residuals is shown as follow.
 
-```{r, echo=FALSE, warning=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=False}
 plot_one <- df %>% 
   ggplot(aes(x = model2_resids)) +
-  stat_bin(bins = 25)
-  geom_histogram()
+  geom_histogram(fill = 'skyblue', color = 'grey30', bins=30) +
+  labs(title =  "Histogram of the Model 2 Residuals", 
+       x = "Model 2 residuals", y = 'Count')
 
@@
 ```
+Limited by the sample size, it is hard to observe a perfect normal distribution from the histogram.
+However, the qqplot implies no deviation from normality.
+Overall, the normally distributed errors assumption is satisfied for model 2.
 
-#### GGPairs
+##### Cook Distance
 
-```{r, echo=FALSE}
-df %>% 
-  select(case_rate_100k, senior_per_100k, poverty_per_100k, unemployed_per_100k) %>% 
-  GGally::ggpairs()
-```
-
-#### Cook Distance 
-10 Florida, 20 Maine, 35 North Dakota
+Additionally, cook distance is used to estimate the influence of each single data point.
 
@@
 ```
+No obvious outlier could be detected here. For reference, the marked data points are Florida(10), Maine(20), and North Dakota(30).
 

### 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd

@@
 Those living below the poverty line may have less access/are unable to afford to preventative controls such as sanitation products and masks that help prevent the spread of the virus, while those who are designated as unemployed may not have the flexibility of sheltering at home as well as access to the aforementioned preventative controls.
+Alternatively, the likelihood of reduced mobility for those living under the poverty line and/or who are unemployed may also play a role in the relationship with case rate.
 As mentioned, these variables operationalize population metrics that may lead to greater case rates, and are already standardized as rates to account for varying absolute populations across states.

### 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Updated regression table analysis and conclusion

--- 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
@@
 With this information, one could suggest whether or not the proliferation of COVID-19 within a state seems to be related to either or both controllable and uncontrollable variables, and to the level at which the effect of these variables differ help.
-The modeling goal of this research question will be one of description, and will be broken up into three phases of investigation
+The modeling goal of this research question will be one of description, and will be broken up into three phases of investigation.
 
@@
 
-Following this, in the second phase of investigation the analysis considers other variables of state-level population that are considered factors of contractability, namely the rate of poverty (defined as the percent of individuals living under the federal poverty line in 2018) and the rate of unemployment (defined as the percent unemployment in 2018).
+Following this, in the second phase of investigation the analysis considers other variables of state-level population that are considered factors of suscesptibility, namely the rate of poverty (defined as the percent of individuals living under the federal poverty line in 2018) and the rate of unemployment (defined as the percent unemployment in 2018).
 Those living below the poverty line may have less access/are unable to afford to preventative controls such as sanitation products and masks that help prevent the spread of the virus, while those who are designated as unemployed may not have the flexibility of sheltering at home as well as access to the aforementioned preventative controls.
+Alternatively, the likelihood of reduced mobility for those living under the poverty line and/or who are unemployed may also play a role in the relationship with case rate.
 As mentioned, these variables operationalize population metrics that may lead to greater case rates, and are already standardized as rates to account for varying absolute populations across states.
@@
 In addition, the coefficients of the covariate variables themselves are also highly significant with p-values of 0.0005 and 0.0059 for poverty rate and unemployment rate respectively.
-However, it should be noted that the magnitudes of these effects are very small, with poverty rate having a positive relationship of 0.002 per 100,000 with case rate while unemployment rate has a negative relationship of 0.005 per 100,000.
+The magnitude of the effect of poverty rate is almost equivalent to that of the senior rate, with a single increase in poverty rate per 100,000 equivalent to 0.232 per 100,000 for case rate, while the unemployment rate has stronger negative coefficient of -0.486 per 100,000 (roughly twice the magnitude of the other two variables).
 
@@
 Conversely, the p-values for the poverty rate and unemployment rate coefficients become less significant, as they increase to 0.002 and 0.049 respectively.
-However, there is almost no detectable change in the coefficient values, as poverty rate is still at 0.002 and unemployment rate increases to -0.004.
-The change in the p-values are likely a direct result of the added variables of mask policy and the logarithmic transform of population density, where these variables introduce some additional variance in the model 2 covariates.
+There are also detectable changes in the coefficient values, as the magnitude for both decreases to 0.197 and -0.364 respectively.
+The changes in the p-values are likely a direct result of the added variables of mask policy and the logarithmic transform of population density, where these variables introduce some additional variance in the model 2 covariates.
 Notably, the coefficient for the mask policy covariate is significant with a p-value of 0.0027 and has by far the greatest effect on the outcome variable with a coefficient of -945.6.
@@
 With a coefficient of -945.6, this means that given the binary mask policy variable, when a mask policy is implemented (value of 1), there is an decrease in case rate of roughly 945 per 100,000 compared to when the policy is not in place.
-The next largest effect is from the senior rate variable, where in the final model, the coefficient of -0.242 translates as a decrease of 0.242 cases per 100,000 with every additional senior per 100,000.
-More intuitively, it can be said that there is an decrease of roughly 1 case per 100,000 with every 4 additional seniors per 100,000.
-This effect is roughly 190 times smaller than the effect of wearing a mask.
-Similarly, the effects of both poverty rate and unemployment rate are nominal.
-For every 500 additional people per 100,000 who are below the poverty line, there is 1 additional case per 100,000, and for every 250 additional people per 100,000 who are unemployed, there is 1 fewer case per 100,000.
-In comparison with likewise minimal effect of the senior rate, these effects are 100 and 50 times smaller respectively.
+The next largest effect is from the unemployment rate variable, where in the final model, the coefficient of -0.364 translates as a decrease of 0.364 cases per 100,000 for every additional increase of 1 in unemployment rate per 100,00.
+More intuitively, it can be said that there is a decrease of roughly 1 case per 100,000 with every 3 additional people unemployed per 100,000.
+This effect is roughly 2,600 times smaller than the effect of implementing a mask policy.
+Similarly, the effects of both poverty rate and senior rate are nominal.
+For every 5 additional people per 100,000 who are below the poverty line, there is roughly 1 additional case per 100,000, and for every 4 additional seniors per 100,000 within the state population, there is roughly 1 fewer case per 100,000.
+In comparison with likewise relatively minimal effect of the unemployment rate, these absolute effects on case rate are approximately one-third (1/3) and one-half (1/2) smaller respectively.
 
@@
 
-In conclusion, although there is no evidence to show that although there is a direct correlation between the COVID case rate and population density as originally hypothesized in our research question, it can be seen from further development of the models that there is some correlation in other variables examined, namely the mask mandate variable.
-As can be seen from model 3, the regression model gives a mask mandate variable coefficient with p-value of 0.003058, which is highly significant.
-The negative coefficient also suggests that the original conjecture that the implementation of mask policies within a state would decrease the COVID case rate was indeed accurate.
+The goal of the analysis was to address the research question to examine the relationship between both the distribution of population demographics and mask-related policy of a state and the COVID-19 case rate within that states.
+In order to address this question, the analysis was broken up into three phases of investigation with corresponding linear models:
+- Phase 1: To investigate the direct relationship between COVID-19 case rate and the population distribution of seniors within a state
+- Phase 2: To build upon Phase 1 by including two additional variables of population demographics, poverty rate and unemployment rate, and investigating the overall effect on COVID-19 case rate
+- Phase 3: To build upon Phase 2 by adding a variable to measure the effect of the implementation of a mask-policy on COVID-19 case rate.
+A variable for population density was also added as an additional measure of population demographics
 
-Conversely, the hypotheses that population and demographics would have an effect on the COVID case rate was unable to be accepted.
-Although upon investigating the poverty percentage variable in model 2 there was a detection of notable adjacency to a significant result (and an actual significant result when using the `stargazer` package to compute the regression table) with a p-value of 0.05755, it was not enough to be truly significant and reject the null hypothesis.
-Similarly, the relationship of race/ethnicity as well as age on case rate was also insignificant with no apparent correlation.
+From the three phases of investigation, there was strong evidence in the data to suggest that senior rate, unemployment rate, and mask policy are all statistically significant in relation to case rate and have a negative relationship with case rate, while poverty rate has a positive relationship.
+The data for population density was not statistically significant enough to suggest a relationship.
+Out of the four significant variables, the greatest effect was from the mask policy variable, followed by the unemployment rate, and then similar absolute effect sizes from the senior rate and povery rate.
 
-Given these results in conjunction with the original aim of the research question, it can be suggested that variables that can be controlled (in this case, mask policy) has a greater effect on stymieing the spread of COVID-19 than the variables that may be inherent to a state, such as population density and age/ethnicity demographics.
-This is a promising insight, as it provides opportunity for states to implement and apply policies to intervene and control the case rate without being worried about its varying effects depending on population characteristics.
-However, it should once again be stressed that the final report may include and analyze alternative variables and apply them to different models based on the findings from this draft report, which as a result may also lead to more concrete or slightly varied insights.
+Starting with the variable with the greatest effect size, it can be seen that the implementation of a mask policy in a state leads to a decrease of 945 cases per 100,000.
+Given that research from leading health organizations shows that the wearing of masks helps to decrease the chance of the transmission of COVID-19, it could be suggested that a mask policy is effective in increasing the adoption of wearing masks among the state population.
+Next, it can be seen that there is a decrease of roughly 1 case per 100,000 with every 3 additional people unemployed per 100,000.
+This suggests that as unemployment increases, the case rate decreases, which could be a result of the fact that those who are unemployed may be less mobile as they are not commuting into work, lowering the likelihood of contracting the virus.
+Additionally, at the time that this data was collected, the Coronavirus Aid, Relief, and Economic Security (CARES) Act was still supporting the unemployed with relief payments so there was less urgency to be out looking for work.
+Following unemployment, when examining the poverty variable there is actually an increase of roughly 1 case per 100,000 for every 5 additional people per 100,000 living under the federal poverty line.
+Although there is some intuition behind the unemployment and poverty rates having similar effects on case rate, it can be seen from the test of collinearity in the Model 2 section that these two variables are not explaining the same effects.
+This difference is likely due to the fact that those living below the poverty line have less access to suitable living conditions and sanitary goods that help to alleviate the possibility of COVID-19 infection, while those who were recently employed up until the pandemic hit may not be in the same economical situation.
+Finally, the effect of the senior rate is negative, with an additional 4 seniors per 100,000 translating to roughly 1 fewer case per 100,000.
+As touched upon in the introduction, this could be largely down to the fact that it is widely broadcasted that seniors fall within the high risk category in terms of having severe complications from COVID-19, and therefore they are more diligent in the precautions they take against contracting the virus.
+
+In conclusion, we see that although the population demographics had varying effects on the COVID-19 case rate, the greatest effect was seen from the mask policy.
+Going back to our original question, it can be suggested within the data that the effect of a mask-related policy is much larger than that of population demographics.
+Therefore, it seems that the COVID-19 case rate at the state-level is much more dependent on masks, and that ultimately, there are effective policy and control measures that can be taken by state governments that have a greater contribution towards combating it's spread relative to factors that cannot be controlled, such as population demographics.
+However, it should be noted that these insights are qualified by the assumptions and limitations laid out throughout the report, namely that there are potential issues in meeting the IID assumption and that there is explanatory information that is lost from both the variables investigated as well as the those that were not included in the dataset.

### 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd

--- 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
@@
 
-Following this, in the second phase of investigation the analysis considers other variables of state-level population that are considered factors of contractability, namely the rate of poverty (defined as the percent of individuals living under the federal poverty line in 2018) and the rate of unemployment (defined as the percent unemployment in 2018).
+Following this, in the second phase of investigation the analysis considers other variables of state-level population that are considered factors of suscesptibility, namely the rate of poverty (defined as the percent of individuals living under the federal poverty line in 2018) and the rate of unemployment (defined as the percent unemployment in 2018).
 Those living below the poverty line may have less access/are unable to afford to preventative controls such as sanitation products and masks that help prevent the spread of the virus, while those who are designated as unemployed may not have the flexibility of sheltering at home as well as access to the aforementioned preventative controls.

### 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Updated Model 1 section to align with structure of model 2 and 3

@@
 
-Following this, in the second phase of investigation the analysis considers other variables of state-level population that are considered factors of suscesptibility, namely the rate of poverty (defined as the percent of individuals living under the federal poverty line in 2018) and the rate of unemployment (defined as the percent unemployment in 2018).
+Following this, in the second phase of investigation the analysis considers other variables of state-level population that are considered factors of susceptibility, namely the rate of poverty (defined as the percent of individuals living under the federal poverty line in 2018) and the rate of unemployment (defined as the percent unemployment in 2018).
 Those living below the poverty line may have less access/are unable to afford to preventative controls such as sanitation products and masks that help prevent the spread of the virus, while those who are designated as unemployed may not have the flexibility of sheltering at home as well as access to the aforementioned preventative controls.
@@
 However, it can be posited that there is a inverse relationship between the variables given the roughly linear relationship in the overall downward trend in the line.
-In order to test this, the following equation is used to create a regression model to determine the true relationship between the case rate and population density variables.
+
+#### Model 1 Regression
+
+In order to test the relationship between the variables, the following equation is used to create a regression model to determine the true relationship between the case rate and population density variables.
 
@@
 The assumption of homoskedasticity is met if the data does not have large variance among the residuals.
+The analysis for this assumption, as well as the other CLM assumptions, is conducted in the following subsection titled "Model 1 CLM Assumptions".
 
-To evaluate this assumption, the square root of the residuals (to remove any negative values) is plotted against the fitted values of the model.
+Below, the t-test of coefficients is run on the regression model above.
+
+```{r}
+coeftest(model1)
+```
+
+As can be seen from the above output, the variable for senior percentage is shown to be significant within the model specification with a p-value of 0.023.
+Furthermore, The coefficient of -0.174 for the variable is negative, supporting the earlier evaluation that there is an inverse relationship between the senior percentage and the case rate.
+In practical significance, this can be interpreted as for every additional senior per 100,000 in the population, there is a decrease of 0.174 COVID-19 cases per 100,000.
+More intuitively, for roughly every 6 additional seniors per 100,000, there is a decrease of 1 case per 100,000.
+
+#### Model 1 CLM Assumptions
+
+##### i. IID Sampling
+
+The first assumption of IID is not detailed within this section given that it is evaluated more generally for all three models within section 3.
+
+##### ii. Linear Conditional Expectation
+
+The second CLM assumption to be evaluated is that of a linear conditional expectation relationship within the data, so that the model is accurately estimating the relationship between the variables.
+To test for this, the residuals of model 1 are plotted against its fitted values.
+
+```{r}
+# Check for linearty in model 1
+plot(model1, which=1)
+```
+
+When examining the plot above for linearity in the conditional expectation, there is a noticeable slightly quadratic shape to the relationship of the conditional expectation.
+This may suggest that the relationship between the variables may not be as linear as the model assumes.
+However, as discussed previously in the scatter plot of the two variables, the relationship can be approximated as roughly linear.
+It is also difficult to explain the complexity behind the COVID-19 case rate with just a single variable that does not fully capture the broader population demographics.
+Additionally, there is hope that given the addition of other input variables within models 2 and 3, the relationship within the data will be better explained and result in more linearity.
+
+##### iii. No Perfect Collinearity
+
+There is no test conducted for the third CLM assumption of no perfect collinearity in model 1 given there is only one input variable.
+
+##### iv. Homoskedastic Errors
+
+To evaluate the fourth CLM assumption of homoskedastic errors, the square root of the residuals (to remove any negative values) is plotted against the fitted values of the model.
 Additionally, a Breusch-Pagan test is run to check the level of heteroskedasticity.
@@
 
-```{r}
-coeftest(model1)
-```
+##### v. Normally Distributed Errors
 
-As can be seen from the above output, the variable for senior percentage is shown to be significant within the model specification with a p-value of 0.023.
-Furthermore, The coefficient of -0.174 for the variable is negative, supporting the earlier evaluation that there is an inverse relationship between the senior percentage and the case rate.
-In practical significance, this can be interpreted as for every additional senior per 100,000 in the population, there is a decrease of 0.174 COVID-19 cases per 100,000.
-More intuitively, for roughly every 6 additional seniors per 100,000, there is a decrease of 1 case per 100,000.
-
-Next, the other CLM assumptions are revisited to determine the whether the coefficients are unbiased and consistent.
-However, the assumption for IID is not detailed here given that it is evaluated more generally for all three models within section 3.
-
-The second CLM assumption to be evaluated here is that of a linear conditional expectation relationship within the data, so that the model is accurately estimating the relationship between the variables.
-To test for this, the residuals of model 1 are plotted against its fitted values.
-
-```{r}
-# Check for linearty in model 1
-plot(model1, which=1)
-```
-
-When examining the plot above for linearity in the conditional expectation, there is a noticeable slightly quadratic shape to the relationship of the conditional expectation.
-This may suggest that the relationship between the variables may not be as linear as the model assumes.
-However, as discussed previously in the scatter plot of the two variables, the relationship can be approximated as roughly linear.
-It is also difficult to explain the complexity behind the COVID-19 case rate with just a single variable that does not fully capture the broader population demographics.
-Additionally, there is hope that given the addition of other input variables within models 2 and 3, the relationship within the data will be better explained and result in more linearity.
-
-The third CLM assumption examined for model 1 is the assumption of normality of errors.
+The final CLM assumption examined for model 1 is the assumption of normality of errors.
 This is a necessary assumption to ascertain that the errors used in the regression are drawn from a normal distribution, so that they can be accepted when used to calculate the significance levels.
@@
 
-Finally the Cook's distance is examined by a plot of the residuals vs the model 1 leverage to look for any outliers in the values. From the plot below, no obvious outliers are detected.
+#### Influence of Data Points in Model 1
+
+In addition to the CLM assumptions, the Cook's distance is also examined by a plot of the residuals vs the model 1 leverage to estimate the influence of the data points and identify any outliers in the values.
+From the plot below, no obvious outliers are detected.
 
@@
 ```
-
-It is also noted that there is no test conducted for collinearity in model 1 given there is only one input variable.
 

### 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Latest

@@
   rename(case_rate_100k = 'Case.Rate.per.100000',
-         death_rate_100k = 'Death.Rate.per.100000',
          population_density = 'Population.density.per.square.miles',
-         white_pct = 'White...of.Cases',
-         black_pct = 'Black...of.Cases',
-         hispanic_pct = 'Hispanic...of.Cases',
-         other_pct = 'Other...of.Cases',
-         state_emergency = 'State.of.emergency',
-         business_closed = 'Closed.other.non.essential.businesses',
-         business_reopen = 'Began.to.reopen.businesses.statewide',
          mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
-         mask_legal='No.legal.enforcement.of.face.mask.mandate',
-         black_population_pct ="Black...of.Total.Population",
-         white_population_pct = "White...of.Total.Population",
          poverty_pct = "Percent.living.under.the.federal.poverty.line..2018.",
@@
          senior_pct = "X65.",
-         population_2018 = "Population.2018", ## extra variable in test
-         non_elderly_pre_existing = "Nonelderly.Adults.Who.Have.A.Pre.Existing.Condition",
-         homeless = "Number.Homeless..2019.") %>%
-  select(State, case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
-         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, 
-         black_population_pct, white_population_pct, poverty_pct, unemployed_pct, senior_pct,
-         population_2018, non_elderly_pre_existing, homeless)
+         population_2018 = "Population.2018" ## extra variable in test
+         ) %>%
+  select(State, case_rate_100k, population_density, mask_public, poverty_pct, 
+         unemployed_pct, senior_pct, population_2018)
 # Assign correct data types
-cols.num <- c("white_pct","black_pct","hispanic_pct","other_pct")
-df[cols.num] <- sapply(df[cols.num],as.numeric)
-df$state_emergency=as.Date(df$state_emergency, format = "%m/%d/%Y")
-df$business_closed=as.Date(df$business_closed, format = "%m/%d/%Y")
-df$business_reopen=as.Date(df$business_reopen, format = "%m/%d/%Y")
 head(df)
@@
 
-## 1. An Introduction
+# 1. An Introduction
 
@@
 
-### Assumptions
+## Assumptions
 
@@
 
-## 2. A Model Building Process
+# 2. A Model Building Process
 
@@
 
-### Model Variables
+## Model Variables
 
@@
 It will also serve to ease the burden on the readers of this report as like-for-like scaled relationships can be drawn between the variables.
-Given that this transformation is constant for all values (scalar multiplication of 100,000 for the seniors variable, and 1,000 for poverty and unemployment), it will have no effect on the distributions of the variables.
+Given that this transformation is constant for all values (scalar multiplication of 100,000 for the seniors variable given it is a ratio, and 1,000 for poverty and unemployment given they are percentages on a scale of 0-100%), it will have no effect on the distributions of the variables.
 
@@
 # Percentage variables are re-scaled to rates per 100k
-
 df <-df %>%
@@
     senior_per_100k = 100000*senior_pct,
-    poverty_per_100k = 100000*poverty_pct/100, 
-    unemployed_per_100k = 100000*unemployed_pct/100,
+    poverty_per_100k = 1000*poverty_pct, 
+    unemployed_per_100k = 1000*unemployed_pct
   )
@@
 
-### Model 1
+## Model 1
 
@@
 
-#### Model 1 Regression
+### Model 1 Regression
 
@@
 
-#### Model 1 CLM Assumptions
+### Model 1 Limitations
 
-##### i. IID Sampling
+#### i. IID Sampling
 
@@
 
-##### ii. Linear Conditional Expectation
+#### ii. Linear Conditional Expectation
 
@@
 
-##### iii. No Perfect Collinearity
+#### iii. No Perfect Collinearity
 
@@
 
-##### iv. Homoskedastic Errors
+#### iv. Homoskedastic Errors
 
@@
 
-##### v. Normally Distributed Errors
+#### v. Normally Distributed Errors
 
@@
 
-#### Influence of Data Points in Model 1
+#### vi. Influence of Data Points
 
@@
 
-### Model 2
+## Model 2
 
-Based on the work of model 1, two extra state demographic information regarding poverty rate (percentage living under the federal poverty line in 2018) and unemployment rate (percentage of unemployed in 2018) would be included in the analysis of model 2.
+Based on the work of model 1, two extra state demographic features regarding poverty rate (percentage living under the federal poverty line in 2018) and unemployment rate (percentage of unemployed in 2018) would be included in the analysis of model 2.
 In addition, both newly-introduced percentage variables would be transformed into rates per 100,000 to align with the case rate variable.
@@
 ```
-As shown in the above distribution, there is a fairly normal distribution in the poverty number 
-per 100,000 across the states. Additionally, there is no outliers observed within the distribution,
-given that the range of rates fall between 7,600 and 19,700. Therefore, there is no further transformation in need for this variable in the following analysis process for both model 2 and model 3.
+As shown in the above distribution, there is a fairly normal distribution in the poverty number per 100,000 across the states.
+Additionally, there is no outlier observed within the distribution, given that the range of rates fall between 7,600 and 19,700.
+Therefore, there is no further transformation in need for this variable in the following analysis process for both model 2 and model 3.
 
@@
 ```
-According to the histogram above, the distribution of unemployed number per 100,000 is of fairly normal distribution.
+According to the histogram above, the distribution of unemployed number per 100,000 is fairly normal distribution.
 Overall, it is not highly skewed or heavily tailed.
-Therefore, there is no further transformation in need for this variable in the following analysis process for both model 2 and model 3.
+Therefore, there is no further transformation needed for this variable in the following analysis process for both model 2 and model 3.
 
-#### Model 2 Regression
+### Model 2 Regression
 
@@
   labs(
-    title = ' Association between Covid-19 ase rate and the poverty rate', 
+    title = ' Association between Covid-19 case rate and the poverty rate', 
     x = 'Number of living under poverty line per 100,000',
-    y = 'Covid-19 care rate per 100,000', 
+    y = 'Covid-19 case rate per 100,000', 
     color = 'Senior num per 100k'
@@
   labs(
-    title = 'Association between Covid-19 ase rate and the unemployment rate', 
+    title = 'Association between Covid-19 case rate and the unemployment rate', 
     x = 'Number of unemployed per 100,000', 
-    y = 'Covid-19 care rate per 100,000',
+    y = 'Covid-19 case rate per 100,000',
     color = 'Senior num per 100k'
@@
 ```
-However, there is an overall increasing trend of COVID-19 care rate with increasing poverty rate, while there is a inverse relationship between the variables of the COVID-19 case rate and the unemployment rate.
+However, there is an overall increasing trend of COVID-19 case rate with increasing poverty rate, while there is an inverse relationship between the variables of the COVID-19 case rate and the unemployment rate.
 
-Based on above data exploratory result, the following equation is used to create a regression model
-to determine the relationship between the case rate and demographic features including the seniors rate, the rate of poverty and the rate of unemployment.
+Based on the above data exploratory result, the following equation is used to create a regression model to determine the relationship between the case rate and demographic features including the seniors rate, the rate of poverty and the rate of unemployment.
 
@@
 
-Accordingt to the assessment of homoskedasticity of model 2, which would be discussed in the following paragraph (Model 2 limits - iv.Homoskedastic Errors), it is plausible to apply the classical standard errors in the t-test of coefficients here.
+According to the assessment of homoskedasticity of model 2, which would be discussed in the following paragraph (Model 2 Limitations - iv.Homoskedastic Errors), it is plausible to apply the classical standard errors in the t-test of coefficients here.
 
@@
 
-Based on the test results of the estimated coefficients in model 2, all three demographic features discussed here are significantly realated to the COVID-19 care rate, especially for the poverty rate (p-value < 0.01).
+Based on the test results of the estimated coefficients in model 2, all three demographic features discussed here are significantly related to the COVID-19 case rate, especially for the poverty rate (p-value < 0.01).
 
-The coefficient of the senior_per_100k is  -0.230970 and statistically significant, which implies that for every one thousand additional seniors per 100,000, there is a decrease of 231 COVID-19 cases per 100,000 holding other demographic features as constant.
-As discussed in mode1 1 section, although seniors are susceptible group, less mobility and social interaction with cautious self procetion may play a important role in this pandemic situation.
+The coefficient of the senior_per_100k is -0.230970 and statistically significant, which implies that for every one thousand additional seniors per 100,000, there is a decrease of 231 COVID-19 cases per 100,000, holding other demographic features as constant.
+As discussed in mode1 1 section, although seniors are susceptible group, less mobility and social interaction with cautious self protection may play an important role in this pandemic situation.
 
-In addition, the coefficient of the poverty_per_100k (0.231750) indicats that for every one thousand additional people living under poverty line per 100,000, there is a increase of 232 COVID-19 case per 100,000 holding other demographic features as constant.
-People living under the poverty line have limited accesss to medical resourses.
-Moreover, most of them could only afford to live in the communites with poor hygiene.
-condition.
+In addition, the coefficient of the poverty_per_100k (0.231750) indicates that for every one thousand additional people living under poverty line per 100,000, there is a increase of 232 COVID-19 cases per 100,000, holding other demographic features as constant.
+People living under the poverty line have limited access to medical resources.
+Moreover, most of them could only afford to live in the communities with poor hygiene conditions.
 
-Finally, the coefficient of the unemployed_per_100k is statiscally significant and negative(-0.486191), which implies that for every one thousand additional unemployed, there is a decrease of 486 COVID-19 case per 100,000 holding other demographic features as constant.
-Although the unemployment leads to less medical insurace coverage and impact on the incoming, the unempolyed people may have more flexibility in obeying stay-at-home.
-In addition, there is less expourse probability when no commuting or in-place working is required for the unemployed.
+Finally, the coefficient of the unemployed_per_100k is statistically significant and negative (-0.486191), which implies that for every one thousand additional unemployed, there is a decrease of 486 COVID-19 case per 100,000, holding other demographic features as constant.
+Although the unemployment leads to less medical insurance coverage and impact on their income, the unemployed people may have more flexibility to obey stay-at-home orders.
+In addition, there is less exposure probability when there is no commuting or in-place working required for the unemployed.
 
-#### Model 2 limits
+### Model 2 Limitations
 
-##### i. IID Sampling 
+#### i. IID Sampling 
 (Would be discussed in the following section - 3. Limitations of the Models)
 
-##### ii.Linear Conditional Expectation
+#### ii.Linear Conditional Expectation
 To assess the linear conditional expectation assumption in higher-dimensional space, the plots of predicted versus residuals of the model 2 is investigated as follow.
@@
 
-#####iii.No Perfect Collinearity
+#### iii.No Perfect Collinearity
 
@@
 
-#####iv.Homoskedastic Errors
+#### iv.Homoskedastic Errors
 To evaluate the homoskedastic errors assumption, the square root of the residuals is plotted against the fitted values of the model 2 (scale-location plot).
@@
 
-#####v.Normally Distributed Errors
+#### v.Normally Distributed Errors
 To access the assumption of normality of error distribution, both the histogram and the qqplot of the residuals is shown as follow.
 
-```{r, echo=FALSE, warning=FALSE, message=False}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 plot_one <- df %>% 
@@
 
-##### Cook Distance
+### Cook Distance
 
@@
 
-### Model 3
+## Model 3
 
@@
 
-#### Model 3 Regression
+### Model 3 Regression
 

### 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Update Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
update model 3 after conflict

@@
 
-Next, the regression model for model 3 is based on the existing work from model 2.
+Next, the regression model for model 3 is created based on the existing work from model 2.
 Two additional variables discussed in Model 3 will be introduced as maximalist approach.
 The first variable is the "mask_public_bool" (whether the state enforced public mask mandate or not) that represents the policy effects.
-The second variable introduced is "population_density" (the value is logged as discussed earlier) which is considered an important factor considering nature of spread in disease.
+The second variable introduced is "population_density" (the value is logged as discussed earlier) which is considered an important factor related to the spread of disease.
 

### 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Latest again

--- 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
@@
 #### i. IID Sampling 
-(Would be discussed in the following section - 3. Limitations of the Models)
+(This will be discussed in the following section - 3. Limitations of the Models)
 
 #### ii.Linear Conditional Expectation
-To assess the linear conditional expectation assumption in higher-dimensional space, the plots of predicted versus residuals of the model 2 is investigated as follow.
+To assess the linear conditional expectation assumption in higher-dimensional space, the plots of predicted versus residuals of the model 2 is investigated as follows.
 
@@
 
-Overall, the residuals maintain around zero across the predicted range. There is no assumption violation phenomenon could be observed here.
+Overall, the residuals maintain around zero across the predicted range. There is no assumption violation phenomenon that can be observed here.
 
@@
 
-Since there are multiple covariates included in this model compared to the model 1, there is more concern about the no perfect collinearity assumption here
-First, the list of variables coefficients shows that no variables were dropped, which mean there is no perfect collinearity detected by the R function.
+Since there are multiple covariates included in this model compared to the model 1, there is more concern about the no perfect collinearity assumption here.
+First, the list of variable coefficients shows that no variables were dropped, which means there is no perfect collinearity detected by the R function.
 
@@
 
-In addition, all the variance inflation factors are less than 4 as follow, which doesn't indicate the exisitence of collinearity.
+In addition, all the variance inflation factors are less than 4 as follows, which doesn't indicate the existence of collinearity.
 
@@
 #### iv.Homoskedastic Errors
-To evaluate the homoskedastic errors assumption, the square root of the residuals is plotted against the fitted values of the model 2 (scale-location plot).
-Although, the curve is not perfect flat smooth, there is no obvious variance of the errors either.
-For reference, the data points marked in the plot represent Florida(10), New York (33) and North Dakota (35).
+To evaluate the homoskedastic errors assumption, the square root of the residuals is plotted against the fitted values of model 2 (scale-location plot).
+Although, the curve is not perfectly flat, there is no obvious variance of the errors either.
+For reference, the data points marked in the plot represent Florida (10), New York (33) and North Dakota (35).
 
@@
 Additionally, a Breusch-Pagan test is run to check the level of heteroskedasticity.
-The test result shows it fails to reject the null hypotheis, which means there is no evidence for heteroskedasticity.
+The test result shows it fails to reject the null hypothesis, which means there is no evidence for heteroskedasticity.
 Overall, the assumption of homoskedastic errors is satisfied for model 2.
@@
 
-#### v.Normally Distributed Errors
-To access the assumption of normality of error distribution, both the histogram and the qqplot of the residuals is shown as follow.
+#### v. Normally Distributed Errors
+To access the assumption of normality of error distribution, both the histogram and the Q-Q plot of the residuals is shown as follow.
 

### 2020-12-05 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Updated as discussed and knitted a PDF

@@
 
-## Assumptions
+## 1-1. Assumptions
 
@@
 
-## Model Variables
-
 The three models aim to investigate the relationship between the COVID-19 case rate, state population demographics, and a mask mandates.
@@
 
-## Model 1
+## 2-1. Model 1
+
+### 2-1-1. Model 1 Exploratory Data Analysis
 
@@
 
-### Model 1 Regression
+### 2-1-2. Model 1 Regression
 
@@
 
-### Model 1 Limitations
+### 2-1-3. Model 1 Limitations
 
@@
 This is a necessary assumption to ascertain that the errors used in the regression are drawn from a normal distribution, so that they can be accepted when used to calculate the significance levels.
-Both a QQ plot and histogram of the distribution of the residuals is plotted to observe the normality.
+Both a Q-Q plot and histogram of the distribution of the residuals is plotted to observe the normality.
 
 ```{r, message = FALSE}
-# Create a QQ plot for the model 1 residuals
-plot(model1, which=2)
-
 # Create a histogram of the distribution of the model 1 residuals
-df %>%
+plot_one <- df %>%
   ggplot(aes(x = resid(model1))) +
@@
        x = "Model 1 residuals", y = 'Count')
+
+# Create a QQ plot for the model 1 residuals
+plot_two <- df %>% 
+  ggplot(aes(sample = resid(model1))) + 
+  stat_qq() + stat_qq_line()
+
+plot_one / plot_two
 ```
@@
 
-## Model 2
+## 2-2. Model 2
+
+### 2-2-1. Model 2 Exploratory Data Analysis
 
@@
 
-### Model 2 Regression
+### 2-2-2. Model 2 Regression
 
@@
 
-### Model 2 Limitations
+### 2-2-3. Model 2 Limitations
 
 #### i. IID Sampling
-(This will be discussed in the following section - 3. Limitations of the Models)
 
-#### ii.Linear Conditional Expectation
+The first assumption of IID is not detailed within this section given that it is evaluated more generally for all three models within section 3.
+
+#### ii. Linear Conditional Expectation
 To assess the linear conditional expectation assumption in higher-dimensional space, the plots of predicted versus residuals of the model 2 is investigated as follows.
@@
 
-#### iii.No Perfect Collinearity
+#### iii. No Perfect Collinearity
 
@@
 
-#### iv.Homoskedastic Errors
+#### iv. Homoskedastic Errors
 To evaluate the homoskedastic errors assumption, the square root of the residuals is plotted against the fitted values of model 2 (scale-location plot).
@@
 Limited by the sample size, it is hard to observe a perfect normal distribution from the histogram.
-However, the qqplot implies no deviation from normality.
+However, the Q-Q plot implies no deviation from normality.
 Overall, the normally distributed errors assumption is satisfied for model 2.
 
-### Cook Distance
+#### vi. Influence of Data Points
 
-Additionally, cook distance is used to estimate the influence of each single data point.
+Additionally, Cook's distance is used to estimate the influence of each single data point.
 
@@
 ```
-No obvious outlier could be detected here. For reference, the marked data points are Florida(10), Maine(20), and North Dakota(30).
 
-## Model 3
+No obvious outliers could be detected here. For reference, the marked data points are Florida (10), Maine (20), and North Dakota (30).
+
+## 2-3. Model 3
+
+### 2-3-1. Model 3 Exploratory Data Analysis
 
@@
 
-### Model 3 Regression
+### 2-3-2. Model 3 Regression
 
@@
 
-### CLM Assumptions Validation for Model 3:
+### 2-3-3. Model 3 Limitations
 
-#### I.I.D Assumption
+#### i. I.I.D Assumption
 
-( Discussed in the common section)
+The first assumption of IID is not detailed within this section given that it is evaluated more generally for all three models within section 3.
 
-#### Zero Conditional Mean
+#### ii. Linear Conditional Expectation
 
@@
 
-#### Normality of Errors
+#### iii. No Perfect Collinearity
 
-From the Normal Q-Q plot below, the points generally stay close to the normal line, with the tails exhibit certain degree of skew.
-There is no significant evidence of concerning normality in the errros.
+Based on the Variance Inflation Factors check, all the variables have variance-inflation factors less than 4. As a result, there's no strong evidence of collinearity concern within the model. 
 ```{r}
-plot(model3, which=2)
+
+vif(model3)>4
 ```
 
-#### Homoskedasticity
+#### iv. Homoskedastic Errors
 
@@
 
-#### Collinearity
+#### v. Normally Distributed Errors
 
-Based on the Variance Inflation Factors check, all the variables have variance-inflation factors less than 4. As a result, there's no strong evidence of collinearity concern within the model. 
-```{r}
+To access the assumption of normality of error distribution, both the histogram and the Q-Q plot of the residuals is shown as follow.
 
-vif(model3)>4
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
+plot_one <- df %>% 
+  ggplot(aes(x = resid(model3))) +
+  geom_histogram(fill = 'skyblue', color = 'grey30', bins=30) +
+  labs(title =  "Histogram of the Model 3 Residuals", 
+       x = "Model 3 residuals", y = 'Count')
+
+plot_two <- df %>% 
+  ggplot(aes(sample = resid(model3))) + 
+  stat_qq() + stat_qq_line()
+
+plot_one / plot_two
 ```
 
-#### Cook's Distance
+There is a fairly normal distribution that can be seen from the histogram above.
+From the Q-Q plot, the points generally stay close to the normal line, with the tails exhibiting a certain degree of skew.
+However, in general, it implies no deviation from normality.
+Overall, the normally distributed errors assumption is satisfied for model 3.
 
-From the Cook's plot, there is no strong evidence of concerning outliers.
+#### vi. Influence of Data Points
+
+From the Cook's distance plot, there is no strong evidence of concerning outliers. For reference, the marked data points are D.C. (9), New Hampshire (30), and North Dakota (35).
 
@@
 
-## 3. Limitations of the Models
+# 3. Limitations of the Models
 
 The most general requirement for all the linear models is that data points be independent and identically distributed (IID) which they are built off of.
-While for the most part the events of one state will primarily affect that single state, travel by individuals spreading the virus across state lines is an inevitable occurence.
+While for the most part the events of one state will primarily affect that single state, travel by individuals spreading the virus across state lines is an inevitable occurrence.
 This may cause the spread of a virus within one state to be influenced by those around it, or generally by its position in the country.
@@
 In the first model, deviations away from certains assumptions are the strongest.
-This is seen mostly with regards to the linear conditonal expectation assumption, where the residuals against the fitted values have a noticeable inverted "V" shape as opposed to being constant around zero.
+This is seen mostly with regards to the linear conditional expectation assumption, where the residuals against the fitted values have a noticeable inverted "V" shape as opposed to being constant around zero.
 In the second model, the same diagnostic shows that the assumption is better met.
@@
 
-## 4. Regression Table
+# 4. Regression Table
 
@@
 Given that the null hypothesis for this test is that the reduced model should be over the full model, this can be safely rejected in favor of the full model.
+Additionally, it can be seen that the standard errors for the significant variables are relatively small and are consistent across the models.
 
@@
 
-In model 1, the coefficient of -0.174 for the senior rate per 100,000 is significant with a p-value of 0.0234 (per the regression results from the Model 1 section).
+In model 1, the coefficient of -0.174 for the senior rate is significant with a p-value of 0.0234 (per the regression results from the Model 1 section).
 When the covariate variables of unemployment rate and poverty rate are added in to model 2, the coefficient for senior rate adjusts to a greater magnitude of -0.231 while the p-value appreciably decreases to a value of 0.0015 (per the regression results from the Model 2 section).
@@
 In addition, the coefficients of the covariate variables themselves are also highly significant with p-values of 0.0005 and 0.0059 for poverty rate and unemployment rate respectively.
-The magnitude of the effect of poverty rate is almost equivalent to that of the senior rate, with a single increase in poverty rate per 100,000 equivalent to 0.232 per 100,000 for case rate, while the unemployment rate has stronger negative coefficient of -0.486 per 100,000 (roughly twice the magnitude of the other two variables).
+The magnitude of the effect of poverty rate is almost equivalent to that of the senior rate, with a coefficient of 0.232, while the unemployment rate has stronger negative coefficient of -0.486 (roughly twice the magnitude of the other two variables).
 
@@
 Notably, the coefficient for the mask policy covariate is significant with a p-value of 0.0027 and has by far the greatest effect on the outcome variable with a coefficient of -945.6.
-The populuation density covariate is not significant with a p-value of 0.3369.
+The population density covariate is not significant with a p-value of 0.3369.
 
@@
 With a coefficient of -945.6, this means that given the binary mask policy variable, when a mask policy is implemented (value of 1), there is an decrease in case rate of roughly 945 per 100,000 compared to when the policy is not in place.
-The next largest effect is from the unemployment rate variable, where in the final model, the coefficient of -0.364 translates as a decrease of 0.364 cases per 100,000 for every additional increase of 1 in unemployment rate per 100,00.
-More intuitively, it can be said that there is a decrease of roughly 1 case per 100,000 with every 3 additional people unemployed per 100,000.
-This effect is roughly 2,600 times smaller than the effect of implementing a mask policy.
+The next largest effect is from the unemployment rate variable, where in the final model, the coefficient of -0.364 translates to a decrease of roughly 1 case per 100,000 with every 3 additional people unemployed per 100,000.
 Similarly, the effects of both poverty rate and senior rate are nominal.
 For every 5 additional people per 100,000 who are below the poverty line, there is roughly 1 additional case per 100,000, and for every 4 additional seniors per 100,000 within the state population, there is roughly 1 fewer case per 100,000.
-In comparison with likewise relatively minimal effect of the unemployment rate, these absolute effects on case rate are approximately one-third (1/3) and one-half (1/2) smaller respectively.
 
-## 5. Omitted Variables
+# 5. Omitted Variables
 
@@
 
-### 5-1: Urban vs Rural Population Ratio
+## 5-1. Urban vs Rural Population Ratio
 
@@
 While population density is potentially useful, it was found to not be significant in the models.
-This may be due to states which have large urban centers, where population density woukd be high, losing their impact due to also containing large areas of land that is sparsely populated.
+This may be due to states which have large urban centers, where population density would be high, losing their impact due to also containing large areas of land that is sparsely populated.
 One example of this would be New York, which contains several of the most densely populated counties in the country, but also large rural areas.
@@
 
-### 5-2: Median Income
+## 5-2. Median Income
 
@@
 
-### 5-3: Gender Ratio
+## 5-3. Gender Ratio
 
-The ratio of gender between women and men may impact the COVID case rate in that men are more suscepitble to viral infections and generally have lower life expectancies from lifestyle differences.
+The ratio of gender between women and men may impact the COVID case rate in that men are more susceptible to viral infections and generally have lower life expectancy from lifestyle differences.
 Greater susceptibility may relate to the case rate in that men may more often become infected when exposed to COVID under similar circumstances as a women.
-Their shorter life expentancy may also impact the case rate in that more COVID cases may result in greater or more serious symptoms, with less occurances of an infection that is asymptomatic or goes undetected.
+Their shorter life expectancy may also impact the case rate in that more COVID cases may result in greater or more serious symptoms, with less occurrences of an infection that is asymptomatic or goes undetected.
 In either of these cases, a higher fraction of men among the population would lead to an overall higher COVID case rate.
@@
 
-### 5-4: Mask Use Among Population
+## 5-4. Mask Use Among Population
 
@@
 
-### 5-5: GDP Percentage of Tourism
+## 5-5. GDP Percentage of Tourism
 
@@
 
-## 6. Conclusion
+# 6. Conclusion
 
@@
 Although there is some intuition behind the unemployment and poverty rates having similar effects on case rate, it can be seen from the test of collinearity in the Model 2 section that these two variables are not explaining the same effects.
-This difference is likely due to the fact that those living below the poverty line have less access to suitable living conditions and sanitary goods that help to alleviate the possibility of COVID-19 infection, while those who were recently employed up until the pandemic hit may not be in the same economical situation.
+This difference is likely due to the fact that those living below the poverty line have less access to suitable living conditions and sanitary goods that help to alleviate the possibility of COVID-19 infection.
+As a caveat, it should be noted that these the poverty and unemployment figures are from a 2018 survey, so it does not reflect the full picture of the state demographics at the time of when the COVID-19 related data was captured.
 Finally, the effect of the senior rate is negative, with an additional 4 seniors per 100,000 translating to roughly 1 fewer case per 100,000.
@@
 In conclusion, we see that although the population demographics had varying effects on the COVID-19 case rate, the greatest effect was seen from the mask policy.
-Going back to our original question, it can be suggested within the data that the effect of a mask-related policy is much larger than that of population demographics.
-Therefore, it seems that the COVID-19 case rate at the state-level is much more dependent on masks, and that ultimately, there are effective policy and control measures that can be taken by state governments that have a greater contribution towards combating it's spread relative to factors that cannot be controlled, such as population demographics.
+Going back to our original question, it can be suggested within the data that the effect of a mask-related policy is something that should not be neglected in comparison to the population demographics.
+Therefore, it seems that the COVID-19 case rate at the state-level is dependent on masks, and that ultimately, there are effective policy and control measures that can be taken by state governments that have some contribution towards combating it's spread relative to factors that cannot be controlled.
 However, it should be noted that these insights are qualified by the assumptions and limitations laid out throughout the report, namely that there are potential issues in meeting the IID assumption and that there is explanatory information that is lost from both the variables investigated as well as the those that were not included in the dataset.
+
+# References

### 2020-12-06 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Updated CLM section.

@@
 
-In examining the results of various diagnostics across different models, generally the CLM assumptions are best met with the second model.
-In the first model, deviations away from certains assumptions are the strongest.
+In examining the results of various diagnostics across different models, generally the CLM assumptions are better met as the model continues to develop.
+In the first model, deviations away from certain assumptions are the strongest.
 This is seen mostly with regards to the linear conditional expectation assumption, where the residuals against the fitted values have a noticeable inverted "V" shape as opposed to being constant around zero.
 In the second model, the same diagnostic shows that the assumption is better met.
+This is further improved upon in the third model, where local variations in the linear conditional expectation diagnostic plot are generally smoother and more centered around zero.
+The assumption with the greatest exception to this is that of no high colinearity within the models, as each coefficient in each model was shown to have VIF values not near the cut off value of four.
 
@@
 For example, the normality of the error term in the model is less well fit than it was in the earlier models.
-It can be seen in the QQ plot for the model how at fitted values towards the edges of the model's range, the is the greatest deviation away from normality.
-This assumption is generally less important than that of the linear conditional expectation, however, where the latter is a fundamental part of CLM.
+It can be seen in the third model's Q-Q plot how at fitted values towards the edges of the model's range, there exists the greatest deviation away from normality.
+This assumption is generally less important than that of the linear conditional expectation, however, which is a fundamental part of CLM.
+Therefore, the third model can still be considered as generally improving the overall analysis.
+
+In addition, with more variables in the third model, outliers appear to have more of an impact than previously.
+This is likely due to the limited number of data points the model is based off of, an entire state only constitutes an individual datapoint.
+With less degrees of freedom in the third model, certain points can be seen to more closely approach the boundaries of the Cook's length plot, although they are still not extreme enough to fully be considered an outlier.
+Increasing the resolution of the data, such as examining COVID-19 cases at the county or city/town level, would help to reduce this by increasing the overall degrees of freedom.
 

### 2020-12-06 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Update Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
update model 3

@@
 Next, the regression model for model 3 is created based on the existing work from model 2.
-Two additional variables discussed in Model 3 will be introduced as maximalist approach.
+Two additional variables discussed in Model 3 will be introduced as the maximalist approach.
 The first variable is the "mask_public_bool" (whether the state enforced public mask mandate or not) that represents the policy effects.
-The second variable introduced is "population_density" (the value is logged as discussed earlier) which is considered an important factor related to the spread of disease.
+The second variable introduced is "population_density" (the variable is transformed with log of 10 as discussed earlier) which is considered an important factor related to the spread of disease.
 
@@
 
-According to the plot of Homoskedasticity, the standardized residual falls into a reasonable range and the line looks smooth without concerning skews.
+According to the plot of Homoskedasticity, the standardized residual falls into a reasonable range and the curve looks smooth without concerning skews.
 There is no significant evidence that the model violates homoskedasticity assumptions.
@@
 
+```
+Additionally, a Breusch-Pagan test is run to check the level of heteroskedasticity.
+The test result shows it fails to reject the null hypothesis, which indicates that there is no strong evidence of heteroskedasticity in model 3.
+
+```{r, echo=FALSE}
+bptest(model3)
 ```
@@
 Going back to our original question, it can be suggested within the data that the effect of a mask-related policy is something that should not be neglected in comparison to the population demographics.
-Therefore, it seems that the COVID-19 case rate at the state-level is dependent on masks, and that ultimately, there are effective policy and control measures that can be taken by state governments that have some contribution towards combating it's spread relative to factors that cannot be controlled.
+Therefore, it seems that the COVID-19 case rate at the state-level is dependent on masks, and that ultimately, there are effective policy and control measures that can be taken by state governments that have some contribution towards combating its spread relative to factors that cannot be controlled.
 However, it should be noted that these insights are qualified by the assumptions and limitations laid out throughout the report, namely that there are potential issues in meeting the IID assumption and that there is explanatory information that is lost from both the variables investigated as well as the those that were not included in the dataset.

### 2020-12-06 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Updated RMD and re-knit per meeting discussions

@@
 
-The research question aims to measure the COVID-19 case rate per 100,000 residents within each state based on the make-up of the population, as well as the policy decisions that were or were not put into place in order to combat the rise of COVID-19 cases.
+The research question aims to analyze the relationship between the COVID-19 case rate per 100,000 residents within each state and the population demographics features, as well as the policy decisions that were or were not put into place in order to combat the rise of COVID-19 cases.
 In this sense, the aim is to examine how dependent the COVID-19 case rate is on that of which cannot be controlled (i.e., population demographics), as well as that of which can be controlled to a certain degree (i.e., implementation of policies to attempt to combat case rate) by each state.
-With this information, one could suggest whether or not the proliferation of COVID-19 within a state seems to be related to either or both controllable and uncontrollable variables, and to the level at which the effect of these variables differ help.
+With this information, one could suggest whether or not the proliferation of COVID-19 within a state seems to be related to either or both controllable and uncontrollable factors and how they differ.
 The modeling goal of this research question will be one of description, and will be broken up into three phases of investigation.
 
-The first and primary phase of investigation involves how the COVID-19 case rate is affected by key population demographics.
-For this, the key variables are COVID-19 case rate per 100,000 residents and the percentage population distribution of seniors (defined as 65 years old or older) within a state.
-The case rate per 100,000 was selected as the key output/dependent variable as it provides a standardized measure of the spread of COVID-19 across the states, taking into account the absolute population of the states.
+The first and primary phase of investigation involves how the COVID-19 case rate is related to key population demographics.
+For this, the key variables are COVID-19 case rate per 100,000 residents and the percentage population distribution of seniors (defined as 65 years old or older in the original dataset) within a state.
+The case rate per 100,000 was selected as the key output/dependent variable as it provides a standardized measure of the spread of COVID-19 across the states, and already takes into account the absolute population of the states.
 The senior percentage was selected as the key input/independent variable given that guidance had been released by the Centers for Disease Control and Prevention (CDC) that seniors belong to the age category for those at higher risk of COVID-19, and thus are a well studied and documented group.
@@
 
-Following this, in the second phase of investigation the analysis considers other variables of state-level population that are considered factors of susceptibility, namely the rate of poverty (defined as the percent of individuals living under the federal poverty line in 2018) and the rate of unemployment (defined as the percent unemployment in 2018).
+Following this, in the second phase of investigation the analysis considers other variables of state-level population demographic features that are considered factors of susceptibility, namely the rate of poverty (defined as the percent of individuals living under the federal poverty line in 2018 in the original dataset) and the rate of unemployment (defined as the percent unemployment in 2018 in the original dataset).
 Those living below the poverty line may have less access/are unable to afford to preventative controls such as sanitation products and masks that help prevent the spread of the virus, while those who are designated as unemployed may not have the flexibility of sheltering at home as well as access to the aforementioned preventative controls.
@@
 
-The tertiary and final phase of investigation includes a variable that measures the implementation of policy as a response to COVID-19, and to understand the added effect of this variable in conjunction with the first phase of investigation that is state population metrics.
+The tertiary and final phase of investigation includes a variable that measures the implementation of policy as a response to COVID-19, and to understand the added effect of this variable in conjunction with the first phase of investigation.
 Specifically, there is a focus on a policy that mandates the wearing of masks within the state.
-Through this, the aim is to analyze whether the implementation of a mask-related policy will have an effect on case rate, as well as the degree of this effect relative to population metrics.
+Through this, the aim is to analyze whether the implementation of a mask-related policy is related to case rate, as well as the strength of the association compared to population metrics.
 To operationalize this policy and simplify the measurements, the models only consider whether or not this policy was implemented through transformed indicator variables (1 = implemented, 0 = not implemented).
@@
 
-Additionally, the final phase of investigation continues to expand on the examination of state-level population demographics via the population density variable (defined as the population density in square miles, population/square miles).
+Additionally, the final phase of investigation continues to expand on the examination of state-level population demographics via the population density variable (defined as the population density in square miles, population/square miles in the original dataset).
 A key factor to the spread of COVID-19 is the idea of social distancing, where the virus is considered to be more likely to spread when people are close in proximity.
 Thus, population density is measured as a conduit to indicate the level of proximity within each state and will operationalize the concept of social distancing.
+It should be noted however that the population density data may be limited in its ability to serve as a direct measurement of social distancing given the it lacks other contextualize information, such as the urban to rural ratio as later discussed in section 5.
 
@@
 Assuming that this has some effect on case rate, this difference in length of time may play a part in the effect.
-There are also other external aspects that cannot be included, such as the level of population density broken out among different population demographics.
+There are also other external aspects that cannot be included, such as the rural to urban ratio as opposed to population density.
 
@@
 The covariates that will be examined in building the models fit into two broad categories.
-The first category is demographic information on the state, namely percentage population distribution of seniors, the unemployment rate, the rate of people living in poverty, and the population density.
-The second general group is that of policy decisions taken by the state, specifically a mask mandate.
+The first category is the demographic features of the state, namely senior rate, the unemployment rate, the poverty rate, and the population density.
+The second category is that of policy decisions taken by the state, specifically a mask mandate.
 Problematic covariates would include any of the other direct measures of COVID-19 severity in the state, namely total infection rate not on a per capita basis and COVID-19 death rates.
-It can be assumed that these variables would be collinear with the primary variable of interest since COVID-19 case rate is simply a linear transformation of total cases not adjusted for population, and death rate is derived from case rate.
+It can be assumed that these variables would be collinear with the primary variable of interest since COVID-19 case rate is simply a linear transformation of total cases not adjusted for population.
 
-As mentioned in the introduction, the modeling goal for the COVID-19 case rate will be one of description.
-A model of causation may follow this research if potential causal relationships are examined and incorporated into the model's structure.
-This may later be expanded to predict the COVID-19 case rate in the past seven days in order for a state to use it to potentially judge what effect a mask policy could have.
-
-The three models aim to investigate the relationship between the COVID-19 case rate, state population demographics, and a mask mandates.
+The three models aim to investigate the relationship between the COVID-19 case rate, state population demographics, and policy (i.e., mask mandate).
 COVID-19 case rate per 100,000 is chosen as the key dependent variable to represent the proliferation of the COVID-19 pandemic in each state.
@@
   labs(
-    title = 'COVID-19 Case Rate due to Senior Population', 
+    title = 'COVID-19 Case Rate vs Senior Population', 
     x = 'Number of seniors per 100,000', 
@@
 
-From the QQ plot, normality in the residuals can be detected given the proximity of the points to the normal line.
+From the Q-Q plot, normality in the residuals can be detected given the proximity of the points to the normal line.
 Additionally, although the shape is a little harder to observe, the histogram of the residuals also gives a fairly normal distribution.
@@
   labs(
-    title = ' Association between Covid-19 case rate and the poverty rate', 
+    title = 'COVID-19 case rate vs poverty rate', 
     x = 'Number of living under poverty line per 100,000',
-    y = 'Covid-19 case rate per 100,000', 
+    y = 'COVID-19 case rate per 100,000', 
     color = 'Senior num per 100k'
@@
   labs(
-    title = 'Association between Covid-19 case rate and the unemployment rate', 
+    title = 'COVID-19 case rate vs unemployment rate', 
     x = 'Number of unemployed per 100,000', 
-    y = 'Covid-19 case rate per 100,000',
+    y = 'COVID-19 case rate per 100,000',
     color = 'Senior num per 100k'
@@
 This is further improved upon in the third model, where local variations in the linear conditional expectation diagnostic plot are generally smoother and more centered around zero.
-The assumption with the greatest exception to this is that of no high colinearity within the models, as each coefficient in each model was shown to have VIF values not near the cut off value of four.
+The assumption with the greatest exception to this is that of no high collinearity within the models, as each coefficient in each model was shown to have VIF values not near the cut off value of four.
 
@@
           star.cutoffs = c(0.05, 0.01, 0.001), 
-          title = "Table 1: The effect of population demographics and mask policy on COVID-19 case rate")
+          title = paste("Table 1: The effect of population demographics and mask",
+          "policy on COVID-19 case rate"))
 ```
@@
 The goal of the analysis was to address the research question to examine the relationship between both the distribution of population demographics and mask-related policy of a state and the COVID-19 case rate within that states.
-In order to address this question, the analysis was broken up into three phases of investigation with corresponding linear models:
+In order to address this question, the analysis was broken up into three phases of investigation with corresponding linear models.
+
  - Phase 1: To investigate the direct relationship between COVID-19 case rate and the population distribution of seniors within a state

### 2020-12-06 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
references update, references.bib created

@@
     df_print: paged
+bibliography: references.bib
+nocite: '@*'
 ---
@@
 
+\pagebreak
 # References

### 2020-12-07 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Added identical distribution information.

@@
 
+The identical distribution of states as individual samples in the model is also questionable.
+Similar to how connections or commonalities between states make them less than truly independent, certain factors would also influence the distribution so that some states are representative of a different underlying distribution than one another.
+For example, if one geographic region as a whole has a factor which influences the spread of the virus differently than another, multiple underlying distributions may be present in the sample.
+If still a new geographic region with many states was added to the sample, and the measures of central tendency of the data continued to evolve, there would be a further indication that the underlying distribution is not identical.
+Without a possibility of refining the data sample however, as states would be immutable in this regard, the assumption must be made for any potential analysis.
+
 In examining the results of various diagnostics across different models, generally the CLM assumptions are better met as the model continues to develop.

### 2020-12-07 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Made proofreading updates to model 1

@@
 
-As can be seen from the distribution above, again there is a fairly normal distribution in the population of seniors per 100,000 across the state.
+As can be seen from the distribution above, again there is a fairly normal distribution in the senior rate per 100,000 across the state.
 Additionally, there do not seem to be any outliers within the distribution, given that the range of rates fall between 11,000 and 21,000.
-Therefore, there are no transformations that need to be made in order to restructure the distribution, and this variable can be used across all three models.
+Therefore, there are no additional transformations that need to be made in order to restructure the distribution, and this variable can be used across all three models.
 
@@
 
-In order to test the relationship between the variables, the following equation is used to create a regression model to determine the true relationship between the case rate and population density variables.
+In order to test the relationship between the variables, the following equation is used to create a regression model for the case rate and senior rate variables.
 
@@
 The assumption of homoskedasticity is met if the data does not have large variance among the residuals.
-The analysis for this assumption, as well as the other CLM assumptions, is conducted in the following subsection titled "Model 1 CLM Assumptions".
+The analysis for this assumption, as well as the other CLM assumptions, is conducted in subsection 2-1-3, titled "Model 1 Limitations".
 
@@
 
-As can be seen from the above output, the variable for senior percentage is shown to be significant within the model specification with a p-value of 0.023.
-Furthermore, The coefficient of -0.174 for the variable is negative, supporting the earlier evaluation that there is an inverse relationship between the senior percentage and the case rate.
-In practical significance, this can be interpreted as for every additional senior per 100,000 in the population, there is a decrease of 0.174 COVID-19 cases per 100,000.
+As can be seen from the above output, the variable for senior rate is shown to be significant within the model specification with a p-value of 0.023.
+Furthermore, the coefficient of -0.174 for the variable is negative, supporting the earlier evaluation that there is an inverse relationship between the senior rate and the case rate.
+In terms of practical significance, this can be interpreted as for every additional senior per 100,000 in the population, there is a decrease of 0.174 COVID-19 cases per 100,000.
 More intuitively, for roughly every 6 additional seniors per 100,000, there is a decrease of 1 case per 100,000.
@@
 
-The second CLM assumption to be evaluated is that of a linear conditional expectation relationship within the data, so that the model is accurately estimating the relationship between the variables.
-To test for this, the residuals of model 1 are plotted against its fitted values.
+The second CLM assumption to be evaluated is that of a linear conditional expectation relationship within the model, to diagnose whether the model is accurately estimating the relationship between the variables.
+To test for this, the residuals of model 1 are plotted against the fitted values.
 
@@
 
-When examining the plot above for linearity in the conditional expectation, there is a noticeable slightly quadratic shape to the relationship of the conditional expectation.
+When examining the plot above for linearity in the conditional expectation, there is a noticeably slightly quadratic shape to the conditional expectation.
 This may suggest that the relationship between the variables may not be as linear as the model assumes.
 However, as discussed previously in the scatter plot of the two variables, the relationship can be approximated as roughly linear.
-It is also difficult to explain the complexity behind the COVID-19 case rate with just a single variable that does not fully capture the broader population demographics.
-Additionally, there is hope that given the addition of other input variables within models 2 and 3, the relationship within the data will be better explained and result in more linearity.
+It is also difficult to explain the complexity behind the COVID-19 case rate with just a single variable that does not fully capture the broader population demographics, so there is hope that given the addition of other input variables within models 2 and 3, the relationship within the data will be better explained and result in more linearity.
 
@@
 This is supported by the p-value of 0.497 from the Breusch-Pagan test, meaning it fails to reject the null hypothesis that the conditional errors are not heteroskedastic.
-With this additional assumption met, the regression model is run using a t-test with classical standard errors.
 
@@
 From the Q-Q plot, normality in the residuals can be detected given the proximity of the points to the normal line.
-Additionally, although the shape is a little harder to observe, the histogram of the residuals also gives a fairly normal distribution.
+Additionally, although the shape is a little harder to observe, the histogram of the residuals gives a fairly normal distribution.
 
@@
 
-In addition to the CLM assumptions, the Cook's distance is also examined by a plot of the residuals vs the model 1 leverage to estimate the influence of the data points and identify any outliers in the values.
+In addition to the CLM assumptions, the Cook's distance is also examined by a plot of the residuals vs the leverage to estimate the influence of the data points and identify any outliers in the values.
 From the plot below, no obvious outliers are detected.

### 2020-12-08 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Also added commentary on reference points in assumption diagnostics for model 1
Pushing latest RMD and PDF per changes discussed during the call

@@
 
- **Initial reassignment of common data used across models:**
+ **Initial reassignment of common variables used across models:**
 ```{r setup, include=FALSE, warning=FALSE, message=FALSE}
@@
          unemployed_pct = "Percent.Unemployed..2018.",
-         senior_pct = "X65.",
-         population_2018 = "Population.2018" ## extra variable in test
+         senior_pct = "X65."
          ) %>%
   select(State, case_rate_100k, population_density, mask_public, poverty_pct, 
-         unemployed_pct, senior_pct, population_2018)
-# Assign correct data types
+         unemployed_pct, senior_pct)
+
 head(df)
@@
 
-Next, the distribution of the senior percentage transformed into the rate per 100,000 (as discussed in the Model Variables section) is examined.
+Next, the distribution of the senior percentage transformed into the rate per 100,000 is examined.
 
@@
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(10000,21000,1000)) +
-  labs(title = 'Histogram of Seniors Distribution', 
+  labs(title = 'Histogram of Senior Rate', 
   x = 'Number of seniors per 100,000', y = 'Count') + 
@@
 
-```{r}
-# Check for linearty in model 1
-plot(model1, which=1)
+```{r, message = FALSE}
+
+df<- df%>% 
+  mutate(
+    model1_preds = predict(model1), 
+    model1_resids = resid(model1)
+  ) 
+df %>% 
+  ggplot(aes(model1_preds, model1_resids)) + 
+  geom_point() + 
+  stat_smooth()
 ```
@@
 This is supported by the p-value of 0.497 from the Breusch-Pagan test, meaning it fails to reject the null hypothesis that the conditional errors are not heteroskedastic.
+For reference, the marked data points are New York (33), North Dakota (35), and South Dakota (42).
 
@@
   geom_histogram(fill = 'skyblue', color = 'grey30', bins=30) +
-  labs(title =  "Histogram of the Model 1 Residuals", 
+  labs(title =  "Histogram and Q-Q Plot of the Model 1 Residuals", 
        x = "Model 1 residuals", y = 'Count')
@@
 From the plot below, no obvious outliers are detected.
+For reference, the marked data points are Alaska (2), Florida (10), and Maine (20).
 
@@
 Based on the work of model 1, two extra state demographic features regarding poverty rate (percentage living under the federal poverty line in 2018) and unemployment rate (percentage of unemployed in 2018) would be included in the analysis of model 2.
+Although the increasing unemployment rate during the pandemic has been widely covered by the media, the potential pandemic causing poverty or unemployment will not be investigated here, according to the designed research question in this work.
+
+Due to the fact that both poverty rate and unemployment rate data represents the status in 2018, it is proper to assume that they represent the pre-existing state demographic characteristics instead of the capturing any influence from the pandemic.
+
 In addition, both newly-introduced percentage variables would be transformed into rates per 100,000 to align with the case rate variable.
@@
   labs(title =  "Histogram of Poverty Rate", 
-       x = "Number of living under poverty line per 100,000", y = 'Count')+
+       x = "Number of people living under poverty line per 100,000", y = 'Count')+
   scale_x_continuous(breaks=seq(6000, 21200, 1600))
@@
 
-Compared to model 1, the relationship investigated here is about variables space with higher dimensions.
-It would be challenging to observe notable variation from variables distribution plots as below.
+Compared to model 1, the relationship investigated here involves a higher dimensional space. It would be challenging to observe notable variation from variables distribution plots as below.
 
@@
     x = 'Number of living under poverty line per 100,000',
-    y = 'COVID-19 case rate per 100,000', 
-    color = 'Senior num per 100k'
+    y = 'Case rate per 100,000', 
+    color = 'Seniors per 100k'
   )
@@
     x = 'Number of unemployed per 100,000', 
-    y = 'COVID-19 case rate per 100,000',
-    color = 'Senior num per 100k'
+    y = 'Case rate per 100,000',
+    color = 'Seniors per 100k'
   )
@@
 
-Based on the above data exploratory result, the following equation is used to create a regression model to determine the relationship between the case rate and demographic features including the seniors rate, the rate of poverty and the rate of unemployment.
+Based on the above data exploratory result, the following equation is used to create a regression model to determine the relationship between the case rate and demographic features including the senior rate, the rate of poverty and the rate of unemployment.
 
@@
 \begin{aligned}
- Case.Rate.Per.100K =\beta_0 &+\beta_1Senior.Rate.Per.100K +\beta_2Poverty.Rate.Per.100K +\beta_3Unemployed.Per.100K
+ case\_rate\_100k =\beta_0 &+\beta_1senior\_per\_100k +\beta_2poverty\_per\_100k +\beta_3unemployed\_per\_100k
 \end{aligned}
@@
 
-The coefficient of the senior_per_100k is -0.230970 and statistically significant, which implies that for every one thousand additional seniors per 100,000, there is a decrease of 231 COVID-19 cases per 100,000, holding other demographic features as constant.
+The coefficient of the senior rate (senior_per_100k) is -0.230970 and statistically significant, which implies that for every one thousand additional seniors per 100,000, there is a decrease of 231 COVID-19 cases per 100,000, holding other demographic features as constant.
 As discussed in mode1 1 section, although seniors are susceptible group, less mobility and social interaction with cautious self protection may play an important role in this pandemic situation.
 
-In addition, the coefficient of the poverty_per_100k (0.231750) indicates that for every one thousand additional people living under poverty line per 100,000, there is a increase of 232 COVID-19 cases per 100,000, holding other demographic features as constant.
+In addition, the coefficient of the poverty rate (poverty_per_100k) (0.231750) indicates that for every one thousand additional people living under poverty line per 100,000, there is a increase of 232 COVID-19 cases per 100,000, holding other demographic features as constant.
 People living under the poverty line have limited access to medical resources.
@@
 
-Finally, the coefficient of the unemployed_per_100k is statistically significant and negative (-0.486191), which implies that for every one thousand additional unemployed, there is a decrease of 486 COVID-19 case per 100,000, holding other demographic features as constant.
+Finally, the coefficient of the unemployment rate (unemployed_per_100k) is statistically significant and negative (-0.486191), which implies that for every one thousand additional unemployed, there is a decrease of 486 COVID-19 case per 100,000, holding other demographic features as constant.
 Although the unemployment leads to less medical insurance coverage and impact on their income, the unemployed people may have more flexibility to obey stay-at-home orders.
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
+```{r, warning=FALSE, message=FALSE}
 df<- df%>% 
@@
 
-In addition, all the variance inflation factors are less than 4 as follows, which doesn't indicate the existence of collinearity.
+In addition, all the variance inflation factors are less than 4 as follows, which doesn't indicate the existence of high collinearity.
 
@@
 
-```{r, echo=FALSE}
+```{r}
 plot(model2, which=3)
@@
 
-```{r, echo=FALSE}
+```{r}
 bptest(model2)
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
+```{r,  warning=FALSE, message=FALSE}
 plot_one <- df %>% 
@@
   geom_histogram(fill = 'skyblue', color = 'grey30', bins=30) +
-  labs(title =  "Histogram of the Model 2 Residuals", 
+  labs(title =  "Histogram and Q-Q Plot of the Model 2 Residuals", 
        x = "Model 2 residuals", y = 'Count')
@@
 
-```{r, echo =FALSE}
+```{r}
 plot(model2, which = 5)
@@
 
-For Model 3, policy related variables will be introduced for the analytics.
+For model 3, policy related variables will be introduced for the analytics.
 Policy variable "Mandate face mask use by all individuals in public spaces" (renamed as mask_public) is selected as control variable to predict the case rate.
@@
 
-There are other variables related to mask such as "No legal enforcement of face mask mandate" and "Mandate face mask use by employees in public-facing businesses", we believe these variables have a lower impact than the policy that is enforced to public.
-Also, it is very likely that if a state enforce public mask policy, they will by default enforce mask policy for employees in public-facing business.
-As a result, we believe mask policy for all individuals in public space can be used to effectively represents the policy impact.
+There are other variables related to mask such as "No legal enforcement of face mask mandate" and "Mandate face mask use by employees in public-facing businesses", it is assume these variables have a lower impact than the policy that is enforced to public.
+Also, it is observed that if a state enforce public mask policy, they will by default enforce mask policy for employees in public-facing business.
+As a result, mask policy for all individuals in public space can be used to represent mask-related policies.
 
-It is worth to note that, the mask_public is in date format, which has value of zero or a actual date when the policy is enforced.
-According to the documentation, zero represents "the absence of an order or directive", which can be interpreted as the policy is not enforced by the state explicitly.
-For linear regression, the date values in mask_public column are transformed as value 1, so that we can distinguish whether states has public mask policy or not.
-It is also noted that by transforming the variable, some important information will be lost because the actual date (early or late) to enforce the policy can also have an impact to the case rate.
-However, it is difficult to measure the time effect as different states that enforced the public mask policy may depend on related situations or different thresholds.
+It is worth to note that, the mask_public is in date format, which has value of 0 or a actual date when the policy is enforced.
+According to the documentation, 0 represents "the absence of an order or directive", which can be interpreted as the policy is not enforced by the state explicitly.
+For linear regression, the date values in mask_public column are transformed as value 1, so that   whether a state has any public mask policy or not can be distinguished.
+It is also noted that by transforming the variable, some important information will be lost because the actual date (early or late) to enforce the policy can also have an impact on case rate.
+However, it is difficult to measure the effect of the timing of implementation across different states.
+The enforcement of public mask policy will vary across states depending on other state related features such as political factors that are not included.
 
-Firstly, "mask_public_bool" column is created from "mask_public", with 1 represents that the state has a pubic mask policy, and 0 means there's no explicit public mask policy from the state.
-The summary shows that there are 35 states that enforced public mask mandate, and 16 states that had no public mask policy.
+Firstly, the mask policy (mask_public_bool) variable is created from the original variable (mask_public), with 1 representing that the state has a pubic mask policy, and 0 means there's no explicit public mask policy from the state.
+The summary shows that there are 35 states that enforced public mask mandate, and 16 states that had no public mask policy at the time the dataset was compiled.
 The ratio looks reasonable for the analysis considering sample size of 51 as there is decent number of samples for each state policy type.
@@
 
-cat('Count for states that enforced mask mandate in public: ', 
+cat('Count of states that enforced mask mandate in public: ', 
     length(df$mask_public_bool[df$mask_public_bool==1]))
 
-cat('\nCount for states did not enforced mask mandate in public: ', 
+cat('\nCount of states did not enforce mask mandate in public: ', 
     length(df$mask_public_bool[df$mask_public_bool==0]))
@@
 
-Histograms below that shows the distribution of states that has public mask policy or not. Both distributions are not very normal with relatively small skews toward both tails.
-
-```{r, message = FALSE}
-
-hist(df$case_rate_100k[df$mask_public_bool == 1], xlab='Case Rate', ylab='Count', 
-     col='skyblue', main='Case Rate - States Enforced Mask Mandate in Public')
-hist(df$case_rate_100k[df$mask_public_bool == 0], xlab='Case Rate', ylab='Count', 
-     col='skyblue', main='Case Rate  - States Enformced No Mask Mandate in Public')
-```
-
-A plot that shows the relationship between different policies vs case rate.
+The plot below shows the relationship between the mask policy and case rate.
+When comparing the states that did not implement a mask policy (value = 0) and those that did (value = 1), it can be seen that the data points for those that have no policy cluster around a higher case rate range, while the data points for the states with a mask policy cluster at a lower case rate range.
 
@@
 df %>% 
-  ggplot(aes(x = mask_public_bool, y = case_rate_100k, color = case_rate_100k)) + 
-  geom_point(aes(size=case_rate_100k)) + 
+  ggplot(aes(x = mask_public_bool, y = case_rate_100k)) + 
+  geom_point(aes(size=5, alpha=0.2), show.legend = FALSE) + 
   labs(
@@
     x = 'Mask Mandate in Public', 
-    y = 'Case Rate', 
-    color = 'Case Rate'
-  )
+    y = 'Case Rate per 100,000')  +
+  scale_x_continuous(breaks = seq(0,1,1), labels = seq(0,1,1))
 ```
@@
   labs(
-    title = 'Distribution of Population Density per Square Miles', 
+    title = 'Histogram of Population Density per Square Mile', 
     x = 'Population density per square mile', y = 'Count')
@@
 
-As can be seen from the histogram, although most of the population density is concentrated in the 0 to 1500 range, there are some grouping of outliers that are very far from this concentration.
+As can be seen from the histogram, although most of the population density clusters in the 0 to 1500 range, there are some grouping of outliers that are very far from this concentration.
 When an analysis is performed, it can be seen that there is only one data sample that is the outlier, which is D.C.
@@
   labs(
-    title = 'Distribution of Population Density per Square Miles', 
-    x = 'Log of population density per square mile', y = 'Count')
+    title = 'Histogram of Population Density per Square Mile', 
+    x = 'Population density per square mile', y = 'Count') + 
+  scale_x_continuous(breaks = seq(0, 5, 1), labels = 10^(seq(0,5,1)))
 
@@
 Two additional variables discussed in Model 3 will be introduced as the maximalist approach.
-The first variable is the "mask_public_bool" (whether the state enforced public mask mandate or not) that represents the policy effects.
-The second variable introduced is "population_density" (the variable is transformed with log of 10 as discussed earlier) which is considered an important factor related to the spread of disease.
+The first variable is the "mask_public_bool" (whether the state enforced public mask mandate or not) that represents the policy.
+The second variable introduced is "population_density" (the variable is transformed by the natural logarithm as discussed earlier) which is considered an important factor related to the spread of disease.
 
@@
 And the result suggests that public mask mandate policy has positive effect in bringing the case number down.
-In detail, the COVID-19 case rate can be reduced by 945 out of every 100k people if the state enforces public mask mandate keeping all other variables the same.
+Specifically, the COVID-19 case rate can be reduced by 945 out of every 100k people if the state enforces public mask mandate keeping all other variables the same.
 
 The population density variable however is not significant in the regression even it is considered as one of the major factors to the spread of disease.
-So according to the regression, we fail to reject the null hypothesis that population density have an direct impact to the COVID-19 case rate.
+According to the regression, there is a failure to reject the null hypothesis that population density is not related to the COVID-19 case rate.
 There could be multiple factors causing the population density to be insignificant.
@@
 As a result the numbers can be diluted for states that have major cities with high population density, but very little urbanization in the majority part of the state suburb area.
-In addition, as stated earlier in the research, Washington, D.C.
+In addition, as stated earlier in subsection 2-3-1 of this report, Washington, D.C.
 is considered as a state however the demography structure and geographical related characteristics are different from other states.
-And due to a small sample we have, individual variables that exhibit unique characteristics could lead to certain degree of bias.
+And due to the small sample size, individual variables that exhibit unique characteristics could lead to a certain degree of bias.
 
-After careful discussion, the team still decided to keep "population_density" variable in model 3 even it is insignificant for the following reasons.
-First, adding population density variable has little effect on the change of coefficients of control variables in model 2.
-Second, we don't see evidence that the variable worsens the validation of CLM assumptions compared to model 2.
-Finally, the variable do offer meaningful information that is relevant to the main research question.
+It was decided that the "population_density" variable should be kept in model 3 even it is insignificant for the following reasons.
+First, adding population density variable has little effect on the change of coefficients of control variables in model 2, which helps to demonstrate the robustness of the results.
+Second, there is no evidence that the variable effects the validation of CLM assumptions compared to model 2.
+Finally, the variable offers meaningful information that is relevant to the main research question.
 
@@
 
-The Residual vs Fitted chat shows a almost flat line that is close to zero, with the tails a bit skewed.
-There is no strong evidence that there are unincoorperated linear relationship that we are not capturing from the control variables.
-```{r}
+The residual versus fitted plot shows an almost flat line that is close to zero, with the tails a bit skewed.
+This implies that there are no strong linear relationships that are not captured beyond the current control variables.
 
-plot(model3, which=1)
+```{r, message = FALSE}
+
+df<- df%>% 
+  mutate(
+    model3_preds = predict(model3), 
+    model3_resids = resid(model3)
+  ) 
+df %>% 
+  ggplot(aes(model3_preds, model3_resids)) + 
+  geom_point() + 
+  stat_smooth()
 ```
@@
 
-Based on the Variance Inflation Factors check, all the variables have variance-inflation factors less than 4. As a result, there's no strong evidence of collinearity concern within the model. 
+First, the list of variable coefficients shows that no variables were dropped, which means there is no perfect collinearity detected by the R function.
+
 ```{r}
+model3$coefficients
+```
 
-vif(model3)>4
+Next, the variance inflation function was run in order to check for strong collinearity.
+Based on the variance inflation factors check, all the variables have variance inflation factors less than 4.
+As a result, there's no strong evidence of collinearity concern within the model.
+
+```{r}
+vif(model3)
 ```
@@
 
-According to the plot of Homoskedasticity, the standardized residual falls into a reasonable range and the curve looks smooth without concerning skews.
-There is no significant evidence that the model violates homoskedasticity assumptions.
+According to the scale-location plot, the standardized residual falls into a reasonable range and the curve looks smooth without strong deviations.
+There is no significant evidence that the model violates the homoskedasticity assumption.
+For reference, the marked data points are New Hampshire (30), New York (33), and North Dakota (35).
+
 ```{r}
@@
 
-```{r, echo=FALSE}
+```{r}
 bptest(model3)
@@
 If still a new geographic region with many states was added to the sample, and the measures of central tendency of the data continued to evolve, there would be a further indication that the underlying distribution is not identical.
-Without a possibility of refining the data sample however, as states would be immutable in this regard, the assumption must be made for any potential analysis.
+Without a possibility of refining the data sample however, as states would be immutable in this regard, the regression analysis proceeds with the recognition of possible violations of the IID assumption.
 
@@
 
-The goal of the analysis was to address the research question to examine the relationship between both the distribution of population demographics and mask-related policy of a state and the COVID-19 case rate within that states.
+The goal of the analysis was to address the research question of how the COVID-19 case rate is related to the distribution of population demographics and mask-related policy of a state.
 In order to address this question, the analysis was broken up into three phases of investigation with corresponding linear models.
 
- - Phase 1: To investigate the direct relationship between COVID-19 case rate and the population distribution of seniors within a state
+ - Phase 1: To investigate the direct relationship between COVID-19 case rate and the senior rate within a state
  - Phase 2: To build upon Phase 1 by including two additional variables of population demographics, poverty rate and unemployment rate, and investigating the overall effect on COVID-19 case rate
@@
 The data for population density was not statistically significant enough to suggest a relationship.
-Out of the four significant variables, the greatest effect was from the mask policy variable, followed by the unemployment rate, and then similar absolute effect sizes from the senior rate and povery rate.
 
-Starting with the variable with the greatest effect size, it can be seen that the implementation of a mask policy in a state leads to a decrease of 945 cases per 100,000.
+Starting with the variable with the largest coefficient magnitude, it can be seen that with the implementation of a mask policy in a state there are 945 fewer cases per 100,000 compared to a state without any implementation of such policy.
 Given that research from leading health organizations shows that the wearing of masks helps to decrease the chance of the transmission of COVID-19, it could be suggested that a mask policy is effective in increasing the adoption of wearing masks among the state population.
@@
 
-In conclusion, we see that although the population demographics had varying effects on the COVID-19 case rate, the greatest effect was seen from the mask policy.
-Going back to our original question, it can be suggested within the data that the effect of a mask-related policy is something that should not be neglected in comparison to the population demographics.
+In conclusion, going back to our original question, it can be suggested within the data that the effect of a mask-related policy is something that should not be neglected in comparison to the population demographics.
 Therefore, it seems that the COVID-19 case rate at the state-level is dependent on masks, and that ultimately, there are effective policy and control measures that can be taken by state governments that have some contribution towards combating its spread relative to factors that cannot be controlled.

### 2020-12-08 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Made grammatical and sentence structure edits across the entire document. Added some qualifying language around the poverty and unemployment rate data, and updated references of sections. Reworded variable effects to avoid causal implications

@@
 
-The research question aims to analyze the relationship between the COVID-19 case rate per 100,000 residents within each state and the population demographics features, as well as the policy decisions that were or were not put into place in order to combat the rise of COVID-19 cases.
+The research question aims to analyze the relationship between the COVID-19 case rate per 100,000 residents within each state (recorded between January 21, 2020 and October 30, 2020) and the population demographics features, as well as the policy decisions that were or were not put into place in order to combat the rise of COVID-19 cases.
 In this sense, the aim is to examine how dependent the COVID-19 case rate is on that of which cannot be controlled (i.e., population demographics), as well as that of which can be controlled to a certain degree (i.e., implementation of policies to attempt to combat case rate) by each state.
-With this information, one could suggest whether or not the proliferation of COVID-19 within a state seems to be related to either or both controllable and uncontrollable factors and how they differ.
+With this information, one could suggest whether or not the proliferation of COVID-19 within a state seems to be related to either or both controllable and uncontrollable factors, and how they differ.
 The modeling goal of this research question will be one of description, and will be broken up into three phases of investigation.
@@
 The senior percentage was selected as the key input/independent variable given that guidance had been released by the Centers for Disease Control and Prevention (CDC) that seniors belong to the age category for those at higher risk of COVID-19, and thus are a well studied and documented group.
-This is likely due to the both the fact that seniors in this age group are more likely or had greater access to be tested given their categorization and the fact that they would have a greater likelihood of exhibiting detectable COVID-19 related symptoms that would prompt them to get tested.
+This is likely due to both the fact that seniors in this age group are more likely or had greater access to be tested given their categorization, and the fact that they would have a greater likelihood of exhibiting detectable COVID-19 related symptoms that would prompt them to get tested.
 An additional rationale for measuring the senior percentage is that they may also be more likely to contract COVID-19 at lower viral loads, making them more susceptible to the virus.
@@
 
-Following this, in the second phase of investigation the analysis considers other variables of state-level population demographic features that are considered factors of susceptibility, namely the rate of poverty (defined as the percent of individuals living under the federal poverty line in 2018 in the original dataset) and the rate of unemployment (defined as the percent unemployment in 2018 in the original dataset).
-Those living below the poverty line may have less access/are unable to afford to preventative controls such as sanitation products and masks that help prevent the spread of the virus, while those who are designated as unemployed may not have the flexibility of sheltering at home as well as access to the aforementioned preventative controls.
-Alternatively, the likelihood of reduced mobility for those living under the poverty line and/or who are unemployed may also play a role in the relationship with case rate.
+Following this, in the second phase of investigation the analysis considers other variables of state-level population demographic features that are considered factors of susceptibility.
+Specifically, these features are the rate of poverty (defined as the percent of individuals living under the federal poverty line in 2018 in the original dataset) and the rate of unemployment (defined as the percent unemployment in 2018 in the original dataset).
+Those living below the poverty line may have less access/are unable to afford preventative controls such as sanitation products and masks that help prevent the spread of the virus, while those who are designated as unemployed may not have the flexibility of sheltering at home nor have access to the aforementioned preventative controls.
+Alternatively, the likelihood of reduced mobility for those living under the poverty line and/or those who are unemployed may also play a role in the relationship with case rate.
 As mentioned, these variables operationalize population metrics that may lead to greater case rates, and are already standardized as rates to account for varying absolute populations across states.
+It should be qualified that the data for poverty rate and unemployment rate are from 2018, and therefore may not provide a fully accurate reflection of the respective rates at the time that the case rate data was recorded.
 
-The tertiary and final phase of investigation includes a variable that measures the implementation of policy as a response to COVID-19, and to understand the added effect of this variable in conjunction with the first phase of investigation.
-Specifically, there is a focus on a policy that mandates the wearing of masks within the state.
-Through this, the aim is to analyze whether the implementation of a mask-related policy is related to case rate, as well as the strength of the association compared to population metrics.
-To operationalize this policy and simplify the measurements, the models only consider whether or not this policy was implemented through transformed indicator variables (1 = implemented, 0 = not implemented).
+The tertiary and final phase of investigation includes a variable that measures the implementation of policy as a response to COVID-19, and to understand the added effect of this variable in conjunction with the other phases of investigation.
+Specifically, there is a focus on a policy that mandates the wearing of masks in public spaces within the state.
+Through this, the aim is to analyze whether the implementation of a mask-related policy is related to case rate, as well as the strength of the association relative to that of population metrics.
+To operationalize this policy and simplify the measurements, the model will only consider whether or not this policy was implemented through a transformed indicator variable (1 = implemented, 0 = not implemented).
 As a result, factors of when and for how long the policy was implemented will be lost.
@@
 
-Based on the work of model 1, two extra state demographic features regarding poverty rate (percentage living under the federal poverty line in 2018) and unemployment rate (percentage of unemployed in 2018) would be included in the analysis of model 2.
+Building upon the work of model 1, two extra state demographic features regarding poverty rate (percentage living under the federal poverty line in 2018) and unemployment rate (percentage of unemployed in 2018) are included in the analysis of model 2.
 Although the increasing unemployment rate during the pandemic has been widely covered by the media, the potential pandemic causing poverty or unemployment will not be investigated here, according to the designed research question in this work.
@@
 Additionally, there is no outlier observed within the distribution, given that the range of rates fall between 7,600 and 19,700.
-Therefore, there is no further transformation in need for this variable in the following analysis process for both model 2 and model 3.
+Therefore, there is no further transformation needed for this variable in the following analysis process for both model 2 and model 3.
 
@@
 ```
-According to the histogram above, the distribution of unemployed number per 100,000 is fairly normal distribution.
+According to the histogram above, the distribution of the unemployed number per 100,000 is fairly normal.
 Overall, it is not highly skewed or heavily tailed.
@@
 
-Compared to model 1, the relationship investigated here involves a higher dimensional space. It would be challenging to observe notable variation from variables distribution plots as below.
+Compared to model 1, the relationship investigated here involves a higher dimensional space. It would be challenging to observe notable variation from the variable distribution plots as below.
 
@@
     title = 'COVID-19 case rate vs poverty rate', 
-    x = 'Number of living under poverty line per 100,000',
+    x = 'Number of people living under poverty line per 100,000',
     y = 'Case rate per 100,000', 
-    color = 'Seniors per 100k'
+    color = 'Seniors per 100,000'
   )
@@
     y = 'Case rate per 100,000',
-    color = 'Seniors per 100k'
+    color = 'Seniors per 100,000'
   )
@@
 
-Based on the above data exploratory result, the following equation is used to create a regression model to determine the relationship between the case rate and demographic features including the senior rate, the rate of poverty and the rate of unemployment.
+Based on the results of the data exploration above, the following equation is used to create a regression model to determine the relationship between the case rate and demographic features including the senior rate, the rate of poverty and the rate of unemployment.
 
@@
 
-According to the assessment of homoskedasticity of model 2, which would be discussed in the following paragraph (Model 2 Limitations - iv.Homoskedastic Errors), it is plausible to apply the classical standard errors in the t-test of coefficients here.
+According to the assessment of homoskedasticity of model 2, which would be discussed in the following subsection (Model 2 Limitations - iv.Homoskedastic Errors), it is plausible to apply the classical standard errors in the t-test of coefficients here.
 
@@
 
-The coefficient of the senior rate (senior_per_100k) is -0.230970 and statistically significant, which implies that for every one thousand additional seniors per 100,000, there is a decrease of 231 COVID-19 cases per 100,000, holding other demographic features as constant.
-As discussed in mode1 1 section, although seniors are susceptible group, less mobility and social interaction with cautious self protection may play an important role in this pandemic situation.
+The coefficient of the senior rate (senior_per_100k) is -0.230970 and is statistically significant, which implies that for every one thousand additional seniors per 100,000, there is a decrease of 231 COVID-19 cases per 100,000, holding other demographic features constant.
+As discussed in the subsection 2-1, although seniors are a susceptible group, their reduced mobility and social interactions with cautious self protection may play an important role in reducing the number of cases among seniors.
 
-In addition, the coefficient of the poverty rate (poverty_per_100k) (0.231750) indicates that for every one thousand additional people living under poverty line per 100,000, there is a increase of 232 COVID-19 cases per 100,000, holding other demographic features as constant.
-People living under the poverty line have limited access to medical resources.
-Moreover, most of them could only afford to live in the communities with poor hygiene conditions.
+In addition, the coefficient of the poverty rate (poverty_per_100k) (0.231750) indicates that for every one thousand additional people living under the poverty line per 100,000, there is an increase of 232 COVID-19 cases per 100,000, holding other demographic features constant.
+People living under the poverty line may have limited access to medical resources.
+Moreover, they may not be able to afford to live in communities with appropriate sanitary conditions.
 
-Finally, the coefficient of the unemployment rate (unemployed_per_100k) is statistically significant and negative (-0.486191), which implies that for every one thousand additional unemployed, there is a decrease of 486 COVID-19 case per 100,000, holding other demographic features as constant.
-Although the unemployment leads to less medical insurance coverage and impact on their income, the unemployed people may have more flexibility to obey stay-at-home orders.
-In addition, there is less exposure probability when there is no commuting or in-place working required for the unemployed.
+Finally, the coefficient of the unemployment rate (unemployed_per_100k) is statistically significant and negative (-0.486191), which implies that for every one thousand additional unemployed per 100,000, there is a decrease of 486 COVID-19 case per 100,000, holding other demographic features constant.
+Although unemployment leads to less medical insurance coverage and impacts income, the unemployed may have more flexibility to obey stay-at-home orders.
+In addition, there is less exposure probability given the unemployed do not commute to work or work in-person.
 
@@
 #### ii. Linear Conditional Expectation
-To assess the linear conditional expectation assumption in higher-dimensional space, the plots of predicted versus residuals of the model 2 is investigated as follows.
+To assess the linear conditional expectation assumption in a higher-dimensional space, the plot of the predicted values versus the residuals of model 2 is investigated as follows.
 
@@
 
-Overall, the residuals maintain around zero across the predicted range. There is no assumption violation phenomenon that can be observed here.
+Overall, the residuals remain around zero across the predicted range. There is no assumption violation phenomenon that can be observed here.
 
@@
 
-Since there are multiple covariates included in this model compared to the model 1, there is more concern about the no perfect collinearity assumption here.
-First, the list of variable coefficients shows that no variables were dropped, which means there is no perfect collinearity detected by the R function.
+Since there are multiple covariates included in this model compared to the model 1, there is concern about the no perfect collinearity assumption.
+First, the list of variable coefficients show that no variables were dropped, which means there was no perfect collinearity detected by the R function.
 
@@
 To evaluate the homoskedastic errors assumption, the square root of the residuals is plotted against the fitted values of model 2 (scale-location plot).
-Although, the curve is not perfectly flat, there is no obvious variance of the errors either.
+Although, the curve is not perfectly flat, there is no obvious variance in the errors either.
 For reference, the data points marked in the plot represent Florida (10), New York (33) and North Dakota (35).
@@
 #### v. Normally Distributed Errors
-To access the assumption of normality of error distribution, both the histogram and the Q-Q plot of the residuals is shown as follow.
+To access the assumption of normality of error distribution, both the histogram and the Q-Q plot of the residuals are shown as follows.
 
@@
 For model 3, policy related variables will be introduced for the analytics.
-Policy variable "Mandate face mask use by all individuals in public spaces" (renamed as mask_public) is selected as control variable to predict the case rate.
-Given it has been proved that mask is effective in preventing the transmission of disease, it is expected that the mask policy mandate in public would have an effect in reducing the case rate.
+The policy variable "Mandate face mask use by all individuals in public spaces" (renamed as mask_public) is selected as control variable to predict the case rate.
+Given it has been proven in external research that masks are effective in preventing the transmission of COVID-19, it is expected that the mask policy mandate in public would have an effect in reducing the case rate.
 
-There are other variables related to mask such as "No legal enforcement of face mask mandate" and "Mandate face mask use by employees in public-facing businesses", it is assume these variables have a lower impact than the policy that is enforced to public.
-Also, it is observed that if a state enforce public mask policy, they will by default enforce mask policy for employees in public-facing business.
-As a result, mask policy for all individuals in public space can be used to represent mask-related policies.
+There are other variables related to masks such as "No legal enforcement of face mask mandate" and "Mandate face mask use by employees in public-facing businesses".
+It is assumed these variables have a lower impact than the policy that is enforced to those in public.
+Also, it is observed that if a state enforces a public mask policy, they will also enforce a mask policy for employees in public-facing businesses.
+As a result, mask policy for all individuals in public spaces can be used to represent mask-related policies.
 
-It is worth to note that, the mask_public is in date format, which has value of 0 or a actual date when the policy is enforced.
+It is worth noting that the mask_public variable is in date format, which has either a value of 0 or an actual date when the policy was enforced.
 According to the documentation, 0 represents "the absence of an order or directive", which can be interpreted as the policy is not enforced by the state explicitly.
-For linear regression, the date values in mask_public column are transformed as value 1, so that   whether a state has any public mask policy or not can be distinguished.
-It is also noted that by transforming the variable, some important information will be lost because the actual date (early or late) to enforce the policy can also have an impact on case rate.
+For the linear regression, the date values in the mask_public variable are transformed into a value of 1, so that whether or not a state has a public mask policy can be distinguished.
+It is also noted that by transforming the variable, some important information will be lost because the actual date (early or late) of enforcement for the policy can also have an impact on case rate.
 However, it is difficult to measure the effect of the timing of implementation across different states.
-The enforcement of public mask policy will vary across states depending on other state related features such as political factors that are not included.
+Additionally, the enforcement of a public mask policy will vary across states depending on other state related features such as political factors that are not included within this analysis.
 
-Firstly, the mask policy (mask_public_bool) variable is created from the original variable (mask_public), with 1 representing that the state has a pubic mask policy, and 0 means there's no explicit public mask policy from the state.
-The summary shows that there are 35 states that enforced public mask mandate, and 16 states that had no public mask policy at the time the dataset was compiled.
-The ratio looks reasonable for the analysis considering sample size of 51 as there is decent number of samples for each state policy type.
+Firstly, the mask policy (mask_public_bool) variable is created from the original variable (mask_public), with 1 representing that the state has a pubic mask policy, and 0 meaning there is no explicit public mask policy in the state.
+The output below shows that there are 35 states that enforced a public mask mandate and 16 states that did not at the time the dataset was compiled.
+This ratio looks reasonable enough for analysis, considering that there is a decent number of samples within each group from the original sample size of 51.
 
@@
 
-cat('Count of states that enforced mask mandate in public: ', 
+cat('Count of states that enforced a mask mandate in public spaces: ', 
     length(df$mask_public_bool[df$mask_public_bool==1]))
 
-cat('\nCount of states did not enforce mask mandate in public: ', 
+cat('\nCount of states did not enforce a mask mandate in public spaces: ', 
     length(df$mask_public_bool[df$mask_public_bool==0]))
@@
 
-```{r}
+```{r, message=FALSE, warning=FALSE}
 df %>% 
@@
   scale_x_continuous(breaks = seq(0,1,1), labels = seq(0,1,1))
+
+df %>%
+  ggplot(aes(x=mask_public_bool, y=case_rate_100k))+
+  geom_jitter(shape=21, size=6, aes( fill = factor(mask_public_bool)), width=0.1)+
+  theme(legend.position = 'none') +
+  labs(
+    title = 'Case Rate vs Mask Public Mandate',
+    x = 'Mask Mandate in Public',
+    y = 'Case Rate per 100,000',
+    color = 'Case Rate'
+  ) +
+  scale_x_continuous(breaks = seq(0,1,1), labels = seq(0,1,1))
+
+df %>%
+  ggplot(aes(case_rate_100k))+
+  geom_histogram(aes(y = ..density.., fill = factor(mask_public_bool)), position = 'identity', alpha = 0.5, color = 'white')+
+  stat_density(geom = 'line', size=1, position = 'identity', aes(color = factor(mask_public_bool)))+
+  scale_fill_brewer(palette = 'Dark2') +
+  scale_color_brewer(palette = 'Dark2') +
+  labs(
+    title = 'Distribution of COVID-19 Case Rate',
+    x = 'Covid-19 case rate per 100,000'
+  ) +
+  scale_fill_discrete(name='Mask public policy',labels=c('No', 'Yes'))+
+  scale_color_discrete(name='Mask public policy',labels=c('No', 'Yes'))
 ```
@@
 
-As can be seen from the histogram, although most of the population density clusters in the 0 to 1500 range, there are some grouping of outliers that are very far from this concentration.
-When an analysis is performed, it can be seen that there is only one data sample that is the outlier, which is D.C.
+As can be seen from the histogram, although most of the population density clusters in the 0 to 1500 range, there is some grouping of outliers that are very far from this concentration.
+When an analysis is performed, it can be seen that there is only one data sample outlier, which is D.C.
 with a value of 11,496.
 This is given due to the fact that D.C.
-is a district that solely consists of a large city, as mentioned as a possibility in the introduction.
+is a district that solely consists of a large city.
 Given that this causes the data to be skewed, the logarithm is taken to scale the variable.
-Alternatively, the data point could have been dropped from the sample, but given the already small sample size, it was determined that a better approach would be to keep it within the sample.
+Alternatively, the data point could have been dropped from the sample, but given the already small sample size, it was determined that a better approach would be to keep it as a data point.
 Once this transformation was performed, it is shown that there is a relatively normal distribution of population densities (see figure below).
@@
 Next, the regression model for model 3 is created based on the existing work from model 2.
-Two additional variables discussed in Model 3 will be introduced as the maximalist approach.
-The first variable is the "mask_public_bool" (whether the state enforced public mask mandate or not) that represents the policy.
+Two additional variables discussed in model 3 will be introduced given the maximalist approach for model 3.
+The first variable is the "mask_public_bool" (whether the state enforced a public mask mandate or not) that represents the policy.
 The second variable introduced is "population_density" (the variable is transformed by the natural logarithm as discussed earlier) which is considered an important factor related to the spread of disease.
 
-The regression coefficient shows that Mask_public_bool variable is highly significant.
-And the result suggests that public mask mandate policy has positive effect in bringing the case number down.
-Specifically, the COVID-19 case rate can be reduced by 945 out of every 100k people if the state enforces public mask mandate keeping all other variables the same.
+The regression coefficient shows that mask_public_bool variable is highly significant.
+The result also suggests that the public mask mandate policy has a negative relationship with the case rate.
+Specifically, the COVID-19 case rate decreases by 945 out of every 100,000 people if the state enforces a public mask mandate, holding all else constant.
 
-The population density variable however is not significant in the regression even it is considered as one of the major factors to the spread of disease.
+Conversely, the population density variable is not significant in the regression even though it is considered as a major factor in stymieing the spread of the virus.
 According to the regression, there is a failure to reject the null hypothesis that population density is not related to the COVID-19 case rate.
@@
 For example, the population density is calculated by population divided by total square miles.
-As a result the numbers can be diluted for states that have major cities with high population density, but very little urbanization in the majority part of the state suburb area.
-In addition, as stated earlier in subsection 2-3-1 of this report, Washington, D.C.
-is considered as a state however the demography structure and geographical related characteristics are different from other states.
-And due to the small sample size, individual variables that exhibit unique characteristics could lead to a certain degree of bias.
+As a result, the numbers can be diluted for states that have major cities with high population density, but very little urbanization in the majority of the rural area.
+In addition, as stated earlier in subsection 2-3-1 of this report, D.C.
+is considered as a state despite the fact that the demographic and geographical characteristics differ from other states.
+And due to the small sample size, individual variables such as this that exhibit unique characteristics could lead to a certain degree of bias.
 
-It was decided that the "population_density" variable should be kept in model 3 even it is insignificant for the following reasons.
-First, adding population density variable has little effect on the change of coefficients of control variables in model 2, which helps to demonstrate the robustness of the results.
-Second, there is no evidence that the variable effects the validation of CLM assumptions compared to model 2.
+It was decided that the "population_density" variable should be kept in model 3 despite its insignificance for several reasons.
+First, adding population density variable has little effect on the coefficients of the control variables in model 2, which helps to demonstrate the robustness of the results.
+Second, there is no evidence that the variable effects the validation of the CLM assumptions compared to those in model 2.
 Finally, the variable offers meaningful information that is relevant to the main research question.
@@
 
-The residual versus fitted plot shows an almost flat line that is close to zero, with the tails a bit skewed.
+The residual versus fitted plot shows an almost flat line that is close to zero, with the tails slightly skewed.
 This implies that there are no strong linear relationships that are not captured beyond the current control variables.
@@
 
-First, the list of variable coefficients shows that no variables were dropped, which means there is no perfect collinearity detected by the R function.
+First, the list of variable coefficients shows that no variables were dropped, which means there was no perfect collinearity detected by the R function.
 
@@
 Next, the variance inflation function was run in order to check for strong collinearity.
-Based on the variance inflation factors check, all the variables have variance inflation factors less than 4.
-As a result, there's no strong evidence of collinearity concern within the model.
+Upon examination of the ouptut below, it can be seen that all the variables have variance inflation factors of less than 4.
+As a result, there is no strong evidence of collinearity within the model.
 
@@
 
-According to the scale-location plot, the standardized residual falls into a reasonable range and the curve looks smooth without strong deviations.
+According to the scale-location plot, the standardized residuals fall into a reasonable range and the curve looks smooth without strong deviations.
 There is no significant evidence that the model violates the homoskedasticity assumption.
@@
 ```
+
 Additionally, a Breusch-Pagan test is run to check the level of heteroskedasticity.
@@
 
-To access the assumption of normality of error distribution, both the histogram and the Q-Q plot of the residuals is shown as follow.
+To assess the assumption of normality of error distribution, both the histogram and the Q-Q plot of the residuals are shown as follows.
 
@@
 There is a fairly normal distribution that can be seen from the histogram above.
-From the Q-Q plot, the points generally stay close to the normal line, with the tails exhibiting a certain degree of skew.
+In the Q-Q plot, the points generally stay close to the normal line, with the tails exhibiting a certain degree of skew.
 However, in general, it implies no deviation from normality.
@@
 
-The most general requirement for all the linear models is that data points be independent and identically distributed (IID) which they are built off of.
+The most general requirement for all the linear models is that data points be independent and identically distributed (IID).
 While for the most part the events of one state will primarily affect that single state, travel by individuals spreading the virus across state lines is an inevitable occurrence.
-This may cause the spread of a virus within one state to be influenced by those around it, or generally by its position in the country.
-In addition, all states are involved in the same market economy system of the country to varying degrees, even though there are economic performance divisions across the states.
+This may cause the spread of a virus within one state to be influenced by those around it, or generally by its geographical location within the country.
+In addition, all states are involved in the same market economy system to varying degrees, even though there are economic performance divisions across the states.
 Likewise, certain regions of the country may also have similar demographics due to a shared history across general geographic regions.
 While this is acknowledged to occur, the effect is likely minor compared to the policy data collected state by state which would affect the spread of the virus.
-Therefore it is relatively safe for this assumption to be met.
+Therefore it is relatively safe to reason that this assumption has been met.
 
 The identical distribution of states as individual samples in the model is also questionable.
-Similar to how connections or commonalities between states make them less than truly independent, certain factors would also influence the distribution so that some states are representative of a different underlying distribution than one another.
+Similar to how connections or commonalities between states make them less than truly independent, certain factors would also influence the distribution so that some states are representative of a different underlying distribution compared to others.
 For example, if one geographic region as a whole has a factor which influences the spread of the virus differently than another, multiple underlying distributions may be present in the sample.
 If still a new geographic region with many states was added to the sample, and the measures of central tendency of the data continued to evolve, there would be a further indication that the underlying distribution is not identical.
-Without a possibility of refining the data sample however, as states would be immutable in this regard, the regression analysis proceeds with the recognition of possible violations of the IID assumption.
+Without any possibility of refining the data sample however, as states would be immutable in this regard, the regression analysis proceeds with the recognition of possible violations in the IID assumption.
 
-In examining the results of various diagnostics across different models, generally the CLM assumptions are better met as the model continues to develop.
+In examining the results of various diagnostics across different models, the CLM assumptions are generally better met as variables continue to be added.
 In the first model, deviations away from certain assumptions are the strongest.
@@
 This is further improved upon in the third model, where local variations in the linear conditional expectation diagnostic plot are generally smoother and more centered around zero.
-The assumption with the greatest exception to this is that of no high collinearity within the models, as each coefficient in each model was shown to have VIF values not near the cut off value of four.
+The assumption with the greatest exception to this is that of no high collinearity within the models, as each coefficient in each model was shown to have VIF values far below the cut off value of four.
 
-However, when adding more variables in the third model, other assumptions begin to be less well met.
+However, when adding additional variables in the third model, other assumptions begin to be less well met.
 For example, the normality of the error term in the model is less well fit than it was in the earlier models.
-It can be seen in the third model's Q-Q plot how at fitted values towards the edges of the model's range, there exists the greatest deviation away from normality.
-This assumption is generally less important than that of the linear conditional expectation, however, which is a fundamental part of CLM.
+It can be seen in the third model's Q-Q plot that for fitted values at the edges of the model's range, there exists the greatest deviation away from normality.
+This assumption is generally less important than that of the linear conditional expectation, however, which is a fundamental part of the CLM.
 Therefore, the third model can still be considered as generally improving the overall analysis.
 
-In addition, with more variables in the third model, outliers appear to have more of an impact than previously.
-This is likely due to the limited number of data points the model is based off of, an entire state only constitutes an individual datapoint.
-With less degrees of freedom in the third model, certain points can be seen to more closely approach the boundaries of the Cook's length plot, although they are still not extreme enough to fully be considered an outlier.
+In addition, with more variables in the third model, outliers appear to have more of an impact than they did previously.
+This is likely due to the limited sample size within the models.
+With less degrees of freedom in the third model, certain points can be seen to more closely approaching the boundaries of the Cook's distance plot, although they are still not extreme enough to be considered outliers.
 Increasing the resolution of the data, such as examining COVID-19 cases at the county or city/town level, would help to reduce this by increasing the overall degrees of freedom.
@@
 Between model 1, 2, and 3, both the R2 and stricter adjusted R2 value increases as variables are added.
-The R2 value reflects the amount of variation that can be explained in the model on a 0-1 scale, while the adjusted R2 value does the same but also takes into account the number of variables measured (as can be seen in the decreasing degrees of freedom, df).
-The adjusted R2 value jumps from 0.082 initially in model 1 to a value of 0.267 in model 2, which is more than a 200% increase.
-The value again increases, albeit by a smaller amount, between model 2 and a value of 0.376 in model 3, which represents a 40% jump.
+The R2 value reflects the amount of variation that can be explained in the model on a 0-1 scale, while the adjusted R2 value does the same but also takes into account the number of variables measured (as can be seen in the decreasing degrees of freedom, or df).
+The adjusted R2 value jumps from a value of 0.082 in model 1 to 0.267 in model 2, which is more than a 200% increase.
+The value again increases, albeit by a smaller amount, from model 2 to a value of 0.376 in model 3, which represents a 40% jump.
 When the R2 value is not adjusted, this value increases to 0.439 for model 3.
@@
 The F-statistic is used to test whether or not to use the full model (inclusive of all input variables) or a reduced one with fewer variables.
-Given that the null hypothesis for this test is that the reduced model should be over the full model, this can be safely rejected in favor of the full model.
+Given that the null hypothesis for this test is that the reduced model should be used over the full model, this can be safely rejected in favor of the full model.
 Additionally, it can be seen that the standard errors for the significant variables are relatively small and are consistent across the models.
@@
 
-In model 1, the coefficient of -0.174 for the senior rate is significant with a p-value of 0.0234 (per the regression results from the Model 1 section).
-When the covariate variables of unemployment rate and poverty rate are added in to model 2, the coefficient for senior rate adjusts to a greater magnitude of -0.231 while the p-value appreciably decreases to a value of 0.0015 (per the regression results from the Model 2 section).
-This explains that not only has the magnitude of effect of the senior rate on case rate increased, but that this increased effect is also more statistically significant.
+In model 1, the coefficient of -0.174 for the senior rate is significant with a p-value of 0.0234 (per the regression results from subsection 2-1-2).
+When the covariate variables of unemployment rate and poverty rate are added in to model 2, the coefficient for senior rate adjusts to a greater magnitude of -0.231 while the p-value appreciably decreases to a value of 0.0015 (per the regression results from subsection 2-2-2).
+This explains that not only has the magnitude of the relationship between senior rate and case rate increased, but that the relationship is also more statistically significant.
 In addition, the coefficients of the covariate variables themselves are also highly significant with p-values of 0.0005 and 0.0059 for poverty rate and unemployment rate respectively.
-The magnitude of the effect of poverty rate is almost equivalent to that of the senior rate, with a coefficient of 0.232, while the unemployment rate has stronger negative coefficient of -0.486 (roughly twice the magnitude of the other two variables).
+The magnitude of the coefficient of poverty rate is almost equivalent to that of the senior rate, with a coefficient of 0.232, while the unemployment rate has stronger negative coefficient of -0.486 (roughly twice the magnitude of the other two variables).
 
-Moving onto model 3, the magnitude of the coefficient for senior rate increases once again to a value of -0.242 while the statistical significance increases with a decreased p-value of 0.0004 (per the regression results from the Model 3 section).
+Moving onto model 3, the magnitude of the coefficient for senior rate increases once again to a value of -0.242, while the statistical significance increases with a decreased p-value of 0.0004 (per the regression results from the subsection 2-3-2).
 As highlighted before, the more complex model does a better job of predicting the effect of the senior rate variable.
@@
 There are also detectable changes in the coefficient values, as the magnitude for both decreases to 0.197 and -0.364 respectively.
-The changes in the p-values are likely a direct result of the added variables of mask policy and the logarithmic transform of population density, where these variables introduce some additional variance in the model 2 covariates.
-Notably, the coefficient for the mask policy covariate is significant with a p-value of 0.0027 and has by far the greatest effect on the outcome variable with a coefficient of -945.6.
+The changes in the p-values are likely a direct result of the added variables of mask policy and the population density, where these variables introduce some additional variance in the model 2 covariates.
+Notably, the coefficient for the mask policy covariate is significant with a p-value of 0.0027 and has by far the largest magnitude of coefficient, with a value of -945.6.
 The population density covariate is not significant with a p-value of 0.3369.
 
-In terms of practical significance, there is great discrepancy between the effect of the mask policy variable compared to the other three significant features.
+In terms of practical significance, there is great discrepancy between the coefficient of the mask policy variable compared to the other three significant features.
 With a coefficient of -945.6, this means that given the binary mask policy variable, when a mask policy is implemented (value of 1), there is an decrease in case rate of roughly 945 per 100,000 compared to when the policy is not in place.
-The next largest effect is from the unemployment rate variable, where in the final model, the coefficient of -0.364 translates to a decrease of roughly 1 case per 100,000 with every 3 additional people unemployed per 100,000.
-Similarly, the effects of both poverty rate and senior rate are nominal.
+The next largest coefficient magnitude is from the unemployment rate variable, where in the final model, the coefficient of -0.364 translates to a decrease of roughly 1 case per 100,000 with every 3 additional people unemployed per 100,000.
+Similarly, the coefficients of both poverty rate and senior rate are relatively small.
 For every 5 additional people per 100,000 who are below the poverty line, there is roughly 1 additional case per 100,000, and for every 4 additional seniors per 100,000 within the state population, there is roughly 1 fewer case per 100,000.
@@
 
-Omitted variables from the model may affect both the internal specifications of included variables as well as the general applicability of the model to explain the output variable.
+Omitted variables from the model may affect both the internal specifications of included variables as well as the general applicability of the model in explaining the output variable.
 For this analysis, the omitted variables mentioned are hypothetical.
-Therefore, without data to include in an follow up analysis with the omitted variable, only the direction of the bias may be estimated rather than the size effect as well.
+Therefore, without the data available to conduct a follow-up analysis on the omitted variables, only the direction of the bias may be estimated, as opposed to the actual effect size.
 
@@
 One of the variables that was hypothesized as being potentially influential is a ratio of a state's urban to rural population.
-Currently, the dataset contains population information as averaged across each state as a whole.
-While population density is potentially useful, it was found to not be significant in the models.
-This may be due to states which have large urban centers, where population density would be high, losing their impact due to also containing large areas of land that is sparsely populated.
-One example of this would be New York, which contains several of the most densely populated counties in the country, but also large rural areas.
+Currently, the dataset contains population density information as averaged across each state as a whole.
+While population density is potentially useful, it was found to not be significant in model 3.
+This may be due to states having large urban centers, where population density would be high, that lose their impact due to the state also containing large areas of land that is sparsely populated.
+One example of this would be New York, which contains several of the most densely populated counties in the country, but also has large rural areas.
 
 An urban to rural ratio would be explanatory of the population density variable as a whole, where a higher urban to rural ratio would result in a higher statewide population density generally.
-With a greater fraction of urban residents, people would also be less likely to be able to social distance as effectively, increasing the per capita case rate.
+With a greater fraction of urban residents, people would also be less likely to be able to social distance as effectively, increasing the case rate.
 Therefore, the coefficient for population density in the original model would be larger and more positive than a model where this ratio is included, resulting in a bias that is positive and away from zero.
-Assuming population density as an effective proxy for an urban to rural ratio variable, however, the original effect described in the models can be assumed to be real.
+Assuming population density is an effective proxy for an urban to rural ratio variable, however, the original effect described in the models can be assumed to be real.
 
@@
 
-Knowledge of the median income for a state would give an indication of the relative job status for a working person in that state.
-This would be important for the case rate to understand about how many people work in low paying jobs that are more frequently in close contact with others (such as the service industry, gig industry, etc.).
+Knowledge of the median income for a state would give an indication of the types of jobs and earnings available to the labor force within that state.
+It could be relevant to case rate to understand how many people work in low paying jobs and thus may more frequently be in close contact with others (such as the service industry, gig industry, etc.).
 Higher median income may imply that more workers are able to function remotely with greater social distancing.
@@
 It would also be explanatory in a way that makes the poverty percentage variable less impactful, where either of these variables may be considered as proxies for one another.
-Because inclusion of the median income variable would led to a lower magnitude poverty percentage variable, and that the coefficient of the poverty percentage variable is positive, this would cause a positive omitted variable bias away from zero.
-Similar to the first omitted variable, however, poverty percentage may also be assumed to proxy median income in that they both describe the economic state of a location and what that implies for its population.
+Because inclusion of the median income variable would lead to a lower magnitude poverty percentage variable, and the coefficient of the poverty percentage variable is positive, this would cause a positive omitted variable bias away from zero.
+Similar to the first omitted variable, however, poverty percentage may also be assumed to be a proxy for median income in that they both describe the economic state of a state and what that implies for its population.
 Because of this, the original effect in the model can be assumed to be real.
@@
 
-The ratio of gender between women and men may impact the COVID case rate in that men are more susceptible to viral infections and generally have lower life expectancy from lifestyle differences.
-Greater susceptibility may relate to the case rate in that men may more often become infected when exposed to COVID under similar circumstances as a women.
-Their shorter life expectancy may also impact the case rate in that more COVID cases may result in greater or more serious symptoms, with less occurrences of an infection that is asymptomatic or goes undetected.
-In either of these cases, a higher fraction of men among the population would lead to an overall higher COVID case rate.
+The ratio of gender between women and men may impact the COVID case rate in that, according to external studies, men are more susceptible to viral infections and generally have lower life expectancy and are less healthy as a result of lifestyle differences.
+Greater susceptibility may relate to the case rate in that men may more often become infected when exposed to COVID under similar circumstances as women.
+Poorer health may also impact the case rate in that a greater proportion of the COVID cases among men may be symptomatic, with fewer occurrences of asymptomatic cases that go undetected.
+In either of these cases, a higher fraction of men among the population would lead to a higher COVID case rate overall.
 
@@
 This would be a negative omitted variable bias towards zero.
-With a negative bias, and no existing variable in the dataset that would serve as a good proxy for the gender ratio, the effects in the original model can be assumed to be real.
+With a negative bias, and no existing variable in the dataset serving as a good proxy for gender ratio, the effects in the original model can be assumed to be real.
 
 However, the magnitude of this change is likely to not be large due to other causal mechanisms between seniors and COVID case rate.
-While men may more frequently present COVID symptoms, older people in general are also more susceptible to the virus due to their physical condition and would also present more symptoms than non seniors.
-When comparing the two, generally seniors are considered as more at-risk of COVID complications than men across all ages, and so the senior percentage would still have a greater impact on the overall COVID case rate than the gender ratio.
+While men may more frequently present COVID symptoms, older people in general are also more susceptible to the virus due to their physical condition and would also present more symptoms than non-seniors.
+When comparing the two, seniors are generally considered to be more at-risk of COVID complications than men across all ages, and so the senior percentage would still have a greater impact on the overall COVID case rate relative to gender ratio.
 
@@
 
-Currently, the model incorporates state policy around masks with a variable indicating whether or not a state mandated a mask policy for the general public.
-While this is useful for describing an action a state government can make and its affect on the case rate, it doesn't full capture whether how much the population actively wears a mask to drive down transmission.
-Instead, a variable which would indicate the fraction of the population actively wearing a mask would better capture the mechanism of how the virus is transmitted among people.
+Currently, the model incorporates state policy around masks, with a variable indicating whether or not a state mandated a mask policy for public spaces.
+While this is useful for describing an action a state government can take against case rate, it does not fully capture the compliance with this policy among the population.
+Instead, a variable which would indicate the fraction of the population that actively wears masks would better capture the mechanism of how the virus is transmitted among people.
 
-A variable capturing the mask use among the population would likely reduce the explanatory power of the state mask mandate variable to be less negative.
+A variable capturing the real mask use among the population would likely reduce the explanatory power of the state mask mandate variable to be less negative.
 This would cause the bias of the omitted variable to be negative and towards from zero.
-It is also reasonable to assume that state mask policy may effectively enough proxy for the mask use among the population, in which case the original effect in the model may be assumed to be real.
+It is also reasonable to assume that state mask policy may be an effective enough proxy for real mask use among the population, in which case the original effect in the model may be assumed to be real.
 
@@
 
-The economic standing of a state not only affects the individual households of its population, but also includes how it interacts with other states and how its peoples' lives might have been economically changed by the pandemic.
-One way that a state may increase its number of COVID cases is both from tourists or others visiting during the pandemic and bringing disease, but also if its own population must work in close contact with others serving them in the tourism industry.
-In addition, if certain key industries to a state, such as tourism, layed off many workers, they would likely have to take lower paying close contact jobs such as in the service or gig economy in order to continue having a livelihood.
-This would further lead to a higher COVID case rate
+The economic standing of a state not only affects the individual households within its population, but may also reflect the economic impact to  their lives as a result of the pandemic.
+One way that a state may suffer from an increase in the number of COVID cases is from tourists or itinerant workers traveling during the pandemic and spreading the virus.
+A by-product of this is that residents of the state who work in the tourism industry may be in close contact with the travelers.
+In addition, if certain key industries in a state, such as tourism, laid off many workers, they would likely have to take lower paying service jobs in the gig economy that force them to be in close contact with others in order to continue to have a livelihood.
+This would further lead to a higher COVID case rate within the state.
 
-These effects may be included with a variable that demonstrates how much of a state's GDP is from tourism.
+These effects may be captured by a variable that demonstrates how much of a state's GDP is derived from tourism.
 It would be related to other economic indicator variables, such as the percent of the population living in poverty.
@@
 This suggests that as unemployment increases, the case rate decreases, which could be a result of the fact that those who are unemployed may be less mobile as they are not commuting into work, lowering the likelihood of contracting the virus.
-Additionally, at the time that this data was collected, the Coronavirus Aid, Relief, and Economic Security (CARES) Act was still supporting the unemployed with relief payments so there was less urgency to be out looking for work.
 Following unemployment, when examining the poverty variable there is actually an increase of roughly 1 case per 100,000 for every 5 additional people per 100,000 living under the federal poverty line.
-Although there is some intuition behind the unemployment and poverty rates having similar effects on case rate, it can be seen from the test of collinearity in the Model 2 section that these two variables are not explaining the same effects.
+Although there is some intuition behind the unemployment and poverty rates having similar effects on case rate, it can be seen from the test of collinearity in subsection 2-2-3 that these two variables are not explaining the same effects.
 This difference is likely due to the fact that those living below the poverty line have less access to suitable living conditions and sanitary goods that help to alleviate the possibility of COVID-19 infection.

### 2020-12-08 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
model 2

@@
 df %>% 
-  ggplot(aes(x = mask_public_bool, y = case_rate_100k)) + 
+  ggplot(aes(x = factor(mask_public_bool), y = case_rate_100k)) + 
   geom_point(aes(size=5, alpha=0.2), show.legend = FALSE) + 
@@
     y = 'Case Rate per 100,000')  +
-  scale_x_continuous(breaks = seq(0,1,1), labels = seq(0,1,1))
+  scale_x_discrete(labels=c("0","1"))
 
 df %>%
-  ggplot(aes(x=mask_public_bool, y=case_rate_100k))+
-  geom_jitter(shape=21, size=6, aes( fill = factor(mask_public_bool)), width=0.1)+
-  theme(legend.position = 'none') +
+  ggplot(mapping = aes(factor(mask_public_bool), case_rate_100k))+
+  geom_jitter(shape=21, size=6, aes( fill = factor(mask_public_bool), alpha=0.5), width=0.1)+
+  theme(legend.position = "none") +
   labs(
@@
     x = 'Mask Mandate in Public', 
-    y = 'Case Rate per 100,000',
-    color = 'Case Rate'
+    y = 'Case Rate'
   ) +
-  scale_x_continuous(breaks = seq(0,1,1), labels = seq(0,1,1))
+  scale_x_discrete(labels=c("No","Yes"))
 

### 2020-12-08 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Updated GDP Tourism section.

--- 2020-12-08 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
@@
 
- **Initial reassignment of common variables used across models:**
+ **Initial reassignment of common data used across models:**
 ```{r setup, include=FALSE, warning=FALSE, message=FALSE}
@@
          unemployed_pct = "Percent.Unemployed..2018.",
-         senior_pct = "X65."
+         senior_pct = "X65.",
+         population_2018 = "Population.2018" ## extra variable in test
          ) %>%
   select(State, case_rate_100k, population_density, mask_public, poverty_pct, 
-         unemployed_pct, senior_pct)
-
+         unemployed_pct, senior_pct, population_2018)
+# Assign correct data types
 head(df)
@@
 
-Next, the distribution of the senior percentage transformed into the rate per 100,000 is examined.
+Next, the distribution of the senior percentage transformed into the rate per 100,000 (as discussed in the Model Variables section) is examined.
 
@@
   geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(10000,21000,1000)) +
-  labs(title = 'Histogram of Senior Rate', 
+  labs(title = 'Histogram of Seniors Distribution', 
   x = 'Number of seniors per 100,000', y = 'Count') + 
@@
 
-```{r, message = FALSE}
-
-df<- df%>% 
-  mutate(
-    model1_preds = predict(model1), 
-    model1_resids = resid(model1)
-  ) 
-df %>% 
-  ggplot(aes(model1_preds, model1_resids)) + 
-  geom_point() + 
-  stat_smooth()
+```{r}
+# Check for linearty in model 1
+plot(model1, which=1)
 ```
@@
 It is also difficult to explain the complexity behind the COVID-19 case rate with just a single variable that does not fully capture the broader population demographics, so there is hope that given the addition of other input variables within models 2 and 3, the relationship within the data will be better explained and result in more linearity.
+For reference, the marked data points are New York (33), North Dakota (35), and South Dakota (42).
 
@@
   geom_histogram(fill = 'skyblue', color = 'grey30', bins=30) +
-  labs(title =  "Histogram and Q-Q Plot of the Model 1 Residuals", 
+  labs(title =  "Histogram of the Model 1 Residuals", 
        x = "Model 1 residuals", y = 'Count')
@@
 Based on the work of model 1, two extra state demographic features regarding poverty rate (percentage living under the federal poverty line in 2018) and unemployment rate (percentage of unemployed in 2018) would be included in the analysis of model 2.
-Although the increasing unemployment rate during the pandemic has been widely covered by the media, the potential pandemic causing poverty or unemployment will not be investigated here, according to the designed research question in this work.
-
-Due to the fact that both poverty rate and unemployment rate data represents the status in 2018, it is proper to assume that they represent the pre-existing state demographic characteristics instead of the capturing any influence from the pandemic.
-
 In addition, both newly-introduced percentage variables would be transformed into rates per 100,000 to align with the case rate variable.
@@
   labs(title =  "Histogram of Poverty Rate", 
-       x = "Number of people living under poverty line per 100,000", y = 'Count')+
+       x = "Number of living under poverty line per 100,000", y = 'Count')+
   scale_x_continuous(breaks=seq(6000, 21200, 1600))
@@
 
-Compared to model 1, the relationship investigated here involves a higher dimensional space. It would be challenging to observe notable variation from variables distribution plots as below.
+Compared to model 1, the relationship investigated here is about variables space with higher dimensions.
+It would be challenging to observe notable variation from variables distribution plots as below.
 
@@
     x = 'Number of living under poverty line per 100,000',
-    y = 'Case rate per 100,000', 
-    color = 'Seniors per 100k'
+    y = 'COVID-19 case rate per 100,000', 
+    color = 'Senior num per 100k'
   )
@@
     x = 'Number of unemployed per 100,000', 
-    y = 'Case rate per 100,000',
-    color = 'Seniors per 100k'
+    y = 'COVID-19 case rate per 100,000',
+    color = 'Senior num per 100k'
   )
@@
 
-Based on the above data exploratory result, the following equation is used to create a regression model to determine the relationship between the case rate and demographic features including the senior rate, the rate of poverty and the rate of unemployment.
+Based on the above data exploratory result, the following equation is used to create a regression model to determine the relationship between the case rate and demographic features including the seniors rate, the rate of poverty and the rate of unemployment.
 
@@
 \begin{aligned}
- case\_rate\_100k =\beta_0 &+\beta_1senior\_per\_100k +\beta_2poverty\_per\_100k +\beta_3unemployed\_per\_100k
+ Case.Rate.Per.100K =\beta_0 &+\beta_1Senior.Rate.Per.100K +\beta_2Poverty.Rate.Per.100K +\beta_3Unemployed.Per.100K
 \end{aligned}
@@
 
-The coefficient of the senior rate (senior_per_100k) is -0.230970 and statistically significant, which implies that for every one thousand additional seniors per 100,000, there is a decrease of 231 COVID-19 cases per 100,000, holding other demographic features as constant.
+The coefficient of the senior_per_100k is -0.230970 and statistically significant, which implies that for every one thousand additional seniors per 100,000, there is a decrease of 231 COVID-19 cases per 100,000, holding other demographic features as constant.
 As discussed in mode1 1 section, although seniors are susceptible group, less mobility and social interaction with cautious self protection may play an important role in this pandemic situation.
 
-In addition, the coefficient of the poverty rate (poverty_per_100k) (0.231750) indicates that for every one thousand additional people living under poverty line per 100,000, there is a increase of 232 COVID-19 cases per 100,000, holding other demographic features as constant.
+In addition, the coefficient of the poverty_per_100k (0.231750) indicates that for every one thousand additional people living under poverty line per 100,000, there is a increase of 232 COVID-19 cases per 100,000, holding other demographic features as constant.
 People living under the poverty line have limited access to medical resources.
@@
 
-Finally, the coefficient of the unemployment rate (unemployed_per_100k) is statistically significant and negative (-0.486191), which implies that for every one thousand additional unemployed, there is a decrease of 486 COVID-19 case per 100,000, holding other demographic features as constant.
+Finally, the coefficient of the unemployed_per_100k is statistically significant and negative (-0.486191), which implies that for every one thousand additional unemployed, there is a decrease of 486 COVID-19 case per 100,000, holding other demographic features as constant.
 Although the unemployment leads to less medical insurance coverage and impact on their income, the unemployed people may have more flexibility to obey stay-at-home orders.
@@
 
-```{r, warning=FALSE, message=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 df<- df%>% 
@@
 
-In addition, all the variance inflation factors are less than 4 as follows, which doesn't indicate the existence of high collinearity.
+In addition, all the variance inflation factors are less than 4 as follows, which doesn't indicate the existence of collinearity.
 
@@
 
-```{r}
+```{r, echo=FALSE}
 plot(model2, which=3)
@@
 
-```{r}
+```{r, echo=FALSE}
 bptest(model2)
@@
 
-```{r,  warning=FALSE, message=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE}
 plot_one <- df %>% 
@@
   geom_histogram(fill = 'skyblue', color = 'grey30', bins=30) +
-  labs(title =  "Histogram and Q-Q Plot of the Model 2 Residuals", 
+  labs(title =  "Histogram of the Model 2 Residuals", 
        x = "Model 2 residuals", y = 'Count')
@@
 
-```{r}
+```{r, echo =FALSE}
 plot(model2, which = 5)
@@
 
-For model 3, policy related variables will be introduced for the analytics.
+For Model 3, policy related variables will be introduced for the analytics.
 Policy variable "Mandate face mask use by all individuals in public spaces" (renamed as mask_public) is selected as control variable to predict the case rate.
@@
 
-There are other variables related to mask such as "No legal enforcement of face mask mandate" and "Mandate face mask use by employees in public-facing businesses", it is assume these variables have a lower impact than the policy that is enforced to public.
-Also, it is observed that if a state enforce public mask policy, they will by default enforce mask policy for employees in public-facing business.
-As a result, mask policy for all individuals in public space can be used to represent mask-related policies.
+There are other variables related to mask such as "No legal enforcement of face mask mandate" and "Mandate face mask use by employees in public-facing businesses", we believe these variables have a lower impact than the policy that is enforced to public.
+Also, it is very likely that if a state enforce public mask policy, they will by default enforce mask policy for employees in public-facing business.
+As a result, we believe mask policy for all individuals in public space can be used to effectively represents the policy impact.
 
-It is worth to note that, the mask_public is in date format, which has value of 0 or a actual date when the policy is enforced.
-According to the documentation, 0 represents "the absence of an order or directive", which can be interpreted as the policy is not enforced by the state explicitly.
-For linear regression, the date values in mask_public column are transformed as value 1, so that   whether a state has any public mask policy or not can be distinguished.
-It is also noted that by transforming the variable, some important information will be lost because the actual date (early or late) to enforce the policy can also have an impact on case rate.
-However, it is difficult to measure the effect of the timing of implementation across different states.
-The enforcement of public mask policy will vary across states depending on other state related features such as political factors that are not included.
+It is worth to note that, the mask_public is in date format, which has value of zero or a actual date when the policy is enforced.
+According to the documentation, zero represents "the absence of an order or directive", which can be interpreted as the policy is not enforced by the state explicitly.
+For linear regression, the date values in mask_public column are transformed as value 1, so that we can distinguish whether states has public mask policy or not.
+It is also noted that by transforming the variable, some important information will be lost because the actual date (early or late) to enforce the policy can also have an impact to the case rate.
+However, it is difficult to measure the time effect as different states that enforced the public mask policy may depend on related situations or different thresholds.
 
-Firstly, the mask policy (mask_public_bool) variable is created from the original variable (mask_public), with 1 representing that the state has a pubic mask policy, and 0 means there's no explicit public mask policy from the state.
-The summary shows that there are 35 states that enforced public mask mandate, and 16 states that had no public mask policy at the time the dataset was compiled.
+Firstly, "mask_public_bool" column is created from "mask_public", with 1 represents that the state has a pubic mask policy, and 0 means there's no explicit public mask policy from the state.
+The summary shows that there are 35 states that enforced public mask mandate, and 16 states that had no public mask policy.
 The ratio looks reasonable for the analysis considering sample size of 51 as there is decent number of samples for each state policy type.
@@
 
-cat('Count of states that enforced mask mandate in public: ', 
+cat('Count for states that enforced mask mandate in public: ', 
     length(df$mask_public_bool[df$mask_public_bool==1]))
 
-cat('\nCount of states did not enforce mask mandate in public: ', 
+cat('\nCount for states did not enforced mask mandate in public: ', 
     length(df$mask_public_bool[df$mask_public_bool==0]))
@@
 
-The plot below shows the relationship between the mask policy and case rate.
-When comparing the states that did not implement a mask policy (value = 0) and those that did (value = 1), it can be seen that the data points for those that have no policy cluster around a higher case rate range, while the data points for the states with a mask policy cluster at a lower case rate range.
+Histograms below that shows the distribution of states that has public mask policy or not. Both distributions are not very normal with relatively small skews toward both tails.
+
+```{r, message = FALSE}
+
+hist(df$case_rate_100k[df$mask_public_bool == 1], xlab='Case Rate', ylab='Count', 
+     col='skyblue', main='Case Rate - States Enforced Mask Mandate in Public')
+hist(df$case_rate_100k[df$mask_public_bool == 0], xlab='Case Rate', ylab='Count', 
+     col='skyblue', main='Case Rate  - States Enformced No Mask Mandate in Public')
+```
+
+A plot that shows the relationship between different policies vs case rate.
 
@@
 df %>% 
-  ggplot(aes(x = mask_public_bool, y = case_rate_100k)) + 
-  geom_point(aes(size=5, alpha=0.2), show.legend = FALSE) + 
+  ggplot(aes(x = mask_public_bool, y = case_rate_100k, color = case_rate_100k)) + 
+  geom_point(aes(size=case_rate_100k)) + 
   labs(
@@
     x = 'Mask Mandate in Public', 
-    y = 'Case Rate per 100,000')  +
-  scale_x_continuous(breaks = seq(0,1,1), labels = seq(0,1,1))
+    y = 'Case Rate', 
+    color = 'Case Rate'
+  )
 ```
@@
   labs(
-    title = 'Histogram of Population Density per Square Mile', 
+    title = 'Distribution of Population Density per Square Miles', 
     x = 'Population density per square mile', y = 'Count')
@@
 
-As can be seen from the histogram, although most of the population density clusters in the 0 to 1500 range, there are some grouping of outliers that are very far from this concentration.
+As can be seen from the histogram, although most of the population density is concentrated in the 0 to 1500 range, there are some grouping of outliers that are very far from this concentration.
 When an analysis is performed, it can be seen that there is only one data sample that is the outlier, which is D.C.
@@
   labs(
-    title = 'Histogram of Population Density per Square Mile', 
-    x = 'Population density per square mile', y = 'Count') + 
-  scale_x_continuous(breaks = seq(0, 5, 1), labels = 10^(seq(0,5,1)))
+    title = 'Distribution of Population Density per Square Miles', 
+    x = 'Log of population density per square mile', y = 'Count')
 
@@
 Two additional variables discussed in Model 3 will be introduced as the maximalist approach.
-The first variable is the "mask_public_bool" (whether the state enforced public mask mandate or not) that represents the policy.
-The second variable introduced is "population_density" (the variable is transformed by the natural logarithm as discussed earlier) which is considered an important factor related to the spread of disease.
+The first variable is the "mask_public_bool" (whether the state enforced public mask mandate or not) that represents the policy effects.
+The second variable introduced is "population_density" (the variable is transformed with log of 10 as discussed earlier) which is considered an important factor related to the spread of disease.
 
@@
 And the result suggests that public mask mandate policy has positive effect in bringing the case number down.
-Specifically, the COVID-19 case rate can be reduced by 945 out of every 100k people if the state enforces public mask mandate keeping all other variables the same.
+In detail, the COVID-19 case rate can be reduced by 945 out of every 100k people if the state enforces public mask mandate keeping all other variables the same.
 
 The population density variable however is not significant in the regression even it is considered as one of the major factors to the spread of disease.
-According to the regression, there is a failure to reject the null hypothesis that population density is not related to the COVID-19 case rate.
+So according to the regression, we fail to reject the null hypothesis that population density have an direct impact to the COVID-19 case rate.
 There could be multiple factors causing the population density to be insignificant.
@@
 As a result the numbers can be diluted for states that have major cities with high population density, but very little urbanization in the majority part of the state suburb area.
-In addition, as stated earlier in subsection 2-3-1 of this report, Washington, D.C.
+In addition, as stated earlier in the research, Washington, D.C.
 is considered as a state however the demography structure and geographical related characteristics are different from other states.
-And due to the small sample size, individual variables that exhibit unique characteristics could lead to a certain degree of bias.
+And due to a small sample we have, individual variables that exhibit unique characteristics could lead to certain degree of bias.
 
-It was decided that the "population_density" variable should be kept in model 3 even it is insignificant for the following reasons.
-First, adding population density variable has little effect on the change of coefficients of control variables in model 2, which helps to demonstrate the robustness of the results.
-Second, there is no evidence that the variable effects the validation of CLM assumptions compared to model 2.
-Finally, the variable offers meaningful information that is relevant to the main research question.
+After careful discussion, the team still decided to keep "population_density" variable in model 3 even it is insignificant for the following reasons.
+First, adding population density variable has little effect on the change of coefficients of control variables in model 2.
+Second, we don't see evidence that the variable worsens the validation of CLM assumptions compared to model 2.
+Finally, the variable do offer meaningful information that is relevant to the main research question.
 
@@
 
-The residual versus fitted plot shows an almost flat line that is close to zero, with the tails a bit skewed.
-This implies that there are no strong linear relationships that are not captured beyond the current control variables.
+The Residual vs Fitted chat shows a almost flat line that is close to zero, with the tails a bit skewed.
+There is no strong evidence that there are unincoorperated linear relationship that we are not capturing from the control variables.
+```{r}
 
-```{r, message = FALSE}
-
-df<- df%>% 
-  mutate(
-    model3_preds = predict(model3), 
-    model3_resids = resid(model3)
-  ) 
-df %>% 
-  ggplot(aes(model3_preds, model3_resids)) + 
-  geom_point() + 
-  stat_smooth()
+plot(model3, which=1)
 ```
@@
 
-First, the list of variable coefficients shows that no variables were dropped, which means there is no perfect collinearity detected by the R function.
+Based on the Variance Inflation Factors check, all the variables have variance-inflation factors less than 4. As a result, there's no strong evidence of collinearity concern within the model. 
+```{r}
 
-```{r}
-model3$coefficients
-```
-
-Next, the variance inflation function was run in order to check for strong collinearity.
-Based on the variance inflation factors check, all the variables have variance inflation factors less than 4.
-As a result, there's no strong evidence of collinearity concern within the model.
-
-```{r}
-vif(model3)
+vif(model3)>4
 ```
@@
 
-According to the scale-location plot, the standardized residual falls into a reasonable range and the curve looks smooth without strong deviations.
-There is no significant evidence that the model violates the homoskedasticity assumption.
-For reference, the marked data points are New Hampshire (30), New York (33), and North Dakota (35).
-
+According to the plot of Homoskedasticity, the standardized residual falls into a reasonable range and the curve looks smooth without concerning skews.
+There is no significant evidence that the model violates homoskedasticity assumptions.
 ```{r}
@@
 
-```{r}
+```{r, echo=FALSE}
 bptest(model3)
@@
 If still a new geographic region with many states was added to the sample, and the measures of central tendency of the data continued to evolve, there would be a further indication that the underlying distribution is not identical.
-Without a possibility of refining the data sample however, as states would be immutable in this regard, the regression analysis proceeds with the recognition of possible violations of the IID assumption.
+Without a possibility of refining the data sample however, as states would be immutable in this regard, the assumption must be made for any potential analysis.
 
@@
 These effects may be included with a variable that demonstrates how much of a state's GDP is from tourism.
-It would be related to other economic indicator variables, such as the percent of the population living in poverty.
-While poverty is a broader phenomena, the number of people in poverty in the pandemic could be potentially higher due to negative growth of the tourism industry.
+It would be related to other economic indicator variables, such as the “real time” poverty rate during the pandemic.
+Although poverty is a broader phenomena, the number of people in poverty in the pandemic could be higher due to negative growth of the tourism industry.
 Therefore, it is expected that states with a greater focus on tourism may also have more people in poverty in the pandemic.
-This would take some of the explanatory power of the poverty variable away, resulting in its model coefficient becoming less positive.
+While poverty rate is included in the model, it is based off of data from 2018 before any pandemic related effects would have occurred.
+Assuming real time poverty was included in the model, a GDP percentage of tourism would take some of the explanatory power of the poverty variable away, resulting in its model coefficient becoming less positive.
 This would result in a positive omitted variable bias away from zero.
-Without a clear proxy in the dataset, and with the nature of this variable in touching on the IID assumption between interaction among states, it should be further investigated to understand how much of the original model may be an effect of omitted variable bias.
+Without a clear proxy in the dataset, and with the nature of this variable in touching on the IID assumption between interaction among states, it should be further investigated to understand a model including only poverty may be an effect of omitted variable bias.
+Again, because the model in this analysis uses 2018 poverty data, this would not be related to a bias from this omitted variable.
 
@@
 
-The goal of the analysis was to address the research question of how the COVID-19 case rate is related to the distribution of population demographics and mask-related policy of a state.
+The goal of the analysis was to address the research question to examine the relationship between both the distribution of population demographics and mask-related policy of a state and the COVID-19 case rate within that states.
 In order to address this question, the analysis was broken up into three phases of investigation with corresponding linear models.
 
- - Phase 1: To investigate the direct relationship between COVID-19 case rate and the senior rate within a state
+ - Phase 1: To investigate the direct relationship between COVID-19 case rate and the population distribution of seniors within a state
  - Phase 2: To build upon Phase 1 by including two additional variables of population demographics, poverty rate and unemployment rate, and investigating the overall effect on COVID-19 case rate
@@
 The data for population density was not statistically significant enough to suggest a relationship.
+Out of the four significant variables, the greatest effect was from the mask policy variable, followed by the unemployment rate, and then similar absolute effect sizes from the senior rate and povery rate.
 
-Starting with the variable with the largest coefficient magnitude, it can be seen that with the implementation of a mask policy in a state there are 945 fewer cases per 100,000 compared to a state without any implementation of such policy.
+Starting with the variable with the greatest effect size, it can be seen that the implementation of a mask policy in a state leads to a decrease of 945 cases per 100,000.
 Given that research from leading health organizations shows that the wearing of masks helps to decrease the chance of the transmission of COVID-19, it could be suggested that a mask policy is effective in increasing the adoption of wearing masks among the state population.
@@
 
-In conclusion, going back to our original question, it can be suggested within the data that the effect of a mask-related policy is something that should not be neglected in comparison to the population demographics.
+In conclusion, we see that although the population demographics had varying effects on the COVID-19 case rate, the greatest effect was seen from the mask policy.
+Going back to our original question, it can be suggested within the data that the effect of a mask-related policy is something that should not be neglected in comparison to the population demographics.
 Therefore, it seems that the COVID-19 case rate at the state-level is dependent on masks, and that ultimately, there are effective policy and control measures that can be taken by state governments that have some contribution towards combating its spread relative to factors that cannot be controlled.

### 2020-12-08 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Reverting changes + new tourism GDP percentage section

--- 2020-12-08 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
@@
 
-The economic standing of a state not only affects the individual households within its population, but may also reflect the economic impact to  their lives as a result of the pandemic.
-One way that a state may suffer from an increase in the number of COVID cases is from tourists or itinerant workers traveling during the pandemic and spreading the virus.
-A by-product of this is that residents of the state who work in the tourism industry may be in close contact with the travelers.
-In addition, if certain key industries in a state, such as tourism, laid off many workers, they would likely have to take lower paying service jobs in the gig economy that force them to be in close contact with others in order to continue to have a livelihood.
-This would further lead to a higher COVID case rate within the state.
+The economic standing of a state not only affects the individual households of its population, but also includes how it interacts with other states and how its peoples’ lives might have been economically changed by the pandemic.
+One way that a state may increase its number of COVID cases is both from tourists or others visiting during the pandemic and bringing disease, but also if its own population must work in close contact with others serving them in the tourism industry.
+In addition, if certain key industries to a state, such as tourism, layed off many workers, they would likely have to take lower paying close contact jobs such as in the service or gig economy in order to continue having a livelihood.
+This would further lead to a higher COVID case rate.
 
-These effects may be captured by a variable that demonstrates how much of a state's GDP is derived from tourism.
-It would be related to other economic indicator variables, such as the percent of the population living in poverty.
-While poverty is a broader phenomena, the number of people in poverty in the pandemic could be potentially higher due to negative growth of the tourism industry.
+These effects may be included with a variable that demonstrates how much of a state’s GDP is from tourism.
+It would be related to other economic indicator variables, such as the “real time” poverty rate during the pandemic.
+Although poverty is a broader phenomena, the number of people in poverty in the pandemic could be higher due to negative growth of the tourism industry.
 Therefore, it is expected that states with a greater focus on tourism may also have more people in poverty in the pandemic.
-This would take some of the explanatory power of the poverty variable away, resulting in its model coefficient becoming less positive.
+While poverty rate is included in the model, it is based off of data from 2018 before any pandemic related effects would have occurred.
+Assuming real time poverty was included in the model, a GDP percentage of tourism would take some of the explanatory power of the poverty variable away, resulting in its model coefficient becoming less positive.
 This would result in a positive omitted variable bias away from zero.
-Without a clear proxy in the dataset, and with the nature of this variable in touching on the IID assumption between interaction among states, it should be further investigated to understand how much of the original model may be an effect of omitted variable bias.
+Without a clear proxy in the dataset, and with the nature of this variable in touching on the IID assumption between interaction among states, it should be further investigated to understand a model including only poverty may be an effect of omitted variable bias.
+Again, because the model in this analysis uses 2018 poverty data, this would not be related to a bias from this omitted variable.
 

### 2020-12-08 Jackson_Liu_Temlock_Zhang_Lab2_Final_Report.Rmd
Updated RMD per meeting discussion and added new file for Final Report

--- 2020-12-08 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
@@
          unemployed_pct, senior_pct)
-
-head(df)
 ```
@@
 
-**Research Question:** How is the COVID-19 case rate related to the distribution of population demographics within a state, and how does the effect of these demographics compare with the effect of mask-related policy decisions made by that state?
+**Research Question:** How is the COVID-19 case rate related to the distribution of population demographics and policy within a state?
 
-For this report, the investigation will be centered around the effect of population demographics and policy on the COVID-19 case rates across the United States (US), which is grouped by the 50 states plus the District of Columbia (D.C.).
+For this report, the investigation will be centered around the relationship of population demographics and policy on the COVID-19 case rates across the United States (US), which is grouped by the 50 states plus the District of Columbia (D.C.).
 Within the report, this collection will be referred to as the "states", and each member a "state", inclusive of D.C..
@@
 The first and primary phase of investigation involves how the COVID-19 case rate is related to key population demographics.
-For this, the key variables are COVID-19 case rate per 100,000 residents and the percentage population distribution of seniors (defined as 65 years old or older in the original dataset) within a state.
+For this, the key variables are COVID-19 case rate per 100,000 residents and the senior rate (defined as 65 years old or older in the original dataset) within a state.
 The case rate per 100,000 was selected as the key output/dependent variable as it provides a standardized measure of the spread of COVID-19 across the states, and already takes into account the absolute population of the states.
@@
 Thus, population density is measured as a conduit to indicate the level of proximity within each state and will operationalize the concept of social distancing.
-It should be noted however that the population density data may be limited in its ability to serve as a direct measurement of social distancing given the it lacks other contextualize information, such as the urban to rural ratio as later discussed in section 5.
+It should be noted however that the population density data may be limited in its ability to serve as a direct measurement of social distancing given it lacks other contextualize information, such as the urban to rural ratio as later discussed in section 5.
 
@@
 Among others, the loss of information on the date of implementation within the mask policy variable strips out any contextual knowledge regarding the length of policy implementation, and precludes the identification of how long a specific policy was implemented.
-Assuming that this has some effect on case rate, this difference in length of time may play a part in the effect.
+Assuming that this has some relationship with case rate, this difference in length of time may play a part.
 There are also other external aspects that cannot be included, such as the rural to urban ratio as opposed to population density.
@@
 Finally, with regard to the policy variables, this report focuses on a form of mask mandate.
-Given this, it does not capture the effect of other policies that may or may not have a greater effect on the case rate (e.g., implementation of stay-at-home orders, closure of non-essential businesses, etc.).
+Given this, it does not capture other policies that may or may not have other relationships with the case rate (e.g., implementation of stay-at-home orders, closure of non-essential businesses, etc.).
 In justification, these choices were largely made due to there either being an appropriate amount of samples in each category (e.g., most states have implemented basic policies such as the closure of non-essential businesses and the implementation of stay-at-home orders, and thus these were not included given the lack of samples for the states that have not implemented them), or that data on other policies were simply not readily available in the dataset.
@@
 
-In order to allow for more simplistic and direct interpretations when analyzing the effect of the input variables on the output, it was determined that all percentage variables (senior, poverty, and unemployment) would be transformed into rates per 100,000 to align with the case rate variable.
+In order to allow for more simplistic and direct interpretations when analyzing the relationships of the input variables on the output, it was determined that all percentage variables (senior, poverty, and unemployment) would be transformed into rates per 100,000 to align with the case rate variable.
 It will also serve to ease the burden on the readers of this report as like-for-like scaled relationships can be drawn between the variables.
@@
 
-For the first model, the relationship between the COVID-19 case rate per 100,000 and the population percentage distribution of seniors was analyzed.
+For the first model, the relationship between the COVID-19 case rate per 100,000 and senior rate was analyzed.
 
@@
 For model 3, policy related variables will be introduced for the analytics.
-The policy variable "Mandate face mask use by all individuals in public spaces" (renamed as mask_public) is selected as control variable to predict the case rate.
-Given it has been proven in external research that masks are effective in preventing the transmission of COVID-19, it is expected that the mask policy mandate in public would have an effect in reducing the case rate.
+The policy variable "Mandate face mask use by all individuals in public spaces" (renamed as mask_public) is selected as an input variable in model 3.
+Given it has been proven in external research that masks are effective in preventing the transmission of COVID-19, it is expected that the mask policy mandate in public is related to a decrease in case rate.
 
@@
 df %>%
-  ggplot(aes(x = factor(mask_public_bool), y = case_rate_100k)) + 
-  geom_point(aes(size=5, alpha=0.2), show.legend = FALSE) + 
-  labs(
-    title = 'Case Rate vs Mask Public Mandate', 
-    x = 'Mask Mandate in Public', 
-    y = 'Case Rate per 100,000')  +
-  scale_x_discrete(labels=c("0","1"))
-
-df %>%
   ggplot(mapping = aes(factor(mask_public_bool), case_rate_100k))+
@@
     x = 'Mask Mandate in Public', 
-    y = 'Case Rate'
+    y = 'Case Rate per 100,000'
   ) +
-  scale_x_discrete(labels=c("No","Yes"))
-
-df %>%
-  ggplot(aes(case_rate_100k))+
-  geom_histogram(aes(y = ..density.., fill = factor(mask_public_bool)), position = 'identity', alpha = 0.5, color = 'white')+
-  stat_density(geom = 'line', size=1, position = 'identity', aes(color = factor(mask_public_bool)))+
-  scale_fill_brewer(palette = 'Dark2') +
-  scale_color_brewer(palette = 'Dark2') +
-  labs(
-    title = 'Distribution of COVID-19 Case Rate',
-    x = 'Covid-19 case rate per 100,000'
-  ) +
-  scale_fill_discrete(name='Mask public policy',labels=c('No', 'Yes'))+
-  scale_color_discrete(name='Mask public policy',labels=c('No', 'Yes'))
+  scale_x_discrete(labels=c("0","1"))
 ```
@@
  - Phase 1: To investigate the direct relationship between COVID-19 case rate and the senior rate within a state
- - Phase 2: To build upon Phase 1 by including two additional variables of population demographics, poverty rate and unemployment rate, and investigating the overall effect on COVID-19 case rate
- - Phase 3: To build upon Phase 2 by adding a variable to measure the effect of the implementation of a mask-policy on COVID-19 case rate.
+ - Phase 2: To build upon Phase 1 by including two additional variables of population demographics, poverty rate and unemployment rate, and investigating the overall relationship on COVID-19 case rate
+ - Phase 3: To build upon Phase 2 by adding a variable to measure the relationship of the implementation of a mask-policy on COVID-19 case rate.
 A variable for population density was also added as an additional measure of population demographics
@@
 Following unemployment, when examining the poverty variable there is actually an increase of roughly 1 case per 100,000 for every 5 additional people per 100,000 living under the federal poverty line.
-Although there is some intuition behind the unemployment and poverty rates having similar effects on case rate, it can be seen from the test of collinearity in subsection 2-2-3 that these two variables are not explaining the same effects.
+Although there is some intuition behind the unemployment and poverty rates having similar relationships with case rate, it can be seen from the test of collinearity in subsection 2-2-3 that these two variables are not explaining the same effects.
 This difference is likely due to the fact that those living below the poverty line have less access to suitable living conditions and sanitary goods that help to alleviate the possibility of COVID-19 infection.
 As a caveat, it should be noted that these the poverty and unemployment figures are from a 2018 survey, so it does not reflect the full picture of the state demographics at the time of when the COVID-19 related data was captured.
-Finally, the effect of the senior rate is negative, with an additional 4 seniors per 100,000 translating to roughly 1 fewer case per 100,000.
+Finally, the coefficient of the senior rate is negative, with an additional 4 seniors per 100,000 translating to roughly 1 fewer case per 100,000.
 As touched upon in the introduction, this could be largely down to the fact that it is widely broadcasted that seniors fall within the high risk category in terms of having severe complications from COVID-19, and therefore they are more diligent in the precautions they take against contracting the virus.

### 2020-12-09 Jackson_Liu_Temlock_Zhang_Lab2_Final_Report.Rmd
Final updates to report

@@
 knitr::opts_chunk$set(echo = TRUE)
-install.packages("GGally")
 library(lmtest)
@@
 library(dplyr)
-library(GGally)
 library(car)
@@
 
+Based on the results of the data exploration above, the following equation is used to generate the regression analysis for model 3.
+
+$$
+\begin{aligned}
+ case\_rate\_100k =\beta_0 &+\beta_1senior\_per\_100k +\beta_2poverty\_per\_100k +\beta_3unemployed\_per\_100k + \beta_4mask\_public\_bool + \beta_5log(population\_density)
+\end{aligned}
+$$
+
 The regression coefficient shows that mask_public_bool variable is highly significant.
@@
 
-#### i. I.I.D Assumption
+#### i. IID Assumption
 
@@
 
+Setting aside the uncertainty provided that the sample is qualified for the IID condition, all three of the models generally meet the CLM assumptions.
+Therefore, the coefficients generated from the regression analyses are unbiased estimates of the actual relationship investigated in this report.
+The uncertainty associated with these coefficients is also unbiased.
+Moreover, it is plausible to apply t-tests based on classical standard errors for these regression coefficients.
+Thus, the discussions and conclusions drawn from these regression results are assumed to be credible.
+However, if the dataset was more granular (e.g., case rate at the county-level rather than at the state-level), the IID assumption would be better satisfied.
+It would lead to greater independence for each data point relative to the rest of the data set.
+It would also take more advantage of the unbiased estimators so far demonstrated in the model, as a higher number of data points would produce more accurate values for the model coefficients.
+Overall, improving the source data would lead to greater reliability in the regression work and conclusions.
+
 # 4. Regression Table
@@
 One way that a state may increase its number of COVID cases is both from tourists or others visiting during the pandemic and bringing disease, but also if its own population must work in close contact with others serving them in the tourism industry.
-In addition, if certain key industries to a state, such as tourism, layed off many workers, they would likely have to take lower paying close contact jobs such as in the service or gig economy in order to continue having a livelihood.
+In addition, if certain key industries to a state, such as tourism, laid off many workers, they would likely have to take lower paying close-contact jobs such as in the service industry or gig economy in order to continue to have a livelihood.
 This would further lead to a higher COVID case rate.
@@
 This would result in a positive omitted variable bias away from zero.
-Without a clear proxy in the dataset, and with the nature of this variable in touching on the IID assumption between interaction among states, it should be further investigated to understand a model including only poverty may be an effect of omitted variable bias.
+Without a clear proxy in the dataset, and with the nature of this variable touching on the IID assumption between interaction among states, it should be further investigated whether a model including only poverty may be affected by omitted variable bias.
 Again, because the model in this analysis uses 2018 poverty data, this would not be related to a bias from this omitted variable.

### 2020-12-10 Jackson_Liu_Temlock_Zhang_Lab2_Final_Report.Rmd
Added titles to the LCE plots

@@
   geom_point() + 
-  stat_smooth()
+  stat_smooth() +
+  labs(
+    title = 'Model 1 Predicted vs Residuals', 
+    x = 'Residuals', 
+    y = 'Predicted'
+  )
 ```
@@
   geom_point() + 
-  stat_smooth()
+  stat_smooth() +
+  labs(
+    title = 'Model 2 Predicted vs Residuals', 
+    x = 'Residuals', 
+    y = 'Predicted'
+  )
 ```
@@
   geom_point() + 
-  stat_smooth()
+  stat_smooth() + 
+  labs(
+    title = 'Model 3 Predicted vs Residuals', 
+    x = 'Residuals', 
+    y = 'Predicted'
+  )
 ```
